### 2018-02-27 doc/manuscript.tex
Started write-up

%---------------------------------------------
% This document is for pdflatex
%---------------------------------------------
\documentclass[11pt]{article}

\usepackage{amsmath,amsfonts,amssymb,graphicx,setspace,authblk}
\usepackage{float}
\usepackage[running]{lineno}
\usepackage[vmargin=1in,hmargin=1in]{geometry}

\usepackage[authoryear]{natbib}

\graphicspath{ {./../genetic-diversity/figures/} }

\usepackage{enumitem}
\setlist{topsep=.125em,itemsep=-0.15em,leftmargin=0.75cm}

\usepackage{gensymb}

\usepackage[compact]{titlesec}

\usepackage{bm,mathrsfs}

\usepackage{ifpdf}
\ifpdf
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{epstopdf}
\else
\DeclareGraphicsExtensions{.eps}
\fi

\renewcommand{\floatpagefraction}{0.98}
\renewcommand{\topfraction}{0.99}
\renewcommand{\textfraction}{0.05}

\clubpenalty = 10000
\widowpenalty = 10000

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%% Just for commenting
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[usenames]{color}
\newcommand{\new}{\textcolor{red}}
\newcommand{\spe}{\textcolor{blue}}
\newcommand{\comment}{\textcolor{black}}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{equation} \begin{aligned}}
\newcommand{\ea}{\end{aligned} \end{equation}}

\def\X{\mathbf{X}}

\floatstyle{boxed}
\newfloat{Box}{tbph}{box}

\title{Integrating spatial and temporal information to make ecological forecasts }

\author[1]{Peter B. Adler\thanks{Corresponding author. Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah Email: peter.adler@usu.edu}}
\author[2]{Ethan White}
\author[3]{others?}
\affil[1]{Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah}
\affil[2]{some shitty Florida joint}
\affil[3]{??}

\renewcommand\Authands{ and }

% \date{Last compile: \today}

\sloppy

\renewcommand{\baselinestretch}{1.25}

\begin{document}

\maketitle

\linenumbers

\section*{Abstract}

stuff

\textbf{\large{Keywords:}} Coexistence, competition, integral projection model, removal experiment, sagebrush steppe.

\section*{Introduction}

Cheer lead for forecasting yada yada

Strengths and weaknesses of spatial approach

Strengths of weaknesses of temporal approach

Let's take the best of both worlds! Combine with weights that change over time. We might learn something interesting from the rate at which the weights change.

\section*{Eco-evo case study}

Consider an annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig. \ref{fig:RxnNorms}).

\begin{figure}[tbp]
\centering
\includegraphics[width=0.6 \textwidth]{reactionnorms.png}
\caption{Reaction norms of the three genotypes.}
\label{fig:RxnNorms}
\end{figure}

Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-loving genotype, warm sites will be dominated by the heat-loving genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig. \ref{fig:SpatialPattern}).

\begin{figure}[tbp]
\centering
\includegraphics[width=0.75 \textwidth]{spatial_pattern.png}
\caption{The spatial pattern of individual genotypes and total population abundances at sites arrayed across a gradient of mean annual temperature. The dashed line shows predictions from the ``spatial model."}
\label{fig:SpatialPattern}
\end{figure}

%\newpage
%\renewcommand{\refname}{Literature cited}
%\bibliographystyle{Ecology}
%\bibliography{RemovalRefs}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% APPENDICES !
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%\clearpage 
%\newpage 
%
%\setcounter{page}{1}
%\setcounter{equation}{0}
%\setcounter{figure}{0}
%\setcounter{section}{0}
%\setcounter{table}{0}
%
%\centerline{\Large \textbf{Appendices}}
%\centerline{Adler et al., ``Weak interspecific interactions''} 
%
%\vspace{0.4in} 
%
%\renewcommand{\theequation}{A-\arabic{equation}}
%\renewcommand{\thetable}{A-\arabic{table}}
%\renewcommand{\thefigure}{A-\arabic{figure}}
%\renewcommand{\thesection}{\Alph{section}}

\end{document}

### 2018-02-27 doc/manuscript.tex
Small additions to manuscript

@@
 
-Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-loving genotype, warm sites will be dominated by the heat-loving genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
+Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-loving homozygous genotype, warm sites will be dominated by the heat-loving homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
 \ref{fig:SpatialPattern}).
@@
 \end{figure}
+
+This spatial pattern is the outcome of steady-state conditions.
+At any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm.
+For example, at a cold site dominated by the cold-loving homozygous genotype, a warmer than normal year would cause a decrease in population size, even though the heat-loving homozygote might perform optimally at that temperature.
+If warmer than normal conditions persist for many years, then genotype frequencies will shift, and the heat-loving homozygote will compensate for the decreases of the cold-loving genotype.
 

### 2018-02-28 doc/manuscript.tex
Combined all figures in one file, finish eco-evo example

--- 2018-02-27 doc/manuscript.tex
@@
 
-\title{Integrating spatial and temporal information to make ecological forecasts }
+\title{Integrating spatial and temporal patterns in ecological forecasts }
 
-\author[1]{Peter B. Adler\thanks{Corresponding author. Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah Email: peter.adler@usu.edu}}
+\author[1]{Peter B. Adler}  %\thanks{Corresponding author. Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah Email: peter.adler@usu.edu}}
 \author[2]{Ethan White}
@@
 
-\textbf{\large{Keywords:}} Coexistence, competition, integral projection model, removal experiment, sagebrush steppe.
+%\textbf{\large{Keywords:}} Coexistence, competition, integral projection model, removal experiment, sagebrush steppe.
 
@@
 
-Cheer lead for forecasting yada yada
+We love ecological forecasting yada yada.
+Common approaches use spatial patterns, temporal patterns, and process-level understanding (i.e., mechanistic models).
+Since we will never know all the mechanisms, we will inevitably need to rely on patterns.
+But they each have limitations.
 
-Strengths and weaknesses of spatial approach
+Strengths and weaknesses of spatial approach...
 
-Strengths of weaknesses of temporal approach
+Strengths of weaknesses of temporal approach...
 
-Let's take the best of both worlds! Combine with weights that change over time. We might learn something interesting from the rate at which the weights change.
+Let's take the best of both worlds! Combine spatial and temporal models with weights that change over time. We might learn something interesting from the rate at which the weights change...
 
-\section*{Eco-evo case study}
+\section*{Eco-evo example}
 
-Consider an annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig. \ref{fig:RxnNorms}).
+Consider a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
+\ref{fig:spatial_model}A).
+Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
+\ref{fig:spatial_model}B).
 
@@
 \centering
-\includegraphics[width=0.6 \textwidth]{reactionnorms.png}
-\caption{Reaction norms of the three genotypes.}
-\label{fig:RxnNorms}
+\includegraphics[width=1 \textwidth]{spatial_model.png}
+\caption{(A) Reaction norms of the three genotypes.
+(B) The spatial pattern of individual genotypes and total population abundances at sites arrayed across a gradient of mean annual temperature.
+The dashed line shows predictions from the ``spatial model." }
+\label{fig:spatial_model}
 \end{figure}
 
-Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-loving genotype, warm sites will be dominated by the heat-loving genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
-\ref{fig:SpatialPattern}).
+This spatial pattern is the outcome of steady-state conditions.
+But any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm.
+For example, at a cold site dominated by the cold-adapted homozygous genotype, a warmer than normal year would cause a decrease in population size, even though the heat-adapted homozygote might perform optimally at that temperature.
+We refer to this as the temporal pattern.
+However, if warmer than normal conditions persist for many years, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
+
+To demonstrate these dynamics, we simulated a population at a colder than average site.
+During the baseline period, the population is dominated by the cold-adapted genotype, and we fit a temporal model that predicts population growth as a function of annual temperature and population size (Fig.
+\ref{fig:temporal_model}).
+We then impose a period of warming, followed by a final period of higher, but stationary, temperature (Fig.
+\ref{fig:forecast} top).
+With the onset of warming, the population crashes as the cold-adapted genotype decreases in abundance.
+Eventually, frequencies of the heterozygous genotype and the warm-adapted homozygous genotype begin to increases and the population recovers (Fig.
+\ref{fig:forecast} bottom).
 
@@
 \centering
-\includegraphics[width=0.75 \textwidth]{spatial_pattern.png}
-\caption{The spatial pattern of individual genotypes and total population abundances at sites arrayed across a gradient of mean annual temperature.
-The dashed line shows predictions from the ``spatial model."}
-\label{fig:SpatialPattern}
+\includegraphics[width=0.6 \textwidth] {temporal_model.png}
+\caption{The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
+Colors show population size, which also influences the population growth rate through density dependence.
+}
+\label{fig:temporal_model}
+\end{figure}
+
+The implications for forecasting are clear: In the initial stages of the warming trend, a temporal model fit to the original, baseline conditions should make better the best predictions about resulting population trends, but in the longer term, as evolutionary change occurs, a spatial model fit will begin to make better predictions.
+To combine the spatial and temporal model into a single forecast, we fit a simple weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
+At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$.
+The temporal model accurately predicts the impact of the initial warming trend, but eventually becomes far too pessimistic, while the spatial model does not handle the initial trend but accurately predicts the eventual, new steady state (Fig.
+\ref{fig:forecast} bottom).
+The combined model initially reflects the temporal model, but then rapidly transitions to reflect the spatial model.
+The rapid transition in the weighting term, $\omega$, occurs during the period of most rapid change in genotype frequencies (Fig.
+\ref{fig:forecast_supp}).
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {forecast.png}
+\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
+(Bottom) Simulated population size and forecasts from the spatial, temporal and combined models.
+}
+\label{fig:forecast}
+\end{figure}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {forecast_supplement.png}
+\caption{Simulated shifts in genotype abundances, and the model weighting term, $\omega$, during the warming phase and the following stationary temperature phase.}
+\label{fig:forecast_supp}
 \end{figure}

### 2018-02-28 genetic-diversity/figures_spatial.R

@@ -1,27 +1,0 @@
-
-# call this script from genetic_diversity_master.R
-
-myCols<-c("blue","darkgrey","red","black")
-
-# plot fecundity as function of temperature for both phenotypes
-png("figures/reactionnorms.png",height=3,width=4,res=300,units="in")
-  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
-  temp=seq(-8,8,0.01)
-  react_norms=getLambdas(temp,fec_Tmu,fec_Tsigma,fec_max)
-  matplot(temp,react_norms,type="l",xlab="Temperature",ylab="Germination rate",col=myCols,lty="solid")
-  text(fec_Tmu[1],0.9*fec_max[1],"AA",col="blue")
-  text(fec_Tmu[2],0.9*fec_max[2],"Aa",col="grey")
-  text(fec_Tmu[3],0.9*fec_max[3],"aa",col="red")
-dev.off()
-
-# plot mean abundances vs mean temperature, along with spatial model "prediction"
-xx <- seq(-5,5,0.01)
-yy <- coef(spatial_model)[1] + coef(spatial_model)[2]*xx + coef(spatial_model)[3]*xx^2
-png("figures/spatial_pattern.png",height=3.5,width=5,units="in",res=400)
-  par(mfrow=c(1,1),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
-  matplot(Tmean,simN_mix[,c(2:5)],xlab="Mean temperature",ylab="N",
-            type="l", col=myCols,
-            lty=c("solid"),lwd=c(1,1,1,2))
-  lines(xx,yy,lty="dashed",lwd=1)
-  #legend("topright",c("AA","Aa","aa","Pop"),fill=myCols,bty="n")
-dev.off()

### 2018-02-28 genetic-diversity/figures_temporal.R

@@ -1,31 +1,0 @@
-
-# show key features of temporal model
-png("figures/temporal_pattern.png",height=4,width=7.5,res=400,units="in")
-  par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
-  # plot temperature vs population growth during baseline period
-  plot(temperature_t1_baseline,r_baseline,xlab="Temperature",ylab="log Population growth rate")
-  # plot observed vs predicted population growth
-  plot(r_baseline,predict(temporal_model),xlab="Observed", ylab= "Predicted")
-dev.off()
-
-
-# plot time series with spatial and temporal predictions!
-png("figures/forecast.png",height=5,width=5,res=400,units="in")
-  
-  layout(matrix(c(1,2),2,1),heights=c(0.25,0.75),widths=1)
-  
-  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(0,4,1,1))
-  plot(temperature,xlab="",ylab="Temperature",type="l",xaxt="n",col="darkgrey")
-  abline(v=baseline_yrs,col="black",lty=3)
-  lines((baseline_yrs+1):sim_yrs,forecast_Tmean,lwd=2,col="black")
-  
-  par(mar=c(3,4,1,1))
-  plot(N,xlab="Time",ylab="N",type="l",col="darkgrey")
-  abline(v=baseline_yrs,col="black",lty=3)
-  lines(1:length(N),spatial_forecast,col="red",lwd=2)
-  lines(1:length(N),temporal_forecast,col="blue",lwd=2)
-  lines(1:length(N),combined_forecast,col="purple",lwd=2,lty=2)
-  legend("bottomleft",c("Observed","Spatial forecast","Temporal forecast","Combined forecast"),
-         col=c("darkgrey","red","blue","purple"),lwd=c(1,2,2,2),lty=c(1,1,1,2),bty="n")
-
-dev.off()

### 2018-02-28 genetic-diversity/genetic_diversity_figures.R

@@ -1,0 +1,81 @@
+
+# call this script from genetic_diversity_master.R
+
+myCols<-c("blue","darkgrey","red","black")
+
+# figure to show spatial relationships
+png("figures/spatial_model.png",height=3.5,width=7,res=400,units="in")
+  
+  par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,2,1))
+  
+  # plot fecundity as function of temperature for all genotypes
+  temp=seq(-7,7,0.01)
+  react_norms=getLambdas(temp,fec_Tmu,fec_Tsigma,fec_max)
+  matplot(temp,react_norms,type="l",xlab="Temperature",ylab="Germination rate",col=myCols,lty="solid")
+  text(fec_Tmu[1],0.9*fec_max[1],"AA",col=myCols[1])
+  text(fec_Tmu[2],0.9*fec_max[2],"Aa",col=myCols[2])
+  text(fec_Tmu[3],0.9*fec_max[3],"aa",col=myCols[3])
+  mtext("(A)",side=3, line=0.5, adj=0)
+  
+  # plot mean abundances vs mean temperature, along with spatial model "prediction"
+  xx <- seq(-5,5,0.01)
+  yy <- coef(spatial_model)[1] + coef(spatial_model)[2]*xx + coef(spatial_model)[3]*xx^2
+  matplot(simN_mix$Tmean,simN_mix[,c(2:5)],xlab="Mean temperature",ylab="N",
+            type="l", col=myCols,
+            lty=c("solid"),lwd=c(1,1,1,2))
+  lines(xx,yy,lty="dashed",lwd=1)
+  #legend("topright",c("AA","Aa","aa","Pop"),fill=myCols,bty="n")
+  mtext("(B)",side=3, line=0.5, adj=0)
+
+dev.off()
+
+# plot temporal pattern: temperature vs population growth during baseline period
+png("figures/temporal_model.png",height=3.5,width=4,res=400,units="in")
+  
+  mytheme <- list(axis.components = list(left = list(tck=0.5), right = list(tck=0.5),
+                                         top = list(tck=0.5), bottom = list(tck=0.5)))
+  trellis.par.set(mytheme)
+  
+  print(levelplot(N_t0_baseline ~ temperature_t1_baseline + r_baseline, 
[…]

### 2018-02-28 genetic-diversity/genetic_diversity_master.R

@@ -8,6 +8,7 @@ setwd("C:/Repos/space-time-forecast/genetic-diversity")
 # load packages and functions
 source("genetic_diversity_functions.R")
 library("boot")
+library("latticeExtra")
 
 ###
 ### 1. Simulate and describe spatial pattern in abundances at sites across a range of mean temperatures
@@ -34,9 +35,6 @@ alpha = matrix(1,3,3)      # all competition coefficients (intra- and inter-) ar
 source("genetic_diversity_sim_spatial.R")
 
 #CHECK: why zeros for least fit genotype's mean abundances at lowest temps but not highest?
-
-# plot spatial figures
-source("figures_spatial.R")
 
 ###
 ### 2. Simulate and describe temporal pattern in abundances at one site experiencing warming
@@ -69,4 +67,4 @@ source("genetic_diversity_sim_temporal.R")
 source("genetic_diversity_forecast.R")
 
 # plot temporal figures
-source("figures_temporal.R")
+source("genetic_diversity_figures.R")

### 2018-02-28 genetic-diversity/genetic_diversity_sim_temporal.R

@@ -22,6 +22,7 @@ for(iYr in 2:(sim_yrs)){
 
 # Fit a "temporal" model describing abundance as a function of annual temperature
 # during the baseline period. Assume Gompertz population growth.
+N_genotypes <- N
 N <-rowSums(N)
 N_t0_baseline <- N[burn_in:(baseline_yrs-1)]
 N_t1_baseline <- N[(burn_in+1):baseline_yrs]

### 2018-02-28 doc/manuscript.tex
Manuscript edits

@@
 \author[2]{Ethan White}
-\author[3]{others?}
+\author[3]{Michael Cortez?}
+\author[4]{others?}
 \affil[1]{Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah}
 \affil[2]{some shitty Florida joint}
-\affil[3]{??}
 
@@
 Since we will never know all the mechanisms, we will inevitably need to rely on patterns.
-But they each have limitations.
+But they have limitations.
 
-Strengths and weaknesses of spatial approach...
+The spatial approach: Correlate mean environment with the mean of an ecological state variable or rate across many sites.
+Examples: population distribution or abundance as a function of climate (SDM refs), primary production as a function of mean precipitation (Lauenroth refs).
+Strengths: broad spatial scales, data are available and analyses are straightforward.
+Weaknesses: many well-recognized issues involving assumptions.
+A less well-recognized issue is that these models are not dynamic; they provide no information about how quickly the system will move from the current state to the predicted, future state.
+These patterns have developed over very long time spans, so they are unlikely to provide much information about dynamic changes on the order of years to decades.
+In addition, transient dynamics could prevent the system from ever reaching the predicted steady state (Urban ref?).
 
-Strengths of weaknesses of temporal approach...
+The temporal approach: Correlate interannual variation in an ecological response with interannual variation in the environment, typically at just one site (but see Kleinhesselink  and Adler 2018, others).
+Examples: population or vtial rate fluctuations as a function of weather, primary production as a function of precipitation (Lauenroth and Sala 1992).
+Strengths: focuses on temporal dynamics, can predict rates of change, easily validated.
+Weaknesses: typically limited spatial extent, ignores slower processes, such as evolutionary adaptation, or turnover in community composition, that could alter the underlying process at longer time scales.
 
-Let's take the best of both worlds! Combine spatial and temporal models with weights that change over time. We might learn something interesting from the rate at which the weights change...
+Our goal is to capitalize on the complementary strengths of the spatial and temporal approaches.
+The idea is to take predictions from a spatial and a temporal model and combine them with a weighted average.
+The model weights can change over time: we expect the temporal model to dominate in the short-term, and the spatial model to dominate in the long-term.
+How rapidly that transition occurs depends on the rate of those ``slow" processes such as evolution and species turnover.
+We might learn something interesting by studying the rate at which the model weights change.
 
@@
 Consider a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
-\ref{fig:spatial_model}A).
+\ref{fig:spatial_model}A; the full model description is provided in the Appendix, and all computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast}).
 Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
@@
 (B) The spatial pattern of individual genotypes and total population abundances at sites arrayed across a gradient of mean annual temperature.
-The dashed line shows predictions from the ``spatial model." }
+The dashed line shows predictions from an empirical ``spatial model," a linear regression that describes mean population size as a function of mean temperature.
+}
 \label{fig:spatial_model}
@@
 This spatial pattern is the outcome of steady-state conditions.
-But any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm.
-For example, at a cold site dominated by the cold-adapted homozygous genotype, a warmer than normal year would cause a decrease in population size, even though the heat-adapted homozygote might perform optimally at that temperature.
-We refer to this as the temporal pattern.
+But at any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm.
+For example, at a cold site dominated by the cold-adapted homozygous genotype, a warmer than average year would cause a decrease in population size, even though the heat-adapted homozygote might perform optimally at that same temperature.
 However, if warmer than normal conditions persist for many years, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
 
-To demonstrate these dynamics, we simulated a population at a colder than average site.
-During the baseline period, the population is dominated by the cold-adapted genotype, and we fit a temporal model that predicts population growth as a function of annual temperature and population size (Fig.
+To demonstrate these dynamics, we simulated a diploid annual plant population at a colder than average site.
+During the baseline period, the population is dominated by the cold-adapted genotype.
+We use the simulated data from this baseline period to fit an empirical model that assume no knowledge of the true underlying process.
+This empirical ``temporal model" predicts population growth rate as a function of annual temperature and population size (Fig.
 \ref{fig:temporal_model}).
-We then impose a period of warming, followed by a final period of higher, but stationary, temperature (Fig.
+We then imposed a period of warming, followed by a final period of higher, but stationary, temperature (Fig.
 \ref{fig:forecast} top).
@@
 
-The implications for forecasting are clear: In the initial stages of the warming trend, a temporal model fit to the original, baseline conditions should make better the best predictions about resulting population trends, but in the longer term, as evolutionary change occurs, a spatial model fit will begin to make better predictions.
-To combine the spatial and temporal model into a single forecast, we fit a simple weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$.
+The implications for forecasting are clear: In the initial stages of the warming trend, the temporal model fit to the original, baseline conditions should make better the best predictions about resulting population trends, but in the longer term, as evolutionary change occurs, the spatial model will begin to make better predictions.
+To combine the spatial and temporal model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
+At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix for full description of the approach).
 The temporal model accurately predicts the impact of the initial warming trend, but eventually becomes far too pessimistic, while the spatial model does not handle the initial trend but accurately predicts the eventual, new steady state (Fig.
@@
 
+Next steps? Play with parameters to show what controls the rate of transition from temporal to spatial model?
+
 %\newpage
@@
 
-%\clearpage 
-%\newpage 
-%
-%\setcounter{page}{1}
-%\setcounter{equation}{0}
-%\setcounter{figure}{0}
-%\setcounter{section}{0}
-%\setcounter{table}{0}
-%
-%\centerline{\Large \textbf{Appendices}}
-%\centerline{Adler et al., ``Weak interspecific interactions''} 
-%
-%\vspace{0.4in} 
-%
-%\renewcommand{\theequation}{A-\arabic{equation}}
-%\renewcommand{\thetable}{A-\arabic{table}}
-%\renewcommand{\thefigure}{A-\arabic{figure}}
-%\renewcommand{\thesection}{\Alph{section}}
+\clearpage 
+\newpage
+
+\setcounter{page}{1}
+\setcounter{equation}{0}
+\setcounter{figure}{0}
+\setcounter{section}{0}
+\setcounter{table}{0}
+
+\centerline{\Large \textbf{Appendices}}
+
+\vspace{0.4in}
+
+\renewcommand{\theequation}{A-\arabic{equation}}
+\renewcommand{\thetable}{A-\arabic{table}}
+\renewcommand{\thefigure}{A-\arabic{figure}}
+\renewcommand{\thesection}{\Alph{section}}
+
+\section*{Description of the eco-evolutionary annual plant model}
+
+Insert Cortez text
+
+\section*{Estimating model weights}
+
+Here is my home brew approach, presented in my typically terrible notation. There must be a better, more formal approach...Help?
+
+The spatial model, which we refer to as $S$, is a linear regression fit to ``observed" data on mean population size, $\bar{N}$ at a series of locations, $x$, which vary in mean temperature, $K$.
+In our case, the observed data are simulated by the model described in the previous section.
+We include a quadratic term to capture the unimodal relationship between  $\bar{N}$ and $K$:
+ \begin{equation}
+ \bar{N}_x = S(K_x) = \beta^S_0 +  \beta^S_1 K_x +\beta^S_2 {K_x}^2 + \varepsilon
+ \label{eqn:spatial_regression}
+ \end{equation}
+
+The temporal model, which we call $T$, starts with a time series of ``observed" population sizes, $\vec{N}$.
+Using data from a baseline period with a stationary distribution of annual temperature at one location, we fit the following regression, which predicts the (log) population growth rate as a function of population size and annual temperature at time $t$:
+ \begin{equation}
+ \ln(\frac{N_{t+1}}{N_t}) = T(N_t,K_t) = \beta^T_0 +  \beta^S_1 ln(N_t) +\beta^S_2 K_t + beta^S_3 {K_t}^2 +  \varepsilon
+ \label{eqn:temporal_regression}
+ \end{equation}
+
+The temporal model returns a per capita growth rate on the log scale.
+To predict population size at the next time step, we simply exponentiate the growth rate and multiply it by the current population size: $exp(T(N_t,K_t)) N_t$.
+
+The combined model is a weighted average of predictions from the spatial and temporal models, with the weights changing as a function of time, here expressed as the forecast horizon.
+The weights change as a function of the square root of the forecast horizon, to allow rapid shifts in the model weights.
+\begin{equation}
+logit(\omega_t)=\beta^C_0 + \beta^C_1 \sqrt{t}
+\label{eqn:weights}
+\end{equation}
+\begin{equation}
+\hat{N}_{t+1} = \omega * exp(T(N_{t},K_t)) N_t + (1-\omega) * S(K_t) 
+\label{eqn:combined_model}
+\end{equation}
+
+I used the \texttt{optim} function to estimate the $\beta^C$'s that minimize the sum of squared errors, $(\hat{N}_{t+1} - N_{t+1})^2$
 

### 2018-02-28 doc/manuscript.tex
Added model description appendix from Cortez

@@
 
-Insert Cortez text
+(from Michael Cortez)
+
+\noindent \textbf{Haploid Model:} Begin with a haploid model that describes the number of seeds present in a seed bank.  $N_{i,t}$ is the number of seeds of species $i$ at time $t$.  The model is
+\begin{align}\begin{split}
+N_{1,t+1} &= s_1 [1-g_1(E_t)]N_{1,t}+\frac{\lambda_1g_1(E_t)N_{1,t}}{1+ \alpha_{11}g_1(E_t)N_{1,t} + \alpha_{12}g_2(E_t)N_{2,t}}\\
+N_{2,t+1} &= s_2 [1-g_2(E_t)]N_{2,t}+\frac{\lambda_2g_2(E_t)N_{2,t}}{1+ \alpha_{21}g_1(E_t)N_{1,t} + \alpha_{22}g_2(E_t)N_{2,t}}
+\end{split}\end{align}
+where $g_{i}(E_t)$ is the probability of germination, $E_t$ is the environmental quality at time $t$, $s_i$ is the seed survival probability for species $i$, and $\lambda_i$ is the seed production rate per plant.
+Below I refer to the $\alpha_{ij}$ as intra- and inter-specific competition coefficients.
+\\
+
+\noindent \textbf{Diploid Model:} Consider a one-species diploid model.
+The genotypes are denoted by $AA$, $Aa$, and $aa$.
+The number of each genotypes at time $t$ is $N_{AA,t}$, $N_{Aa,t}$, and $N_{aa,t}$.
+The germination rates for each genotype are $g_{AA}(E_t)$, $g_{Aa}(E_t)$, and $g_{aa}(E_t)$.
+The seed survival probability and seed production rate for genotype $AA$ are $s_{AA}$ and $\lambda_{AA}$, respectively.
+The analogous parameters for the other genotypes are similarly denoted.
+The competition coefficients are denoted by $\alpha_{i,j}$, e.g., $\alpha_{AA,AA}$ or $\alpha_{AA,Aa}$.
+Throughout we assume that gametes mix randomly in the population.
+
+First consider the case where the competition coefficients are zero ($\alpha_{i,j}=0$).  Let $T$ denote the total number of gamete-pairs produced in a given year,
+\begin{equation}
+T = \lambda_{AA}N_{AA,t}g_{AA}(E_t)+ \lambda_{Aa}N_{Aa,t}g_{Aa}(E_t)+\lambda_{aa}N_{aa,t}g_{aa}(E_t).
+\end{equation}
+The first term is the number of gamete-pairs produced by $AA$ individuals.
+The second and third terms are the numbers of gamete-pairs produced by $Aa$ and $aa$ individuals, respectively.
+The proportion of $A$ gametes ($\phi_A$) and the proportion of $a$ gametes ($\phi_a$) are given by
+\begin{align}\begin{split}
+\phi_{A} &= \frac{\lambda_{AA}N_{AA,t}g_{AA}(E_t)+ \frac{1}{2}\lambda_{Aa}N_{Aa,t}g_{Aa}(E_t)}{T} \hspace{10pt} \text{and} \hspace{10pt} \phi_a = 1-\phi_{A}.
+\end{split}\end{align}
+Note that the $T$ in the denominator of $\phi_A$ shows up because we are computing proportions.  Combining all of these we get the dynamics for each genotype,
+\begin{align}\begin{split}
+N_{AA,t+1} &= s_{AA}[1-g_{AA}(E_t)]N_{AA,t} + \phi_A^2T\\
+N_{Aa,t+1} &= s_{Aa}[1-g_{Aa}(E_t)]N_{Aa,t} + \phi_A\phi_aT\\
+N_{aa,t+1} &= s_{aa}[1-g_{aa}(E_t)]N_{aa,t} + \phi_a^2T
+\end{split}\end{align}
+
+Now consider the case where the competition coefficients are non-zero ($\alpha_{i,j}\neq0$).
+Including competition changes the way in which we compute $T$, $\phi_A$, and $\phi_a$.
+Specifically, because the total number of seeds produced per year by each genotypes is reduced based on intra- and inter-specific competition, the total number of gamete-pairs becomes
+\begin{align}\begin{split}
+T &=  \frac{\lambda_{AA}N_{AA,t}g_{AA}(E_t)}{1+ \alpha_{AA,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{AA,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{AA,aa}g_{aa}(E_t)N_{aa,t}} \\ 
+&+ \frac{\lambda_{Aa}N_{Aa,t}g_{Aa}(E_t)}{1+ \alpha_{Aa,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{Aa,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{Aa,aa}g_{aa}(E_t)N_{aa,t}}\\
+&+\frac{\lambda_{aa}N_{aa,t}g_{aa}(E_t)}{1+ \alpha_{aa,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{aa,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{aa,aa}g_{aa}(E_t)N_{aa,t}}.
+\label{eqn:defineT}
+\end{split}\end{align}
+The first line is the number of gamete-pairs produced by $AA$ individuals after accounting for the effects of competition.
+The second and third lines are the numbers of gamete-pairs produced by $Aa$ and $aa$ individuals, respectively.
+The proportions of $A$ gametes and $a$ gametes are
+\begin{align}\begin{split}
+\phi_A &= \frac{1}{T}\frac{\lambda_{AA}N_{AA,t}g_{AA}(E_t)}{1+ \alpha_{AA,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{AA,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{AA,aa}g_{aa}(E_t)N_{aa,t}} \\
+&+ \frac{1}{2T}\frac{\lambda_{Aa}N_{Aa,t}g_{Aa}(E_t)}{1+ \alpha_{Aa,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{Aa,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{Aa,aa}g_{aa}(E_t)N_{aa,t}}\\
+\phi_a &= 1- \phi_A
+\label{eqn:definePhi}
+\end{split}\end{align}
+Combining all of this results in the same model as above,
+\begin{align}\begin{split}
+N_{AA,t+1} &= s_{AA}[1-g_{AA}(E_t)]N_{AA,t} + \phi_A^2T\\
+N_{Aa,t+1} &= s_{Aa}[1-g_{Aa}(E_t)]N_{Aa,t} + 2 \phi_A\phi_aT\\
+N_{aa,t+1} &= s_{aa}[1-g_{aa}(E_t)]N_{aa,t} + \phi_a^2T,
+\end{split}\end{align}
+but the definitions of $T$, $\phi_A$, and $\phi_a$ are given by equations (\ref{eqn:defineT}) and (\ref{eqn:definePhi}) .
 

### 2018-07-02 doc/manuscript.tex
Started community example write-up

@@
 
-\graphicspath{ {./../genetic-diversity/figures/} }
+\graphicspath{ {./../genetic-diversity/figures/} {./../community-turnover/figures/} }
 
@@
 
+\subsection{Community turnover example}
+
+In the previous example, evolutionary change was the slow process whose rate determined the shift in weight from the temporal to spatial forecast of the state of a population.
+The process of compositional turnover--change in the abundances and identities of species interacting at a site--can have a similar effect at the community level.
+We demonstrate this using a metacommunity model developed by Alexander et al.
+(ref).
 %\newpage

### 2018-07-02 doc/manuscript.tex
Writing up community example, intermediate save

@@
 
-\title{Integrating spatial and temporal patterns in ecological forecasts }
+\title{\textbf{Integrating spatial and temporal patterns in ecological forecasts }}
 
@@
 
-\subsection{Community turnover example}
+\subsection*{Community turnover example}
 
@@
 (ref).
+The model describes Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
+Composition varies along the gradient because of a trade-off between growth rate and cold tolerance: cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by fast-growing species that are cold intolerant.
+Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics.
+Sites are linked by dispersal: a specified fraction of offspring of each species leaves the site where they were produced and reaches all other sites with equal probability.
+
+As in our eco-evolutionary example, we simulate a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase and then a final period of stationary temperature.
+Interannual variation in temperature is the same at all sites, but mean temperature varies among sites.
+All sites experienced the same increase in mean temperature.
+We focus on two response variables: 1) total community biomass and 2) the biomass of one focal species that dominates the baseline period at the site in the middle of the gradient.
+
+Dynamics during the baseline period reveal very strong spatial patterns across the mean temperature gradient contrasted by weak responses to interannual variation in temperature at one site.
+Individual species show classic ``Whittaker" patterns of abundances across the gradient (Fig.
+\ref{fig:community-composition}).
+Time-averaged total community biomass increases strongly with temperature, because the fast-growing species at the warm sites produce more biomass, while the average biomass of our focal species shows a classic unimodal pattern across the temperature gradient (Fig.
+\ref{fig:community-baseline}).
+These spatial patterns are the basis for our "spatial models" of the temperature-biomass relationship.
+However, at our focal site in the center of the gradient, both total biomass and the biomass of the focal species are quite insensitive to interannnual variation in temperature (Fig.
+\ref{fig:community-baseline}).
+Our "temporal models" estimate this weak temperature effect, along with strong effects of biomass in the previous year (e.g., lags).
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {mean_biomass_spp_by_site.png}
+\caption{Mean biomass by species (colors) across the temperature gradient during the baseline period. The focal species, dominant at the site in the center of the gradient, is shown in dark blue.  }
+\label{fig:community-composition}
+\end{figure}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_models.png}
+\caption{Time-averaged spatial relationships between temperature and both total community biomass and the focal species' biomass are shown in blue, while interannual variations in total and focal species biomass at one site in the center of the gradient are shown in red.
+Symbols show simulated values, lines show predictions from statistical models fit to the simulated data.
+}
+\label{fig:community-baseline}
+\end{figure}
+
+We can forecast the effect of temperature increase on total biomass and the focal species' biomass using the spatial models and the temporal models.
+Once again, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time.
+For total biomass, the spatial and temporal models both predict immediate increases in biomass with warming, because all species respond positively to temperature increases.
+However, over time, as faster growing species increase in abundance at the focal site, biomass continues to increase, approaching the predictions of the spatial model (Fig.
+\ref{fig:community-forecasts} middle).
+The predictions of the spatial and temporal model for focal species biomass contrast more dramatically, with the temporal model predicting a large increase in biomass and the spatial model predicting a decrease.
+Initially, the simulated abundances follow the increase predicted by the temporal model, but as faster-growing species increase in abundance, the biomass of the focal species then decreases, eventually falling below its baseline level (Fig.
+\ref{fig:community-forecasts} bottom).
+For total biomass (the sum across all species), the weights for the combined model shift rapidly to favor the spatial model, whereas for focal species biomass, the weights shift more slowly towards the spatial model, as the biomass of the focal species decreases and its competitors' biomass increase (Fig.
+\ref{fig:community-weights}).
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_forecast.png}
+\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
+(Middle) Simulated total biomass and forecasts from the spatial, temporal and combined models.
+(Bottom) Simulated focal species' biomass and forecasts from the spatial, temporal and combined models.
+}
+\label{fig:community-forecasts}
+\end{figure}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_change_plus_weights.png}
+\caption{Simulated changes in biomass of the focal species (black) and all other species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue) and focal species biomass (red).
+}
+\label{fig:community-weights}
+\end{figure}
+
 %\newpage

### 2018-07-02 community-turnover/comm_turn_figures.R

@@ -8,7 +8,7 @@ myCols<-alpha(c("blue3","red3"),0.4)
 ###
 
 # plot spatial and temporal models
-png("figures/spatial_temporal_models.png",height=3.5,width=4,res=400,units="in")
+png("figures/community_models.png",height=3.5,width=4,res=400,units="in")
 
   par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
   
@@ -50,7 +50,7 @@ my_obs_temp <- temperature[(burnin_yrs+1):sim_yrs,site]
 my_biomass <- rowSums(spxp[(burnin_yrs+1):sim_yrs,site,])
 my_biomass_spp <- spxp[(burnin_yrs+1):sim_yrs,site,my_species]
 
-png("figures/forecast.png",height=7.5,width=4,res=400,units="in")
+png("figures/community_forecast.png",height=7.5,width=4,res=400,units="in")
   
   layout(matrix(c(1,2,3),3,1),heights=c(0.1,0.4,0.4),widths=1)
   

### 2018-07-02 doc/manuscript.tex
Writing up community example, starting Appendix B

@@
 \usepackage[vmargin=1in,hmargin=1in]{geometry}
+\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
 
@@
 
-\title{\textbf{Integrating spatial and temporal patterns in ecological forecasts }}
+\title{\textbf{Integrating spatial and temporal patterns \\ in ecological forecasts }}
 
@@
 
-\section*{Description of the eco-evolutionary annual plant model}
+\section{Description of the eco-evolutionary annual plant model}
 
@@
 
-\section*{Estimating model weights}
+\subsection{Estimating model weights}
 
@@
 
+\section{Description of the metacommunity model}
+
+[How much detail to include from Alexander et al?]
+
+-Describe spatial model
+
+-Describe temporal model
+
+-Model weights
+
 \end{document}

### 2018-07-03 doc/manuscript.tex
Finished writing up community example

@@
 \author[2]{Ethan White}
-\author[3]{Michael Cortez?}
-\author[4]{others?}
+\author[1]{Michael Cortez?}
+\author[3]{Heather Lynch?}
 \affil[1]{Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah}
@@
 
-In the previous example, evolutionary change was the slow process whose rate determined the shift in weight from the temporal to spatial forecast of the state of a population.
+In the previous example, evolutionary change was the slow process that determined how quickly the combined model shifted from a temporal to spatial forecast.
 The process of compositional turnover--change in the abundances and identities of species interacting at a site--can have a similar effect at the community level.
@@
  \begin{equation}
- \ln(\frac{N_{t+1}}{N_t}) = T(N_t,K_t) = \beta^T_0 +  \beta^S_1 ln(N_t) +\beta^S_2 K_t + beta^S_3 {K_t}^2 +  \varepsilon
+ \ln(\frac{N_{t+1}}{N_t}) = T(N_t,K_t) = \beta^T_0 +  \beta^T_1 ln(N_t) +\beta^T_2 K_t + beta^T_3 {K_t}^2 +  \varepsilon
  \label{eqn:temporal_regression}
@@
 
--Describe spatial model
-
--Describe temporal model
-
--Model weights
+The spatial model for total community biomass is a simple linear regression.
+The spatial model for the biomass of the focal species is a multiple regression with a quadratic term, identical to Eqn.
+\ref{eqn:spatial_regression}.
+The temporal model for total biomass, $M$, is:
+ \begin{equation}
+ \ln(M_{t+1} = T(M_{t},K_t) = \beta^T_0 +  \beta^T_1 M_t +\beta^T_2 K_t  +  \varepsilon .
+ \label{eqn:community_temporal_regression}
+ \end{equation}
+The temporal model for focal species biomass has the same functional form.
+The combined models are based on the same approach that we used in the eco-evolutionary example (eqn.
+\ref{eqn:combined_model}).
 

### 2019-04-22 doc/manuscript.tex
writing intro

@@
 
-\title{\textbf{Integrating spatial and temporal patterns \\ in ecological forecasts }}
+\title{\textbf{ Short and long-term forecasts will require models incorporating different processes and covariates }}
 
@@
 
-We love ecological forecasting yada yada.
-Common approaches use spatial patterns, temporal patterns, and process-level understanding (i.e., mechanistic models).
-Since we will never know all the mechanisms, we will inevitably need to rely on patterns.
-But they have limitations.
+Recently, increased interest in ecological forecasting
+Primary goal is applied: inform decisions (Clark 2001, Dietze et al. 2017, Mouquet et al. 2017)
+Important secondary goal is to advance basic research (Houlahan et al. 2017)
+Emerging consensus that we need to do this
+
+How to do it is not so clear
+Lots of different approaches
+Lots of argument about the virtues of process-based models and phenomenological, or correlative models (refs?)
+We often see basic research justified with the argument that improved process-level understanding will improve predictions, in fact we have made this
+argument ourselves (cite our own stuff?)
+It is easy to find examples of how extrapolation of linear relationships gets us in trouble (examples? refs?)
+But in many cases simple phenomenological models make surprisingly good predictions (examples? refs?) and they can avoid the parameter error often associated
+with complex process-based models
+
+Pragmatism is the way forward: use the models that make the best predictions
+This requires making testable predictions
+Thus (Dietze et al 2017) emphasis on near-term forecasting (emphasize that we are on board?)
+Longer-term forecasts harder to test, hindcasts may be best bet (Williams refs? Blois?)
+
+What current literature and the argument for near term forecasting has not explicitly stated is that different models, incorporating different kinds of processes will
+work better for short vs long-term forecasts. We demonstrate this using simulation models. Our analyses show that: 
+1. For short-term forecasts, phenomenological time series approaches are hard to beat
+2. Long-term forecasts require understanding slow processes such as evolutionary and ecological selection as well as dispersal
+3. Different kinds of data reflect the operation of different processes:
+	longitudinal data capture autocorrelation and responses of current assemblage;
+	data across spatial gradients capture long-term outcome of selection and dispersal;
+	as the time-scale of the forecast increases, the importance of these processes shifts.
+4. A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.
+
+\section*{Modeling approach}
+
+Simulate effects of environmental change on a system with known dynamics
+Assume fully mechanistic, process-based model is impossible
+Fit two phenomenological models, each representing processes operating at different time scales, and make forecasts with them
 
@@
 
-Our goal is to capitalize on the complementary strengths of the spatial and temporal approaches.
-The idea is to take predictions from a spatial and a temporal model and combine them with a weighted average.
+We show that responses to environmental change reflect both kinds of processes but at different time scales.
+We demonstrate that  by taking predictions from a spatial and a temporal model and combine them with a weighted average.
 The model weights can change over time: we expect the temporal model to dominate in the short-term, and the spatial model to dominate in the long-term.
 How rapidly that transition occurs depends on the rate of those ``slow" processes such as evolution and species turnover.
-We might learn something interesting by studying the rate at which the model weights change.
 

### 2019-04-22 doc/manuscript.tex
finished intro 1st draft

--- 2018-07-03 doc/manuscript.tex
@@
 
-\title{\textbf{Integrating spatial and temporal patterns \\ in ecological forecasts }}
+\title{\textbf{ Forecast time-scale determines which processes to model }}
 
@@
 
-%\textbf{\large{Keywords:}} Coexistence, competition, integral projection model, removal experiment, sagebrush steppe.
+\textbf{\large{Keywords:}} dispersal, ecological forecasting, eco-evolutionary dynamics, global change, selection
 
@@
 
-We love ecological forecasting yada yada.
-Common approaches use spatial patterns, temporal patterns, and process-level understanding (i.e., mechanistic models).
-Since we will never know all the mechanisms, we will inevitably need to rely on patterns.
-But they have limitations.
+A recent flurry of synthesis papers in high impact journals attests to surging interest in ecological forecasting. Many of the papers 
+argue that a commitment to ecological forecasting is necessary to guide the environmental policy and management
+decisions that will become increasingly critical under global change (Clark 2001, Mouquet et al. 2017, Dietze et al. 2018).
+A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery (Houlahan et al. 2017, Dietz 2017, Dietz et al. 2018).
+Taken together, these papers, along with a new textbook (Dietze 2017) represent an emerging consensus that forecasting should be a priority for ecological research.
 
-The spatial approach: Correlate mean environment with the mean of an ecological state variable or rate across many sites.
-Examples: population distribution or abundance as a function of climate (SDM refs), primary production as a function of mean precipitation (Lauenroth refs).
-Strengths: broad spatial scales, data are available and analyses are straightforward.
-Weaknesses: many well-recognized issues involving assumptions.
-A less well-recognized issue is that these models are not dynamic; they provide no information about how quickly the system will move from the current state to the predicted, future state.
-These patterns have developed over very long time spans, so they are unlikely to provide much information about dynamic changes on the order of years to decades.
-In addition, transient dynamics could prevent the system from ever reaching the predicted steady state (Urban ref?).
+No such consensus is available for exactly how to tackle this challenge.
+Ecologists rely on many different approaches and models for making predictions, and 
+frequently argue about the virtues of process-based vs. phenomenological models (refs?).
+We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g., Adler et al. 2012),
+and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions (refs?).
+But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
+uncertainty (refs? Tredennick et al. 2017?).
 
-The temporal approach: Correlate interannual variation in an ecological response with interannual variation in the environment, typically at just one site (but see Kleinhesselink  and Adler 2018, others).
-Examples: population or vtial rate fluctuations as a function of weather, primary production as a function of precipitation (Lauenroth and Sala 1992).
-Strengths: focuses on temporal dynamics, can predict rates of change, easily validated.
-Weaknesses: typically limited spatial extent, ignores slower processes, such as evolutionary adaptation, or turnover in community composition, that could alter the underlying process at longer time scales.
+The pragmatic solution to these debates is clear: we should forecast using the models that make the best predictions.
+Identifying these models requires rigorously testing predictions against observations. However, many common ecological predictions,
+such as a species' geographic range in the year 2100, are virtually impossible to test (but see paleo hindcasting refs,Williams refs? Blois?), 
+leading Dietze et al (2017) to call for greater emphasis on near-term forecasting. The assumption is that building 
+our capacity to make short-term forecasts will ultimately improve our long-term forecasts.
 
-Our goal is to capitalize on the complementary strengths of the spatial and temporal approaches.
-The idea is to take predictions from a spatial and a temporal model and combine them with a weighted average.
-The model weights can change over time: we expect the temporal model to dominate in the short-term, and the spatial model to dominate in the long-term.
-How rapidly that transition occurs depends on the rate of those ``slow" processes such as evolution and species turnover.
-We might learn something interesting by studying the rate at which the model weights change.
+Our goal is to consider a hypothesis that, in our view, has not been carefully considered by either the push for near-term forecasting or
+ongoing debates about competing modeling approaches. We expect that short and long-term forecasts will require models
+incorporating different kinds of processes and fit to different kinds of data sets. We demonstrate this idea using two simulation case studies, the first for an 
+eco-evolutionary scenario, the second concerned with feedbacks between community composition and ecosystem functioning.  
+Our analyses show that: 
+1. For short-term forecasts, phenomenological time-series approaches are hard to beat.
+2. Long-term forecasts require understanding slow processes such as evolutionary and ecological selection as well as dispersal.
+3. Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while 
+data spanning spatial gradients capture the long-term outcome of selection and dispersal.  As the time-scale of the forecast increases, the importance of these processes shifts.
+4. A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.
 
-\section*{Eco-evo example}
+\section*{Modeling approach}
+
+In both of our case studies, we simulate the effects of a hypothetical temperature increase on a simple system with known dynamics.
+The ``truth'' is represented by a model that is fully mechanistic for at least one important process, but we assume that recovering 
+this true, process-based model would be impossible in practice (we return to this point in the Discussion). We begin each simulation
+under stationary temperature, allowing the system to equilibrate; we call this the baseline phase. We then increase temperature progressively
+over a period of time, followed by a second period of stationary, now elevated, temperature. The objective is to forecast the response of the 
+system to temperature increase based on data gathered during the baseline period.
+
+We make two forecasts based on two different phenomenological models, each representing processes operating at different time scales. 
+The first phenomenological model represents what we call a "spatial approach." We correlate the mean environment with the mean of an ecological state or rate 
+ across many sites. This is the approach commonly used to predict population distribution or abundance as a function of climate (SDM refs) or
+ primary production as a function of mean precipitation (Lauenroth refs). The approach is attractive because it addresses broad spatial scales relevant to 
+ management, and because the necessary data are widely available and statistical analyses are straightforward. The approach also has well-known weaknesses.
+Criticism often focuses on the assumption that current spatial patterns are in equilibrium with climate (refs?). A less well-recognized issue is that these models are not dynamic; 
+they provide no information about how quickly the system will move from the current state to the predicted, future state. Because the patterns in the data have developed 
+over long time spans, they are unlikely to provide much information about dynamic changes on the order of years to decades. Moreover, transient dynamics could 
+prevent the system from ever reaching the predicted steady state (Urban ref?).
+
+The second model represents what we call a ``temporal approach.'' We correlate interannual variation in an ecological response with interannual variation in the environment, 
+typically at just one site. This approach is often used to study population or vital rate fluctuations as a function of weather, 
+or primary production as a function of precipitation (Lauenroth and Sala 1992). The advantages of this approach include is focus on temporal dynamics, meaning it 
+can predict rates of change, and validation of the resulting near-term predictions are straightforward. However, models built using this approach typically cover 
+a limited spatial extent (but see Kleinhesselink  and Adler 2018, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
+that could alter the underlying process at longer time scales.
+
+We compare forecasts from both models to the true, simulated dynamics to show that responses to environmental change reflect both kinds of processes, but at different time scales.
+To quantify this pattern, we combine forecasts from the spatial and temporal approaches with a weighted average and then show how the optimal model weights change over time: we expect the temporal model to best predict short-term dynamics, and the spatial model to best predict long-term dynamics.
+How rapidly that transition occurs depends on the rate of  ``slow" processes such as evolution and species turnover.
+
+\section*{Eco-evolutionary example}
 

### 2019-04-25 doc/manuscript.tex
more ms additions

@@
 In both of our case studies, we simulate the effects of a hypothetical temperature increase on a simple system with known dynamics.
-The ``truth'' is represented by a model that is fully mechanistic for at least one important process, but we assume that recovering 
+The ``truth'' is represented by a model that is mechanistic for at least one important process, but we assume that recovering 
 this true, process-based model would be impossible in practice (we return to this point in the Discussion). We begin each simulation
@@
 We make two forecasts based on two different phenomenological models, each representing processes operating at different time scales. 
-The first phenomenological model represents what we call a "spatial approach." We correlate the mean environment with the mean of an ecological state or rate 
+The first phenomenological model represents what we call a ``spatial approach.'' We correlate the mean environment with the mean of an ecological state or rate 
  across many sites. This is the approach commonly used to predict population distribution or abundance as a function of climate (SDM refs) or
@@
  management, and because the necessary data are widely available and statistical analyses are straightforward. The approach also has well-known weaknesses.
-Criticism often focuses on the assumption that current spatial patterns are in equilibrium with climate (refs?). A less well-recognized issue is that these models are not dynamic; 
+Criticism often focuses on the assumption that current spatial patterns are in equilibrium with climate (refs?). Another issue is that these models are not dynamic; 
 they provide no information about how quickly the system will move from the current state to the predicted, future state. Because the patterns in the data have developed 
@@
 
-Consider a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
+Evolutionary adaptation is a key uncertainty in predicting how environmental change will impact a focal population at a given location (refs).
+To explore how this process influences dynamics and forecasts at different time scales, we consider a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
 \ref{fig:spatial_model}A; the full model description is provided in the Appendix, and all computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast}).
@@
 
-This spatial pattern is the outcome of steady-state conditions.
+The spatial pattern shown in Fig.
+\ref{fig:spatial_model} is the outcome of steady-state conditions.
 But at any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm.
@@
 During the baseline period, the population is dominated by the cold-adapted genotype.
-We use the simulated data from this baseline period to fit an empirical model that assume no knowledge of the true underlying process.
+We use the simulated data from this baseline period to fit an empirical model that assumes no knowledge of the underlying eco-evolutionary process.
 This empirical ``temporal model" predicts population growth rate as a function of annual temperature and population size (Fig.
@@
 With the onset of warming, the population crashes as the cold-adapted genotype decreases in abundance.
-Eventually, frequencies of the heterozygous genotype and the warm-adapted homozygous genotype begin to increases and the population recovers (Fig.
+Eventually, frequencies of the heterozygous genotype and the warm-adapted homozygous genotype begin to increase and the population recovers (Fig.
 \ref{fig:forecast} bottom).
@@
 
-The implications for forecasting are clear: In the initial stages of the warming trend, the temporal model fit to the original, baseline conditions should make better the best predictions about resulting population trends, but in the longer term, as evolutionary change occurs, the spatial model will begin to make better predictions.
+The implications for forecasting are clear: In the initial stages of the warming trend, the temporal model fit to the original, baseline conditions should make better the best predictions 
+about resulting population trends, but in the longer term, as evolutionary change occurs, the spatial model ultimately makes better predictions. 
 To combine the spatial and temporal model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
@@
 \end{figure}
+
+The combined model's predictions look impressively accurate, but that is because we used the full, simulated time series to fit the weighting term. 
+Unlike our spatial and temporal model forecasts, we did not generate out-of-sample predictions from the combined model; it merely provides 
+a convenient way to quantify how rapidly dynamics shift from from being dominated by interannual 
+variation to being dominated by steady-state equilibrium. To generate a true forecast from the combined model, we would need an independent 
+method to determine how the model weights shift over time.
 

### 2019-04-25 doc/manuscript.tex
even more ms additions

@@
 In the previous example, evolutionary change was the slow process that determined how quickly the combined model shifted from a temporal to spatial forecast.
-The process of compositional turnover--change in the abundances and identities of species interacting at a site--can have a similar effect at the community level.
+The process of compositional turnover--change in the abundances and identities of species interacting at a site--can have a similar effect at the level of communities and even ecosystems.
 We demonstrate this using a metacommunity model developed by Alexander et al.
@@
 All sites experienced the same increase in mean temperature.
-We focus on two response variables: 1) total community biomass and 2) the biomass of one focal species that dominates the baseline period at the site in the middle of the gradient.
+We focus on two response variables at one site in the middle of the gradient.
+: 1) total community biomass and 2) the biomass of one focal species that dominates the baseline period
 
@@
 \ref{fig:community-baseline}).
-These spatial patterns are the basis for our "spatial models" of the temperature-biomass relationship.
+These spatial patterns are the basis for our ``spatial models" of the temperature-biomass relationship.
 However, at our focal site in the center of the gradient, both total biomass and the biomass of the focal species are quite insensitive to interannnual variation in temperature (Fig.
 \ref{fig:community-baseline}).
-Our "temporal models" estimate this weak temperature effect, along with strong effects of biomass in the previous year (e.g., lags).
+Our ``temporal models" estimate this weak temperature effect, along with strong effects of biomass in the previous year (e.g., lags).
 
@@
 We can forecast the effect of temperature increase on total biomass and the focal species' biomass using the spatial models and the temporal models.
-Once again, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time.
 For total biomass, the spatial and temporal models both predict immediate increases in biomass with warming, because all species respond positively to temperature increases.
@@
 \ref{fig:community-forecasts} bottom).
+Once again, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time.
 For total biomass (the sum across all species), the weights for the combined model shift rapidly to favor the spatial model, whereas for focal species biomass, the weights shift more slowly towards the spatial model, as the biomass of the focal species decreases and its competitors' biomass increase (Fig.

### 2019-05-07 doc/manuscript.tex
Finished very rough draft

@@
 
-\title{\textbf{ Forecast time-scale determines which processes to model }}
+\title{\textbf{ Matching forecast time-scale with ecological process }}
 
@@
 incorporating different kinds of processes and fit to different kinds of data sets. We demonstrate this idea using two simulation case studies, the first for an 
-eco-evolutionary scenario, the second concerned with feedbacks between community composition and ecosystem functioning.  
-Our analyses show that: 
-1. For short-term forecasts, phenomenological time-series approaches are hard to beat.
-2. Long-term forecasts require understanding slow processes such as evolutionary and ecological selection as well as dispersal.
-3. Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while 
-data spanning spatial gradients capture the long-term outcome of selection and dispersal.  As the time-scale of the forecast increases, the importance of these processes shifts.
-4. A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.
+eco-evolutionary scenario, the second concerned with feedbacks between community composition and ecosystem functioning. Our analyses show that: 
+\begin{enumerate}
+	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow 
+	processes such as evolutionary and ecological selection as well as dispersal.
+	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while 
+	data spanning spatial gradients capture the long-term outcome of selection and dispersal.  Whether predictive models should be trained using longitudinal or
+	spatial data sets, or both, depends on the time-scale of the desired forecast.
+	\item A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.
+\end{enumerate}
 
@@
 \ref{fig:forecast} top).
-With the onset of warming, the population crashes as the cold-adapted genotype decreases in abundance.
-Eventually, frequencies of the heterozygous genotype and the warm-adapted homozygous genotype begin to increase and the population recovers (Fig.
+With the onset of warming, the population crashed as the cold-adapted genotype decreased in abundance.
+Eventually, frequencies of the heterozygous genotype and the warm-adapted homozygous genotype began to increase and the population recovered (Fig.
 \ref{fig:forecast} bottom).
@@
 At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix for full description of the approach).
-The temporal model accurately predicts the impact of the initial warming trend, but eventually becomes far too pessimistic, while the spatial model does not handle the initial trend but accurately predicts the eventual, new steady state (Fig.
+The temporal model accurately predicted the impact of the initial warming trend, but eventually became far too pessimistic, while the spatial model did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
 \ref{fig:forecast} bottom).
-The combined model initially reflects the temporal model, but then rapidly transitions to reflect the spatial model.
-The rapid transition in the weighting term, $\omega$, occurs during the period of most rapid change in genotype frequencies (Fig.
+The combined model initially reflected the temporal model, but then rapidly transitioned to reflect the spatial model.
+The rapid transition in the weighting term, $\omega$, occured during the period of most rapid change in genotype frequencies (Fig.
 \ref{fig:forecast_supp}).
@@
 variation to being dominated by steady-state equilibrium. To generate a true forecast from the combined model, we would need an independent 
-method to determine how the model weights shift over time.
+method to predict how the model weights shift over time.
 
@@
 In the previous example, evolutionary change was the slow process that determined how quickly the combined model shifted from a temporal to spatial forecast. 
-The process of compositional turnover--change in the abundances and identities of species interacting at a site--can have a similar effect at the level of communities and even ecosystems.
-We demonstrate this using a metacommunity model developed by Alexander et al.
-(ref).
-The model describes Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
-Composition varies along the gradient because of a trade-off between growth rate and cold tolerance: cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by fast-growing species that are cold intolerant.
-Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics.
-Sites are linked by dispersal: a specified fraction of offspring of each species leaves the site where they were produced and reaches all other sites with equal probability.
+The process of compositional turnover--change in the abundances and identities of species interacting at a site--can have a similar effect at the level of 
+communities and even ecosystems. We demonstrate this using a metacommunity model developed by Alexander et al. (ref). The model describes Lotka-Volterra 
+competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient. Composition varies along the gradient because of a 
+trade-off between growth rate and cold tolerance: cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by 
+fast-growing species that are cold intolerant. Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics. 
+Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
 
-As in our eco-evolutionary example, we simulate a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase and then a final period of stationary temperature.
-Interannual variation in temperature is the same at all sites, but mean temperature varies among sites.
-All sites experienced the same increase in mean temperature.
-We focus on two response variables at one site in the middle of the gradient.
-: 1) total community biomass and 2) the biomass of one focal species that dominates the baseline period
+As in our eco-evolutionary example, we simulated a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase, and then a 
+final period of stationary temperature. Interannual variation in temperature is the same at all sites, but mean temperature varies among sites. All sites experienced the 
+same absolute increase in mean temperature. We focused on two response variables at one site in the middle of the gradient: 1) total community biomass and 
+2) the biomass of one focal species that dominated the baseline period.
 
-Dynamics during the baseline period reveal very strong spatial patterns across the mean temperature gradient contrasted by weak responses to interannual variation in temperature at one site.
-Individual species show classic ``Whittaker" patterns of abundances across the gradient (Fig.
-\ref{fig:community-composition}).
-Time-averaged total community biomass increases strongly with temperature, because the fast-growing species at the warm sites produce more biomass, while the average biomass of our focal species shows a classic unimodal pattern across the temperature gradient (Fig.
+Dynamics during the baseline period revealed very strong spatial patterns across the mean temperature gradient. 
+Individual species showed classic ``Whittaker" patterns of abundances across the gradient (Fig. \ref{fig:community-composition}), while time-averaged total community 
+biomass increased strongly with temperature, because the fast-growing species at the warm sites produced more biomass. 
+The average biomass of our focal species showed a unimodal pattern across the temperature gradient (Fig. \ref{fig:community-baseline}). 
+These spatial patterns are the basis for our ``spatial models" of the temperature-biomass relationship.
+
+In contrast to the strong spatial patterns, population and community responses to interannual variation in temperature within sites were weak.
+At our focal site in the center of the gradient, both total biomass and the biomass of the focal species were quite insensitive to interannnual variation in temperature (Fig.
 \ref{fig:community-baseline}).
-These spatial patterns are the basis for our ``spatial models" of the temperature-biomass relationship.
-However, at our focal site in the center of the gradient, both total biomass and the biomass of the focal species are quite insensitive to interannnual variation in temperature (Fig.
-\ref{fig:community-baseline}).
-Our ``temporal models" estimate this weak temperature effect, along with strong effects of biomass in the previous year (e.g., lags).
+Our ``temporal models" estimate this weak temperature effect, along with strong lag effects of biomass in the previous year.
 
@@
 
-We can forecast the effect of temperature increase on total biomass and the focal species' biomass using the spatial models and the temporal models.
-For total biomass, the spatial and temporal models both predict immediate increases in biomass with warming, because all species respond positively to temperature increases.
-However, over time, as faster growing species increase in abundance at the focal site, biomass continues to increase, approaching the predictions of the spatial model (Fig.
-\ref{fig:community-forecasts} middle).
-The predictions of the spatial and temporal model for focal species biomass contrast more dramatically, with the temporal model predicting a large increase in biomass and the spatial model predicting a decrease.
-Initially, the simulated abundances follow the increase predicted by the temporal model, but as faster-growing species increase in abundance, the biomass of the focal species then decreases, eventually falling below its baseline level (Fig.
-\ref{fig:community-forecasts} bottom).
-Once again, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time.
-For total biomass (the sum across all species), the weights for the combined model shift rapidly to favor the spatial model, whereas for focal species biomass, the weights shift more slowly towards the spatial model, as the biomass of the focal species decreases and its competitors' biomass increase (Fig.
-\ref{fig:community-weights}).
+We forecasted the effect of temperature increase on total biomass and the focal species' biomass using the spatial models and the temporal models. 
+For total biomass, the spatial and temporal models both predicted immediate increases in biomass with warming, because all species respond positively to temperature increases. 
+However, over time, as faster growing species increased in abundance at the focal site, biomass continued to increase, approaching the predictions of the spatial model 
+(Fig. \ref{fig:community-forecasts} middle). The predictions of the spatial and temporal model for focal species biomass contrasted more dramatically, with the temporal 
+model predicting a large increase in biomass and the spatial model predicting a decrease. Initially, the simulated abundances followed the increase predicted by the 
+temporal model, but as faster-growing species increased in abundance, the biomass of the focal species then decreased, eventually falling below its baseline level 
+(Fig. \ref{fig:community-forecasts} bottom).
+
+As in the eco-evolutionary example, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time. 
+For total biomass (the sum across all species), the weights for the combined model shifted rapidly to favor the spatial model, whereas for focal species biomass, 
+the weights shifted more slowly towards the spatial model, as the biomass of the focal species decreased and its competitors' biomass increased (Fig. \ref{fig:community-weights}). 
+Again, the combined model matches the simulated data because the weights were determined by fitting directly to those data. A true, combined forecast would require
+a method to determine the model weights \emph{a priori}.
 
@@
 \end{figure}
+
+\subsection*{Discussion}
+
+Our simulations demonstrate three lessons:
+
+\emph{Lesson 1: For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow 
+processes such as evolutionary and ecological selection as well as dispersal.} [Ethan, I am hoping you can fix and add to this paragraph, perhaps
+drawing from examples in other fields?] In the short-term, time-series approaches work well because lag effects are strong: the state of the system in the next 
+time step depends heavily on the current state. Furthermore, in the short-term we usually are not extrapolating far into novel conditions; the historical range of variation 
+may be a good guide when there has been little time for genotype frequencies or species composition to change. In contrast, farther we forecast into the future, 
+the weaker the influence of the current state, and the more likely that slow processes like selection and dispersal will alter genotype frequencies and
+species composition.
+
+This lesson has implications for the emerging field of ecological forecasting. If we only focus on iterative, near-term forecasting, we are likely to find 
+that time-series approaches work well, and that consideration of slow processes like evolution, dispersal, and compositional turnover rarely improves
+forecast skill. Assuming that the same models will make good longer-term forecasts would be a mistake.
+
+We do not mean to undermine the call for an emphasis on near-term forecasting. Building capacity to make near-term forecasts is extremely important, not
+only because of the many critical applications (examples?) but because it is a necessary step to building a disciplinary culture and practice of ecological 
+forecasting (cite Dietze?). It is also feasible, because we have the data to validate forecasts at short time-scales. Rather, the point is that near-term forecasting
+alone is not enough. We also need to build capacity to make forecasts at time-scales where slow ecological processes cannot be ignored. The challenge 
+will be finding ways to validate long-term predictions. Possibilities include hindcasting with paleoecological and long-term historical data sets, use of model
+systems with short-generation times, and....simulations...? (or move this last bit to Lesson 3?).
+
+\emph{Lesson 2. Different kinds of data reflect the operation of different processes; the choice of data for model training should depend on the time-scale of the desired forecast}.
+Longitudinal data capture autocorrelation and responses of current assemblages, while data spanning spatial gradients capture the long-term outcome of 
+selection and dispersal. For short-term forecasts, building models using longitudinal data is clearly the best choice. As our simulations show, models fit to 
+steady-state, spatial patterns make terrible predictions in the short term. Conversely, models based on longitudinal data make poor predictions in the long-term,
+while the models fit to spatial patterns can perform well. Although the temporal and spatial models we fit are phenomenological, they capture different kinds
+ of underlying processes. The traditional debate about phenomenological vs. process-based models obscures the fact that, depending on the choice of the
+ training data, phenomenological models can represent different kinds of processes.
+
+ Why not simply build a process-based model? Perhaps something like the process-based models we used to generate the simulated time series in the first place? 
+ That kind of model could make good predictions at both short and long time-scales. Unfortunately, this approach is just not feasible in most cases.
+ The first challenge is simply determining the critical process. Take our eco-evolutionary example: should we simply assume that every annual plant species
+ has warm and cold-adapted genotypes? And that temperature adaptation will be the key trait driving the response to warming, as opposed to the effects of 
+ compositional turnover highlighted in our second example? Second, even if the key process is known in advance, fitting a mechanistic model presents its
+ own set of challenges. The model for our eco-evolutionary example requires, in order of increasing empirical difficulty, demographic rates for the full life
+ cycle, including density dependence, a quantitative mapping of genotype-by-environment interactions onto fitness, and dispersal kernels. The model for 
+ our community and ecosystem example requires demography, environment-dependent intra- and interspecific competition coefficients, and dispersal kernels.
+ We can think of empirical models that capture some of these features, but not all of them. Finally, even were it possible to fit such a model, the high number of 
+ parameters and associated parameter error might lead to poor predictive skill (refs?).
+
+Inevitably, models used for ecological forecasting will be partially if not entirely phenomenological. But that does not mean that they will not rely on process-based
+understanding in some way. The point of our simulations is that different processes should be considered for different forecast time-scales, and this can be done 
+by fitting phenomenological models to different kinds of data sets. Even when process-level understanding does not enable a fully mechanistic model, it can
+improve the specification of phenomenological models.
+
+\emph{Lesson 3. A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.}
+Our simulated examples suggest the potential for a combined forecast, which weights predictions from spatial and temporal models, to 
+accurately forecast both short and long-term dynamics. But we cheated: our combined ``forecasts'' were fit to data, not generated prospectively
+using \emph{a priori} model weights. Finding a general solution for determining those weights poses a fascinating research problem whose 
+solution would lead to important applications in ecological forecasting.
+
+We see a role for new theory and new empirical tests. On the theoretical side, what are the parameters that determine how quickly slow processes
+influence dynamics? We can play with our simulation models to provide some qualitative answers: the stronger the selection (at the level of genotype 
+or species), the shorter the lifespan of the focal organisms, and the greater the dispersal, the more rapid the shift from the temporal to the spatial model
+(should we actually DO these runs?). But providing general, analytical solutions would be difficult, even for our relatively simple models.
+
+On the empirical side, the key problem is finding or creating data sets to estimate the weighting function. As described under Lesson 1,
+this requires data sets appropriate for validating long-term forecasts.
+(Move material from end of  Lesson 1 down here?). With enough case studies, we might be able to infer patterns
+that could guide applications in new systems.
+
+Anything else to add?
 

### 2019-05-07 doc/manuscript.tex
started adding references

@@
 argue that a commitment to ecological forecasting is necessary to guide the environmental policy and management
-decisions that will become increasingly critical under global change (Clark 2001, Mouquet et al. 2017, Dietze et al. 2018).
-A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery (Houlahan et al. 2017, Dietz 2017, Dietz et al. 2018).
-Taken together, these papers, along with a new textbook (Dietze 2017) represent an emerging consensus that forecasting should be a priority for ecological research.
+decisions that will become increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
+A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery 
+\citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
+Taken together, these papers, along with a new textbook \citep{dietze_ecological_2017} represent an emerging consensus that 
+forecasting should be a priority for ecological research.
 
@@
 But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
-uncertainty (refs? Tredennick et al. 2017?).
+uncertainty (refs? \cite{tredennick_we_2017}?).
 
@@
 such as a species' geographic range in the year 2100, are virtually impossible to test (but see paleo hindcasting refs,Williams refs? Blois?), 
-leading Dietze et al (2017) to call for greater emphasis on near-term forecasting. The assumption is that building 
+leading \citep{dietze_iterative_2018} to call for greater emphasis on near-term forecasting. The assumption is that building 
 our capacity to make short-term forecasts will ultimately improve our long-term forecasts.
@@
 The second model represents what we call a ``temporal approach.'' We correlate interannual variation in an ecological response with interannual variation in the environment, 
-typically at just one site. This approach is often used to study population or vital rate fluctuations as a function of weather, 
-or primary production as a function of precipitation (Lauenroth and Sala 1992). The advantages of this approach include is focus on temporal dynamics, meaning it 
+typically at just one site. This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, 
+or primary production as a function of precipitation \citep{lauenroth_long-term_1992}. The advantages of this approach include is focus on temporal dynamics, meaning it 
 can predict rates of change, and validation of the resulting near-term predictions are straightforward. However, models built using this approach typically cover 
-a limited spatial extent (but see Kleinhesselink and Adler 2018, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
+a limited spatial extent (but see \cite{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
 that could alter the underlying process at longer time scales.
@@
 
-%\newpage
-%\renewcommand{\refname}{Literature cited}
-%\bibliographystyle{Ecology}
-%\bibliography{RemovalRefs}
+\newpage
+\renewcommand{\refname}{Literature cited}
+\bibliographystyle{Ecology}
+\bibliography{references}
 

### 2019-05-09 community-turnover/comm_turn_figures.R
Added rate experiments

@@ -1,168 +1,0 @@
-
-# call this script from comm_turn_master.R
-
-myCols<-alpha(c("blue3","red3"),0.4)
-
-###
-### Total biomass figures
-###
-
-# plot spatial and temporal models
-png("figures/community_models.png",height=3.5,width=4,res=400,units="in")
-
-  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
-  
-  # set up empty axes
-  plot(x=Tmean,y_spatial,type="n",xlab="Temperature",ylab="Biomass")
-  abline(v=Tmean[site], lty=2,col="darkgray")  # show focal site
-  
-  # total biomass temporal
-  xx <- seq(min(x_temporal),max(x_temporal),0.1)
-  points(x_temporal,y_temporal,pch=16,col=myCols[2])
-  y_hat <- coef(temporal_model)[1] + coef(temporal_model)[2]*mean(y_temporal) + coef(temporal_model)[3]*xx
-  lines(xx,y_hat, col=myCols[2], lwd=2 ) # plot marginal effect of temperature for temporal model
-  
-  # focal species temporal
-  points(x_temporal,y_temporal_spp,pch=1,col=myCols[2])
-  y_hat <- coef(temporal_model_spp)[1] + coef(temporal_model_spp)[2]*mean(y_temporal_spp) + coef(temporal_model_spp)[3]*xx
-  lines(xx,y_hat, col=myCols[2], lwd=1, lty="dotted" ) # plot marginal effect of temperature for temporal model
-  
-  # total biomass spatial
-  points(x_spatial,y_spatial,pch=16,col=myCols[1])
-  abline(spatial_model,col=myCols[1],lwd=2)
-  
-  # focal species spatial
-  points(x_spatial,y_spatial_spp,pch=1,col=myCols[1])
-  y_hat <-  coef(spatial_model_spp)[1] + coef(spatial_model_spp)[2]*x_spatial + coef(spatial_model_spp)[3]*x_spatial^2
-  lines(x_spatial,y_hat,col=myCols[1],lwd=1, lty="dotted")
-  
-  legend("topleft",c("Spatial (total)","Temporal (total)","Spatial (species)","Temporal (species)"),
-         lty=c("solid","solid","dotted","dotted"),pch=c(16,16,1,1),lwd=c(2,2,1,1),
[…]

### 2019-05-09 community-turnover/comm_turn_master.R

@@ -34,7 +34,7 @@ sim_yrs <- burnin_yrs+ baseline_yrs + warming_yrs + final_yrs # total number of 
 
 # parameters to generate species pool
 N <- 40      # number of species
-Gmax <- 0.5   #R max growth rate 
+Gmax <- 0.5   # max growth rate 
 Gmin <- 0.2   # min growth rate
 Lmax <- 1.5   # max sensitivity to competition
 Lmin <- 0.7   # min sensitivity to competition 
@@ -52,13 +52,20 @@ source("comm_turn_sim.R")
 source("comm_turn_forecast.R")
 
 ###
-### 3. Make figures
+### 3. Make first batch of figures
 ###
 
 if(file.exists("figures")==F) dir.create("figures")
 
-source("comm_turn_figures.R")
+source("comm_turn_figures_1.R")
 
+###
+### 4. Run dispersal and strength of selection experiments
+###
+
+source("comm_turn_expts.R")
+
+source("comm_turn_figures_2.R")
 
 
 

### 2019-05-09 genetic-diversity/genetic_diversity_figures.R

@@ -1,81 +1,0 @@
-
-# call this script from genetic_diversity_master.R
-
-myCols<-c("blue","darkgrey","red","black")
-
-# figure to show spatial relationships
-png("figures/spatial_model.png",height=3.5,width=7,res=400,units="in")
-  
-  par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,2,1))
-  
-  # plot fecundity as function of temperature for all genotypes
-  temp=seq(-7,7,0.01)
-  react_norms=getLambdas(temp,fec_Tmu,fec_Tsigma,fec_max)
-  matplot(temp,react_norms,type="l",xlab="Temperature",ylab="Germination rate",col=myCols,lty="solid")
-  text(fec_Tmu[1],0.9*fec_max[1],"AA",col=myCols[1])
-  text(fec_Tmu[2],0.9*fec_max[2],"Aa",col=myCols[2])
-  text(fec_Tmu[3],0.9*fec_max[3],"aa",col=myCols[3])
-  mtext("(A)",side=3, line=0.5, adj=0)
-  
-  # plot mean abundances vs mean temperature, along with spatial model "prediction"
-  xx <- seq(-5,5,0.01)
-  yy <- coef(spatial_model)[1] + coef(spatial_model)[2]*xx + coef(spatial_model)[3]*xx^2
-  matplot(simN_mix$Tmean,simN_mix[,c(2:5)],xlab="Mean temperature",ylab="N",
-            type="l", col=myCols,
-            lty=c("solid"),lwd=c(1,1,1,2))
-  lines(xx,yy,lty="dashed",lwd=1)
-  #legend("topright",c("AA","Aa","aa","Pop"),fill=myCols,bty="n")
-  mtext("(B)",side=3, line=0.5, adj=0)
-
-dev.off()
-
-# plot temporal pattern: temperature vs population growth during baseline period
-png("figures/temporal_model.png",height=3.5,width=4,res=400,units="in")
-  
-  mytheme <- list(axis.components = list(left = list(tck=0.5), right = list(tck=0.5),
-                                         top = list(tck=0.5), bottom = list(tck=0.5)))
-  trellis.par.set(mytheme)
-  
-  print(levelplot(N_t0_baseline ~ temperature_t1_baseline + r_baseline, 
[…]

### 2019-05-09 genetic-diversity/genetic_diversity_master.R

@@ -26,7 +26,7 @@ for(iT in 1:length(Tmean)){
 
 # diploid model
 fec_Tmu = c(-1,0,1)  # optimal temperature for genotypes AA, Aa, and aa
-fec_Tsigma = rep(5,3)     # standard deviation in temperature response for genotypes AA, Aa, and aa
+fec_Tsigma = rep(10,3)     # standard deviation in temperature response for genotypes AA, Aa, and aa
 fec_max = c(100,100,100)  # fecundity for genotypes AA, Aa, and aa
 G <- 1
 seedSurv = 0.5  # survival of ungerminated seeds (same for both genotypes)
@@ -66,5 +66,18 @@ source("genetic_diversity_sim_temporal.R")
 
 source("genetic_diversity_forecast.R")
 
-# plot temporal figures
-source("genetic_diversity_figures.R")
+###
+### 4. Plot first set of figures
+###
+
+if(file.exists("figures")==F) dir.create("figures")
+
+source("genetic_diversity_figures_1.R")
+
+###
+### 5. Run niche width experiment
+###
+
+source("genetic_diversity_expt.R")
+
+source("genetic_diversity_figures_2.R")

### 2019-05-09 community-turnover/comm_turn_expts.R
Adding rate expts take 2

@@ -1,0 +1,84 @@
+# Call from comm_turn_master.R. 
+# Assumes many objects already created.
+# Completes a couple more runs at higher dispersal rates
+# and compares rate of change in model weights for the focal species.
+
+# Uses existing species pool
+
+###
+### 1. Increase dispersal rates
+###
+
+d_baseline <- 0.01   # this should match value specified on comm_turn_master.R
+d_vals <- c(d_baseline,0.05,0.1) # specify additional values of d to try
+saved_weights_d <- matrix(NA, nrow=length(weight_spp),ncol=length(d_vals)) # matrix to store fitted model weights
+saved_weights_d[,1] <- weight_spp  # store existing run
+
+for(iRun in 2:length(d_vals)){
+  
+  d <- d_vals[iRun]
+  
+  # dispersal matrix (seeds disperse only to nearest neighbor locations)
+  d <- d/dt
+  seed_rain <- as.matrix(dist(landscape[,1:2],diag=T,upper=T))
+  seed_rain[seed_rain > 1] <- 0
+  seed_rain <- seed_rain*(d*dt*0.5)
+  diag(seed_rain) <- 1-d*dt
+  seed_rain[1,1] = 1-dt*d/2  
+  seed_rain[L_land,L_land] = 1-dt*d/2
+  
+  # Initialize community
+  for(t in 2:sim_yrs){
+    yr_temp <- temperature[t,]
+    spxp[t,,] <- CommunityTempDis(spxp[t-1,,], tr, seed_rain, yr_temp, dt)
+  } 
+  spxp_mean_baseline <- apply(spxp[burnin_yrs:(burnin_yrs+baseline_yrs),,],MARGIN = c(2,3),FUN=mean)
+  
+  source("comm_turn_forecast.R")
+  
+  saved_weights_d[,iRun] <- weight_spp
+  
+}
+
+###
+### 2. Increase strength of selection by increasing sensitivity to interspp competition
+###
+
+# reset dispersal to baseline value
+d <- d_baseline  
+d <- d/dt
+seed_rain <- as.matrix(dist(landscape[,1:2],diag=T,upper=T))
+seed_rain[seed_rain > 1] <- 0
+seed_rain <- seed_rain*(d*dt*0.5)
+diag(seed_rain) <- 1-d*dt
+seed_rain[1,1] = 1-dt*d/2  
+seed_rain[L_land,L_land] = 1-dt*d/2
+
[…]

### 2019-05-09 community-turnover/comm_turn_figures_1.R

@@ -1,0 +1,168 @@
+
+# call this script from comm_turn_master.R
+
+myCols<-alpha(c("blue3","red3"),0.4)
+
+###
+### Total biomass figures
+###
+
+# plot spatial and temporal models
+png("figures/community_models.png",height=3.5,width=4,res=400,units="in")
+
+  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
+  
+  # set up empty axes
+  plot(x=Tmean,y_spatial,type="n",xlab="Temperature",ylab="Biomass")
+  abline(v=Tmean[site], lty=2,col="darkgray")  # show focal site
+  
+  # total biomass temporal
+  xx <- seq(min(x_temporal),max(x_temporal),0.1)
+  points(x_temporal,y_temporal,pch=16,col=myCols[2])
+  y_hat <- coef(temporal_model)[1] + coef(temporal_model)[2]*mean(y_temporal) + coef(temporal_model)[3]*xx
+  lines(xx,y_hat, col=myCols[2], lwd=2 ) # plot marginal effect of temperature for temporal model
+  
+  # focal species temporal
+  points(x_temporal,y_temporal_spp,pch=1,col=myCols[2])
+  y_hat <- coef(temporal_model_spp)[1] + coef(temporal_model_spp)[2]*mean(y_temporal_spp) + coef(temporal_model_spp)[3]*xx
+  lines(xx,y_hat, col=myCols[2], lwd=1, lty="dotted" ) # plot marginal effect of temperature for temporal model
+  
+  # total biomass spatial
+  points(x_spatial,y_spatial,pch=16,col=myCols[1])
+  abline(spatial_model,col=myCols[1],lwd=2)
+  
+  # focal species spatial
+  points(x_spatial,y_spatial_spp,pch=1,col=myCols[1])
+  y_hat <-  coef(spatial_model_spp)[1] + coef(spatial_model_spp)[2]*x_spatial + coef(spatial_model_spp)[3]*x_spatial^2
+  lines(x_spatial,y_hat,col=myCols[1],lwd=1, lty="dotted")
+  
+  legend("topleft",c("Spatial (total)","Temporal (total)","Spatial (species)","Temporal (species)"),
+         lty=c("solid","solid","dotted","dotted"),pch=c(16,16,1,1),lwd=c(2,2,1,1),
[…]

### 2019-05-09 community-turnover/comm_turn_figures_2.R

@@ -1,0 +1,14 @@
+
+
+# call this script from comm_turn_master.R
+
+# Figure to compare rate of change in weights for different
+# dispersal rates
+myCols<-c("black","gray40","gray80")
+png("figures/compare_dispersal.png",height=3.5,width=5,res=400,units="in")
+par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
+matplot(saved_weights_d,type="l",lwd=2,lty=1,col=myCols,
+        xlab="Time step",ylab="Weight")
+legend("topright",legend=d_vals,
+       title="dispersal",lwd=2,lty=1,col=myCols,bty="n")
+dev.off()

### 2019-05-09 genetic-diversity/genetic_diversity_expt.R

@@ -1,0 +1,42 @@
+
+# Call from genetic_diversity_master.R, with all objects in memory
+
+# See how rate of shift in models weights responds to
+# an increased temperature niche of each genotype.
+
+fec_Tsigma_baseline <- fec_Tsigma # baseline temperature niche widths
+delta_fec_sigma <- c(0.5,1,2)  
+saved_weights <- matrix(NA,nrow=length(weight),ncol=length(delta_fec_sigma))
+saved_weights[,1] <- weight
+
+for(iRun in 1:length(delta_fec_sigma)){
+  
+  fec_Tsigma <- fec_Tsigma_baseline*delta_fec_sigma[iRun]
+  
+  # simulate baseline populations across temperature gradient
+  sim_yrs<-2000
+  Tmean = seq(-5,5,0.25)   # set of mean temperatures to explore
+  Tstdev <- 1   # st dev of temperature
+  temperature<-matrix(NA,sim_yrs,length(Tmean))
+  for(iT in 1:length(Tmean)){
+    temperature[,iT] = rnorm(sim_yrs,Tmean[iT],Tstdev)  # generate temperature time series
+  } 
+  source("genetic_diversity_sim_spatial.R")
+  
+  # simulate long-term run for focal population
+  baseline_yrs<-500  # number of yrs at baseline temperature
+  warming_yrs <- 100  # number of yrs with warming occurring
+  final_yrs <- 200  # number of yrs at steadty-state, warmed temperature
+  sim_yrs <- baseline_yrs + warming_yrs + final_yrs # total number of years
+  Tmean <- baseT + c(rep(0,baseline_yrs),seq(deltaT/warming_yrs,deltaT,deltaT/warming_yrs),
+                     rep(deltaT,final_yrs))
+  temperature = rnorm(sim_yrs,Tmean,Tstdev)  # generate temperature time series
+  source("genetic_diversity_sim_temporal.R")
+  
+  # generate forecasts
+  source("genetic_diversity_forecast.R")
+  
+  saved_weights[,iRun] <- weight
+  
+}
+

### 2019-05-09 genetic-diversity/genetic_diversity_figures_2.R

@@ -1,0 +1,15 @@
+
+# call this script from genetic_diversity_master.R
+
+# Figure to compare rate of change in weights for different
+# widths of temperature niche
+myCols<-c("black","gray40","gray80")
+png("figures/compare_Tsigmas.png",height=3.5,width=5,res=400,units="in")
+  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
+  matplot(saved_weights,type="l",lwd=2,lty=1,col=myCols,
+          xlab="Time step",ylab="Weight")
+  legend("topright",legend=fec_Tsigma_baseline[1]*delta_fec_sigma,
+         title=expression(sigma[T]),lwd=2,lty=1,col=myCols,bty="n")
+dev.off()
+
+

### 2019-05-09 doc/manuscript.tex
Added expts to ms

@@
 
-Next steps? Play with parameters to show what controls the rate of transition from temporal to spatial model?
-
 \subsection*{Community turnover example}
@@
 We see a role for new theory and new empirical tests. On the theoretical side, what are the parameters that determine how quickly slow processes
-influence dynamics? We can play with our simulation models to provide some qualitative answers: the stronger the selection (at the level of genotype 
-or species), the shorter the lifespan of the focal organisms, and the greater the dispersal, the more rapid the shift from the temporal to the spatial model
-(should we actually DO these runs?). But providing general, analytical solutions would be difficult, even for our relatively simple models.
+influence dynamics?
+We can play with our simulation models to provide some qualitative answers.
+The effects of some parameters are intuitive: in the community turnover example, increasing the fraction of dispersing individuals causes a more rapid shift in species composition and in model weights (Fig.
+\ref{fig:dispersal}).
+But other parameters have less straightforward effects.
+In the eco-evolutionary model, a potentially important parameter is the temperature tolerance of each genotype.
+We initially thought that increasing the temperature tolerance might accelerate shifts in genotype frequencies by increasing genotypic diversity within populations, but it had the opposite effect (Fig.
+\ref{fig:thermal_niche}), probably because it decreased the strength of selection.
+Other factors to consider  might include organism lifespans, and the magnitude of directional environmental change relative to interannaul variation.
+While simulation experiments are easy and informative, they cannot provide the generality of analytical theory.
+Unfortunately, deriving analytical solutions would be difficult, even for our relatively simple models.
 
@@
 
+\section{Supplementary Figures}
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.7 \textwidth] {compare_Tsigmas.png}
+	\caption{In our eco-evolutionary model, the rate of change in the weight of the temporal forecast (y-axis) depends on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches).
+}
+	\label{fig:thermal_niche}
+\end{figure}
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.7 \textwidth] {compare_dispersal.png}
+	\caption{In our community turnover model, the rate of change in the weight of the temporal forecast (y-axis) depends on fraction of propagules dispersing each time step. }
+	\label{fig:dispersal}
+\end{figure}
+
 \end{document}

### 2019-05-09 genetic-diversity/genetic_diversity_master.R
Reset Tfec sigma

@@ -26,7 +26,7 @@ for(iT in 1:length(Tmean)){
 
 # diploid model
 fec_Tmu = c(-1,0,1)  # optimal temperature for genotypes AA, Aa, and aa
-fec_Tsigma = rep(10,3)     # standard deviation in temperature response for genotypes AA, Aa, and aa
+fec_Tsigma = rep(5,3)     # standard deviation in temperature response for genotypes AA, Aa, and aa
 fec_max = c(100,100,100)  # fecundity for genotypes AA, Aa, and aa
 G <- 1
 seedSurv = 0.5  # survival of ungerminated seeds (same for both genotypes)

### 2019-05-21 community-turnover/comm_turn_figures_1.R
Redoing figures

@@ -8,64 +8,72 @@ myCols<-alpha(c("blue3","red3"),0.4)
 ###
 
 # plot spatial and temporal models
-png("figures/community_models.png",height=3.5,width=4,res=400,units="in")
+png("figures/community_models_species.png",height=3.5,width=4,res=400,units="in")
 
   par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
   
   # set up empty axes
-  plot(x=Tmean,y_spatial,type="n",xlab="Temperature",ylab="Biomass")
+  plot(x=x_temporal,y_temporal_spp,type="n",xlab="Temperature",ylab="Biomass",
+       ylim=c(0,max(y_temporal_spp)*1.2))
   abline(v=Tmean[site], lty=2,col="darkgray")  # show focal site
   
-  # total biomass temporal
-  xx <- seq(min(x_temporal),max(x_temporal),0.1)
-  points(x_temporal,y_temporal,pch=16,col=myCols[2])
-  y_hat <- coef(temporal_model)[1] + coef(temporal_model)[2]*mean(y_temporal) + coef(temporal_model)[3]*xx
-  lines(xx,y_hat, col=myCols[2], lwd=2 ) # plot marginal effect of temperature for temporal model
-  
   # focal species temporal
+  xx <- seq(min(x_temporal),max(x_temporal),0.1)
   points(x_temporal,y_temporal_spp,pch=1,col=myCols[2])
   y_hat <- coef(temporal_model_spp)[1] + coef(temporal_model_spp)[2]*mean(y_temporal_spp) + coef(temporal_model_spp)[3]*xx
-  lines(xx,y_hat, col=myCols[2], lwd=1, lty="dotted" ) # plot marginal effect of temperature for temporal model
-  
-  # total biomass spatial
-  points(x_spatial,y_spatial,pch=16,col=myCols[1])
-  abline(spatial_model,col=myCols[1],lwd=2)
+  lines(xx,y_hat, col=myCols[2], lwd=2, lty="solid" ) # plot marginal effect of temperature for temporal model
   
   # focal species spatial
-  points(x_spatial,y_spatial_spp,pch=1,col=myCols[1])
+  points(x_spatial,y_spatial_spp,pch=16,col=myCols[1])
[…]

### 2019-05-22 community-turnover/comm_turn_expts.R
Finished redoing figures

@@ -77,8 +77,6 @@ for(iRun in 1:length(c_factor)){
 }
 # this seems to slow the rate of change; initial communities less diverse
 
-
-
-
-
-
+# write results to file
+colnames(saved_weights_c) <- d_vals
+write.csv(saved_weights_c,"simulations/saved_weights.csv",row.names=F)

### 2019-05-22 community-turnover/comm_turn_figures.R

@@ -3,37 +3,42 @@
 
 myCols<-alpha(c("blue3","red3"),0.4)
 
-###
-### Total biomass figures
-###
-
-# plot spatial and temporal models
-png("figures/community_models_species.png",height=3.5,width=4,res=400,units="in")
-
-  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
+# plot spatial and temporal patterns and models for the focal species
+png("figures/species_patterns_models.png",height=3.4,width=8,res=400,units="in")
   
-  # set up empty axes
-  plot(x=x_temporal,y_temporal_spp,type="n",xlab="Temperature",ylab="Biomass",
-       ylim=c(0,max(y_temporal_spp)*1.2))
-  abline(v=Tmean[site], lty=2,col="darkgray")  # show focal site
+  # panel A--spatila
+  par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),oma=c(0,2,0,0),mar=c(3,1,3,1))
+  y <-spxp_mean_baseline[,colSums(spxp_mean_baseline)>0]
+  jet.colors <-colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan",
+                       "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"),bias=2.5)
+  tmp <- jet.colors(n=NCOL(y))
+  matplot(Tmean,y,type="l",xlab="Baseline temperature",ylab="Biomass",
+          ylim=c(0,max(y_temporal_spp)*1.1),lty=1,col=tmp)
+  
+  y_hat <-  coef(spatial_model_spp)[1] + coef(spatial_model_spp)[2]*x_spatial + coef(spatial_model_spp)[3]*x_spatial^2
+  lines(x_spatial,y_hat,col=tmp[1],lwd=1, lty="dotted")
+  abline(v=Tmean[site], lty=1,col="darkgray")
+  
+  mtext("(A)",side=3,line=0.5,adj=0)
+  mtext("Biomass",side=2,line=1,outer=T)
+
+  # panel B--temporal
+    # set up empty axes
+  plot(x=x_temporal,y_temporal_spp,type="n",xlab="Annual temperature",ylab="",
+       ylim=c(0,max(y_temporal_spp)*1.1))
+  #abline(v=Tmean[site], lty=1,col="darkgray")  # show focal site
   
   # focal species temporal
   xx <- seq(min(x_temporal),max(x_temporal),0.1)
[…]

### 2019-05-22 community-turnover/comm_turn_master.R

@@ -57,7 +57,7 @@ source("comm_turn_forecast.R")
 
 if(file.exists("figures")==F) dir.create("figures")
 
-source("comm_turn_figures_1.R")
+source("comm_turn_figures.R")
 
 ###
 ### 4. Run dispersal and strength of selection experiments
@@ -65,7 +65,6 @@ source("comm_turn_figures_1.R")
 
 source("comm_turn_expts.R")
 
-source("comm_turn_figures_2.R")
 
 
 

### 2019-05-22 genetic-diversity/final_figure.R

@@ -1,0 +1,31 @@
+
+# call this script from genetic_diversity_master.R
+
+
+# read in weights from commuynity turnover example
+saved_weights_comm <- read.csv("./../community-turnover/simulations/saved_weights.csv",header=T)
+d_vals <- colnames(saved_weights_comm)
+d_vals <- gsub("X","",d_vals)
+
+# Figure to compare rate of change in weights for different
+# widths of temperature niche
+myCols<-c("black","gray40","gray80")
+png("figures/compare_Tsigmas.png",height=3.5,width=8.5,res=400,units="in")
+  
+  par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
+  
+  #community example
+  matplot(saved_weights_comm,type="l",lwd=2,lty=1,col=myCols,
+          xlab="Time step",ylab="Weight")
+  legend("topright",legend=d_vals,
+       title="dispersal",lwd=2,lty=1,col=myCols,bty="n")
+  
+  #eco-evo example
+  matplot(saved_weights,type="l",lwd=2,lty=1,col=myCols,
+          xlab="Time step",ylab="Weight")
+  legend("topright",legend=fec_Tsigma_baseline[1]*delta_fec_sigma,
+         title=expression(sigma[T]),lwd=2,lty=1,col=myCols,bty="n")
+  
+dev.off()
+
+

### 2019-05-22 genetic-diversity/genetic_diversity_expt.R

@@ -5,7 +5,7 @@
 # an increased temperature niche of each genotype.
 
 fec_Tsigma_baseline <- fec_Tsigma # baseline temperature niche widths
-delta_fec_sigma <- c(0.5,1,2)  
+delta_fec_sigma <- c(0.5,1,1.5)  
 saved_weights <- matrix(NA,nrow=length(weight),ncol=length(delta_fec_sigma))
 saved_weights[,1] <- weight
 
@@ -24,9 +24,9 @@ for(iRun in 1:length(delta_fec_sigma)){
   source("genetic_diversity_sim_spatial.R")
   
   # simulate long-term run for focal population
-  baseline_yrs<-500  # number of yrs at baseline temperature
-  warming_yrs <- 100  # number of yrs with warming occurring
-  final_yrs <- 200  # number of yrs at steadty-state, warmed temperature
+  # baseline_yrs<-500  # number of yrs at baseline temperature
+  # warming_yrs <- 100  # number of yrs with warming occurring
+  # final_yrs <- 300  # number of yrs at steadty-state, warmed temperature
   sim_yrs <- baseline_yrs + warming_yrs + final_yrs # total number of years
   Tmean <- baseT + c(rep(0,baseline_yrs),seq(deltaT/warming_yrs,deltaT,deltaT/warming_yrs),
                      rep(deltaT,final_yrs))

### 2019-05-22 genetic-diversity/genetic_diversity_figures.R

@@ -3,10 +3,10 @@
 
 myCols<-c("blue","darkgrey","red","black")
 
-# figure to show spatial relationships
-png("figures/spatial_model.png",height=3.5,width=7,res=400,units="in")
+# figure to show spatial and temporal relationships
+png("figures/spatial&temporal_model.png",height=10,width=4,res=400,units="in")
   
-  par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,2,1))
+  par(mfrow=c(3,1),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,2,1),cex=1)
   
   # plot fecundity as function of temperature for all genotypes
   temp=seq(-7,7,0.01)
@@ -26,21 +26,20 @@ png("figures/spatial_model.png",height=3.5,width=7,res=400,units="in")
   lines(xx,yy,lty="dashed",lwd=1)
   #legend("topright",c("AA","Aa","aa","Pop"),fill=myCols,bty="n")
   mtext("(B)",side=3, line=0.5, adj=0)
+  
+  # plot population growth and density vs annual temperature
+  # first make colors based on N0 density
+  rr <- range(N_t0_baseline)
+  svals <- (N_t0_baseline-rr[1])/diff(rr)
+  f <- colorRamp(c("red", "blue"), 0.4)
+  density_colors <- alpha(rgb(f(svals)/255),0.6)
+  # now plot
+  plot(temperature_t1_baseline,r_baseline,xlab="Annual temperature",ylab="log Population growth rate",
+       col=density_colors,pch=16)
+  mtext("(C)",side=3, line=0.5, adj=0)
 
 dev.off()
 
-# plot temporal pattern: temperature vs population growth during baseline period
-png("figures/temporal_model.png",height=3.5,width=4,res=400,units="in")
-  
-  mytheme <- list(axis.components = list(left = list(tck=0.5), right = list(tck=0.5),
-                                         top = list(tck=0.5), bottom = list(tck=0.5)))
-  trellis.par.set(mytheme)
-  
-  print(levelplot(N_t0_baseline ~ temperature_t1_baseline + r_baseline, 
-            xlab="Temperature",ylab="log Population growth rate",
-            panel = panel.levelplot.points, col.regions = topo.colors(25)))
-
-dev.off()
 
 # plot forecast
 png("figures/forecast.png",height=5,width=5,res=400,units="in")

### 2019-05-22 genetic-diversity/genetic_diversity_figures_2.R

@@ -1,15 +1,0 @@
-
-# call this script from genetic_diversity_master.R
-
-# Figure to compare rate of change in weights for different
-# widths of temperature niche
-myCols<-c("black","gray40","gray80")
-png("figures/compare_Tsigmas.png",height=3.5,width=5,res=400,units="in")
-  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
-  matplot(saved_weights,type="l",lwd=2,lty=1,col=myCols,
-          xlab="Time step",ylab="Weight")
-  legend("topright",legend=fec_Tsigma_baseline[1]*delta_fec_sigma,
-         title=expression(sigma[T]),lwd=2,lty=1,col=myCols,bty="n")
-dev.off()
-
-

### 2019-05-22 genetic-diversity/genetic_diversity_master.R

@@ -42,7 +42,7 @@ source("genetic_diversity_sim_spatial.R")
 
 baseline_yrs<-500  # number of yrs at baseline temperature
 warming_yrs <- 100  # number of yrs with warming occurring
-final_yrs <- 200  # number of yrs at steadty-state, warmed temperature
+final_yrs <- 300  # number of yrs at steadty-state, warmed temperature
 sim_yrs <- baseline_yrs + warming_yrs + final_yrs # total number of years
 
 Tstdev <- 1   # st dev of temperature (stationary)
@@ -72,7 +72,7 @@ source("genetic_diversity_forecast.R")
 
 if(file.exists("figures")==F) dir.create("figures")
 
-source("genetic_diversity_figures_1.R")
+source("genetic_diversity_figures.R")
 
 ###
 ### 5. Run niche width experiment
@@ -80,4 +80,4 @@ source("genetic_diversity_figures_1.R")
 
 source("genetic_diversity_expt.R")
 
-source("genetic_diversity_figures_2.R")
+source("final_figure.R")

### 2019-05-22 genetic-diversity/final_figure.R
Added panel labels

@@ -12,19 +12,21 @@ d_vals <- gsub("X","",d_vals)
 myCols<-c("black","gray40","gray80")
 png("figures/compare_Tsigmas.png",height=3.5,width=8.5,res=400,units="in")
   
-  par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
+  par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,2,1))
   
   #community example
   matplot(saved_weights_comm,type="l",lwd=2,lty=1,col=myCols,
           xlab="Time step",ylab="Weight")
   legend("topright",legend=d_vals,
        title="dispersal",lwd=2,lty=1,col=myCols,bty="n")
+  mtext("(A)",side=3,line=0.5,adj=0)
   
   #eco-evo example
   matplot(saved_weights,type="l",lwd=2,lty=1,col=myCols,
           xlab="Time step",ylab="Weight")
   legend("topright",legend=fec_Tsigma_baseline[1]*delta_fec_sigma,
          title=expression(sigma[T]),lwd=2,lty=1,col=myCols,bty="n")
+  mtext("(B)",side=3,line=0.5,adj=0)
   
 dev.off()
 

### 2019-05-22 genetic-diversity/genetic_diversity_master.R

@@ -26,7 +26,7 @@ for(iT in 1:length(Tmean)){
 
 # diploid model
 fec_Tmu = c(-1,0,1)  # optimal temperature for genotypes AA, Aa, and aa
-fec_Tsigma = rep(5,3)     # standard deviation in temperature response for genotypes AA, Aa, and aa
+fec_Tsigma = rep(8,3)     # standard deviation in temperature response for genotypes AA, Aa, and aa
 fec_max = c(100,100,100)  # fecundity for genotypes AA, Aa, and aa
 G <- 1
 seedSurv = 0.5  # survival of ungerminated seeds (same for both genotypes)

### 2019-05-23 doc/manuscript.tex
Major reorganization

@@
 
-\title{\textbf{ Matching forecast time-scale with ecological process }}
+\title{\textbf{ Matching the forecast horizon with the relevant ecological processes }}
 
@@
 ongoing debates about competing modeling approaches. We expect that short and long-term forecasts will require models
-incorporating different kinds of processes and fit to different kinds of data sets. We demonstrate this idea using two simulation case studies, the first for an 
-eco-evolutionary scenario, the second concerned with feedbacks between community composition and ecosystem functioning. Our analyses show that: 
+incorporating different kinds of processes and fit to different kinds of data sets. We demonstrate this idea using two simulation case studies, the first 
+focused on how competitive interactions affect population dynamics of a focal species, the second for  an 
+eco-evolutionary scenario. Our analyses show that: 
 \begin{enumerate}
@@
 
-\section*{Eco-evolutionary example}
+\subsection*{Community turnover example}
 
-Evolutionary adaptation is a key uncertainty in predicting how environmental change will impact a focal population at a given location (refs).
-To explore how this process influences dynamics and forecasts at different time scales, we consider a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
-\ref{fig:spatial_model}A; the full model description is provided in the Appendix, and all computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast}).
-Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
-\ref{fig:spatial_model}B).
+A common problem for conservation biologists and natural resource managers is anticipating the impact of environmental change on the abundance of 
+endangered species, biological invaders, and harvested species. Although the managers may be primarily interested in just the focal species, 
+skillful prediction might require considering interactions with many other species, greatly complicating the problem. But at what forecast horizon do
+altered species interactions become impossible to ignore? We explored this question using a metacommunity model developed by Alexander et al. (ref). 
+The model describes Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
+Composition varies along the gradient because of a trade-off between growth rate and cold tolerance: 
+cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by 
+fast-growing species that are cold intolerant. Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics. 
+Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
+All computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast})
+
+We first simulated a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase, and then a 
+final period of stationary temperature. Interannual variation in temperature is the same at all sites, but mean temperature varies among sites. All sites experienced the 
+same absolute increase in mean temperature. We focused on the biomass dynamics of one focal species that dominated the baseline period.
+
+Dynamics during the baseline period revealed very strong spatial patterns across the mean temperature gradient. 
+Individual species, including our focal species, showed classic, unimodal ``Whittaker" patterns of abundances across the gradient (Fig. \ref{fig:species-patterns-models}A).
+These spatial patterns are the basis for our ``spatial model" of the temperature-biomass relationship for our focal species (Fig. \ref{fig:species-patterns-models}A). 
+In contrast to the strong spatial patterns, population and community responses to interannual variation in temperature within sites were weak.
+At our focal site in the center of the gradient, the biomass of the focal species was quite insensitive to interannnual variation in temperature (Fig.
+\ref{fig:species-patterns-models}B).
+Our ``temporal model" estimates this weak, linear temperature effect, along with strong lag effects of biomass in the previous year.
+The spatial and temporal statistical
+models are described in Appendix \ref{metacomm}.
 
@@
 \centering
-\includegraphics[width=1 \textwidth]{spatial_model.png}
+\includegraphics[width=1 \textwidth] {species_patterns_models.png}
+\caption{(A) Mean biomass by species (colors) across the temperature gradient during the baseline period.
+The focal species, dominant at the site in the center of the gradient (vertical gray line), is shown in dark blue.
+The dashed blue line shows predictions from the spatial model.
+(B) Annual biomass of the focal species at the central site during the baseline period.
+The dashed line shows predictions from the temporal model.
+}
+\label{fig:species-patterns-models}
+\end{figure}
+
+We forecasted the effect of temperature increase on the focal species' biomass at one location in the center of the temperature gradient 
+using the spatial models and the temporal models (Appendix \ref{metacomm}). 
+The predictions from the spatial and temporal model contrasted markedly, with the temporal 
+model predicting a large increase in biomass and the spatial model predicting a decrease. Initially, the simulated abundances followed the increase predicted by the 
+temporal model, but as faster-growing species colonized and increased in abundance at the study site, the biomass of the focal species then decreased, eventually falling below its baseline level (Fig.
+\ref{fig:community-forecast-species}).
+
+To combine the spatial and temporal model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
+At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{weights} for full description of the approach).
+This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-weights-spp}) 
+because the weights were determined by fitting directly to those data. Unlike our spatial and temporal model forecasts, 
+we did not generate out-of-sample predictions from the combined model; it merely provides 
+a convenient way to quantify how rapidly dynamics shift from from being dominated by the interannual 
+variations captured in the temporal model to being dominated by the steady-state equilibrium captured by the spatial model.
+The most rapid shifts in model weights occur during the period when warm-adapted, faster growing species are increasing most rapidly
+in abundance (Fig. \ref{fig:community-weights-spp}). A true, combined forecast would require a method to determine the model weights \emph{a priori}.
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_forecast_species.png}
+\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
+(Bottom) Simulated focal species biomass and forecasts from the spatial, temporal and combined models.
+}
+\label{fig:community-forecast-species}
+\end{figure}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_change_plus_weights_spp.png}
+\caption{Simulated changes in biomass of the focal species (black) and all other species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for focal species biomass (blue).
+}
+\label{fig:community-weights-spp}
+\end{figure}
+
+The compositional turnover affecting our focal species also influences total biomass, linking community and ecosystem dynamics. We repeated our focal species analysis 
+for total community biomass, and the results were similar: the temporal model initially made the best forecasts immediately following the onset of the temperature
+increase, but as the identify and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (Appendix \ref{biomass}).
+
+\section*{Eco-evolutionary example}
+
+Evolutionary adaptation is a key uncertainty in predicting how environmental change will impact a focal population at a given location (refs). Like the shifts
+in species composition illustrated in the previous example, shifts in genotype frequencies can also influence dynamics and forecasts at different time scales.
+We demonstrate this process with a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
+\ref{fig:evo_evo_spatial_temporal_models}A; the full model description is provided in Appendix \ref{eco-evo}, and all computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast}).
+Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
+\ref{fig:evo_evo_spatial_temporal_models}B).
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.4 \textwidth]{spatial&temporal_model.png}
 \caption{(A) Reaction norms of the three genotypes.
-(B) The spatial pattern of individual genotypes and total population abundances at sites arrayed across a gradient of mean annual temperature.
+(B) The spatial pattern of individual genotypes (colors) and total population abundance (black) at sites arrayed across a gradient of mean annual temperature.
 The dashed line shows predictions from an empirical ``spatial model," a linear regression that describes mean population size as a function of mean temperature.
+(C) The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
+Colors show population size (the blue to red gradient depicting low to high population density), which influences the population growth rate through density dependence.
 }
-\label{fig:spatial_model}
+\label{fig:evo_evo_spatial_temporal_models}
 \end{figure}
@@
 The spatial pattern shown in Fig.
-\ref{fig:spatial_model} is the outcome of steady-state conditions.
+\ref{fig:evo_evo_spatial_temporal_models}B is the outcome of steady-state conditions.
 But at any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm.
@@
 During the baseline period, the population is dominated by the cold-adapted genotype.
-We use the simulated data from this baseline period to fit an empirical model that assumes no knowledge of the underlying eco-evolutionary process.
+We used the simulated data from this baseline period to fit an empirical model that assumes no knowledge of the underlying eco-evolutionary process.
 This empirical ``temporal model" predicts population growth rate as a function of annual temperature and population size (Fig.
-\ref{fig:temporal_model}).
+\ref{fig:evo_evo_spatial_temporal_models}C).
+
 We then imposed a period of warming, followed by a final period of higher, but stationary, temperature (Fig.
@@
 \ref{fig:forecast} bottom).
+The implications for forecasting are clear: In the initial stages of the warming trend, the temporal model fit to the original, baseline conditions should make better the best predictions about resulting population trends, but in the longer term, as evolutionary change occurs, the spatial model ultimately makes better predictions.
 
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.6 \textwidth] {temporal_model.png}
-\caption{The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
-Colors show population size, which also influences the population growth rate through density dependence.
-}
-\label{fig:temporal_model}
-\end{figure}
-
-The implications for forecasting are clear: In the initial stages of the warming trend, the temporal model fit to the original, baseline conditions should make better the best predictions 
-about resulting population trends, but in the longer term, as evolutionary change occurs, the spatial model ultimately makes better predictions. 
-To combine the spatial and temporal model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix for full description of the approach).
+As in the eco-evolutionary example, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time. 
 The temporal model accurately predicted the impact of the initial warming trend, but eventually became far too pessimistic, while the spatial model did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
@@
 \ref{fig:forecast_supp}).
+The combined model's predictions look impressively accurate, but, as in the community turnover example, that is because we used the full, simulated time series to fit the weighting term.
+To generate a true forecast from the combined model, we would need an independent method to predict how the model weights shift over time.
 
@@
 \label{fig:forecast}
-\end{figure}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {forecast_supplement.png}
-\caption{Simulated shifts in genotype abundances, and the model weighting term, $\omega$, during the warming phase and the following stationary temperature phase.}
-\label{fig:forecast_supp}
-\end{figure}
-
-The combined model's predictions look impressively accurate, but that is because we used the full, simulated time series to fit the weighting term. 
-Unlike our spatial and temporal model forecasts, we did not generate out-of-sample predictions from the combined model; it merely provides 
-a convenient way to quantify how rapidly dynamics shift from from being dominated by interannual 
-variation to being dominated by steady-state equilibrium. To generate a true forecast from the combined model, we would need an independent 
-method to predict how the model weights shift over time.
-
-\subsection*{Community turnover example}
-
-In the previous example, evolutionary change was the slow process that determined how quickly the combined model shifted from a temporal to spatial forecast. 
-The process of compositional turnover--change in the abundances and identities of species interacting at a site--can have a similar effect at the level of 
-communities and even ecosystems. We demonstrate this using a metacommunity model developed by Alexander et al. (ref). The model describes Lotka-Volterra 
-competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient. Composition varies along the gradient because of a 
-trade-off between growth rate and cold tolerance: cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by 
-fast-growing species that are cold intolerant. Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics. 
-Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
-
-As in our eco-evolutionary example, we simulated a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase, and then a 
-final period of stationary temperature. Interannual variation in temperature is the same at all sites, but mean temperature varies among sites. All sites experienced the 
-same absolute increase in mean temperature. We focused on two response variables at one site in the middle of the gradient: 1) total community biomass and 
-2) the biomass of one focal species that dominated the baseline period.
-
-Dynamics during the baseline period revealed very strong spatial patterns across the mean temperature gradient. 
-Individual species showed classic ``Whittaker" patterns of abundances across the gradient (Fig. \ref{fig:community-composition}), while time-averaged total community 
-biomass increased strongly with temperature, because the fast-growing species at the warm sites produced more biomass. 
-The average biomass of our focal species showed a unimodal pattern across the temperature gradient (Fig. \ref{fig:community-baseline}). 
-These spatial patterns are the basis for our ``spatial models" of the temperature-biomass relationship.
-
-In contrast to the strong spatial patterns, population and community responses to interannual variation in temperature within sites were weak.
-At our focal site in the center of the gradient, both total biomass and the biomass of the focal species were quite insensitive to interannnual variation in temperature (Fig.
-\ref{fig:community-baseline}).
-Our ``temporal models" estimate this weak temperature effect, along with strong lag effects of biomass in the previous year.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {mean_biomass_spp_by_site.png}
-\caption{Mean biomass by species (colors) across the temperature gradient during the baseline period. The focal species, dominant at the site in the center of the gradient, is shown in dark blue.  }
-\label{fig:community-composition}
-\end{figure}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_models.png}
-\caption{Time-averaged spatial relationships between temperature and both total community biomass and the focal species' biomass are shown in blue, while interannual variations in total and focal species biomass at one site in the center of the gradient are shown in red.
-Symbols show simulated values, lines show predictions from statistical models fit to the simulated data.
-}
-\label{fig:community-baseline}
-\end{figure}
-
-We forecasted the effect of temperature increase on total biomass and the focal species' biomass using the spatial models and the temporal models. 
-For total biomass, the spatial and temporal models both predicted immediate increases in biomass with warming, because all species respond positively to temperature increases. 
-However, over time, as faster growing species increased in abundance at the focal site, biomass continued to increase, approaching the predictions of the spatial model 
-(Fig. \ref{fig:community-forecasts} middle). The predictions of the spatial and temporal model for focal species biomass contrasted more dramatically, with the temporal 
-model predicting a large increase in biomass and the spatial model predicting a decrease. Initially, the simulated abundances followed the increase predicted by the 
-temporal model, but as faster-growing species increased in abundance, the biomass of the focal species then decreased, eventually falling below its baseline level 
-(Fig. \ref{fig:community-forecasts} bottom).
-
-As in the eco-evolutionary example, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time. 
-For total biomass (the sum across all species), the weights for the combined model shifted rapidly to favor the spatial model, whereas for focal species biomass, 
-the weights shifted more slowly towards the spatial model, as the biomass of the focal species decreased and its competitors' biomass increased (Fig. \ref{fig:community-weights}). 
-Again, the combined model matches the simulated data because the weights were determined by fitting directly to those data. A true, combined forecast would require
-a method to determine the model weights \emph{a priori}.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_forecast.png}
-\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(Middle) Simulated total biomass and forecasts from the spatial, temporal and combined models.
-(Bottom) Simulated focal species' biomass and forecasts from the spatial, temporal and combined models.
-}
-\label{fig:community-forecasts}
-\end{figure}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_change_plus_weights.png}
-\caption{Simulated changes in biomass of the focal species (black) and all other species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue) and focal species biomass (red).
-}
-\label{fig:community-weights}
 \end{figure}
@@
 alone is not enough. We also need to build capacity to make forecasts at time-scales where slow ecological processes cannot be ignored. The challenge 
-will be finding ways to validate long-term predictions. Possibilities include hindcasting with paleoecological and long-term historical data sets, use of model
-systems with short-generation times, and....simulations...? (or move this last bit to Lesson 3?).
+will be finding ways to validate long-term predictions.
 
@@
 The effects of some parameters are intuitive: in the community turnover example, increasing the fraction of dispersing individuals causes a more rapid shift in species composition and in model weights (Fig.
-\ref{fig:dispersal}).
+\ref{fig:dispersal_niche_width}A).
 But other parameters have less straightforward effects.
@@
 We initially thought that increasing the temperature tolerance might accelerate shifts in genotype frequencies by increasing genotypic diversity within populations, but it had the opposite effect (Fig.
-\ref{fig:thermal_niche}), probably because it decreased the strength of selection.
-Other factors to consider  might include organism lifespans, and the magnitude of directional environmental change relative to interannaul variation.
+\ref{fig:dispersal_niche_width}B), probably because it decreased the strength of selection, slowing the rate of change in genotype frequencies.
+Other factors to consider include organism lifespans and the magnitude of directional environmental change relative to interannaul variation.
 While simulation experiments are easy and informative, they cannot provide the generality of analytical theory.
@@
 
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.7 \textwidth] {dispersal_niche_width.png}
+	\caption{(The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
+}
+	\label{fig:dispersal_niche_width}
+\end{figure}
+
 On the empirical side, the key problem is finding or creating data sets to estimate the weighting function. As described under Lesson 1,
 this requires data sets appropriate for validating long-term forecasts.
-(Move material from end of  Lesson 1 down here?). With enough case studies, we might be able to infer patterns
+Possibilities include hindcasting with paleoecological and long-term historical data sets, use of model
+systems with short-generation times, and....more simulation studies...? . With enough case studies, we might be able to infer patterns
 that could guide applications in new systems.
@@
 
-\section{Description of the eco-evolutionary annual plant model}
+\section{Spatial and temporal models for the community turnover example}\label{metacomm}
 
-(from Michael Cortez)
+[How much detail to include from Alexander et al?]
+
+The spatial model for total community biomass is a simple linear regression.
+The spatial model for the biomass of the focal species is a multiple regression with a quadratic term, identical to Eqn.
+\ref{eqn:spatial_regression}.
+The temporal model for total biomass, $M$, is:
+ \begin{equation}
+ \ln(M_{t+1} = T(M_{t},K_t) = \beta^T_0 +  \beta^T_1 M_t +\beta^T_2 K_t  +  \varepsilon .
+ \label{eqn:community_temporal_regression}
+ \end{equation}
+The temporal model for focal species biomass has the same functional form.
+The combined models are based on the same approach that we used in the eco-evolutionary example (eqn.
+\ref{eqn:combined_model}).
+
+\subsection{Estimating model weights}\label{weights}
+
+Here is my home brew approach, presented in my typically terrible notation. There must be a better, more formal approach...Help?
+
+The spatial model, which we refer to as $S$, is a linear regression fit to ``observed" data on mean population size, $\bar{N}$ at a series of locations, $x$, which vary in mean temperature, $K$.
+In our case, the observed data are simulated by the model described in the previous section.
+We include a quadratic term to capture the unimodal relationship between  $\bar{N}$ and $K$:
+ \begin{equation}
+ \bar{N}_x = S(K_x) = \beta^S_0 +  \beta^S_1 K_x +\beta^S_2 {K_x}^2 + \varepsilon
+ \label{eqn:spatial_regression}
+ \end{equation}
+
+The temporal model, which we call $T$, starts with a time series of ``observed" population sizes, $\vec{N}$.
+Using data from a baseline period with a stationary distribution of annual temperature at one location, we fit the following regression, which predicts the (log) population growth rate as a function of population size and annual temperature at time $t$:
+ \begin{equation}
+ \ln(\frac{N_{t+1}}{N_t}) = T(N_t,K_t) = \beta^T_0 +  \beta^T_1 ln(N_t) +\beta^T_2 K_t + beta^T_3 {K_t}^2 +  \varepsilon
+ \label{eqn:temporal_regression}
+ \end{equation}
+
+The temporal model returns a per capita growth rate on the log scale.
+To predict population size at the next time step, we simply exponentiate the growth rate and multiply it by the current population size: $exp(T(N_t,K_t)) N_t$.
+
+The combined model is a weighted average of predictions from the spatial and temporal models, with the weights changing as a function of time, here expressed as the forecast horizon.
+The weights change as a function of the square root of the forecast horizon, to allow rapid shifts in the model weights.
+\begin{equation}
+logit(\omega_t)=\beta^C_0 + \beta^C_1 \sqrt{t}
+\label{eqn:weights}
+\end{equation}
+\begin{equation}
+\hat{N}_{t+1} = \omega * exp(T(N_{t},K_t)) N_t + (1-\omega) * S(K_t) 
+\label{eqn:combined_model}
+\end{equation}
+
+I used the \texttt{optim} function to estimate the $\beta^C$'s that minimize the sum of squared errors, $(\hat{N}_{t+1} - N_{t+1})^2$
+
+\section{Total biomass results}\label{biomass}
+
+Put total biomass results here
+
+For total biomass, the spatial and temporal models both predicted immediate increases in biomass with warming, because all species respond positively to temperature increases. 
+However, over time, as faster growing species increased in abundance at the focal site, biomass continued to increase, approaching the predictions of the spatial model 
+(Fig. \ref{fig:community-forecasts-total} middle).
+
+For total biomass (the sum across all species), the weights for the combined model shifted rapidly to favor the spatial model, whereas for focal species biomass, 
+the weights shifted more slowly towards the spatial model, as the biomass of the focal species decreased and its competitors' biomass increased (Fig. \ref{fig:community-weights}).
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_forecast_total.png}
+\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
+(Bottom) Simulated total biomass and forecasts from the spatial, temporal and combined models.
+}
+\label{fig:community-forecasts-total}
+\end{figure}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_change_plus_weights_total.png}
+\caption{Simulated changes in biomass all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).  }
+\label{fig:community-weights-total}
+\end{figure}
+
+\section{Description of the eco-evolutionary annual plant model}\label{eco-evo}
 
@@
 
-\subsection{Estimating model weights}
-
-Here is my home brew approach, presented in my typically terrible notation. There must be a better, more formal approach...Help?
-
-The spatial model, which we refer to as $S$, is a linear regression fit to ``observed" data on mean population size, $\bar{N}$ at a series of locations, $x$, which vary in mean temperature, $K$.
-In our case, the observed data are simulated by the model described in the previous section.
-We include a quadratic term to capture the unimodal relationship between  $\bar{N}$ and $K$:
- \begin{equation}
- \bar{N}_x = S(K_x) = \beta^S_0 +  \beta^S_1 K_x +\beta^S_2 {K_x}^2 + \varepsilon
- \label{eqn:spatial_regression}
- \end{equation}
-
-The temporal model, which we call $T$, starts with a time series of ``observed" population sizes, $\vec{N}$.
-Using data from a baseline period with a stationary distribution of annual temperature at one location, we fit the following regression, which predicts the (log) population growth rate as a function of population size and annual temperature at time $t$:
- \begin{equation}
- \ln(\frac{N_{t+1}}{N_t}) = T(N_t,K_t) = \beta^T_0 +  \beta^T_1 ln(N_t) +\beta^T_2 K_t + beta^T_3 {K_t}^2 +  \varepsilon
- \label{eqn:temporal_regression}
- \end{equation}
-
-The temporal model returns a per capita growth rate on the log scale.
-To predict population size at the next time step, we simply exponentiate the growth rate and multiply it by the current population size: $exp(T(N_t,K_t)) N_t$.
-
-The combined model is a weighted average of predictions from the spatial and temporal models, with the weights changing as a function of time, here expressed as the forecast horizon.
-The weights change as a function of the square root of the forecast horizon, to allow rapid shifts in the model weights.
-\begin{equation}
-logit(\omega_t)=\beta^C_0 + \beta^C_1 \sqrt{t}
-\label{eqn:weights}
-\end{equation}
-\begin{equation}
-\hat{N}_{t+1} = \omega * exp(T(N_{t},K_t)) N_t + (1-\omega) * S(K_t) 
-\label{eqn:combined_model}
-\end{equation}
-
-I used the \texttt{optim} function to estimate the $\beta^C$'s that minimize the sum of squared errors, $(\hat{N}_{t+1} - N_{t+1})^2$
-
-\section{Description of the metacommunity model}
-
-[How much detail to include from Alexander et al?]
-
-The spatial model for total community biomass is a simple linear regression.
-The spatial model for the biomass of the focal species is a multiple regression with a quadratic term, identical to Eqn.
-\ref{eqn:spatial_regression}.
-The temporal model for total biomass, $M$, is:
- \begin{equation}
- \ln(M_{t+1} = T(M_{t},K_t) = \beta^T_0 +  \beta^T_1 M_t +\beta^T_2 K_t  +  \varepsilon .
- \label{eqn:community_temporal_regression}
- \end{equation}
-The temporal model for focal species biomass has the same functional form.
-The combined models are based on the same approach that we used in the eco-evolutionary example (eqn.
-\ref{eqn:combined_model}).
-
 \section{Supplementary Figures}
@@
 \centering
-	\includegraphics[width=0.7 \textwidth] {compare_Tsigmas.png}
-	\caption{In our eco-evolutionary model, the rate of change in the weight of the temporal forecast (y-axis) depends on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches).
-}
-	\label{fig:thermal_niche}
-\end{figure}
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.7 \textwidth] {compare_dispersal.png}
-	\caption{In our community turnover model, the rate of change in the weight of the temporal forecast (y-axis) depends on fraction of propagules dispersing each time step. }
-	\label{fig:dispersal}
+\includegraphics[width=0.7 \textwidth] {forecast_supplement.png}
+\caption{Simulated shifts in genotype abundances, and the model weighting term, $\omega$, during the warming phase and the following stationary temperature phase.}
+\label{fig:forecast_supp}
 \end{figure}

### 2019-05-23 community-turnover/comm_turn_figures.R

@@ -6,13 +6,13 @@ myCols<-alpha(c("blue3","red3"),0.4)
 # plot spatial and temporal patterns and models for the focal species
 png("figures/species_patterns_models.png",height=3.4,width=8,res=400,units="in")
   
-  # panel A--spatila
+  # panel A--spatial
   par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),oma=c(0,2,0,0),mar=c(3,1,3,1))
   y <-spxp_mean_baseline[,colSums(spxp_mean_baseline)>0]
   jet.colors <-colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan",
                        "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"),bias=2.5)
   tmp <- jet.colors(n=NCOL(y))
-  matplot(Tmean,y,type="l",xlab="Baseline temperature",ylab="Biomass",
+  matplot(Tmean,y,type="l",xlab="Mean temperature",ylab="Biomass",
           ylim=c(0,max(y_temporal_spp)*1.1),lty=1,col=tmp)
   
   y_hat <-  coef(spatial_model_spp)[1] + coef(spatial_model_spp)[2]*x_spatial + coef(spatial_model_spp)[3]*x_spatial^2

### 2019-05-23 community-turnover/comm_turn_figures_2.R

@@ -1,14 +1,0 @@
-
-
-# call this script from comm_turn_master.R
-
-# Figure to compare rate of change in weights for different
-# dispersal rates
-myCols<-c("black","gray40","gray80")
-png("figures/compare_dispersal.png",height=3.5,width=5,res=400,units="in")
-par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,1,1))
-matplot(saved_weights_d,type="l",lwd=2,lty=1,col=myCols,
-        xlab="Time step",ylab="Weight")
-legend("topright",legend=d_vals,
-       title="dispersal",lwd=2,lty=1,col=myCols,bty="n")
-dev.off()

### 2019-05-23 genetic-diversity/final_figure.R

@@ -10,7 +10,7 @@ d_vals <- gsub("X","",d_vals)
 # Figure to compare rate of change in weights for different
 # widths of temperature niche
 myCols<-c("black","gray40","gray80")
-png("figures/compare_Tsigmas.png",height=3.5,width=8.5,res=400,units="in")
+png("figures/dispersal_niche_width.png",height=3.5,width=8.5,res=400,units="in")
   
   par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,2,1))
   

### 2019-05-23 doc/manuscript.tex
intermediate changes

@@
 \caption{Simulated changes in biomass of the focal species (black) and all other species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for focal species biomass (blue).
+Year 0 in this figure corresponds to the start of the temperature increase.
 }
@@
 	\caption{(The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
+Year 0 in these figures corresponds to the start of the temperature increase.
 }
@@
 
-\section{Spatial and temporal models for the community turnover example}\label{metacomm}
-
-[How much detail to include from Alexander et al?]
+\section{Spatial, temporal, and combined models}\label{metacomm}
 
@@
 \includegraphics[width=0.7 \textwidth] {community_change_plus_weights_total.png}
-\caption{Simulated changes in biomass all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).  }
+\caption{Simulated changes in biomass all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).
+Year 0 in this figure corresponds to the start of the temperature increase.
+}
 \label{fig:community-weights-total}
@@
 \includegraphics[width=0.7 \textwidth] {forecast_supplement.png}
-\caption{Simulated shifts in genotype abundances, and the model weighting term, $\omega$, during the warming phase and the following stationary temperature phase.}
+\caption{Simulated shifts in genotype abundances, and the model weighting term, $\omega$, during the warming phase and the following stationary temperature phase.
+Year 0 in this figure corresponds to the start of the temperature increase.}
 \label{fig:forecast_supp}

### 2019-05-24 doc/manuscript.tex
More reorg of SI

@@
 But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
-uncertainty (refs? \cite{tredennick_we_2017}?).
+uncertainty (refs? \citealt{tredennick_we_2017}?).
 
@@
 The spatial and temporal statistical
-models are described in Appendix \ref{metacomm}.
+models are described in Appendix \ref{models}.
 
@@
 We forecasted the effect of temperature increase on the focal species' biomass at one location in the center of the temperature gradient 
-using the spatial models and the temporal models (Appendix \ref{metacomm}). 
+using the spatial models and the temporal models. 
 The predictions from the spatial and temporal model contrasted markedly, with the temporal 
@@
 To combine the spatial and temporal model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{weights} for full description of the approach).
+At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{models} for full description of the approach).
 This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-weights-spp}) 
@@
 for total community biomass, and the results were similar: the temporal model initially made the best forecasts immediately following the onset of the temperature
-increase, but as the identify and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (Appendix \ref{biomass}).
+increase, but as the identify and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (Figs.
+\ref{fig:community-models-total} - \ref{fig:community-weights-total}).
 
@@
 We used the simulated data from this baseline period to fit an empirical model that assumes no knowledge of the underlying eco-evolutionary process.
-This empirical ``temporal model" predicts population growth rate as a function of annual temperature and population size (Fig.
+This empirical ``temporal model" (Appendix \ref{models}) predicts population growth rate as a function of annual temperature and population size (Fig.
 \ref{fig:evo_evo_spatial_temporal_models}C).
@@
 
-\vspace{0.4in}
-
-\renewcommand{\theequation}{A-\arabic{equation}}
-\renewcommand{\thetable}{A-\arabic{table}}
-\renewcommand{\thefigure}{A-\arabic{figure}}
 \renewcommand{\thesection}{\Alph{section}}
 
-\section{Spatial, temporal, and combined models}\label{metacomm}
-
-The spatial model for total community biomass is a simple linear regression.
-The spatial model for the biomass of the focal species is a multiple regression with a quadratic term, identical to Eqn.
-\ref{eqn:spatial_regression}.
-The temporal model for total biomass, $M$, is:
- \begin{equation}
- \ln(M_{t+1} = T(M_{t},K_t) = \beta^T_0 +  \beta^T_1 M_t +\beta^T_2 K_t  +  \varepsilon .
- \label{eqn:community_temporal_regression}
- \end{equation}
-The temporal model for focal species biomass has the same functional form.
-The combined models are based on the same approach that we used in the eco-evolutionary example (eqn.
-\ref{eqn:combined_model}).
-
-\subsection{Estimating model weights}\label{weights}
-
-Here is my home brew approach, presented in my typically terrible notation. There must be a better, more formal approach...Help?
+\section{Spatial, temporal and combined models}\label{models}
 
@@
 In our case, the observed data are simulated by the model described in the previous section.
-We include a quadratic term to capture the unimodal relationship between  $\bar{N}$ and $K$:
+For both examples in the main text, we include a quadratic term to capture the unimodal relationship between  $\bar{N}$ and $K$:
  \begin{equation}
@@
  \end{equation}
+For the total biomass example from the community turnover model (results in Appendix \ref{biomass}), we dropped the quadratic term.
 
-The temporal model, which we call $T$, starts with a time series of ``observed" population sizes, $\vec{N}$.
-Using data from a baseline period with a stationary distribution of annual temperature at one location, we fit the following regression, which predicts the (log) population growth rate as a function of population size and annual temperature at time $t$:
+The temporal model, which we call $T$, starts with a time series of ``observed" population sizes, or total biomasses, $\vec{N}$ at one location.
+In the community turnover example, we use data from a baseline period with a stationary distribution of annual temperature at the focal location, we fit the following regression, which predicts biomass at time $t+1$ as a function of biomass and annual temperature at time $t$:
  \begin{equation}
- \ln(\frac{N_{t+1}}{N_t}) = T(N_t,K_t) = \beta^T_0 +  \beta^T_1 ln(N_t) +\beta^T_2 K_t + beta^T_3 {K_t}^2 +  \varepsilon
- \label{eqn:temporal_regression}
+ \ln(N_{t+1}) = T(N_t,K_t) = \beta^T_0 +  \beta^T_1 ln(N_t) +\beta^T_2 K_t  +  \varepsilon
+ \label{eqn:temporal_regression_community}
  \end{equation}
-
-The temporal model returns a per capita growth rate on the log scale.
+ For the eco-evolutionary example, the response variable is the log of the population growth rate, and we include a quadratic effect of annual temperature:
+  \begin{equation}
+  \ln(\frac{N_{t+1}}{N_t}) = T(N_t,K_t) = \beta^T_0 +  \beta^T_1 ln(N_t) +\beta^T_2 K_t  +\beta^T_3 K_t^2 +  \varepsilon
+  \label{eqn:temporal_regression_ecoevo}
+  \end{equation}
+ This version of the temporal model returns a per capita growth rate on the log scale.
 To predict population size at the next time step, we simply exponentiate the growth rate and multiply it by the current population size: $exp(T(N_t,K_t)) N_t$.
@@
 \end{equation}
+For the community turnover example, the predicted biomass from the combined model is:
+\begin{equation}
+\hat{N}_{t+1} = \omega *T(N_{t},K_t) + (1-\omega) * S(K_t) 
+\label{eqn:combined_model}
+\end{equation}
+For the eco-evolutionary example, the predicted population size from the combined model is:
 \begin{equation}
@@
 
-I used the \texttt{optim} function to estimate the $\beta^C$'s that minimize the sum of squared errors, $(\hat{N}_{t+1} - N_{t+1})^2$
-
-\section{Total biomass results}\label{biomass}
-
-Put total biomass results here
-
-For total biomass, the spatial and temporal models both predicted immediate increases in biomass with warming, because all species respond positively to temperature increases. 
-However, over time, as faster growing species increased in abundance at the focal site, biomass continued to increase, approaching the predictions of the spatial model 
-(Fig. \ref{fig:community-forecasts-total} middle).
-
-For total biomass (the sum across all species), the weights for the combined model shifted rapidly to favor the spatial model, whereas for focal species biomass, 
-the weights shifted more slowly towards the spatial model, as the biomass of the focal species decreased and its competitors' biomass increased (Fig. \ref{fig:community-weights}).
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_forecast_total.png}
-\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(Bottom) Simulated total biomass and forecasts from the spatial, temporal and combined models.
-}
-\label{fig:community-forecasts-total}
-\end{figure}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_change_plus_weights_total.png}
-\caption{Simulated changes in biomass all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).
-Year 0 in this figure corresponds to the start of the temperature increase.
-}
-\label{fig:community-weights-total}
-\end{figure}
+We used the \texttt{optim} function to estimate the $\beta^C$s that minimize the sum of squared errors, $(\hat{N}_{t+1} - N_{t+1})^2$
 
 \section{Description of the eco-evolutionary annual plant model}\label{eco-evo}
+
+\renewcommand{\theequation}{B-\arabic{equation}}
+\renewcommand{\thetable}{B-\arabic{table}}
+\renewcommand{\thefigure}{B-\arabic{figure}}
 
@@
 
+\renewcommand{\thefigure}{C-\arabic{figure}}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_models_total.png}
+\caption{(Results for total biomass from the community turnover model.
+Blue points show mean total biomass during the baseline period at locations across the temperature gradient, and the blue line shows predictions from the spatial model.
+Red points show annual total biomass during the baseline period as a function of annual temperature at the central site on the gradient.
+The red line shows predictions from the temporal model.
+}
+\label{fig:community-models-total}
+\end{figure}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_forecast_total.png}
+\caption{Results for total biomass from the community turnover model.
+(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
+(Bottom) Simulated total biomass and forecasts from the spatial, temporal and combined models.
+}
+\label{fig:community-forecasts-total}
+\end{figure}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=0.7 \textwidth] {community_change_plus_weights_total.png}
+\caption{Simulated changes in biomass of all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).
+Year 0 in this figure corresponds to the start of the temperature increase.
+}
+\label{fig:community-weights-total}
+\end{figure}
+
 \begin{figure}[tbp]

### 2019-05-24 doc/manuscript.tex
Finished this draft

@@
 frequently argue about the virtues of process-based vs. phenomenological models (refs?).
-We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g., Adler et al. 2012),
-and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions (refs?).
+We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g.,\citealt{Adler2012}),
+and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions \citep{Fitzpatrick,Veloz2012}.
 But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
@@
 Identifying these models requires rigorously testing predictions against observations. However, many common ecological predictions,
-such as a species' geographic range in the year 2100, are virtually impossible to test (but see paleo hindcasting refs,Williams refs? Blois?), 
+such as a species' geographic range in the year 2100, are virtually impossible to test (but see \citealt{Maguire2015}), 
 leading \citep{dietze_iterative_2018} to call for greater emphasis on near-term forecasting. The assumption is that building 
@@
 The first phenomenological model represents what we call a ``spatial approach.'' We correlate the mean environment with the mean of an ecological state or rate 
- across many sites. This is the approach commonly used to predict population distribution or abundance as a function of climate (SDM refs) or
- primary production as a function of mean precipitation (Lauenroth refs). The approach is attractive because it addresses broad spatial scales relevant to 
+ across many sites. This is the approach commonly used to predict population distribution or abundance as a function of climate \citep{elith_species_2009} or
+ primary production as a function of mean precipitation \citep{Sala1988}. The approach is attractive because it addresses broad spatial scales relevant to 
  management, and because the necessary data are widely available and statistical analyses are straightforward. The approach also has well-known weaknesses.
@@
 over long time spans, they are unlikely to provide much information about dynamic changes on the order of years to decades. Moreover, transient dynamics could 
-prevent the system from ever reaching the predicted steady state (Urban ref?).
+prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
 
@@
 can predict rates of change, and validation of the resulting near-term predictions are straightforward. However, models built using this approach typically cover 
-a limited spatial extent (but see \cite{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
+a limited spatial extent (but see \citealt{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
 that could alter the underlying process at longer time scales.
@@
 To quantify this pattern, we combine forecasts from the spatial and temporal approaches with a weighted average and then show how the optimal model weights change over time: we expect the temporal model to best predict short-term dynamics, and the spatial model to best predict long-term dynamics.
-How rapidly that transition occurs depends on the rate of  ``slow" processes such as evolution and species turnover.
+How rapidly that transition occurs depends on the rate of  ``slow" processes such as species turnover and evolution.
 
@@
 skillful prediction might require considering interactions with many other species, greatly complicating the problem. But at what forecast horizon do
-altered species interactions become impossible to ignore? We explored this question using a metacommunity model developed by Alexander et al. (ref). 
+altered species interactions become impossible to ignore? We explored this question using a metacommunity model developed by \cite{alexander_lags_2018}. 
 The model describes Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
@@
 Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
-All computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast})
+All computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast}).
 
@@
 At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{models} for full description of the approach).
-This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-weights-spp}) 
+This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}) 
 because the weights were determined by fitting directly to those data. Unlike our spatial and temporal model forecasts, 
@@
 increase, but as the identify and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (Figs.
-\ref{fig:community-models-total} - \ref{fig:community-weights-total}).
+\ref{fig:community-models-total} to \ref{fig:community-weights-total}).
 
@@
 We demonstrate this process with a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
-\ref{fig:evo_evo_spatial_temporal_models}A; the full model description is provided in Appendix \ref{eco-evo}, and all computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast}).
+\ref{fig:evo_evo_spatial_temporal_models}A).
+The full model description is provided in Appendix \ref{eco-evo}, and all computer code is available on Github  (\texttt{https://github.com/pbadler/space-time-forecast}).
 Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.

### 2019-07-18 doc/manuscript.tex
Rework introduction to center the space-time distinction in ecological forecasting
Since this paper focuses on time-series vs space-for-time approaches this revision
highlights both the distinction between those two approaches and the ideas behind
why the two approaches may behave differently at different time horizons.

@@
 
-A recent flurry of synthesis papers in high impact journals attests to surging interest in ecological forecasting. Many of the papers 
-argue that a commitment to ecological forecasting is necessary to guide the environmental policy and management
-decisions that will become increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
-A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery 
+* Forecasting is important
+* Two different approaches to this time-series & space-for-time
+	* Not clear which is best
+	* Some limited comparisons
+* Which approach is best may depend on the time-horizon
+* Here we explore this possibility using simulation based approach
+
+Forecasting is increasingly recognized as important to the advancement and application of ecological research.
+Ecological forecasting is necessary to guide the environmental policy and management
+decisions that are increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
+In addition, evaluating the accuracy of ecological forecasts can provide rigorous tests of predictions from ecological models that can
+advance understanding of the processes governing ecological systems  
 \citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
-Taken together, these papers, along with a new textbook \citep{dietze_ecological_2017} represent an emerging consensus that 
-forecasting should be a priority for ecological research.
+This combined value of advancing our fundamental understanding of ecology and applying that understanding to support management and decision making 
+makes that forecasting is an important priority for ecological research.
 
-No such consensus is available for exactly how to tackle this challenge.
-Ecologists rely on many different approaches and models for making predictions, and 
-frequently argue about the virtues of process-based vs. phenomenological models (refs?).
-We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g.,\citealt{Adler2012}),
-and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions \citep{Fitzpatrick,Veloz2012}.
-But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
-uncertainty (refs? \citealt{tredennick_we_2017}?).
+Ecological models used for forecasting have traditionally been based on either time-series modeling or space-for-time substitutions.
+Time-series approaches use the historical behavior of a system to fit models that describe the temporal dynamics of that system and use those
+dynamic models to make predictions about what will happen in the future. This is a common approach for population based forecasts and includes both
+process based population models and data-driven models capturing empirical patterns in data (e.g., \citep{ward_complexity_2014}) 
+Alternatively, space-for-time modeling approaches develop models of how and ecological property varies in across space in response to environmental
+conditions. These spatial relationships are assumed to also hold for changes at a site through time, and forecasts are made by forecasting future
+environmental conditions and determining the associated ecological response. Most forecasts for future species distributions and biodiversity
+are based on this type of approach, with a models built to describe the environmental conditions where species occur across space being applied to
+forecast environmental conditions to determine where they will occur in the future (e.g., citation). Despite both of these approaches being
+widely used there has been little comparison of how they compare for forecasting and understanding the dynamics of ecological systems \citep{harris_forecasting_2018}.
 
-The pragmatic solution to these debates is clear: we should forecast using the models that make the best predictions.
-Identifying these models requires rigorously testing predictions against observations. However, many common ecological predictions,
-such as a species' geographic range in the year 2100, are virtually impossible to test (but see \citealt{Maguire2015}), 
-leading \citep{dietze_iterative_2018} to call for greater emphasis on near-term forecasting. The assumption is that building 
-our capacity to make short-term forecasts will ultimately improve our long-term forecasts.
+Whether historical dynamics, contemporary spatial patterns, or a combination of the two will serve as the best source of information for forecasting
+may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. This potential dependency on the "forecast horizon"; ()\textit{sensu} citation)
+results from lags in the response of ecological conditions to environmental change, changes in the importance of ecological processes with
+time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of conditions present in the data.
+At short forecast horizons (days to years) conditions are likely to stay within the range of the historical data at a site, the time scales of change will be similar to
+the time scales present in that historical data, and the current state of the system is likely to capture the influence of unmeasured processes.
+As a result, for near-term forecasts time-series approaches may effectively capture the key dynamics and result in accurate forecasts. In contrast,
+at long forecast horizons (decades) an ecosystem is likely to experience conditions that have not been historically observed () and be responding to
+processes operating at scales difficult to detect in even long ecological time-series. Therefore, for the kinds of far-term forecasts typically conducted
+in biodiversity and species-distribution modeling space-for-time based approaches may include important information about how ecosystems respond to major
+shifts in climate over long periods and as a results produce better long-term forecasts. Using different approaches for different time horizons is common
+in other areas of forecasting, including forecasts for environmental conditions where models for short-term changes (weather) differ substantially from
+those for long-term changes (climate).
 
-Our goal is to consider a hypothesis that, in our view, has not been carefully considered by either the push for near-term forecasting or
-ongoing debates about competing modeling approaches. We expect that short and long-term forecasts will require models
-incorporating different kinds of processes and fit to different kinds of data sets. We demonstrate this idea using two simulation case studies, the first 
-focused on how competitive interactions affect population dynamics of a focal species, the second for  an 
-eco-evolutionary scenario. Our analyses show that: 
+Here we investigate the possibility that the best approaches for building models for ecological forecasting may depend on the time horizon of the forecast
+and explore how time-series and space-for-time approaches might be combined to address mid-term ecological forecasting. We do this using two simulation case studies, one 
+focused on how competitive interactions affect population dynamics of a focal species and one eco-evolutionary scenario. Our analyses show that: 
 \begin{enumerate}

### 2019-07-22 doc/manuscript.tex
Emphasize biological processes in Intro

@@
 dynamic models to make predictions about what will happen in the future. This is a common approach for population based forecasts and includes both
-process based population models and data-driven models capturing empirical patterns in data (e.g., \citep{ward_complexity_2014}) 
+process based population models and data-driven models capturing empirical patterns in data (e.g., \citep{ward_complexity_2014}). Time-series models
+capture processes like birth, death, small scale dispersal events, and short-term responses to environmental conditions.
 Alternatively, space-for-time modeling approaches develop models of how and ecological property varies in across space in response to environmental
@@
 are based on this type of approach, with a models built to describe the environmental conditions where species occur across space being applied to
-forecast environmental conditions to determine where they will occur in the future (e.g., citation). Despite both of these approaches being
+forecast environmental conditions to determine where they will occur in the future (e.g., citation). Space-for-time models capture processes like
+immigration, extinction, and responses to large changes in environmental conditions over long time periods. Despite both of these approaches being
 widely used there has been little comparison of how they compare for forecasting and understanding the dynamics of ecological systems \citep{harris_forecasting_2018}.
@@
 time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of conditions present in the data.
-At short forecast horizons (days to years) conditions are likely to stay within the range of the historical data at a site, the time scales of change will be similar to
-the time scales present in that historical data, and the current state of the system is likely to capture the influence of unmeasured processes.
-As a result, for near-term forecasts time-series approaches may effectively capture the key dynamics and result in accurate forecasts. In contrast,
-at long forecast horizons (decades) an ecosystem is likely to experience conditions that have not been historically observed () and be responding to
-processes operating at scales difficult to detect in even long ecological time-series. Therefore, for the kinds of far-term forecasts typically conducted
-in biodiversity and species-distribution modeling space-for-time based approaches may include important information about how ecosystems respond to major
-shifts in climate over long periods and as a results produce better long-term forecasts. Using different approaches for different time horizons is common
-in other areas of forecasting, including forecasts for environmental conditions where models for short-term changes (weather) differ substantially from
-those for long-term changes (climate).
+At short forecast horizons (days to years) processes driving individual demographics tend to dominate (similar to the time scales present in that
+historical data), conditions are likely to stay within the range of the historical data at a site, and the current state of the system is likely
+to capture the influence of unmeasured processes. As a result, for near-term forecasts time-series approaches may effectively capture the key dynamics
+and result in accurate forecasts. In contrast, at long forecast horizons (decades) an ecosystem is likely to experience conditions that have not been
+historically observed () and be responding to large scale processes large scale climate and land use change that are difficult to detect in even long
+ecological time-series. Therefore, for the kinds of long-term forecasts typically conducted in biodiversity and species-distribution modeling
+space-for-time based approaches may include important information about how ecosystems respond to major shifts in climate over long periods and as a
+results produce better long-term forecasts. Using different approaches for different time horizons is common in other areas of forecasting,
+including forecasts for environmental conditions where models for short-term changes (weather) differ substantially from those for long-term changes
+(climate).
 

### 2019-07-22 doc/manuscript.tex
Minor editorial work on Modeling Approach section

--- 2019-05-24 doc/manuscript.tex
@@
 
-In both of our case studies, we simulate the effects of a hypothetical temperature increase on a simple system with known dynamics.
-The ``truth'' is represented by a model that is mechanistic for at least one important process, but we assume that recovering 
-this true, process-based model would be impossible in practice (we return to this point in the Discussion). We begin each simulation
+In two case studies, we simulate the effects of an increase in temperature on simple systems with known dynamics.
+The ``truth'' is represented by a model that is mechanistic for at least one important process, but treat the model as unknown when
+analyzing the data and assume that perfectly recovering this model would not be possible in practice. We begin each simulation
 under stationary temperature, allowing the system to equilibrate; we call this the baseline phase. We then increase temperature progressively
 over a period of time, followed by a second period of stationary, now elevated, temperature. The objective is to forecast the response of the 
-system to temperature increase based on data gathered during the baseline period.
+system to the temperature increase based on data gathered during the baseline period.
 
-We make two forecasts based on two different phenomenological models, each representing processes operating at different time scales. 
+We make forecasts based on two phenomenological models, each representing processes operating at different time scales. 
 The first phenomenological model represents what we call a ``spatial approach.'' We correlate the mean environment with the mean of an ecological state or rate 
@@
 they provide no information about how quickly the system will move from the current state to the predicted, future state. Because the patterns in the data have developed 
-over long time spans, they are unlikely to provide much information about dynamic changes on the order of years to decades. Moreover, transient dynamics could 
+over long time spans, they are may not provide good information about dynamic changes on the order of a few years. Moreover, transient dynamics could 
 prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
@@
 a limited spatial extent (but see \citealt{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
-that could alter the underlying process at longer time scales.
+that could alter the underlying process at longer time scales and improve forecastability \citep{clark_ecological_2001}.
 
-We compare forecasts from both models to the true, simulated dynamics to show that responses to environmental change reflect both kinds of processes, but at different time scales.
-To quantify this pattern, we combine forecasts from the spatial and temporal approaches with a weighted average and then show how the optimal model weights change over time: we expect the temporal model to best predict short-term dynamics, and the spatial model to best predict long-term dynamics.
-How rapidly that transition occurs depends on the rate of  ``slow" processes such as species turnover and evolution.
+We compare forecasts from both models to the simulated dynamics to determine how well the two approaches to forecasting perform at different time scales.
+We also assess the potential for combining the information available in both spatial and temporal patters by combining forecasts from the spatial and temporal approaches using a weighted average and assess how the optimal model weights change over time.
+Based on the hypothesis that spatial patterns capture longer-term processes we expect the temporal model to best predict short-term dynamics, and the spatial model to best predict long-term dynamics, and that a combination of the two approaches will provide the best forecasts at transitional intermediate time scales.
 

### 2019-07-22 doc/manuscript.tex
Editoral work on Community Turnover example

@@
 
-In two case studies, we simulate the effects of an increase in temperature on simple systems with known dynamics.
-The ``truth'' is represented by a model that is mechanistic for at least one important process, but treat the model as unknown when
-analyzing the data and assume that perfectly recovering this model would not be possible in practice. We begin each simulation
+In both of our case studies, we simulate the effects of a hypothetical temperature increase on a simple system with known dynamics.
+The ``truth'' is represented by a model that is mechanistic for at least one important process, but we assume that recovering 
+this true, process-based model would be impossible in practice (we return to this point in the Discussion). We begin each simulation
 under stationary temperature, allowing the system to equilibrate; we call this the baseline phase. We then increase temperature progressively
 over a period of time, followed by a second period of stationary, now elevated, temperature. The objective is to forecast the response of the 
-system to the temperature increase based on data gathered during the baseline period.
+system to temperature increase based on data gathered during the baseline period.
 
-We make forecasts based on two phenomenological models, each representing processes operating at different time scales. 
+We make two forecasts based on two different phenomenological models, each representing processes operating at different time scales. 
 The first phenomenological model represents what we call a ``spatial approach.'' We correlate the mean environment with the mean of an ecological state or rate 
@@
 they provide no information about how quickly the system will move from the current state to the predicted, future state. Because the patterns in the data have developed 
-over long time spans, they are may not provide good information about dynamic changes on the order of a few years. Moreover, transient dynamics could 
+over long time spans, they are unlikely to provide much information about dynamic changes on the order of years to decades. Moreover, transient dynamics could 
 prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
@@
 a limited spatial extent (but see \citealt{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
-that could alter the underlying process at longer time scales and improve forecastability \citep{clark_ecological_2001}.
+that could alter the underlying process at longer time scales.
 
-We compare forecasts from both models to the simulated dynamics to determine how well the two approaches to forecasting perform at different time scales.
-We also assess the potential for combining the information available in both spatial and temporal patters by combining forecasts from the spatial and temporal approaches using a weighted average and assess how the optimal model weights change over time.
-Based on the hypothesis that spatial patterns capture longer-term processes we expect the temporal model to best predict short-term dynamics, and the spatial model to best predict long-term dynamics, and that a combination of the two approaches will provide the best forecasts at transitional intermediate time scales.
+We compare forecasts from both models to the true, simulated dynamics to show that responses to environmental change reflect both kinds of processes, but at different time scales.
+To quantify this pattern, we combine forecasts from the spatial and temporal approaches with a weighted average and then show how the optimal model weights change over time: we expect the temporal model to best predict short-term dynamics, and the spatial model to best predict long-term dynamics.
+How rapidly that transition occurs depends on the rate of  ``slow" processes such as species turnover and evolution.
+Code for simulating dynamics and reproducing the analyses in this paper is available on Github \texttt{https://github.com/pbadler/space-time-forecast}) and
+archived on Zenodo \texttt{https://zenodo.org}).
 
@@
 
-A common problem for conservation biologists and natural resource managers is anticipating the impact of environmental change on the abundance of 
+A common need for conservation biologists and natural resource managers is anticipating the impact of environmental change on the abundance of 
 endangered species, biological invaders, and harvested species. Although the managers may be primarily interested in just the focal species, 
@@
 Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
-All computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast}).
 
@@
 
-Dynamics during the baseline period revealed very strong spatial patterns across the mean temperature gradient. 
+During the baseline period there were strong spatial patterns across the mean temperature gradient. 
 Individual species, including our focal species, showed classic, unimodal ``Whittaker" patterns of abundances across the gradient (Fig. \ref{fig:species-patterns-models}A).
@@
 
-We forecasted the effect of temperature increase on the focal species' biomass at one location in the center of the temperature gradient 
-using the spatial models and the temporal models. 
-The predictions from the spatial and temporal model contrasted markedly, with the temporal 
-model predicting a large increase in biomass and the spatial model predicting a decrease. Initially, the simulated abundances followed the increase predicted by the 
-temporal model, but as faster-growing species colonized and increased in abundance at the study site, the biomass of the focal species then decreased, eventually falling below its baseline level (Fig.
+We forecast the effect of a temperature increase on the focal species' biomass at one location in the center of the temperature gradient 
+using both spatial models and the temporal models. 
+The predictions from the spatial and temporal models contrasted markedly, with the temporal 
+model predicting a large increase in biomass and the spatial model predicting a decrease.Initially, the simulated abundances followed the increase predicted by the 
+temporal model, but as faster-growing species colonized and increased in abundance at the focal site, the biomass of the focal species decreased, eventually falling below its baseline level (Fig.
 \ref{fig:community-forecast-species}).
@@
 At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{models} for full description of the approach).
-This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}) 
-because the weights were determined by fitting directly to those data. Unlike our spatial and temporal model forecasts, 
+This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}), but 
+it works so effectively because the weights were determined by fitting directly to those data. Unlike our spatial and temporal model forecasts, 
 we did not generate out-of-sample predictions from the combined model; it merely provides 
-a convenient way to quantify how rapidly dynamics shift from from being dominated by the interannual 
-variations captured in the temporal model to being dominated by the steady-state equilibrium captured by the spatial model.
-The most rapid shifts in model weights occur during the period when warm-adapted, faster growing species are increasing most rapidly
-in abundance (Fig. \ref{fig:community-weights-spp}). A true, combined forecast would require a method to determine the model weights \emph{a priori}.
+a convenient way to quantify how rapidly dynamics shift from being dominated by interannual 
+variation captured in the temporal model to being dominated by the steady-state equilibrium captured by the spatial model.
+A true, combined forecast would require a method to determine the model weights \emph{a priori}.
+The most rapid shifts in the model weights occurred during the period when warm-adapted, faster growing species were increasing most rapidly
+in abundance (Fig. \ref{fig:community-weights-spp}).
 

### 2019-07-22 doc/manuscript.tex
Editorial work on Eco-Evo example

@@
 How rapidly that transition occurs depends on the rate of  ``slow" processes such as species turnover and evolution.
-Code for simulating dynamics and reproducing the analyses in this paper is available on Github \texttt{https://github.com/pbadler/space-time-forecast}) and
-archived on Zenodo \texttt{https://zenodo.org}).
 
@@
 
-A common need for conservation biologists and natural resource managers is anticipating the impact of environmental change on the abundance of 
+A common problem for conservation biologists and natural resource managers is anticipating the impact of environmental change on the abundance of 
 endangered species, biological invaders, and harvested species. Although the managers may be primarily interested in just the focal species, 
@@
 Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
+All computer code is available on Github: \texttt{https://github.com/pbadler/space-time-forecast}).
 
@@
 
-During the baseline period there were strong spatial patterns across the mean temperature gradient. 
+Dynamics during the baseline period revealed very strong spatial patterns across the mean temperature gradient. 
 Individual species, including our focal species, showed classic, unimodal ``Whittaker" patterns of abundances across the gradient (Fig. \ref{fig:species-patterns-models}A).
@@
 
-We forecast the effect of a temperature increase on the focal species' biomass at one location in the center of the temperature gradient 
-using both spatial models and the temporal models. 
-The predictions from the spatial and temporal models contrasted markedly, with the temporal 
-model predicting a large increase in biomass and the spatial model predicting a decrease.Initially, the simulated abundances followed the increase predicted by the 
-temporal model, but as faster-growing species colonized and increased in abundance at the focal site, the biomass of the focal species decreased, eventually falling below its baseline level (Fig.
+We forecasted the effect of temperature increase on the focal species' biomass at one location in the center of the temperature gradient 
+using the spatial models and the temporal models. 
+The predictions from the spatial and temporal model contrasted markedly, with the temporal 
+model predicting a large increase in biomass and the spatial model predicting a decrease. Initially, the simulated abundances followed the increase predicted by the 
+temporal model, but as faster-growing species colonized and increased in abundance at the study site, the biomass of the focal species then decreased, eventually falling below its baseline level (Fig.
 \ref{fig:community-forecast-species}).
@@
 At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{models} for full description of the approach).
-This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}), but 
-it works so effectively because the weights were determined by fitting directly to those data. Unlike our spatial and temporal model forecasts, 
+This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}) 
+because the weights were determined by fitting directly to those data. Unlike our spatial and temporal model forecasts, 
 we did not generate out-of-sample predictions from the combined model; it merely provides 
-a convenient way to quantify how rapidly dynamics shift from being dominated by interannual 
-variation captured in the temporal model to being dominated by the steady-state equilibrium captured by the spatial model.
-A true, combined forecast would require a method to determine the model weights \emph{a priori}.
-The most rapid shifts in the model weights occurred during the period when warm-adapted, faster growing species were increasing most rapidly
-in abundance (Fig. \ref{fig:community-weights-spp}).
+a convenient way to quantify how rapidly dynamics shift from from being dominated by the interannual 
+variations captured in the temporal model to being dominated by the steady-state equilibrium captured by the spatial model.
+The most rapid shifts in model weights occur during the period when warm-adapted, faster growing species are increasing most rapidly
+in abundance (Fig. \ref{fig:community-weights-spp}). A true, combined forecast would require a method to determine the model weights \emph{a priori}.
 
@@
 \ref{fig:evo_evo_spatial_temporal_models}A).
-The full model description is provided in Appendix \ref{eco-evo}, and all computer code is available on Github  (\texttt{https://github.com/pbadler/space-time-forecast}).
+The full model description is provided in Appendix \ref{eco-evo}.
 Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
@@
 \ref{fig:evo_evo_spatial_temporal_models}C).
+We then imposed a period of warming, followed by a final period of higher stationary temperature (Fig.
+\ref{fig:forecast} top).
 
-We then imposed a period of warming, followed by a final period of higher, but stationary, temperature (Fig.
-\ref{fig:forecast} top).
 With the onset of warming, the population crashed as the cold-adapted genotype decreased in abundance.
@@
 \ref{fig:forecast} bottom).
-The implications for forecasting are clear: In the initial stages of the warming trend, the temporal model fit to the original, baseline conditions should make better the best predictions about resulting population trends, but in the longer term, as evolutionary change occurs, the spatial model ultimately makes better predictions.
+The temporal model accurately predicted the impact of the initial warming trend, but eventually became too pessimistic, while the spatial model did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
+\ref{fig:forecast} bottom).
 
-As in the eco-evolutionary example, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time. 
-The temporal model accurately predicted the impact of the initial warming trend, but eventually became far too pessimistic, while the spatial model did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
-\ref{fig:forecast} bottom).
+As in the community turnover example, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time.
 The combined model initially reflected the temporal model, but then rapidly transitioned to reflect the spatial model.
-The rapid transition in the weighting term, $\omega$, occured during the period of most rapid change in genotype frequencies (Fig.
+The rapid transition in the weighting term, $\omega$, occurred during the period of most rapid change in genotype frequencies (Fig.
 \ref{fig:forecast_supp}).

### 2019-07-24 doc/manuscript.tex
Editorial changes to Discussion

@@
 \ref{fig:evo_evo_spatial_temporal_models}A).
-The full model description is provided in Appendix \ref{eco-evo}.
+The full model description is provided in Appendix \ref{eco-evo}, and all computer code is available on Github  (\texttt{https://github.com/pbadler/space-time-forecast}).
 Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
@@
 \ref{fig:evo_evo_spatial_temporal_models}C).
-We then imposed a period of warming, followed by a final period of higher stationary temperature (Fig.
+
+We then imposed a period of warming, followed by a final period of higher, but stationary, temperature (Fig.
 \ref{fig:forecast} top).
-
 With the onset of warming, the population crashed as the cold-adapted genotype decreased in abundance.
@@
 \ref{fig:forecast} bottom).
-The temporal model accurately predicted the impact of the initial warming trend, but eventually became too pessimistic, while the spatial model did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
+The implications for forecasting are clear: In the initial stages of the warming trend, the temporal model fit to the original, baseline conditions should make better the best predictions about resulting population trends, but in the longer term, as evolutionary change occurs, the spatial model ultimately makes better predictions.
+
+As in the eco-evolutionary example, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time. 
+The temporal model accurately predicted the impact of the initial warming trend, but eventually became far too pessimistic, while the spatial model did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
 \ref{fig:forecast} bottom).
-
-As in the community turnover example, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time.
 The combined model initially reflected the temporal model, but then rapidly transitioned to reflect the spatial model.
-The rapid transition in the weighting term, $\omega$, occurred during the period of most rapid change in genotype frequencies (Fig.
+The rapid transition in the weighting term, $\omega$, occured during the period of most rapid change in genotype frequencies (Fig.
 \ref{fig:forecast_supp}).
@@
 \subsection*{Discussion}
+Ecological forecasts are typically made using models fit to either time-series data or spatial data based on the concept of space-for-time substitutions.
+Our results demonstrate that forecasts based on these two approaches can lead to very different predictions for the future state of ecological systems.
+Which approach provides the most accurate forecasts depends on the time-horizon, or how far into the future we are attempting to forecast.
+For short-term forecasts, time-series approaches perform best, whereas longer-term forecasts are more accurately made using models from spatial data.
+This stems from the different data types capturing information on the operation of different processes.
+Local time-series data capture autocorrelation (including lagged responses) and responses of current assemblages to small changes in environmental conditions.
+In addition, the state of the system in the next time step depends heavily on the current state (i.e., the initial condition).  
+Since short-term forecasts do not typically require extrapolating into novel conditions; the historical range of variation, incorporating lags and accurate
+initial conditions present in local time-series data is likely to be the most generally userful approach for forecasting.
+In contrast, data spanning spatial gradients captures the long-term outcome of slower processes, like selection and dispersal. This makes models based on these
+data more likely to accurately predict long-term steady state outcomes of ecological change.
 
-Our simulations demonstrate three lessons:
-
-\emph{Lesson 1: For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow 
-processes such as evolutionary and ecological selection as well as dispersal.} [Ethan, I am hoping you can fix and add to this paragraph, perhaps
-drawing from examples in other fields?] In the short-term, time-series approaches work well because lag effects are strong: the state of the system in the next 
-time step depends heavily on the current state. Furthermore, in the short-term we usually are not extrapolating far into novel conditions; the historical range of variation 
-may be a good guide when there has been little time for genotype frequencies or species composition to change. In contrast, farther we forecast into the future, 
-the weaker the influence of the current state, and the more likely that slow processes like selection and dispersal will alter genotype frequencies and
-species composition.
-
-This lesson has implications for the emerging field of ecological forecasting. If we only focus on iterative, near-term forecasting, we are likely to find 
-that time-series approaches work well, and that consideration of slow processes like evolution, dispersal, and compositional turnover rarely improves
-forecast skill. Assuming that the same models will make good longer-term forecasts would be a mistake.
-
-We do not mean to undermine the call for an emphasis on near-term forecasting. Building capacity to make near-term forecasts is extremely important, not
-only because of the many critical applications (examples?) but because it is a necessary step to building a disciplinary culture and practice of ecological 
-forecasting (cite Dietze?). It is also feasible, because we have the data to validate forecasts at short time-scales. Rather, the point is that near-term forecasting
-alone is not enough. We also need to build capacity to make forecasts at time-scales where slow ecological processes cannot be ignored. The challenge 
-will be finding ways to validate long-term predictions.
-
-\emph{Lesson 2. Different kinds of data reflect the operation of different processes; the choice of data for model training should depend on the time-scale of the desired forecast}.
-Longitudinal data capture autocorrelation and responses of current assemblages, while data spanning spatial gradients capture the long-term outcome of 
-selection and dispersal. For short-term forecasts, building models using longitudinal data is clearly the best choice. As our simulations show, models fit to 
-steady-state, spatial patterns make terrible predictions in the short term. Conversely, models based on longitudinal data make poor predictions in the long-term,
-while the models fit to spatial patterns can perform well. Although the temporal and spatial models we fit are phenomenological, they capture different kinds
- of underlying processes. The traditional debate about phenomenological vs. process-based models obscures the fact that, depending on the choice of the
- training data, phenomenological models can represent different kinds of processes.
-
- Why not simply build a process-based model? Perhaps something like the process-based models we used to generate the simulated time series in the first place? 
- That kind of model could make good predictions at both short and long time-scales. Unfortunately, this approach is just not feasible in most cases.
- The first challenge is simply determining the critical process. Take our eco-evolutionary example: should we simply assume that every annual plant species
- has warm and cold-adapted genotypes? And that temperature adaptation will be the key trait driving the response to warming, as opposed to the effects of 
- compositional turnover highlighted in our second example? Second, even if the key process is known in advance, fitting a mechanistic model presents its
- own set of challenges. The model for our eco-evolutionary example requires, in order of increasing empirical difficulty, demographic rates for the full life
- cycle, including density dependence, a quantitative mapping of genotype-by-environment interactions onto fitness, and dispersal kernels. The model for 
- our community and ecosystem example requires demography, environment-dependent intra- and interspecific competition coefficients, and dispersal kernels.
- We can think of empirical models that capture some of these features, but not all of them. Finally, even were it possible to fit such a model, the high number of 
- parameters and associated parameter error might lead to poor predictive skill (refs?).
-
-Inevitably, models used for ecological forecasting will be partially if not entirely phenomenological. But that does not mean that they will not rely on process-based
-understanding in some way. The point of our simulations is that different processes should be considered for different forecast time-scales, and this can be done 
-by fitting phenomenological models to different kinds of data sets. Even when process-level understanding does not enable a fully mechanistic model, it can
-improve the specification of phenomenological models.
-
-\emph{Lesson 3. A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.}
-Our simulated examples suggest the potential for a combined forecast, which weights predictions from spatial and temporal models, to 
-accurately forecast both short and long-term dynamics. But we cheated: our combined ``forecasts'' were fit to data, not generated prospectively
-using \emph{a priori} model weights. Finding a general solution for determining those weights poses a fascinating research problem whose 
-solution would lead to important applications in ecological forecasting.
-
-We see a role for new theory and new empirical tests. On the theoretical side, what are the parameters that determine how quickly slow processes
-influence dynamics?
-We can play with our simulation models to provide some qualitative answers.
-The effects of some parameters are intuitive: in the community turnover example, increasing the fraction of dispersing individuals causes a more rapid shift in species composition and in model weights (Fig.
-\ref{fig:dispersal_niche_width}A).
-But other parameters have less straightforward effects.
-In the eco-evolutionary model, a potentially important parameter is the temperature tolerance of each genotype.
-We initially thought that increasing the temperature tolerance might accelerate shifts in genotype frequencies by increasing genotypic diversity within populations, but it had the opposite effect (Fig.
-\ref{fig:dispersal_niche_width}B), probably because it decreased the strength of selection, slowing the rate of change in genotype frequencies.
-Other factors to consider include organism lifespans and the magnitude of directional environmental change relative to interannaul variation.
-While simulation experiments are easy and informative, they cannot provide the generality of analytical theory.
-Unfortunately, deriving analytical solutions would be difficult, even for our relatively simple models.
+These results suggest that near-term forecasts will likely be most effective when made using time-series models, a result backed by recent empirical
+analyses of biodiversity forecasts at time scales from one to ten years (Harris et al. 2018). For long-term forecasts our results suggest the use
+of models built on spatial data assuming. However, determining what is "near-term" and what is "long-term" in this context is an open question, and
+many important forecasts in ecology will be needed at intermediate time horizons (on the scale of one to two decades) where neither approach will be
+fully sufficient. Our simulation results demonstrate that there are often extended transitional period where neither approach is effective on its own,
+but where a weighted combination of time-series and spatial models can produce highly accurate predictions. Unfortunately there is currently no clear
+way to determine how to set the weights for the different approaches for different time horizons. Our combined ``forecasts'' used the performance of
+the forecasts on simulated data to select the weights and therefore our approach cannot be used for actual forecasting. Determining how to set these
+weights could be pursued using both theoretical and empirical approaches. On the theoretical side, it should be possible to explore the influence of different
+parameters on how quickly slow processes (represented in spatial data) begin to influence dynamics. The effects of some parameters are intuitive:
+in the community turnover example, increasing the fraction of dispersing individuals causes a more rapid shift in species composition and in model weights
+(Fig. \ref{fig:dispersal_niche_width}A). Other factors to consider include organism lifespans and the magnitude of directional environmental change relative
+to interannaul variation. From an empirical perspective, we can address this issue by creating data sets to estimate the weighting function. By evaluating both time-series and
+spatial models for large numbers of forecasts across a range of time-horizons it should be possible to be begin to develop rules of thumb for ensembling
+these two modeling approaches. This requires data sets that can support the evaluation of long-term forecasts, which could include paleoecological data (e.g.,),
+long-term historical data sets (e.g., ), and the use of model systems with short-generation times (e.g., Petchy), With enough case studies, we might be able
+to infer patterns that could guide applications in new systems.
 
@@
 
-On the empirical side, the key problem is finding or creating data sets to estimate the weighting function. As described under Lesson 1,
-this requires data sets appropriate for validating long-term forecasts.
-Possibilities include hindcasting with paleoecological and long-term historical data sets, use of model
-systems with short-generation times, and....more simulation studies...? . With enough case studies, we might be able to infer patterns
-that could guide applications in new systems.
+Similar challenges with intermediate time horizons occur in other forecasting contexts. Weather forecasts based one type of models are very effective for
+forecasting a week to ten days in advance, but then become largely uninformative. Climate models provide accurate predictions for similar variables, but
+generally at decadal time-scales. Forecasting in between these two time scales has been challenging.
 
-Anything else to add?
+For short-term forecasts, building models using time-series data will often be the best choice when this type of data is available. As our simulations show,
+
+require understanding slow 
+processes such as evolutionary and ecological selection as well as dispersal.} [Ethan, I am hoping you can fix and add to this paragraph, perhaps
+drawing from examples in other fields?] In contrast, the further we forecast into the future, 
+the weaker the influence of the current state, and the more likely that slow processes like selection and dispersal will alter genotype frequencies and
+species composition.
+
+; the choice of data for model training should depend on the time-scale of the desired forecast.
+
+One potential solution to these challenges is to build fully process based models of ecological systems. If it is possible to accurately characterize all
+of the processes governing a system then a model based on that understanding should make accurate predictions at all time-horizons. For example, if we had
+included the process-based models that we used to generate the simulated time series, these models would have been able to accurately capture
+the full dynamics and forecasts for all time horizons. In concept, process based models should also be more robust for making predictions outside of historically
+observed conditions (or beyond the conditions observed across spatial gradients). This is particularly important for making predictions in a future that is
+predicted to be increasingly "non-analog", where combinations of environment and species interactions are operating that have never occured before (Williams et al.)
+Unfortunately, this approach is not currently feasible in most cases, because we lack a detailed
+knowledge of all of the complex and interacting processes influencing the dynamics of real systems...
+The first challenge is simply determining the critical process. Take our eco-evolutionary example: should we simply assume that every annual plant species
+has warm and cold-adapted genotypes? And that temperature adaptation will be the key trait driving the response to warming, as opposed to the effects of 
+compositional turnover highlighted in our second example? Second, even if the key process is known in advance, fitting a mechanistic model presents its
+own set of challenges. The model for our eco-evolutionary example requires, in order of increasing empirical difficulty, demographic rates for the full life
+cycle, including density dependence, a quantitative mapping of genotype-by-environment interactions onto fitness, and dispersal kernels. The model for 
+our community and ecosystem example requires demography, environment-dependent intra- and interspecific competition coefficients, and dispersal kernels.
+We can think of empirical models that capture some of these features, but not all of them. Finally, even were it possible to fit such a model, the high number of 
+parameters and associated parameter error might lead to poor predictive skill (refs?)... In contrast, for our simulations models based on time-series data make poor predictions in the long-term,
+while the models fit to spatial patterns can perform well. Although the temporal and spatial models we fit are phenomenological, they capture different kinds
+ of underlying processes. The traditional debate about phenomenological vs. process-based models obscures the fact that, depending on the choice of the
+ training data, phenomenological models can represent different kinds of processes.
+
+Inevitably, models used for ecological forecasting will be partially if not entirely phenomenological. But that does not mean that they will not rely on process-based
+understanding in some way. The point of our simulations is that different processes should be considered for different forecast time-scales, and this can be done 
+by fitting phenomenological models to different kinds of data sets. Even when process-level understanding does not enable a fully mechanistic model, it can
+improve the specification of phenomenological models.
+
+These results have important implications for the emerging field of ecological forecasting. If we only focus on iterative, near-term forecasting, we are likely to find 
+that time-series approaches work well, and that consideration of slow processes like evolution, dispersal, and compositional turnover rarely improves
+forecast skill. Assuming that the same models will make good longer-term forecasts would be a mistake. We do not mean to undermine the call for an emphasis
+on near-term forecasting. Building capacity to make near-term forecasts is extremely important, not
+only because of the many critical applications (examples?) but because it is a necessary step to building a disciplinary culture and practice of ecological 
+forecasting (cite Dietze?). It is also feasible, because we have the data to validate forecasts at short time-scales. Rather, the point is that near-term forecasting
+alone is not enough. We also need to build capacity to make forecasts at time-scales where slow ecological processes cannot be ignored. The challenge 
+will be finding ways to validate long-term predictions.
 

### 2019-07-29 doc/manuscript.tex

@@
 \subsection*{Discussion}
+
 Ecological forecasts are typically made using models fit to either time-series data or spatial data based on the concept of space-for-time substitutions.
@@
 For short-term forecasts, time-series approaches perform best, whereas longer-term forecasts are more accurately made using models from spatial data.
-This stems from the different data types capturing information on the operation of different processes.
-Local time-series data capture autocorrelation (including lagged responses) and responses of current assemblages to small changes in environmental conditions.
-In addition, the state of the system in the next time step depends heavily on the current state (i.e., the initial condition).  
-Since short-term forecasts do not typically require extrapolating into novel conditions; the historical range of variation, incorporating lags and accurate
-initial conditions present in local time-series data is likely to be the most generally userful approach for forecasting.
-In contrast, data spanning spatial gradients captures the long-term outcome of slower processes, like selection and dispersal. This makes models based on these
-data more likely to accurately predict long-term steady state outcomes of ecological change.
+This stems from differences between data types in the types and scales of processes they capture.
+Local time-series data capture demographic processes, autocorrelation (including lagged responses) and responses of current assemblages to small changes
+in environmental conditions. In addition, the state of the system a short period into the future depends heavily on the current state (i.e., the initial
+condition). Since short-term forecasts do not typically require extrapolating into novel conditions; the historical range of variation, incorporating lags
+and accurate initial conditions present in local time-series data is likely to be the most generally useful approach for forecasting.
+In contrast, data spanning spatial gradients captures the long-term outcome of slower processes including selection, dispersal, and responses to large
+changes in the environment. This makes models based on these data more likely to accurately predict long-term steady state outcomes of ecological change.
 
@@
 analyses of biodiversity forecasts at time scales from one to ten years (Harris et al. 2018). For long-term forecasts our results suggest the use
-of models built on spatial data assuming. However, determining what is "near-term" and what is "long-term" in this context is an open question, and
-many important forecasts in ecology will be needed at intermediate time horizons (on the scale of one to two decades) where neither approach will be
+of models built on spatial data. However, determining what is "near-term" and what is "long-term" in this context is an open question, and
+many important forecasts in ecology will be needed at intermediate time horizons where neither approach will be
 fully sufficient. Our simulation results demonstrate that there are often extended transitional period where neither approach is effective on its own,
-but where a weighted combination of time-series and spatial models can produce highly accurate predictions. Unfortunately there is currently no clear
-way to determine how to set the weights for the different approaches for different time horizons. Our combined ``forecasts'' used the performance of
-the forecasts on simulated data to select the weights and therefore our approach cannot be used for actual forecasting. Determining how to set these
-weights could be pursued using both theoretical and empirical approaches. On the theoretical side, it should be possible to explore the influence of different
-parameters on how quickly slow processes (represented in spatial data) begin to influence dynamics. The effects of some parameters are intuitive:
+but where a weighted combination of time-series and spatial models can produce highly accurate predictions. Unfortunately there is no clear
+way to determine how to weight these two approaches for different time horizons. Our combined ``forecasts'' used the performance of
+the forecasts on simulated data to select the weights. This approach cannot be used for actual forecasting because it is effectively an empirical
+quantification of the observed performance of the forecasts at different time horizons. Determining how to set these
+weights a priori could be pursued using both theoretical and empirical approaches. On the theoretical side, it should be possible to explore the influence
+of different parameters on how quickly slow processes (represented in spatial data) begin to influence dynamics. The effects of some parameters are intuitive:
 in the community turnover example, increasing the fraction of dispersing individuals causes a more rapid shift in species composition and in model weights
-(Fig. \ref{fig:dispersal_niche_width}A). Other factors to consider include organism lifespans and the magnitude of directional environmental change relative
-to interannaul variation. From an empirical perspective, we can address this issue by creating data sets to estimate the weighting function. By evaluating both time-series and
-spatial models for large numbers of forecasts across a range of time-horizons it should be possible to be begin to develop rules of thumb for ensembling
-these two modeling approaches. This requires data sets that can support the evaluation of long-term forecasts, which could include paleoecological data (e.g.,),
-long-term historical data sets (e.g., ), and the use of model systems with short-generation times (e.g., Petchy), With enough case studies, we might be able
-to infer patterns that could guide applications in new systems.
+(Fig. \ref{fig:dispersal_niche_width}A). Other factors to consider include organism lifespans, niche widths, and the magnitude of directional environmental
+change relative to interannual variation. From an empirical perspective, this challenge could be addressed by using synthetic analyses to estimate the
+weighting function. By evaluating both time-series and spatial models for large numbers of forecasts across a range of time-horizons it should be possible
+to develop rules of thumb for ensembling models encompassing different scales of processes. This requires data sets that can support the evaluation of
+long-term forecasts, which could include paleoecological data (e.g.,), long-term historical data sets (e.g., ), and the use of model systems with
+short-generation times (e.g., Petchy), With enough case studies, we might be able to infer patterns that could guide applications in new systems.
 
@@
 
-Similar challenges with intermediate time horizons occur in other forecasting contexts. Weather forecasts based one type of models are very effective for
-forecasting a week to ten days in advance, but then become largely uninformative. Climate models provide accurate predictions for similar variables, but
-generally at decadal time-scales. Forecasting in between these two time scales has been challenging.
+Instead of trying to weight distinct approaches an alternative approach is to build fully process based models of ecological systems  for forecasting.
+If it is possible to accurately characterize all of the processes governing a system then a model based on that understanding should make accurate
+predictions at all time-horizons. For example, if we had evaluated the exact process-based models that we used to generate the simulated time series,
+these models would have been able to accurately capture the full dynamics and forecasts for all time horizons. In concept, process based models should
+also be more robust for making predictions outside of historically observed conditions (or beyond the conditions observed across spatial gradients).
+This is particularly important for making predictions in a future that is predicted to be increasingly "non-analog", where combinations of environment
+and species interactions exist that have never occurred before (Williams et al.) Unfortunately, this approach is not currently feasible in most cases,
+because we lack a detailed knowledge of all of the complex and interacting processes influencing the dynamics of real ecological systems. Even if the
+general form of the models was understood estimating the high number of parameters and quantifying how they vary across
+ecosystems typically require more data than is currently available for even well studied systems. As a result, models used for ecological forecasting
+will include at least some phenomenological components. But that does not mean that these forecasts do not rely on process-based understanding in some way.
+The message from our simulations is that different processes should be considered for different forecast time-scales, and this can be done by fitting
+models to different kinds of data sets. Even when process-level understanding does not enable a fully mechanistic model, it can improve the specification
+of phenomenological models.
 
-For short-term forecasts, building models using time-series data will often be the best choice when this type of data is available. As our simulations show,
+Another approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is joint spatiotemporal modeling...
 
-require understanding slow 
-processes such as evolutionary and ecological selection as well as dispersal.} [Ethan, I am hoping you can fix and add to this paragraph, perhaps
-drawing from examples in other fields?] In contrast, the further we forecast into the future, 
-the weaker the influence of the current state, and the more likely that slow processes like selection and dispersal will alter genotype frequencies and
-species composition.
-
-; the choice of data for model training should depend on the time-scale of the desired forecast.
-
-One potential solution to these challenges is to build fully process based models of ecological systems. If it is possible to accurately characterize all
-of the processes governing a system then a model based on that understanding should make accurate predictions at all time-horizons. For example, if we had
-included the process-based models that we used to generate the simulated time series, these models would have been able to accurately capture
-the full dynamics and forecasts for all time horizons. In concept, process based models should also be more robust for making predictions outside of historically
-observed conditions (or beyond the conditions observed across spatial gradients). This is particularly important for making predictions in a future that is
-predicted to be increasingly "non-analog", where combinations of environment and species interactions are operating that have never occured before (Williams et al.)
-Unfortunately, this approach is not currently feasible in most cases, because we lack a detailed
-knowledge of all of the complex and interacting processes influencing the dynamics of real systems...
-The first challenge is simply determining the critical process. Take our eco-evolutionary example: should we simply assume that every annual plant species
-has warm and cold-adapted genotypes? And that temperature adaptation will be the key trait driving the response to warming, as opposed to the effects of 
-compositional turnover highlighted in our second example? Second, even if the key process is known in advance, fitting a mechanistic model presents its
-own set of challenges. The model for our eco-evolutionary example requires, in order of increasing empirical difficulty, demographic rates for the full life
-cycle, including density dependence, a quantitative mapping of genotype-by-environment interactions onto fitness, and dispersal kernels. The model for 
-our community and ecosystem example requires demography, environment-dependent intra- and interspecific competition coefficients, and dispersal kernels.
-We can think of empirical models that capture some of these features, but not all of them. Finally, even were it possible to fit such a model, the high number of 
-parameters and associated parameter error might lead to poor predictive skill (refs?)... In contrast, for our simulations models based on time-series data make poor predictions in the long-term,
-while the models fit to spatial patterns can perform well. Although the temporal and spatial models we fit are phenomenological, they capture different kinds
- of underlying processes. The traditional debate about phenomenological vs. process-based models obscures the fact that, depending on the choice of the
- training data, phenomenological models can represent different kinds of processes.
-
-Inevitably, models used for ecological forecasting will be partially if not entirely phenomenological. But that does not mean that they will not rely on process-based
-understanding in some way. The point of our simulations is that different processes should be considered for different forecast time-scales, and this can be done 
-by fitting phenomenological models to different kinds of data sets. Even when process-level understanding does not enable a fully mechanistic model, it can
-improve the specification of phenomenological models.
-
-These results have important implications for the emerging field of ecological forecasting. If we only focus on iterative, near-term forecasting, we are likely to find 
-that time-series approaches work well, and that consideration of slow processes like evolution, dispersal, and compositional turnover rarely improves
-forecast skill. Assuming that the same models will make good longer-term forecasts would be a mistake. We do not mean to undermine the call for an emphasis
-on near-term forecasting. Building capacity to make near-term forecasts is extremely important, not
-only because of the many critical applications (examples?) but because it is a necessary step to building a disciplinary culture and practice of ecological 
-forecasting (cite Dietze?). It is also feasible, because we have the data to validate forecasts at short time-scales. Rather, the point is that near-term forecasting
-alone is not enough. We also need to build capacity to make forecasts at time-scales where slow ecological processes cannot be ignored. The challenge 
-will be finding ways to validate long-term predictions.
+These results have important implications for the emerging field of ecological forecasting. They suggest that evaluating both near-term and long-term
+forecasts is essential to fully understanding the best methods for forecasting and how well they perform. Our results further suggest that while
+single approaches may perform reasonably well at short and long time horizons, that intermediate time horizons may require a combination of information
+from spatial and temporal patterns. Related challenges with intermediate time horizons occur in other forecasting contexts. Weather forecasts based one
+type of models are very effective for forecasting a week to ten days in advance, but then become largely uninformative. Climate models provide accurate
+predictions for similar variables, but generally at decadal time-scales. Forecasting in between these two time scales has been challenging in meteorology
+and will likely be challenging in ecology as well. In combination that means that while the recent emphasis on near-term iterative forecasting (Dietze 
+et al. 2018), is important, we also need to build understanding and capacity for mkaing forecasts across the full range of scales of interest in
+ecology, from hours to decades.
 

### 2019-07-29 doc/manuscript.tex

@@
 How rapidly that transition occurs depends on the rate of  ``slow" processes such as species turnover and evolution.
+Code for simulating dynamics and reproducing the analyses in this paper is available on Github \texttt{https://github.com/pbadler/space-time-forecast}) and
+archived on Zenodo \texttt{https://zenodo.org}).
 

### 2019-07-29 doc/manuscript.tex

@@
 a limited spatial extent (but see \citealt{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
-that could alter the underlying process at longer time scales.
+that could alter the underlying process at longer time scales and improve forecastability \citep{clark_ecological_2001}.
 
-We compare forecasts from both models to the true, simulated dynamics to show that responses to environmental change reflect both kinds of processes, but at different time scales.
-To quantify this pattern, we combine forecasts from the spatial and temporal approaches with a weighted average and then show how the optimal model weights change over time: we expect the temporal model to best predict short-term dynamics, and the spatial model to best predict long-term dynamics.
-How rapidly that transition occurs depends on the rate of  ``slow" processes such as species turnover and evolution.
-Code for simulating dynamics and reproducing the analyses in this paper is available on Github \texttt{https://github.com/pbadler/space-time-forecast}) and
-archived on Zenodo \texttt{https://zenodo.org}).
+We compare forecasts from both models to the simulated dynamics to determine how well the two approaches to forecasting perform at different time scales.
+We also assess the potential for combining the information available in both spatial and temporal patters by combining forecasts from the spatial and temporal approaches using a weighted average and assess how the optimal model weights change over time.
+Based on the hypothesis that spatial patterns capture longer-term processes we expect the temporal model to best predict short-term dynamics, and the spatial model to best predict long-term dynamics, and that a combination of the two approaches will provide the best forecasts at transitional intermediate time scales.
 

### 2019-07-29 doc/manuscript.tex

@@
 
-A recent flurry of synthesis papers in high impact journals attests to surging interest in ecological forecasting. Many of the papers 
-argue that a commitment to ecological forecasting is necessary to guide the environmental policy and management
-decisions that will become increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
-A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery 
+* Forecasting is important
+* Two different approaches to this time-series & space-for-time
+	* Not clear which is best
+	* Some limited comparisons
+* Which approach is best may depend on the time-horizon
+* Here we explore this possibility using simulation based approach
+
+Forecasting is increasingly recognized as important to the advancement and application of ecological research.
+Ecological forecasting is necessary to guide the environmental policy and management
+decisions that are increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
+In addition, evaluating the accuracy of ecological forecasts can provide rigorous tests of predictions from ecological models that can
+advance understanding of the processes governing ecological systems  
 \citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
-Taken together, these papers, along with a new textbook \citep{dietze_ecological_2017} represent an emerging consensus that 
-forecasting should be a priority for ecological research.
+This combined value of advancing our fundamental understanding of ecology and applying that understanding to support management and decision making 
+makes that forecasting is an important priority for ecological research.
 
-No such consensus is available for exactly how to tackle this challenge.
-Ecologists rely on many different approaches and models for making predictions, and 
-frequently argue about the virtues of process-based vs. phenomenological models (refs?).
-We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g.,\citealt{Adler2012}),
-and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions \citep{Fitzpatrick,Veloz2012}.
-But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
-uncertainty (refs? \citealt{tredennick_we_2017}?).
+Ecological models used for forecasting have traditionally been based on either time-series modeling or space-for-time substitutions.
+Time-series approaches use the historical behavior of a system to fit models that describe the temporal dynamics of that system and use those
+dynamic models to make predictions about what will happen in the future. This is a common approach for population based forecasts and includes both
+process based population models and data-driven models capturing empirical patterns in data (e.g., \citep{ward_complexity_2014}). Time-series models
+capture processes like birth, death, small scale dispersal events, and short-term responses to environmental conditions.
+Alternatively, space-for-time modeling approaches develop models of how and ecological property varies in across space in response to environmental
+conditions. These spatial relationships are assumed to also hold for changes at a site through time, and forecasts are made by forecasting future
+environmental conditions and determining the associated ecological response. Most forecasts for future species distributions and biodiversity
+are based on this type of approach, with a models built to describe the environmental conditions where species occur across space being applied to
+forecast environmental conditions to determine where they will occur in the future (e.g., citation). Space-for-time models capture processes like
+immigration, extinction, and responses to large changes in environmental conditions over long time periods. Despite both of these approaches being
+widely used there has been little comparison of how they compare for forecasting and understanding the dynamics of ecological systems \citep{harris_forecasting_2018}.
 
-The pragmatic solution to these debates is clear: we should forecast using the models that make the best predictions.
-Identifying these models requires rigorously testing predictions against observations. However, many common ecological predictions,
-such as a species' geographic range in the year 2100, are virtually impossible to test (but see \citealt{Maguire2015}), 
-leading \citep{dietze_iterative_2018} to call for greater emphasis on near-term forecasting. The assumption is that building 
-our capacity to make short-term forecasts will ultimately improve our long-term forecasts.
+Whether historical dynamics, contemporary spatial patterns, or a combination of the two will serve as the best source of information for forecasting
+may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. This potential dependency on the "forecast horizon"; ()\textit{sensu} citation)
+results from lags in the response of ecological conditions to environmental change, changes in the importance of ecological processes with
+time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of conditions present in the data.
+At short forecast horizons (days to years) processes driving individual demographics tend to dominate (similar to the time scales present in that
+historical data), conditions are likely to stay within the range of the historical data at a site, and the current state of the system is likely
+to capture the influence of unmeasured processes. As a result, for near-term forecasts time-series approaches may effectively capture the key dynamics
+and result in accurate forecasts. In contrast, at long forecast horizons (decades) an ecosystem is likely to experience conditions that have not been
+historically observed () and be responding to large scale processes large scale climate and land use change that are difficult to detect in even long
+ecological time-series. Therefore, for the kinds of long-term forecasts typically conducted in biodiversity and species-distribution modeling
+space-for-time based approaches may include important information about how ecosystems respond to major shifts in climate over long periods and as a
+results produce better long-term forecasts. Using different approaches for different time horizons is common in other areas of forecasting,
+including forecasts for environmental conditions where models for short-term changes (weather) differ substantially from those for long-term changes
+(climate).
 
-Our goal is to consider a hypothesis that, in our view, has not been carefully considered by either the push for near-term forecasting or
-ongoing debates about competing modeling approaches. We expect that short and long-term forecasts will require models
-incorporating different kinds of processes and fit to different kinds of data sets. We demonstrate this idea using two simulation case studies, the first 
-focused on how competitive interactions affect population dynamics of a focal species, the second for  an 
-eco-evolutionary scenario. Our analyses show that: 
+Here we investigate the possibility that the best approaches for building models for ecological forecasting may depend on the time horizon of the forecast
+and explore how time-series and space-for-time approaches might be combined to address mid-term ecological forecasting. We do this using two simulation case studies, one 
+focused on how competitive interactions affect population dynamics of a focal species and one eco-evolutionary scenario. Our analyses show that: 
 \begin{enumerate}

### 2019-07-30 doc/manuscript.tex
Edited Intro

--- 2019-07-29 doc/manuscript.tex
@@
 
-A recent flurry of synthesis papers in high impact journals attests to surging interest in ecological forecasting. Many of the papers 
-argue that a commitment to ecological forecasting is necessary to guide the environmental policy and management
-decisions that will become increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
-A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery 
-\citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
-Taken together, these papers, along with a new textbook \citep{dietze_ecological_2017} represent an emerging consensus that 
-forecasting should be a priority for ecological research.
+Forecasting is increasingly recognized as important to the application and advancement of ecological research.
+Forecasts are necessary to guide environmental policy and management
+decisions about mitigation and adaption to global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
+But forecasts can also advance understanding of the processes governing ecological systems by providing rigorous tests of 
+model predictions \citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
+The combined benefits of informing policy and management and advancing basic research 
+makes forecasting an important priority for ecological research.
 
-No such consensus is available for exactly how to tackle this challenge.
-Ecologists rely on many different approaches and models for making predictions, and 
-frequently argue about the virtues of process-based vs. phenomenological models (refs?).
-We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g.,\citealt{Adler2012}),
-and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions \citep{Fitzpatrick,Veloz2012}.
-But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
-uncertainty (refs? \citealt{tredennick_we_2017}?).
+The models used for ecological forecasting have traditionally been based on either time-series approaches or space-for-time substitutions.
+The time-series approach involves fitting models to long-term datasets to describe the temporal dynamics of a system.  We then use those
+dynamic models to make predictions about what will happen in the future. This is the common approach for population forecasts and includes 
+both process-based and data-driven models (e.g., \citep{ward_complexity_2014}). Whether process-based or
+data-driven, time-series models typically capture processes operating on interannual time-scales, such as birth, death, 
+small scale dispersal events, and short-term responses to environmental conditions.
+Space-for-time substitution approaches begin by describing how an ecological variable of interest, such as occupancy or productivity, 
+varies across sites experiencing different environmental conditions. These spatial relationships between environment and ecological 
+response are assumed to also hold for changes at a site through time. To make a forecast, we first predict the future
+environmental conditions and then determinie the associated ecological response, based on the observed spatial relationship. 
+Most forecasts for future species distributions and biodiversity are based on this approach. 
+Space-for-time models capture processes operating over long time periods, such as
+immigration, extinction, and responses to large or prolonged environmental changes. Although both of these approaches are
+widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or 
+advancing our understanding of ecological dynamics \citep{harris_forecasting_2018} (but see Renwick et al ??).
 
-The pragmatic solution to these debates is clear: we should forecast using the models that make the best predictions.
-Identifying these models requires rigorously testing predictions against observations. However, many common ecological predictions,
-such as a species' geographic range in the year 2100, are virtually impossible to test (but see \citealt{Maguire2015}), 
-leading \citep{dietze_iterative_2018} to call for greater emphasis on near-term forecasting. The assumption is that building 
-our capacity to make short-term forecasts will ultimately improve our long-term forecasts.
+Whether historical dynamics, contemporary spatial patterns, or spme combination of the two will serve as the best source of 
+information for forecasting may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. 
+This potential dependency on the "forecast horizon"; ()\textit{sensu} citation)
+reflects lags in the response of ecological conditions to environmental change, shifts in the importance of ecological processes with
+time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of 
+environmental conditions represented in observed data.
+At short forecast horizons (days to years), dynamics will reflect the physiological and demographic responses of 
+the organisms present at a site more than than turnover of genotypes or species, 
+environmental conditions are likely to stay within the range of historical variation, 
+and the current state of the system is likely to capture the influence of unmeasured processes. As a result, 
+for near-term forecasts time-series approaches may capture the key dynamics and provide accurate predictions.
 
-Our goal is to consider a hypothesis that, in our view, has not been carefully considered by either the push for near-term forecasting or
-ongoing debates about competing modeling approaches. We expect that short and long-term forecasts will require models
-incorporating different kinds of processes and fit to different kinds of data sets. We demonstrate this idea using two simulation case studies, the first 
-focused on how competitive interactions affect population dynamics of a focal species, the second for  an 
-eco-evolutionary scenario. Our analyses show that: 
+In contrast, at long forecast horizons (decades to centuries),  environmental conditions that have not been 
+historically observed are likely to not only occur but to persist long enough to drive significant turnover of genotypes and species 
+along with changes in the flux of energy and nutrients.  At these long scales, the current state of the system may be 
+little help in predicting the future state. For the century-scale forecasts often featured in biodiversity and 
+species-distribution modeling, space-for-time approaches may effectively capture the response of ecosystems to major shifts 
+in climate over long periods, producing better long-term forecasts than time-series approaches. 
+Using different modeling approaches for different forecast horizons is common in other disciplines.
+For example, meteorological models for short-term weather forecasts differ substantially in spatial
+and temporal resolution and extent from the global circulation models used to predict long-term changes
+in climate [IS THIS REALLY TRUE? REFERENCE?].
+
+Here we use simulation models to 1) demonstrate that the best model-building approaches for ecological forecasting 
+depend on the time horizon of the forecast, and 2) explore how time-series and space-for-time approaches might be 
+combined to make better forecasts at intermediate time scales. We conduct two simulation case studies, one 
+focused on how interspecific interactions affect the population dynamics of a focal species, and the second focused on
+an eco-evolutionary scenario. Our analyses show that: 
 \begin{enumerate}
-	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow 
-	processes such as evolutionary and ecological selection as well as dispersal.
-	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while 
-	data spanning spatial gradients capture the long-term outcome of selection and dispersal.  Whether predictive models should be trained using longitudinal or
-	spatial data sets, or both, depends on the time-scale of the desired forecast.
+	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow processes such as evolutionary and ecological selection as well as dispersal.
+	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while data spanning spatial gradients capture the long-term outcome of selection and dispersal.
+Whether predictive models should be trained using longitudinal or spatial data sets, or both, depends on the time-scale of the desired forecast.
 	\item A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.

### 2019-07-31 doc/manuscript.tex
Intermediate save

@@
 small scale dispersal events, and short-term responses to environmental conditions.
+
 Space-for-time substitution approaches begin by describing how an ecological variable of interest, such as occupancy or productivity, 
@@
 In two case studies, we simulate the effects of an increase in temperature on simple systems with known dynamics.
-The ``truth'' is represented by a model that is mechanistic for at least one important process, but treat the model as unknown when
-analyzing the data and assume that perfectly recovering this model would not be possible in practice. We begin each simulation
-under stationary temperature, allowing the system to equilibrate; we call this the baseline phase. We then increase temperature progressively
-over a period of time, followed by a second period of stationary, now elevated, temperature. The objective is to forecast the response of the 
-system to the temperature increase based on data gathered during the baseline period.
+The ``truth'' is represented by a model that is mechanistic for at least one important process, but we treat the model as unknown when
+analyzing the data and we assume that perfectly recovering this model would not be possible in practice. We begin each simulation
+under stationary temperature, allowing the system to equilibrate; we call this the baseline phase. 
+We then increase temperature progressively over a period of time, followed by a second period of stationary, now elevated, temperature. 
+The objective is to forecast the response of the system to the temperature increase based on data gathered during the baseline period.
 
 We make forecasts based on two phenomenological models, each representing processes operating at different time scales. 
-The first phenomenological model represents what we call a ``spatial approach.'' We correlate the mean environment with the mean of an ecological state or rate 
- across many sites. This is the approach commonly used to predict population distribution or abundance as a function of climate \citep{elith_species_2009} or
- primary production as a function of mean precipitation \citep{Sala1988}. The approach is attractive because it addresses broad spatial scales relevant to 
- management, and because the necessary data are widely available and statistical analyses are straightforward. The approach also has well-known weaknesses.
-Criticism often focuses on the assumption that current spatial patterns are in equilibrium with climate (refs?). Another issue is that these models are not dynamic; 
-they provide no information about how quickly the system will move from the current state to the predicted, future state. Because the patterns in the data have developed 
-over long time spans, they are may not provide good information about dynamic changes on the order of a few years. Moreover, transient dynamics could 
-prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
+The first phenomenological model represents the space-for-time substiution approach, which we call the ``spatial approach'' for brevity. 
+We correlate the mean environment with the mean of an ecological state or rate across many sites. 
+This is the approach commonly used to predict population distribution or abundance as a function of climate \citep{elith_species_2009} 
+or primary production as a function of mean precipitation \citep{Sala1988}.
+The approach is attractive because it addresses broad spatial scales relevant to  management, and because the necessary data are widely available and the statistical analyses straightforward.
+The approach also has well-known weaknesses.
+Criticism often focuses on the assumption that current spatial patterns are in equilibrium with climate (refs?).
+Another issue is that these models are not dynamic;  they provide no information about how quickly the system will move from the current state to the predicted, future state.
+Because the patterns in the data have developed  over long time spans, they may not provide good information about dynamic changes on the order of a few years.
+Moreover, transient dynamics could prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
 
-The second model represents what we call a ``temporal approach.'' We correlate interannual variation in an ecological response with interannual variation in the environment, 
-typically at just one site. This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, 
-or primary production as a function of precipitation \citep{lauenroth_long-term_1992}. The advantages of this approach include is focus on temporal dynamics, meaning it 
-can predict rates of change, and validation of the resulting near-term predictions are straightforward. However, models built using this approach typically cover 
-a limited spatial extent (but see \citealt{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
-that could alter the underlying process at longer time scales and improve forecastability \citep{clark_ecological_2001}.
+The second model represents the time-series or ``temporal approach.'' We correlate interannual variation in an ecological response with interannual variation in the environment at just one site.
+This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, or primary production as a function of precipitation \citep{lauenroth_long-term_1992}.
+The advantages of this approach include its focus on temporal dynamics, meaning it can predict rates of change, and validation of the resulting near-term predictions are straightforward.
+However, models built using this approach typically cover a limited spatial extent (but see \citealt{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could alter the underlying process at longer time scales \citep{clark_ecological_2001}.
 
-We compare forecasts from both models to the simulated dynamics to determine how well the two approaches to forecasting perform at different time scales.
-We also assess the potential for combining the information available in both spatial and temporal patters by combining forecasts from the spatial and temporal approaches using a weighted average and assess how the optimal model weights change over time.
-Based on the hypothesis that spatial patterns capture longer-term processes we expect the temporal model to best predict short-term dynamics, and the spatial model to best predict long-term dynamics, and that a combination of the two approaches will provide the best forecasts at transitional intermediate time scales.
+We compare forecasts from both models to the simulated dynamics to determine how well the two approaches perform at different forecast horizons.
+We also assess the potential for combining the information available in both spatial and temporal patterns by using a weighted average of the forecasts from the spatial and temporal approaches optimized to best match the (simulated) observations.
+We then study how the optimal model weights change over time.
+We expect the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, and that a combination of the two approaches will provide the best forecasts at transitional intermediate time scales.
 
@@
 
-A common need for conservation biologists and natural resource managers is anticipating the impact of environmental change on the abundance of 
-endangered species, biological invaders, and harvested species. Although the managers may be primarily interested in just the focal species, 
-skillful prediction might require considering interactions with many other species, greatly complicating the problem. But at what forecast horizon do
-altered species interactions become impossible to ignore? We explored this question using a metacommunity model developed by \cite{alexander_lags_2018}. 
+Conservation biologists and natural resource managers often need to anticipate the impact of environmental change on the abundance of 
+endangered species, biological invaders, and harvested species.
+Although the managers may be primarily interested in just the focal species, skillful prediction might require considering interactions with many other species, greatly complicating the problem.
+But at what forecast horizon do altered species interactions become impossible to ignore?
+We explored this question using a metacommunity model developed by \cite{alexander_lags_2018}.
 The model describes Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
-Composition varies along the gradient because of a trade-off between growth rate and cold tolerance: 
-cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by 
-fast-growing species that are cold intolerant. Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics. 
+Composition varies along the gradient because of a trade-off between growth rate and cold tolerance: cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by  fast-growing species that are cold intolerant.
+Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics.
 Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
 
-We first simulated a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase, and then a 
-final period of stationary temperature. Interannual variation in temperature is the same at all sites, but mean temperature varies among sites. All sites experienced the 
-same absolute increase in mean temperature. We focused on the biomass dynamics of one focal species that dominated the baseline period.
+We first simulated a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase, and then a final period of stationary temperature.
+Interannual variation in temperature is the same at all sites, but mean temperature varies among sites.
+All sites experienced the same absolute increase in mean temperature.
+We focused on the biomass dynamics of one focal species that dominated the central site during the baseline period.
 
@@
 The predictions from the spatial and temporal models contrasted markedly, with the temporal 
-model predicting a large increase in biomass and the spatial model predicting a decrease.Initially, the simulated abundances followed the increase predicted by the 
+model predicting a large increase in biomass and the spatial model predicting a decrease. Initially, the simulated abundances followed the increase predicted by the 
 temporal model, but as faster-growing species colonized and increased in abundance at the focal site, the biomass of the focal species decreased, eventually falling below its baseline level (Fig.
@@
 This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}), but 
-it works so effectively because the weights were determined by fitting directly to those data. Unlike our spatial and temporal model forecasts, 
+the reason it works so well is that the weights were determined by fitting directly to the data. Unlike our spatial and temporal model forecasts, 
 we did not generate out-of-sample predictions from the combined model; it merely provides 
@@
 But at any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm.
-For example, at a cold site dominated by the cold-adapted homozygous genotype, a warmer than average year would cause a decrease in population size, even though the heat-adapted homozygote might perform optimally at that same temperature.
+For example, at a cold site dominated by the cold-adapted homozygous genotype, a warmer than average year would cause a decrease in population size, even though the heat-adapted homozygote might perform optimally at that temperature.
 However, if warmer than normal conditions persist for many years, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
@@
 We used the simulated data from this baseline period to fit an empirical model that assumes no knowledge of the underlying eco-evolutionary process.
-This empirical ``temporal model" (Appendix \ref{models}) predicts population growth rate as a function of annual temperature and population size (Fig.
+This empirical temporal model (Appendix \ref{models}) predicts population growth rate as a function of annual temperature and population size (Fig.
 \ref{fig:evo_evo_spatial_temporal_models}C).

### 2019-08-06 doc/manuscript.tex
working on discussion

@@
 
-Ecological forecasts are typically made using models fit to either time-series data or spatial data based on the concept of space-for-time substitutions.
-Our results demonstrate that forecasts based on these two approaches can lead to very different predictions for the future state of ecological systems.
-Which approach provides the most accurate forecasts depends on the time-horizon, or how far into the future we are attempting to forecast.
-For short-term forecasts, time-series approaches perform best, whereas longer-term forecasts are more accurately made using models from spatial data.
-This stems from differences between data types in the types and scales of processes they capture.
-Local time-series data capture demographic processes, autocorrelation (including lagged responses) and responses of current assemblages to small changes
-in environmental conditions. In addition, the state of the system a short period into the future depends heavily on the current state (i.e., the initial
-condition). Since short-term forecasts do not typically require extrapolating into novel conditions; the historical range of variation, incorporating lags
-and accurate initial conditions present in local time-series data is likely to be the most generally useful approach for forecasting.
-In contrast, data spanning spatial gradients captures the long-term outcome of slower processes including selection, dispersal, and responses to large
-changes in the environment. This makes models based on these data more likely to accurately predict long-term steady state outcomes of ecological change.
+Ecological forecasts are typically made using either a space-for-time substitution approach based on models fit to spatial data or using dynamic models fit to time-series data.
+Our results demonstrate that these two approaches can make very different predictions about the future state of ecological systems.
+Which approach provides the most accurate forecasts depends on the forecast-horizon.
+For short-term forecasts, time-series approaches perform best, whereas models based on spatial data made more accurate long-term forecasts.
+The change in model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
+Local time-series data capture demographic processes, autocorrelation (including lagged responses) and responses of current assemblages to small changes in environmental conditions.
+In addition, the state of the system in the near future depends heavily on the current state (i.e., the initial condition).
+Since short-term forecasts do not typically require extrapolating into novel conditions, a model based on the historical range of variation and which incorporates lags and accurate initial conditions is likely to be the most generally useful approach for forecasting.
+The hypothesis that time-series models will be most effective for near-term forecasts is supported backed by recent empirical analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
+In contrast, data sets spanning spatial gradients capture the long-term outcome of slower processes including ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
+Models based on data covering spatial gradients may be a better choice for predicting long-term steady state outcomes of ecological change.
 
-These results suggest that near-term forecasts will likely be most effective when made using time-series models, a result backed by recent empirical
-analyses of biodiversity forecasts at time scales from one to ten years (Harris et al. 2018). For long-term forecasts our results suggest the use
-of models built on spatial data. However, determining what is "near-term" and what is "long-term" in this context is an open question, and
-many important forecasts in ecology will be needed at intermediate time horizons where neither approach will be
+The real challenge is determining what is ``near-term'' and what is ``long-term.'' Furthermore,
+many of the forecasts we need in ecology may fall in the intermediate time horizons where neither the time-series
+START HERE
+approach will be
 fully sufficient. Our simulation results demonstrate that there are often extended transitional period where neither approach is effective on its own,

### 2019-08-07 doc/manuscript.tex
More edits on discussion

@@
 
-The real challenge is determining what is ``near-term'' and what is ``long-term.'' Furthermore,
-many of the forecasts we need in ecology may fall in the intermediate time horizons where neither the time-series
+The challenge is determining what is ``near-term'' and what is ``long-term,'' and how to handle the many forecasts we need in ecology which fall in  intermediate time horizons.
+Our simulations demonstrate extended transitional periods during which neither the time-series or the spatial approach is effective on its own, but a weighted combination of the two can produce highly accurate predictions.
+Unfortunately, we currently have no way to determine how to weight these two approaches for different time horizons.
+Our combined models used weights fitted directly to the simulated data, which is impossible to do for actual forecasting when the future is unknown.
+We need new theory or empirical case studies in order to set these weights \emph{a priori}.
+Theory could explore the influence of different parameters on on the rate at which slow processes (represented in spatial data) begin to influence dynamics.
+The effects of some parameters are intuitive: in the community turnover example, increasing the fraction of dispersing individuals causes a more rapid shift in species composition and in model weights (Fig.
+\ref{fig:dispersal_niche_width}A).
+Other factors to consider include organism lifespans, niche widths, and the magnitude of directional environmental change relative to historical interannual variation.
+Empirical research could address this challenge by accumulating enough case studies to infer patterns in the weight functions and guide applications in new systems.
+Developing  rules of thumb would require testing many forecasts from both time-series and spatial models across a range of time-horizons.
+It will require data sets that can support the evaluation of long-term forecasts, which could include paleoecological data (e.g.,), long-term historical data sets (e.g., ), and the use of model systems with short-generation times (e.g., Petchy),
+
 START HERE
-approach will be
-fully sufficient. Our simulation results demonstrate that there are often extended transitional period where neither approach is effective on its own,
-but where a weighted combination of time-series and spatial models can produce highly accurate predictions. Unfortunately there is no clear
-way to determine how to weight these two approaches for different time horizons. Our combined ``forecasts'' used the performance of
-the forecasts on simulated data to select the weights. This approach cannot be used for actual forecasting because it is effectively an empirical
-quantification of the observed performance of the forecasts at different time horizons. Determining how to set these
-weights a priori could be pursued using both theoretical and empirical approaches. On the theoretical side, it should be possible to explore the influence
-of different parameters on how quickly slow processes (represented in spatial data) begin to influence dynamics. The effects of some parameters are intuitive:
-in the community turnover example, increasing the fraction of dispersing individuals causes a more rapid shift in species composition and in model weights
-(Fig. \ref{fig:dispersal_niche_width}A). Other factors to consider include organism lifespans, niche widths, and the magnitude of directional environmental
-change relative to interannual variation. From an empirical perspective, this challenge could be addressed by using synthetic analyses to estimate the
-weighting function. By evaluating both time-series and spatial models for large numbers of forecasts across a range of time-horizons it should be possible
-to develop rules of thumb for ensembling models encompassing different scales of processes. This requires data sets that can support the evaluation of
-long-term forecasts, which could include paleoecological data (e.g.,), long-term historical data sets (e.g., ), and the use of model systems with
-short-generation times (e.g., Petchy), With enough case studies, we might be able to infer patterns that could guide applications in new systems.
 

### 2019-08-07 doc/manuscript.tex
PBA finished ms edits

@@
 \newcommand{\spe}{\textcolor{blue}}
-\newcommand{\comment}{\textcolor{black}}
+\newcommand{\comment}{\textcolor{red}}
 
@@
 The models used for ecological forecasting have traditionally been based on either time-series approaches or space-for-time substitutions.
-The time-series approach involves fitting models to long-term datasets to describe the temporal dynamics of a system.  We then use those
-dynamic models to make predictions about what will happen in the future. This is the common approach for population forecasts and includes 
-both process-based and data-driven models (e.g., \citep{ward_complexity_2014}). Whether process-based or
-data-driven, time-series models typically capture processes operating on interannual time-scales, such as birth, death, 
-small scale dispersal events, and short-term responses to environmental conditions.
+The time-series approach involves fitting models to long-term datasets to describe the temporal dynamics of a system.
+We then use those dynamic models to make predictions about what will happen in the future.
+This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, or primary production as a function of precipitation \citep{lauenroth_long-term_1992}.
+Whether process-based or data-driven (e.g., \citep{ward_complexity_2014}), time-series models typically capture processes operating on interannual time-scales, such as birth, death, small scale dispersal events, and short-term responses to environmental conditions.
+However, models built using this approach typically cover a limited spatial extent (but see \citealt{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could alter the underlying process at longer time scales \citep{clark_ecological_2001}.
 
-Space-for-time substitution approaches begin by describing how an ecological variable of interest, such as occupancy or productivity, 
-varies across sites experiencing different environmental conditions. These spatial relationships between environment and ecological 
-response are assumed to also hold for changes at a site through time. To make a forecast, we first predict the future
-environmental conditions and then determinie the associated ecological response, based on the observed spatial relationship. 
-Most forecasts for future species distributions and biodiversity are based on this approach. 
-Space-for-time models capture processes operating over long time periods, such as
-immigration, extinction, and responses to large or prolonged environmental changes. Although both of these approaches are
-widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or 
-advancing our understanding of ecological dynamics \citep{harris_forecasting_2018} (but see Renwick et al ??).
+Space-for-time substitution approaches begin by describing how an ecological variable of interest, such as occupancy or productivity,  varies across sites experiencing different environmental conditions.
+These spatial relationships between environment and ecological response are assumed to also hold for changes at a site through time.
+To make a forecast, we first predict the future environmental conditions and then determine the associated ecological response, based on the observed spatial relationship.
+This is the approach commonly used to predict population distribution or abundance as a function of climate \citep{elith_species_2009} or primary production as a function of mean precipitation \citep{Sala1988}.
+Space-for-time models capture processes operating over long time periods, such as immigration, extinction, and responses to large or prolonged environmental changes.
+however, these models provide no information about how quickly the system will move from the current state to the predicted, future state.
+Moreover, transient dynamics could prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
+Although both time-series and space-for-time approaches are widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics \citep{harris_forecasting_2018} (but see Renwick et al ??).
 
-Whether historical dynamics, contemporary spatial patterns, or spme combination of the two will serve as the best source of 
+Whether historical dynamics, contemporary spatial patterns, or some combination of the two will serve as the best source of 
 information for forecasting may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. 
@@
 and temporal resolution and extent from the global circulation models used to predict long-term changes
-in climate [IS THIS REALLY TRUE? REFERENCE?].
+in climate \comment{Great point, but this shows up in Conclusion too. I think it is better here? Also, do we need to fact check this?}.
 
@@
 
-In two case studies, we simulate the effects of an increase in temperature on simple systems with known dynamics.
+In two case studies, we simulated the effects of an increase in temperature on simple systems with known dynamics.
 The ``truth'' is represented by a model that is mechanistic for at least one important process, but we treat the model as unknown when
-analyzing the data and we assume that perfectly recovering this model would not be possible in practice. We begin each simulation
+analyzing the data and we assumed that perfectly recovering this model would not be possible in practice. We began each simulation
 under stationary temperature, allowing the system to equilibrate; we call this the baseline phase. 
-We then increase temperature progressively over a period of time, followed by a second period of stationary, now elevated, temperature. 
-The objective is to forecast the response of the system to the temperature increase based on data gathered during the baseline period.
+We then increased temperature progressively over a period of time, followed by a second period of stationary, now elevated, temperature. 
+The objective was to forecast the response of the system to the temperature increase based on data gathered during the baseline period.
 
-We make forecasts based on two phenomenological models, each representing processes operating at different time scales. 
+We made forecasts based on two phenomenological models, each representing processes operating at different time scales. 
 The first phenomenological model represents the space-for-time substiution approach, which we call the ``spatial approach'' for brevity. 
-We correlate the mean environment with the mean of an ecological state or rate across many sites. 
-This is the approach commonly used to predict population distribution or abundance as a function of climate \citep{elith_species_2009} 
-or primary production as a function of mean precipitation \citep{Sala1988}.
-The approach is attractive because it addresses broad spatial scales relevant to  management, and because the necessary data are widely available and the statistical analyses straightforward.
-The approach also has well-known weaknesses.
-Criticism often focuses on the assumption that current spatial patterns are in equilibrium with climate (refs?).
-Another issue is that these models are not dynamic;  they provide no information about how quickly the system will move from the current state to the predicted, future state.
-Because the patterns in the data have developed  over long time spans, they may not provide good information about dynamic changes on the order of a few years.
-Moreover, transient dynamics could prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
-
-The second model represents the time-series or ``temporal approach.'' We correlate interannual variation in an ecological response with interannual variation in the environment at just one site.
-This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, or primary production as a function of precipitation \citep{lauenroth_long-term_1992}.
-The advantages of this approach include its focus on temporal dynamics, meaning it can predict rates of change, and validation of the resulting near-term predictions are straightforward.
-However, models built using this approach typically cover a limited spatial extent (but see \citealt{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could alter the underlying process at longer time scales \citep{clark_ecological_2001}.
-
-We compare forecasts from both models to the simulated dynamics to determine how well the two approaches perform at different forecast horizons.
-We also assess the potential for combining the information available in both spatial and temporal patterns by using a weighted average of the forecasts from the spatial and temporal approaches optimized to best match the (simulated) observations.
-We then study how the optimal model weights change over time.
-We expect the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, and that a combination of the two approaches will provide the best forecasts at transitional intermediate time scales.
+We correlated the mean environment with the mean of an ecological state or rate across many sites.
+The second model represents the time-series or ``temporal approach.'' We correlated interannual variation in an ecological response with interannual variation in the environment at just one site.
+We compared forecasts from both models to the simulated dynamics to determine how well the two approaches perform at different forecast horizons.
+We also assessed the potential for combining the information available in both spatial and temporal patterns by using a weighted average of the forecasts from the spatial and temporal approaches optimized to best match the (simulated) observations.
+We then studied how the optimal model weights changed over time.
+We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, and that a combination of the two approaches would provide the best forecasts at transitional intermediate time scales.
 
@@
 We need new theory or empirical case studies in order to set these weights \emph{a priori}.
-Theory could explore the influence of different parameters on on the rate at which slow processes (represented in spatial data) begin to influence dynamics.
+Theory could explore the influence of different parameters on the rate at which slow processes (represented in spatial data) begin to influence dynamics.
 The effects of some parameters are intuitive: in the community turnover example, increasing the fraction of dispersing individuals causes a more rapid shift in species composition and in model weights (Fig.
@@
 Developing  rules of thumb would require testing many forecasts from both time-series and spatial models across a range of time-horizons.
-It will require data sets that can support the evaluation of long-term forecasts, which could include paleoecological data (e.g.,), long-term historical data sets (e.g., ), and the use of model systems with short-generation times (e.g., Petchy),
-
-START HERE
+This effort may require a novel integration of typically disparate approaches, such as analyses of paleoecological data (e.g.,), long-term historical data (e.g., ), and model systems with short-generation times (e.g., Petchy).
 
@@
 
-Instead of trying to weight distinct approaches an alternative approach is to build fully process based models of ecological systems  for forecasting.
-If it is possible to accurately characterize all of the processes governing a system then a model based on that understanding should make accurate
-predictions at all time-horizons. For example, if we had evaluated the exact process-based models that we used to generate the simulated time series,
-these models would have been able to accurately capture the full dynamics and forecasts for all time horizons. In concept, process based models should
-also be more robust for making predictions outside of historically observed conditions (or beyond the conditions observed across spatial gradients).
-This is particularly important for making predictions in a future that is predicted to be increasingly "non-analog", where combinations of environment
-and species interactions exist that have never occurred before (Williams et al.) Unfortunately, this approach is not currently feasible in most cases,
-because we lack a detailed knowledge of all of the complex and interacting processes influencing the dynamics of real ecological systems. Even if the
-general form of the models was understood estimating the high number of parameters and quantifying how they vary across
-ecosystems typically require more data than is currently available for even well studied systems. As a result, models used for ecological forecasting
-will include at least some phenomenological components. But that does not mean that these forecasts do not rely on process-based understanding in some way.
-The message from our simulations is that different processes should be considered for different forecast time-scales, and this can be done by fitting
-models to different kinds of data sets. Even when process-level understanding does not enable a fully mechanistic model, it can improve the specification
-of phenomenological models.
+An alternative to a weighted combination of predictions from time-series and space-for-time models is to rely on fully process-based models.
+If we could accurately characterize all of the processes governing a system, then a model based on that understanding should make accurate
+predictions at all time-horizons.
+For example, rather than fitting phenomenological models to our simulated time-series, we could have fit the exact process-based models that we used to generate those time-series.
+Assuming reasonable estimates of the parameter, those models would have accurately predicted the dynamics at all forecast horizons.
+Process based models should also be more robust for making predictions outside of historically observed conditions and even beyond the conditions observed across spatial gradients, which will be especially important for making predictions in a future with increasingly novel combinations of environment and species interactions (Williams et al.) Unfortunately, in most cases this approach is not currently feasible because we lack a detailed knowledge of all the complex and interacting processes influencing the dynamics of real ecological systems.
+Even if the general form of the models was understood, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well studied systems.
+As a result, models used for ecological forecasting will include at least some phenomenological components.
+But that does not mean that these forecasts do not rely on process-based understanding in some way.
+The message from our simulations is that different processes should be considered for different forecast time-scales, and this can be done by fitting models to different kinds of data sets.
+Even when process-level understanding does not enable a fully mechanistic model, it can improve the specification of phenomenological models.
 
-Another approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is joint spatiotemporal modeling...
+Another approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is joint spatiotemporal modeling...\comment{Ethan, what do you have in mind here?}
 
-These results have important implications for the emerging field of ecological forecasting. They suggest that evaluating both near-term and long-term
-forecasts is essential to fully understanding the best methods for forecasting and how well they perform. Our results further suggest that while
-single approaches may perform reasonably well at short and long time horizons, that intermediate time horizons may require a combination of information
-from spatial and temporal patterns. Related challenges with intermediate time horizons occur in other forecasting contexts. Weather forecasts based one
-type of models are very effective for forecasting a week to ten days in advance, but then become largely uninformative. Climate models provide accurate
-predictions for similar variables, but generally at decadal time-scales. Forecasting in between these two time scales has been challenging in meteorology
-and will likely be challenging in ecology as well. In combination that means that while the recent emphasis on near-term iterative forecasting (Dietze 
-et al. 2018), is important, we also need to build understanding and capacity for mkaing forecasts across the full range of scales of interest in
-ecology, from hours to decades.
+Our results have important implications for the emerging field of ecological forecasting. First, they suggest that evaluating both near-term and long-term
+forecasts will be essential as research on forecasting methods accelerates.
+Second, while single approaches may perform reasonably well at either short or long forecast horizons, skillful predictions at intermediate time horizons may require a combination of information from spatial and temporal patterns.
+Intermediate time horizons pose challenges in other forecasting contexts as well.
+Weather forecasts based on regional-scale meteorological models are very effective for forecasting a week to ten days in advance, but then become largely uninformative.
+Global circulation models provide predictions for similar variables but at decadal time-scales \comment{PBA: Is this right?
+Are these models really different, or do they just differ in scake and resolution?
+I love this point, but should we move it out of last paragraph?}.
+Forecasting in between these two time scales has been challenging in meteorology and will likely be challenging in ecology as well.
+While the recent emphasis on near-term iterative forecasting (Dietze et al.
+2018) is the logical and tractable starting point, we also need to build understanding and capacity for forecasting ecological dynamics across all temporal scales of interest.
 

### 2019-09-05 doc/manuscript.tex
Added concept figure to Intro

@@
 \author[2]{Ethan White}
-\author[1]{Michael Cortez?}
-\author[3]{Heather Lynch?}
+\author[3]{Michael Cortez}
 \affil[1]{Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah}
 \affil[2]{some shitty Florida joint}
+\affil[3]{some shitty Florida joint}
 
@@
 for near-term forecasts time-series approaches may capture the key dynamics and provide accurate predictions.
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.5 \textwidth] {fast-slow-figure.png}
+	\caption{Fast and slow processes operate at different time scales, and are reflected in different kinds of datasets.
+Fast processes, such as births, deaths, and individual growth, operate at all time scales, but are the exclusive drivers of the short-term dynamics captured in most time series datasets.
+Slower processes, such as evolutionary selection on genotype frequencies, ecological selection on species abundances, and colonization and extinction, interact with fast processes to drive dynamics over the long-term.
+The influence of these slow processes is seen in very long time series, or in spatial gradients.
+Spatiotemporal datasets with long enough records to capture changes in spatial patterns are rare, but would be extremely valuable for understanding dynamics at intermediate time scales.
+The time scales shown here were chosen with vascular plants in mind; the same concepts would apply for much shorter-lived organisms but at shorter time scales.
+}
+	\label{fig:fast-slow-concept}
+\end{figure}
 

### 2019-09-05 doc/manuscript.tex
Finished minor Cortez edits

@@
 \author[1]{Peter B. Adler}  %\thanks{Corresponding author. Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah Email: peter.adler@usu.edu}}
-\author[2]{Ethan White}
-\author[3]{Michael Cortez}
+\author[2]{Ethan P. White}
+\author[3]{Michael H. Cortez}
 \affil[1]{Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah}
-\affil[2]{some shitty Florida joint}
-\affil[3]{some shitty Florida joint}
+\affil[2]{Department of Wildlife Ecology and Conservation, University of Florida, Gainesville, Florida}
+\affil[3]{Department of Biological Science, Florida State University, Tallahasee, Florida}
 
@@
 This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, or primary production as a function of precipitation \citep{lauenroth_long-term_1992}.
-Whether process-based or data-driven (e.g., \citep{ward_complexity_2014}), time-series models typically capture processes operating on interannual time-scales, such as birth, death, small scale dispersal events, and short-term responses to environmental conditions.
-However, models built using this approach typically cover a limited spatial extent (but see \citealt{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could alter the underlying process at longer time scales \citep{clark_ecological_2001}.
+Whether process-based or data-driven (e.g., \citealt{ward_complexity_2014}), time-series models typically capture ``fast processes'' operating on interannual time-scales, such as birth, death, individual growth, small-scale dispersal events, and short-term responses to environmental conditions (Fig.
+\ref{fig:fast-slow-concept}).
+However, models built using this approach typically cover a limited spatial extent (but see \citealt{Hefley2017,kleinhesselink_response_2018}), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could influence dynamics at longer time scales \citep{clark_ecological_2001}.
 
@@
 This is the approach commonly used to predict population distribution or abundance as a function of climate \citep{elith_species_2009} or primary production as a function of mean precipitation \citep{Sala1988}.
-Space-for-time models capture processes operating over long time periods, such as immigration, extinction, and responses to large or prolonged environmental changes.
-however, these models provide no information about how quickly the system will move from the current state to the predicted, future state.
+Space-for-time models capture the outcome of interactions between fast processes and slower processes operating over long time periods, such as immigration, extinction, and responses to large or prolonged environmental changes (Fig.
+\ref{fig:fast-slow-concept}).
+However, space-for-time models provide no information about how quickly the system will move from the current state to the predicted, future state.
 Moreover, transient dynamics could prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
-Although both time-series and space-for-time approaches are widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics \citep{harris_forecasting_2018} (but see Renwick et al ??).
+Although both time-series and space-for-time approaches are widely used, there has Rebeen little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics (\citealt{harris_forecasting_2018} but see \citealt{Renwick2018}).
 
@@
 information for forecasting may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. 
-This potential dependency on the "forecast horizon"; ()\textit{sensu} citation)
+This potential dependency on the ``forecast horizon'' (\textit{sensu} \citealt{Hyndman2018})
 reflects lags in the response of ecological conditions to environmental change, shifts in the importance of ecological processes with
@@
 At short forecast horizons (days to years), dynamics will reflect the physiological and demographic responses of 
-the organisms present at a site more than than turnover of genotypes or species, 
+the organisms present at a site more than turnover of genotypes or species, 
 environmental conditions are likely to stay within the range of historical variation, 
@@
 and temporal resolution and extent from the global circulation models used to predict long-term changes
-in climate \comment{Great point, but this shows up in Conclusion too. I think it is better here? Also, do we need to fact check this?}.
+in climate.
 
@@
 \begin{enumerate}
-	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow processes such as evolutionary and ecological selection as well as dispersal.
-	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while data spanning spatial gradients capture the long-term outcome of selection and dispersal.
+	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require accounting for the influence of slow processes such as evolutionary and ecological selection as well as dispersal.
+	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and fast responses of current assemblages to interannual environmental variation, while data spanning spatial gradients capture the long-term outcome of interactions between fast and slow processes.
 Whether predictive models should be trained using longitudinal or spatial data sets, or both, depends on the time-scale of the desired forecast.
-	\item A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.
+	\item A key challenge for future research is determining the rate at which slow processes begin to influence dynamics.
 \end{enumerate}
@@
 To combine the spatial and temporal model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, the combined population forecast is $\omega * T(N_{t-1},K_t) + (1-\omega) * S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$ and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{models} for full description of the approach).
+At any time point, $t$, the combined population forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{models} for full description of the approach).
 This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}), but 
@@
 a convenient way to quantify how rapidly dynamics shift from being dominated by interannual 
-variation captured in the temporal model to being dominated by the steady-state equilibrium captured by the spatial model.
+variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
+\ref{fig:community-forecast-species}) to being dominated by the steady-state equilibrium captured by the spatial model (time $t \geq 2500$).
 A true, combined forecast would require a method to determine the model weights \emph{a priori}.
@@
 for total community biomass, and the results were similar: the temporal model initially made the best forecasts immediately following the onset of the temperature
-increase, but as the identify and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (Figs.
+increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (Figs.
 \ref{fig:community-models-total} to \ref{fig:community-weights-total}).
@@
 \ref{fig:evo_evo_spatial_temporal_models}A).
+Our model describes how the local density of each genotype changes between years,  which depends on temperature and genotype densities in the previous year.
+Transient temporal dynamics are computed directly from the model; these dynamics define our temporal forecast.
+To create ``spatial data'', we simulated the equilibrium density of each genotype under different mean temperatures.
+The pattern of equilibrium densities across a gradient in mean annual temperature defines our spatial forecast: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
+\ref{fig:evo_evo_spatial_temporal_models}B).
 The full model description is provided in Appendix \ref{eco-evo}.
-Genotype frequencies will shift across a gradient of mean annual temperature: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
-\ref{fig:evo_evo_spatial_temporal_models}B).
 
@@
 \ref{fig:evo_evo_spatial_temporal_models}B is the outcome of steady-state conditions.
-But at any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm.
-For example, at a cold site dominated by the cold-adapted homozygous genotype, a warmer than average year would cause a decrease in population size, even though the heat-adapted homozygote might perform optimally at that temperature.
+But at any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm (Fig.
+\ref{fig:evo_evo_spatial_temporal_models}A).
+For example, at a cold site dominated by the cold-adapted homozygous genotype, a warmer than average year would cause a decrease in population size due to decreases in fecundity (blue line in Fig.
+\ref{fig:evo_evo_spatial_temporal_models}A), even though the heat-adapted homozygote might perform optimally at that temperature.
 However, if warmer than normal conditions persist for many years, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
@@
 \ref{fig:forecast} bottom).
-The temporal model accurately predicted the impact of the initial warming trend, but eventually became too pessimistic, while the spatial model did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
+The temporal model (solid blue line in Fig.
+\ref{fig:forecast}) accurately predicted the impact of the initial warming trend, but eventually became too pessimistic, while the spatial model (solid red line in Fig.
+\ref{fig:forecast}) did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
 \ref{fig:forecast} bottom).
@@
 As in the community turnover example, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time.
-The combined model initially reflected the temporal model, but then rapidly transitioned to reflect the spatial model.
+The combined model initially reflected the temporal model (decrease from $t=500$ to $t=600$), but then rapidly transitioned to reflect the spatial model ($t \geq 700$).
 The rapid transition in the weighting term, $\omega$, occurred during the period of most rapid change in genotype frequencies (Fig.

### 2019-09-05 community-turnover/comm_turn_figures.R
combined forecsast and weights figures

@@ -70,25 +70,43 @@ my_obs_temp <- temperature[(burnin_yrs+1):sim_yrs,site]
 my_biomass <- rowSums(spxp[(burnin_yrs+1):sim_yrs,site,])
 my_biomass_spp <- spxp[(burnin_yrs+1):sim_yrs,site,my_species]
 
-png("figures/community_forecast_species.png",height=4.5,width=4,res=400,units="in")
+png("figures/community_forecast_species.png",height=7.5,width=4.5,res=400,units="in")
   
-  layout(matrix(c(1,2),2,1),heights=c(0.1,0.4),widths=1)
+  layout(matrix(c(1,2,3),3,1),heights=c(0.1,0.4,0.4),widths=1)
   
-  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(1,4,1,1),oma=c(2,0,0,0),cex=0.8)
+  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(1,4,1,4),oma=c(2,0,0,0),cex=0.8)
   
+  # top panel: temperature 
   plot(my_obs_temp,xlab="",ylab="Temperature",type="l",col="darkgrey")
   abline(v=baseline_yrs,col="black",lty=3)
   lines(my_expected_temp,lwd=2,col="black")
-
-  # focal species
+  mtext(side=3," A",line=-1.1,adj=0,cex=0.8)
+  
+  # middle panel: focal species and forecasts
   plot(my_biomass_spp,xlab="",ylab="Focal species biomass",type="l",col="darkgrey")
   abline(v=baseline_yrs,col="black",lty=3)
   lines(1:length(my_biomass_spp),spatial_forecast_spp,col="red",lwd=2)
   lines(1:length(my_biomass_spp),temporal_forecast_spp,col="blue",lwd=2)
   lines(1:length(my_biomass_spp),combined_forecast_spp,col="purple",lwd=2,lty=2)
+  mtext(side=3," B",line=-1.1,adj=0,cex=0.8)
   
   legend("right",c("Observed","Spatial forecast","Temporal forecast","Combined forecast"),
          col=c("darkgrey","red","blue","purple"),lwd=c(1,2,2,2),lty=c(1,1,1,2),bty="n",cex=0.8)
+  
+  # bottom panel: weights and species
+  spp_colors <- rep("darkgrey",N)
[…]

### 2019-09-05 genetic-diversity/genetic_diversity_figures.R

@@ -12,9 +12,9 @@ png("figures/spatial&temporal_model.png",height=10,width=4,res=400,units="in")
   temp=seq(-7,7,0.01)
   react_norms=getLambdas(temp,fec_Tmu,fec_Tsigma,fec_max)
   matplot(temp,react_norms,type="l",xlab="Temperature",ylab="Germination rate",col=myCols,lty="solid")
-  text(fec_Tmu[1],0.9*fec_max[1],"AA",col=myCols[1])
-  text(fec_Tmu[2],0.9*fec_max[2],"Aa",col=myCols[2])
-  text(fec_Tmu[3],0.9*fec_max[3],"aa",col=myCols[3])
+  text(fec_Tmu[1]-0.4,0.95*fec_max[1],"AA",col=myCols[1])
+  text(fec_Tmu[2],0.95*fec_max[2],"Aa",col=myCols[2])
+  text(fec_Tmu[3]+0.4,0.95*fec_max[3],"aa",col=myCols[3])
   mtext("(A)",side=3, line=0.5, adj=0)
   
   # plot mean abundances vs mean temperature, along with spatial model "prediction"
@@ -32,7 +32,7 @@ png("figures/spatial&temporal_model.png",height=10,width=4,res=400,units="in")
   rr <- range(N_t0_baseline)
   svals <- (N_t0_baseline-rr[1])/diff(rr)
   f <- colorRamp(c("red", "blue"), 0.4)
-  density_colors <- alpha(rgb(f(svals)/255),0.6)
+  density_colors <- rgb(f(svals)/255, alpha=0.6)
   # now plot
   plot(temperature_t1_baseline,r_baseline,xlab="Annual temperature",ylab="log Population growth rate",
        col=density_colors,pch=16)

### 2019-09-05 doc/manuscript.tex
combined forecsast and weights figures
New Fig 3 incorporated in text

@@
 
-We forecast the effect of a temperature increase on the focal species' biomass at one location in the center of the temperature gradient 
+We forecast the effect of a temperature increase (Fig. \ref{fig:community-forecast-species}A) on the focal species' biomass at one location in the center of the temperature gradient 
 using both spatial models and the temporal models. 
@@
 temporal model, but as faster-growing species colonized and increased in abundance at the focal site, the biomass of the focal species decreased, eventually falling below its baseline level (Fig.
-\ref{fig:community-forecast-species}).
+\ref{fig:community-forecast-species}B).
 
@@
 At any time point, $t$, the combined population forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{models} for full description of the approach).
-This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}), but 
+This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}B), but 
 the reason it works so well is that the weights were determined by fitting directly to the data. Unlike our spatial and temporal model forecasts, 
@@
 variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
-\ref{fig:community-forecast-species}) to being dominated by the steady-state equilibrium captured by the spatial model (time $t \geq 2500$).
+\ref{fig:community-forecast-species}B) to being dominated by the steady-state equilibrium captured by the spatial model (time $t \geq 2500$).
 A true, combined forecast would require a method to determine the model weights \emph{a priori}.
 The most rapid shifts in the model weights occurred during the period when warm-adapted, faster growing species were increasing most rapidly
-in abundance (Fig. \ref{fig:community-weights-spp}).
+in abundance (Fig. \ref{fig:community-forecast-species}C).
 
@@
 \includegraphics[width=0.7 \textwidth] {community_forecast_species.png}
-\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(Bottom) Simulated focal species biomass and forecasts from the spatial, temporal and combined models.
+\caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
+(B) Simulated focal species biomass and forecasts from the spatial, temporal and combined models at the focal site in the metacommunity model.
+(C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal model for focal species biomass (blue).
+Year 1000 in each panel corresponds to the start of the temperature increase.
 }
 \label{fig:community-forecast-species}
-\end{figure}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_change_plus_weights_spp.png}
-\caption{Simulated changes in biomass of the focal species (black) and all other species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for focal species biomass (blue).
-Year 0 in this figure corresponds to the start of the temperature increase.
-}
-\label{fig:community-weights-spp}
 \end{figure}
@@
 (C) The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
-Colors show population size (the blue to red gradient depicting low to high population density), which influences the population growth rate through density dependence.
+Colors show population size (the blue to red gradient depicting high to low population density), which influences the population growth rate through density dependence.
 }
@@
 	\includegraphics[width=0.7 \textwidth] {dispersal_niche_width.png}
-	\caption{(The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
+	\caption{The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
 Year 0 in these figures corresponds to the start of the temperature increase.

### 2019-09-06 doc/manuscript.tex
Added refs and fixed notation

@@
 Which approach provides the most accurate forecasts depends on the forecast-horizon.
-For short-term forecasts, time-series approaches perform best, whereas models based on spatial data made more accurate long-term forecasts.
+In our simulations, time-series approaches performed best for short-term forecasts, whereas models based on spatial data made more accurate long-term forecasts.
 The change in model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
-Local time-series data capture demographic processes, autocorrelation (including lagged responses) and responses of current assemblages to small changes in environmental conditions.
-In addition, the state of the system in the near future depends heavily on the current state (i.e., the initial condition).
-Since short-term forecasts do not typically require extrapolating into novel conditions, a model based on the historical range of variation and which incorporates lags and accurate initial conditions is likely to be the most generally useful approach for forecasting.
-The hypothesis that time-series models will be most effective for near-term forecasts is supported backed by recent empirical analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
-In contrast, data sets spanning spatial gradients capture the long-term outcome of slower processes including ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
-Models based on data covering spatial gradients may be a better choice for predicting long-term steady state outcomes of ecological change.
+Local time-series data capture demographic processes, lagged effects, and responses of current assemblages to small changes in environmental conditions.
+In addition, the state of the system in the near future depends heavily on the current state.
+Since short-term forecasts do not typically require extrapolating into novel conditions, a model based on the historical range of variation which incorporates lags and accurate initial conditions is likely to be the most generally useful approach for forecasting.
+The hypothesis that time-series models will be most effective for near-term forecasts is supported by recent empirical analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
+In contrast, datasets spanning spatial gradients capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
+Models based on data from spatial gradients may be a better choice for predicting long-term, steady-state outcomes of ecological change.
 
-The challenge is determining what is ``near-term'' and what is ``long-term,'' and how to handle the many forecasts we need in ecology which fall in  intermediate time horizons.
+The challenge is determining what is ``near-term'' and what is ``long-term,'' and how to handle the many forecasts we need in ecology which fall in between.
 Our simulations demonstrate extended transitional periods during which neither the time-series or the spatial approach is effective on its own, but a weighted combination of the two can produce highly accurate predictions.
@@
 Our combined models used weights fitted directly to the simulated data, which is impossible to do for actual forecasting when the future is unknown.
-We need new theory or empirical case studies in order to set these weights \emph{a priori}.
-Theory could explore the influence of different parameters on the rate at which slow processes (represented in spatial data) begin to influence dynamics.
-The effects of some parameters are intuitive: in the community turnover example, increasing the fraction of dispersing individuals causes a more rapid shift in species composition and in model weights (Fig.
+We need new theory or empirical case studies in order to assign these weights \emph{a priori}.
+Theory could explore the influence of different parameters on the rate at which slow processes begin to influence dynamics.
+The effects of some parameters are intuitive: in the community turnover example, increasing the fraction of dispersing individuals caused a more rapid shift in species composition and in model weights (Fig.
 \ref{fig:dispersal_niche_width}A).
 Other factors to consider include organism lifespans, niche widths, and the magnitude of directional environmental change relative to historical interannual variation.
-Empirical research could address this challenge by accumulating enough case studies to infer patterns in the weight functions and guide applications in new systems.
+Empirical research could address this challenge by accumulating enough case studies to infer patterns in the weighting functions and guide applications in new systems.
 Developing  rules of thumb would require testing many forecasts from both time-series and spatial models across a range of time-horizons.
-This effort may require a novel integration of typically disparate approaches, such as analyses of paleoecological data (e.g.,), long-term historical data (e.g., ), and model systems with short-generation times (e.g., Petchy).
+This effort may require a novel integration of typically disparate approaches, such as analyses of paleoecological data (e.g., \citealt{Worth2014}), long-term observational (e.g., \citealt{Nice2019}) or experimental data (e.g., \citealt{Silvertown2006}), and model systems with short-generation times (e.g., \citealt{Good2017}).
 
@@
 An alternative to a weighted combination of predictions from time-series and space-for-time models is to rely on fully process-based models.
-If we could accurately characterize all of the processes governing a system, then a model based on that understanding should make accurate
-predictions at all time-horizons.
+If we could accurately characterize all of the processes governing a system, then a model based on that understanding should make accurate predictions at all time-horizons.
 For example, rather than fitting phenomenological models to our simulated time-series, we could have fit the exact process-based models that we used to generate those time-series.
 Assuming reasonable estimates of the parameter, those models would have accurately predicted the dynamics at all forecast horizons.
-Process based models should also be more robust for making predictions outside of historically observed conditions and even beyond the conditions observed across spatial gradients, which will be especially important for making predictions in a future with increasingly novel combinations of environment and species interactions (Williams et al.) Unfortunately, in most cases this approach is not currently feasible because we lack a detailed knowledge of all the complex and interacting processes influencing the dynamics of real ecological systems.
-Even if the general form of the models was understood, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well studied systems.
+Process-based models should also be more robust for making predictions outside of historically observed conditions and even beyond the conditions observed across spatial gradients, which will be especially important for making predictions in a future with increasingly novel combinations of environment and species interactions \citep{Williams2007}.
+Unfortunately, in most cases this approach is not currently feasible because we lack a detailed knowledge of all the complex and interacting processes influencing the dynamics of real ecological systems.
+Even if the general form of the models was known, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well studied systems.
 As a result, models used for ecological forecasting will include at least some phenomenological components.
-But that does not mean that these forecasts do not rely on process-based understanding in some way.
-The message from our simulations is that different processes should be considered for different forecast time-scales, and this can be done by fitting models to different kinds of data sets.
+But that does not mean that phenomenological forecast models do not benefit from process-based understanding.
+The message from our simulations is that different processes should be considered for different forecast time-scales, and this can be done by fitting models to different kinds of datasets.
 Even when process-level understanding does not enable a fully mechanistic model, it can improve the specification of phenomenological models.
 
-Another approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is joint spatiotemporal modeling...\comment{Ethan, what do you have in mind here?}
+Another approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is joint spatiotemporal modeling...\comment{Ethan, what do you have in mind here?
+Probably something more than just working with spatiotemporal datasets and accounting for autocorrelation.
+Mevin has lots of papers with some sort of mechanistic modeling of spatiotemoral processes like dispersal and colonization.
+Should we say that the datasets needed to train this kind of model are rare, but that might be changing with more and more remote sensing data that is extensive in space and time?
+}
 
@@
 Weather forecasts based on regional-scale meteorological models are very effective for forecasting a week to ten days in advance, but then become largely uninformative.
-Global circulation models provide predictions for similar variables but at decadal time-scales \comment{PBA: Is this right?
-Are these models really different, or do they just differ in scake and resolution?
-I love this point, but should we move it out of last paragraph?}.
-Forecasting in between these two time scales has been challenging in meteorology and will likely be challenging in ecology as well.
-While the recent emphasis on near-term iterative forecasting (Dietze et al.
-2018) is the logical and tractable starting point, we also need to build understanding and capacity for forecasting ecological dynamics across all temporal scales of interest.
+Forecasting these intermediate scales has been challenging in meteorology and will likely be challenging in ecology as well.
+While the recent emphasis on near-term iterative forecasting \citep{dietze_iterative_2018} is the logical and tractable starting point, we also need to build understanding and capacity for forecasting ecological dynamics across all temporal scales of interest.
 
@@
 
-The spatial model, which we refer to as $S$, is a linear regression fit to ``observed" data on mean population size, $\bar{N}$ at a series of locations, $x$, which vary in mean temperature, $K$.
-In our case, the observed data are simulated by the model described in the previous section.
-For both examples in the main text, we include a quadratic term to capture the unimodal relationship between  $\bar{N}$ and $K$:
+The two simulation models in the main text describe how population size, $N(x,t)$, at location $x$ changes over time ($t$).
+We assume that the temperature, $K(x,t)$, at each location can vary in time and space.
+
+The spatial model, which we refer to as $S$, is a quadratic regression of the mean long-term population density at a location ($\bar{N}(x)$) against the mean temperature at that location ($\bar{K}(x)$).
+The quadratic term describes the unimodal relationship between $\bar{N}$ and $\bar{K}$.
+The spatial statistical model is
 \begin{equation}
- \bar{N}_x = S(K_x) = \beta^S_0 +  \beta^S_1 K_x +\beta^S_2 {K_x}^2 + \varepsilon
+	\bar{N}(x) = S(\bar{K}(x)) = \beta^S_0 +  \beta^S_1 \bar{K}(x) +\beta^S_2 {\bar{K}(x)}^2 + \varepsilon
 	\label{eqn:spatial_regression}
 \end{equation}
-For the total biomass example from the community turnover model (results in Appendix \ref{biomass}), we dropped the quadratic term.
 
-The temporal model, which we call $T$, starts with a time series of ``observed" population sizes, or total biomasses, $\vec{N}$ at one location.
-In the community turnover example, we use data from a baseline period with a stationary distribution of annual temperature at the focal location, we fit the following regression, which predicts biomass at time $t+1$ as a function of biomass and annual temperature at time $t$:
+The temporal model, which we call $T$, starts with a time-series of ``observed" population sizes, or total biomasses, at one location, $N(t)$, for $t=1...n$ (the spatial index is suppressed because we only focus on one location at a time).
+In the community turnover example, we fit the following regression, which predicts biomass at time $t+1$ as a function of biomass ($N(t)$) and annual temperature ($K(t)$) at time $t$,
 \begin{equation}
- \ln(N_{t+1}) = T(N_t,K_t) = \beta^T_0 +  \beta^T_1 ln(N_t) +\beta^T_2 K_t  +  \varepsilon
+	\ln(N(t+1)) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 ln(N(t)) +\beta^T_2 K(t)  +  \varepsilon
 	\label{eqn:temporal_regression_community}
 \end{equation}
- For the eco-evolutionary example, the response variable is the log of the population growth rate, and we include a quadratic effect of annual temperature:
+In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression is 
 \begin{equation}
-  \ln(\frac{N_{t+1}}{N_t}) = T(N_t,K_t) = \beta^T_0 +  \beta^T_1 ln(N_t) +\beta^T_2 K_t  +\beta^T_3 K_t^2 +  \varepsilon
+	\ln(\frac{N(t+1)}{N(t)}) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 ln(N(t)) +\beta^T_2 K(t)  +\beta^T_3 K(t)^2 +  \varepsilon
 	\label{eqn:temporal_regression_ecoevo}
@@
 This version of the temporal model returns a per capita growth rate on the log scale.
-To predict population size at the next time step, we simply exponentiate the growth rate and multiply it by the current population size: $exp(T(N_t,K_t)) N_t$.
+To predict population size at the next time step, we exponentiate the growth rate and multiply it by the current population size: $\exp(T(N(t),K(t))) N(t)$.
 
@@
 \begin{equation}
-\hat{N}_{t+1} = \omega *T(N_{t},K_t) + (1-\omega) * S(K_t) 
+\hat{N}(t+1)= \omega \cdot T(N(t),K(t)) + (1-\omega) \cdot S(K(t)) 
 \label{eqn:combined_model}
 \end{equation}
+Again, we suppress the spatial subscript ($x$) here because we are focused on densities at just one location.
 For the eco-evolutionary example, the predicted population size from the combined model is:
 \begin{equation}
-\hat{N}_{t+1} = \omega * exp(T(N_{t},K_t)) N_t + (1-\omega) * S(K_t) 
+	\hat{N}(t+1) = \omega \cdot exp(T(N(t),K(t))) N(t) + (1-\omega) \cdot S(K(t)) 
 	\label{eqn:combined_model}
@@
 
-We used the \texttt{optim} function to estimate the $\beta^C$s that minimize the sum of squared errors, $(\hat{N}_{t+1} - N_{t+1})^2$
+We used the \texttt{optim} function to estimate the $\beta^C$s that minimize the sum of squared errors, $(\hat{N}(t+1) - N(t+1))^2$.
+
+\section{Description of the metacommunity model}\label{metacomm}
 
@@
 \begin{align}\begin{split}
-N_{1,t+1} &= s_1 [1-g_1(E_t)]N_{1,t}+\frac{\lambda_1g_1(E_t)N_{1,t}}{1+ \alpha_{11}g_1(E_t)N_{1,t} + \alpha_{12}g_2(E_t)N_{2,t}}\\
-N_{2,t+1} &= s_2 [1-g_2(E_t)]N_{2,t}+\frac{\lambda_2g_2(E_t)N_{2,t}}{1+ \alpha_{21}g_1(E_t)N_{1,t} + \alpha_{22}g_2(E_t)N_{2,t}}
+N_{1,t+1} &= s_1 [1-g_1(K(t))]N_{1,t}+\frac{\lambda_1g_1(K(t))N_{1,t}}{1+ \alpha_{11}g_1(K(t))N_{1,t} + \alpha_{12}g_2(K(t))N_{2,t}}\\
+N_{2,t+1} &= s_2 [1-g_2(K(t))]N_{2,t}+\frac{\lambda_2g_2(K(t))N_{2,t}}{1+ \alpha_{21}g_1(K(t))N_{1,t} + \alpha_{22}g_2(K(t))N_{2,t}}
 \end{split}\end{align}
-where $g_{i}(E_t)$ is the probability of germination, $E_t$ is the environmental quality at time $t$, $s_i$ is the seed survival probability for species $i$, and $\lambda_i$ is the seed production rate per plant.
-Below I refer to the $\alpha_{ij}$ as intra- and inter-specific competition coefficients.
-\\
+where $g_{i}(K(t))$ is the probability of germination, $K(t)$ is the temperature at time $t$, $s_i$ is the seed survival probability for species $i$, and $\lambda_i$ is the seed production rate per plant.
+Below we refer to the $\alpha_{ij}$ as intra- and inter-specific competition coefficients.
 
@@
 The genotypes are denoted by $AA$, $Aa$, and $aa$.
-The number of each genotypes at time $t$ is $N_{AA,t}$, $N_{Aa,t}$, and $N_{aa,t}$.
-The germination rates for each genotype are $g_{AA}(E_t)$, $g_{Aa}(E_t)$, and $g_{aa}(E_t)$.
+The number of each genotypes at time $t$ is $N_{AA}(t)$, $N_{Aa}(t)$, and $N_{aa}(t)$.
+The germination rates for each genotype are $g_{AA}(K(t))$, $g_{Aa}(K(t))$, and $g_{aa}(K(t))$.
 The seed survival probability and seed production rate for genotype $AA$ are $s_{AA}$ and $\lambda_{AA}$, respectively.
@@
 \begin{equation}
-T = \lambda_{AA}N_{AA,t}g_{AA}(E_t)+ \lambda_{Aa}N_{Aa,t}g_{Aa}(E_t)+\lambda_{aa}N_{aa,t}g_{aa}(E_t).
+T = \lambda_{AA}N_{AA}(t)g_{AA}(K(t))+ \lambda_{Aa}N_{Aa}(t)g_{Aa}(K(t))+\lambda_{aa}N_{aa}(t)g_{aa}(K(t)).
 \end{equation}
@@
 \begin{align}\begin{split}
-\phi_{A} &= \frac{\lambda_{AA}N_{AA,t}g_{AA}(E_t)+ \frac{1}{2}\lambda_{Aa}N_{Aa,t}g_{Aa}(E_t)}{T} \hspace{10pt} \text{and} \hspace{10pt} \phi_a = 1-\phi_{A}.
+\phi_{A} &= \frac{\lambda_{AA}N_{AA}(t)g_{AA}(K(t))+ \frac{1}{2}\lambda_{Aa}N_{Aa}(t)g_{Aa}(K(t))}{T} \hspace{10pt} \text{and} \hspace{10pt} \phi_a = 1-\phi_{A}.
 \end{split}\end{align}
@@
 \begin{align}\begin{split}
-N_{AA,t+1} &= s_{AA}[1-g_{AA}(E_t)]N_{AA,t} + \phi_A^2T\\
-N_{Aa,t+1} &= s_{Aa}[1-g_{Aa}(E_t)]N_{Aa,t} + \phi_A\phi_aT\\
-N_{aa,t+1} &= s_{aa}[1-g_{aa}(E_t)]N_{aa,t} + \phi_a^2T
+N_{AA}(t+1) &= s_{AA}[1-g_{AA}(K(t))]N_{AA}(t) + \phi_A^2T\\
+N_{Aa}(t+1) &= s_{Aa}[1-g_{Aa}(K(t))]N_{Aa}(t) + \phi_A\phi_aT\\
+N_{aa}(t+1) &= s_{aa}[1-g_{aa}(K(t))]N_{aa}(t) + \phi_a^2T
 \end{split}\end{align}
@@
 \begin{align}\begin{split}
-T &=  \frac{\lambda_{AA}N_{AA,t}g_{AA}(E_t)}{1+ \alpha_{AA,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{AA,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{AA,aa}g_{aa}(E_t)N_{aa,t}} \\ 
-&+ \frac{\lambda_{Aa}N_{Aa,t}g_{Aa}(E_t)}{1+ \alpha_{Aa,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{Aa,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{Aa,aa}g_{aa}(E_t)N_{aa,t}}\\
-&+\frac{\lambda_{aa}N_{aa,t}g_{aa}(E_t)}{1+ \alpha_{aa,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{aa,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{aa,aa}g_{aa}(E_t)N_{aa,t}}.
+T &=  \frac{\lambda_{AA}N_{AA}(t)g_{AA}(K(t))}{1+ \alpha_{AA,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{AA,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{AA,aa}g_{aa}(K(t))N_{aa}(t)} \\ 
+&+ \frac{\lambda_{Aa}N_{Aa}(t)g_{Aa}(K(t))}{1+ \alpha_{Aa,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{Aa,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{Aa,aa}g_{aa}(K(t))N_{aa}(t)}\\
+&+\frac{\lambda_{aa}N_{aa}(t)g_{aa}(K(t))}{1+ \alpha_{aa,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{aa,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{aa,aa}g_{aa}(K(t))N_{aa}(t)}.
 \label{eqn:defineT}
@@
 \begin{align}\begin{split}
-\phi_A &= \frac{1}{T}\frac{\lambda_{AA}N_{AA,t}g_{AA}(E_t)}{1+ \alpha_{AA,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{AA,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{AA,aa}g_{aa}(E_t)N_{aa,t}} \\
-&+ \frac{1}{2T}\frac{\lambda_{Aa}N_{Aa,t}g_{Aa}(E_t)}{1+ \alpha_{Aa,AA}g_{AA}(E_t)N_{AA,t} + \alpha_{Aa,Aa}g_{Aa}(E_t)N_{Aa,t}+ \alpha_{Aa,aa}g_{aa}(E_t)N_{aa,t}}\\
+\phi_A &= \frac{1}{T}\frac{\lambda_{AA}N_{AA}(t)g_{AA}(K(t))}{1+ \alpha_{AA,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{AA,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{AA,aa}g_{aa}(K(t))N_{aa}(t)} \\
+&+ \frac{1}{2T}\frac{\lambda_{Aa}N_{Aa}(t)g_{Aa}(K(t))}{1+ \alpha_{Aa,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{Aa,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{Aa,aa}g_{aa}(K(t))N_{aa}(t)}\\
 \phi_a &= 1- \phi_A
@@
 \begin{align}\begin{split}
-N_{AA,t+1} &= s_{AA}[1-g_{AA}(E_t)]N_{AA,t} + \phi_A^2T\\
-N_{Aa,t+1} &= s_{Aa}[1-g_{Aa}(E_t)]N_{Aa,t} + 2 \phi_A\phi_aT\\
-N_{aa,t+1} &= s_{aa}[1-g_{aa}(E_t)]N_{aa,t} + \phi_a^2T,
+N_{AA}(t+1) &= s_{AA}[1-g_{AA}(K(t))]N_{AA}(t) + \phi_A^2T\\
+N_{Aa}(t+1) &= s_{Aa}[1-g_{Aa}(K(t))]N_{Aa}(t) + 2 \phi_A\phi_aT\\
+N_{aa}(t+1) &= s_{aa}[1-g_{aa}(K(t))]N_{aa}(t) + \phi_a^2T,
 \end{split}\end{align}

### 2019-09-06 doc/manuscript.tex
Finished major round of edits

@@
 
-stuff
+Most models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
+Here we consider how the forecast horizon determines whether the most accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
+We use two simulation case studies to show that forecasts for short and long-time scales need to focus on different ecological processes, which are reflected in different kinds of data.
+In the short-term, dynamics reflect initial conditions and fast processes such as birth and death, and the phenomenological time-series approach makes the best predictions.
+In the long-term, dynamics reflect the additional influence of slower processes such as evolutionary and ecological selection, colonization and extinction, which the space-for-time approach can effectively capture.
+At intermediate time-scales, a weighted average of the two approaches shows promise.
+However, making this combined approach operational will require new research to predict the rate at which slow processes begin to influence dynamics.
+
+\vspace{0.2in}
 
@@
 
-The models used for ecological forecasting have traditionally been based on either time-series approaches or space-for-time substitutions.
+Most models used for ecological forecasting have traditionally been based on either time-series approaches or space-for-time substitutions.
 The time-series approach involves fitting models to long-term datasets to describe the temporal dynamics of a system.
@@
 However, space-for-time models provide no information about how quickly the system will move from the current state to the predicted, future state.
-Moreover, transient dynamics could prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
-Although both time-series and space-for-time approaches are widely used, there has Rebeen little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics (\citealt{harris_forecasting_2018} but see \citealt{Renwick2018}).
+In fact, transient dynamics could prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
+Although both time-series and space-for-time approaches are widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics (\citealt{harris_forecasting_2018}, but see \citealt{Renwick2018}).
 
@@
 We then studied how the optimal model weights changed over time.
-We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, and that a combination of the two approaches would provide the best forecasts at transitional intermediate time scales.
+These spatial, temporal and combined statistical
+models are described in Appendix \ref{models}.
+We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, and that a combination of the two approaches would provide the best forecasts at transitional, intermediate time scales.
 
@@
 endangered species, biological invaders, and harvested species.
-Although the managers may be primarily interested in just the focal species, skillful prediction might require considering interactions with many other species, greatly complicating the problem.
+Although the managers may be primarily interested in just one focal species, skillful prediction might require considering interactions with many other species, greatly complicating the problem.
 But at what forecast horizon do altered species interactions become impossible to ignore?
@@
 Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
+We provide a more detailed description of the model in Appendix \ref{metacomm}.
 
@@
 Our ``temporal model" estimates this weak, linear temperature effect, along with strong lag effects of biomass in the previous year.
-The spatial and temporal statistical
-models are described in Appendix \ref{models}.
 
@@
 We forecast the effect of a temperature increase (Fig. \ref{fig:community-forecast-species}A) on the focal species' biomass at one location in the center of the temperature gradient 
-using both spatial models and the temporal models. 
+using both the spatial model and the temporal model. 
 The predictions from the spatial and temporal models contrasted markedly, with the temporal 
@@
 
-Evolutionary adaptation is a key uncertainty in predicting how environmental change will impact a focal population at a given location (refs). Like the shifts
+Evolutionary adaptation is a key uncertainty in predicting how environmental change will impact a focal population at a given location \citep{Hoffmann2011}. Like the shifts
 in species composition illustrated in the previous example, shifts in genotype frequencies can also influence dynamics and forecasts at different time scales.
@@
 In our simulations, time-series approaches performed best for short-term forecasts, whereas models based on spatial data made more accurate long-term forecasts.
+
 The change in model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
@@
 We need new theory or empirical case studies in order to assign these weights \emph{a priori}.
+
 Theory could explore the influence of different parameters on the rate at which slow processes begin to influence dynamics.
@@
 \ref{fig:dispersal_niche_width}A).
-Other factors to consider include organism lifespans, niche widths, and the magnitude of directional environmental change relative to historical interannual variation.
+Other parameters have less intuitive effects: we expected that increasing the temperature tolerance of genotypes in the evo-evolutionary example would accelerate the shift in model weights by maintaining higher genetic diversity.
+Our simulations showed the opposite effect, with wider tolerances slowing the shift in model weights, presumably by decreasing the strength of selection.
+Additional factors to consider include organism lifespans and the magnitude of directional environmental change relative to historical interannual variation.
+
 Empirical research could address this challenge by accumulating enough case studies to infer patterns in the weighting functions and guide applications in new systems.
@@
 	\ln(N(t+1)) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 ln(N(t)) +\beta^T_2 K(t)  +  \varepsilon
-	\label{eqn:temporal_regression_community}
+	\label{eqn:temporal_regression_community1}
 \end{equation}
@@
 	\ln(\frac{N(t+1)}{N(t)}) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 ln(N(t)) +\beta^T_2 K(t)  +\beta^T_3 K(t)^2 +  \varepsilon
-	\label{eqn:temporal_regression_ecoevo}
+	\label{eqn:temporal_regression_ecoevo2}
 \end{equation}
@@
 \hat{N}(t+1)= \omega \cdot T(N(t),K(t)) + (1-\omega) \cdot S(K(t)) 
-\label{eqn:combined_model}
+\label{eqn:combined_model1}
 \end{equation}
@@
 	\hat{N}(t+1) = \omega \cdot exp(T(N(t),K(t))) N(t) + (1-\omega) \cdot S(K(t)) 
-	\label{eqn:combined_model}
+	\label{eqn:combined_model2}
 \end{equation}
@@
 
-\section{Description of the metacommunity model}\label{metacomm}
+\section{Description of the meta-community model}\label{metacomm}
+
+\cite{alexander_lags_2018} developed a meta-community model to represent dynamics of local communities arrayed along a one-dimensional elevation gradient, as influenced by three main processes: temperature-dependent growth, competition, and dispersal.
+Here we adapt their notation to be consistent our own.
+
+The population size of species $i$ in cell $x$ at time $t + 1$, $N_{i}(x,t+1)$, is computed in two steps.
+The first step accounts for changes in local population sizes due to dispersal.
+In each local community, all species export a fraction ($d$) of their local population to the two adjacent communities in the 1-dimensional landscape:
+\begin{equation}
+	N'_{i}(x,t) = (1-d) \cdot N_{i}(x,t) + \frac{d}{2} \cdot (N_{i}(x+1,t) + N_{i}(x-1,t))
+\end{equation}
+Here $N'$ distinguishes the post-dispersal population size from the pre-dispersal population size.
+
+The second step computes population growth, taking into account competition:
+\begin{equation}
+	N_{i}(x,t+1) = N'_{i}(x,t) + N'_{i}(x,t)[g_i(K(x) - Kmin_i) - c_i N'_{i}(x,t) - l_i \sum_{k} N'_{k}(x,t)]
+\end{equation}
+\comment{They have a delta t term right before the second P', but since this is a discrete time model and our time step always = 1, it seems confusing to add that...}
+In the absence of competition, the growth rate ($g$) is determined by the difference between the temperature ($K$) at site $x$ and the focal species' minimum temperature tolerance, $Kmin$, the lowest temperature at which a species can maintain a positive growth rate.
+Growth is further reduced by intraspecific ($c$) and interspecific ($l$) competition.
+All species are assigned the same value of $c$, which represents the additional limitation each species experiences from intraspecific competition.
+This offset also stabilizes coexistence, since every species will exert stronger intra- than interspecific competition.
+However, values of $l$ vary among species to create a trade-off between growth rates and competitive ability and low temperature tolerance: fast-growing species (high $g$) are more tolerant of interspecific competition (low $l$) but are more limited by temperature (high $Kmin$).
 
@@
 
-\renewcommand{\theequation}{B-\arabic{equation}}
-\renewcommand{\thetable}{B-\arabic{table}}
-\renewcommand{\thefigure}{B-\arabic{figure}}
+%\renewcommand{\theequation}{B-\arabic{equation}}
+%\renewcommand{\thetable}{B-\arabic{table}}
+%\renewcommand{\thefigure}{B-\arabic{figure}}
 

### 2019-09-10 doc/manuscript.tex
Revisions to appendices

@@
 
-The compositional turnover affecting our focal species also influences total biomass, linking community and ecosystem dynamics. We repeated our focal species analysis 
-for total community biomass, and the results were similar: the temporal model initially made the best forecasts immediately following the onset of the temperature
-increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (Figs.
-\ref{fig:community-models-total} to \ref{fig:community-weights-total}).
+The compositional turnover affecting our focal species also influences total biomass, linking community and ecosystem dynamics.
+We repeated our focal species analysis for total community biomass, and the results were similar: the temporal model initially made the best forecasts immediately following the onset of the temperature increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (Figs.
+\ref{fig:community-models-total} and \ref{fig:community-forecasts-total}).
 
@@
 \begin{equation}
-	\ln(N(t+1)) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 ln(N(t)) +\beta^T_2 K(t)  +  \varepsilon
+\ln(N(t+1)) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 \ln(N(t)) +\beta^T_2 K(t)  +  \varepsilon
 \label{eqn:temporal_regression_community1}
@@
 \begin{equation}
-	\ln(\frac{N(t+1)}{N(t)}) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 ln(N(t)) +\beta^T_2 K(t)  +\beta^T_3 K(t)^2 +  \varepsilon
+\ln\left(\frac{N(t+1)}{N(t)}\right) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 \ln(N(t)) +\beta^T_2 K(t)  +\beta^T_3 K(t)^2 +  \varepsilon
 \label{eqn:temporal_regression_ecoevo2}
@@
 \begin{equation}
-	\hat{N}(t+1) = \omega \cdot exp(T(N(t),K(t))) N(t) + (1-\omega) \cdot S(K(t)) 
+\hat{N}(t+1) = \omega \cdot \exp(T(N(t),K(t))) N(t) + (1-\omega) \cdot S(K(t)) 
 \label{eqn:combined_model2}
@@
 \end{equation}
-\comment{They have a delta t term right before the second P', but since this is a discrete time model and our time step always = 1, it seems confusing to add that...}
-In the absence of competition, the growth rate ($g$) is determined by the difference between the temperature ($K$) at site $x$ and the focal species' minimum temperature tolerance, $Kmin$, the lowest temperature at which a species can maintain a positive growth rate.
-Growth is further reduced by intraspecific ($c$) and interspecific ($l$) competition.
-All species are assigned the same value of $c$, which represents the additional limitation each species experiences from intraspecific competition.
-This offset also stabilizes coexistence, since every species will exert stronger intra- than interspecific competition.
-However, values of $l$ vary among species to create a trade-off between growth rates and competitive ability and low temperature tolerance: fast-growing species (high $g$) are more tolerant of interspecific competition (low $l$) but are more limited by temperature (high $Kmin$).
+In the absence of competition, the growth rate ($g_i$) is determined by the difference between the temperature at site $x$ ($K(x)$) and the focal species' minimum temperature tolerance, $Kmin_i$, the lowest temperature at which a species can maintain a positive growth rate.
+Growth is further reduced by intraspecific and interspecific competition, parameterized by $c_i$ and $l_i$.
+All species are assigned the same value of $c_i$, which represents an additional effect of intraspecific competition on top of interspecific competition.
+This stabilizes coexistence, since every species will exert stronger intra- than interspecific competition.
+However, values of $l$ vary among species to create a trade-off between growth rates and competitive ability versus low temperature tolerance: fast-growing species (high $g_i$) are more tolerant of interspecific competition (low $l_i$) but are more limited by temperature (high $Kmin_i$).
 
@@
 
+\newpage
+
 \section{Supplementary Figures}
 
-\renewcommand{\thefigure}{C-\arabic{figure}}
+\renewcommand{\thefigure}{S-\arabic{figure}}
 
-\begin{figure}[tbp]
+\begin{figure}[!ht]
 \centering
@@
 \caption{Results for total biomass from the community turnover model.
-(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(Bottom) Simulated total biomass and forecasts from the spatial, temporal and combined models.
+(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
+(B) Simulated total biomass and forecasts from the spatial, temporal and combined models.
+(C) Simulated changes in biomass of all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).
+Year 1000 in this figure corresponds to the start of the temperature increase.
 }
 \label{fig:community-forecasts-total}
-\end{figure}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_change_plus_weights_total.png}
-\caption{Simulated changes in biomass of all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).
-Year 0 in this figure corresponds to the start of the temperature increase.
-}
-\label{fig:community-weights-total}
 \end{figure}

### 2019-09-10 community-turnover/comm_turn_figures.R

@@ -107,21 +107,22 @@ png("figures/community_forecast_species.png",height=7.5,width=4.5,res=400,units=
   mtext(side=4,line=2,"Weight",cex=0.8)
   mtext(side=3," C",line=-1.1,adj=0,cex=0.8)
   
-  
   mtext(side=1,"Time",outer=T,line=1,cex=0.8)
   
 dev.off()
 
 # now total biomass
-png("figures/community_forecast_total.png",height=4.5,width=4,res=400,units="in")
+png("figures/community_forecast_total.png",height=7.5,width=4.5,res=400,units="in")
   
-  layout(matrix(c(1,2),2,1),heights=c(0.1,0.4,0.4),widths=1)
+  layout(matrix(c(1,2,3),3,1),heights=c(0.1,0.4,0.4),widths=1)
   
-  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(1,4,1,1),oma=c(2,0,0,0),cex=0.8)
+  par(tcl=-0.2,mgp=c(2,0.5,0),mar=c(1,4,1,4),oma=c(2,0,0,0),cex=0.8)
   
+  # temperature trend
   plot(my_obs_temp,xlab="",ylab="Temperature",type="l",col="darkgrey")
   abline(v=baseline_yrs,col="black",lty=3)
   lines(my_expected_temp,lwd=2,col="black")
+  mtext(side=3," A",line=-1.1,adj=0,cex=0.8)
   
   # total biomass
   plot(my_biomass,xlab="",ylab="Total biomass",type="l",col="darkgrey")
@@ -129,61 +130,30 @@ png("figures/community_forecast_total.png",height=4.5,width=4,res=400,units="in"
   lines(1:length(my_biomass),spatial_forecast,col="red",lwd=2)
   lines(1:length(my_biomass),temporal_forecast,col="blue",lwd=2)
   lines(1:length(my_biomass),combined_forecast,col="purple",lwd=2,lty=2)
-  
+  mtext(side=3," B",line=-1.1,adj=0,cex=0.8)
   legend("right",c("Observed","Spatial forecast","Temporal forecast","Combined forecast"),
          col=c("darkgrey","red","blue","purple"),lwd=c(1,2,2,2),lty=c(1,1,1,2),bty="n",cex=0.8)
+  
+  # weights
+  xx = 1:(sim_yrs-burnin_yrs+1)
[…]

### 2019-09-13 doc/manuscript.tex

@@
 
-Another approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is joint spatiotemporal modeling...\comment{Ethan, what do you have in mind here?
-Probably something more than just working with spatiotemporal datasets and accounting for autocorrelation.
-Mevin has lots of papers with some sort of mechanistic modeling of spatiotemoral processes like dispersal and colonization.
-Should we say that the datasets needed to train this kind of model are rare, but that might be changing with more and more remote sensing data that is extensive in space and time?
-}
+Another potential approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is to use models that incorporate both spatial and temporal information in data simultaneously.
+Spatiotemporal modeling approaches are increasingly common in ecology for patterns and processes that are of interest to forecsating (e.g., https://doi.org/10.1002/ecy.1643, https://doi.org/10.1890/0012-9658(2003)084[1382:HBMFPT]2.0.CO;2, https://doi.org/10.1111/geb.12666).
+However these models are rarely used in forecasting contexts.
+This is due in part to a combination of data limitation and computational challenges related to fitting and predicting from these models.
+Many data sources contain either spatial of temporal variation, but not both, and when spatiotemporal data is available it often involves irregular sampling, which creates challenges for modeling.
+Spatiotemporal models are also computationally intensive to fit and make predictions from, especially with large datasets (https://doi.org/10.1002/sta4.160).
+Fortunately thanks to large scale monitoring efforts like the National Ecological Observatory Network (citation), community science projects (e.g., eBird), and remote sensing large scale spatiotemporal data is increasingly available.
+In addition, new methods for spatiotemporal forecasting are being developed that address existing computational challenges (https://onlinelibrary.wiley.com/doi/full/10.1002/sta4.160) and access to high performance computing resources are increasingly common.
+As such, future ecological forecasting efforts should explore spatiotemporal approaches when data is available and assess whether these approaches produce improved predictions at intermediate time scales relative to spatial or temporal approaches alone.
 

### 2019-09-13 doc/manuscript.tex
Edited Ethan's spatiotemporal addition

--- 2019-09-10 doc/manuscript.tex
@@
 
-Another approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is joint spatiotemporal modeling...\comment{Ethan, what do you have in mind here?
-Probably something more than just working with spatiotemporal datasets and accounting for autocorrelation.
-Mevin has lots of papers with some sort of mechanistic modeling of spatiotemoral processes like dispersal and colonization.
-Should we say that the datasets needed to train this kind of model are rare, but that might be changing with more and more remote sensing data that is extensive in space and time?
-}
+Another potential approach for intermediate scale forecast horizons is to use models that incorporate both spatial and temporal information simultaneously.
+Spatiotemporal modeling approaches are being developed to study patterns and processes of interest to ecological forecasters, such the spread of an invasive species or population status of a threatened species  \citep{Wikle2003,Williams2017,Schliep2018}.
+Because these models include both fast processes, such as births and deaths, and slower processes, such as colonization and extinction dynamics, they have the potential to make better predictions at intermediate forecast horizons than purely spatial or temporal models.
+However, these spatiotemporal models have rarely been used in a forecasting context, due to a combination of data limitation and computational challenges.
+Many data sources contain either spatial or temporal variation, but not both, and when spatiotemporal datasets are available they often involve irregular sampling, creating challenges for modeling.
+Fitting and generating predictions from spatiotemporal models is also computationally intensive, especially with large datasets \citep{McDermott2017}.
+Fortunately, thanks to large-scale monitoring efforts from remote sensing platforms, the National Ecological Observatory Network (https://www.neonscience.org/), and community science projects (e.g., eBird), large scale spatiotemporal data is increasingly available.
+In addition, new methods for spatiotemporal forecasting are being developed that address existing computational challenges \citep{McDermott2017}, and access to high performance computing resources is increasingly common.
+Given these developments, future ecological forecasting efforts should explore spatiotemporal approaches and assess whether they improve predictions at intermediate time scales relative to traditional spatial or temporal models.
 

### 2019-09-13 doc/manuscript.tex
Acknowledgements

@@
 We then studied how the optimal model weights changed over time.
-These spatial, temporal and combined statistical
-models are described in Appendix \ref{models}.
 We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, and that a combination of the two approaches would provide the best forecasts at transitional, intermediate time scales.
+The spatial, temporal and combined statistical models are described in Appendix \ref{models} and all code for both case studies is available at Github (https://github.com/pbadler/space-time-forecast).
 
@@
 
+\subsubsection*{Acknowledgements}
+
+We thank Heather Lynch and Juan Manuel Morales for comments that improved early drafts of the paper.
+PBA was supported by NSF DEB-1927282 and the Utah Agriculture Experiment Station, EPW was supported by ??, and MHC was supported by ??.
+
 \newpage

### 2019-09-16 doc/manuscript.tex
Addressed Morales' comments

--- 2019-09-13 doc/manuscript.tex
@@
 \author[1]{Peter B. Adler}  %\thanks{Corresponding author. Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah Email: peter.adler@usu.edu}}
-\author[2]{Ethan P. White}
-\author[3]{Michael H. Cortez}
+\author[2,3,4]{Ethan P. White}
+\author[5]{Michael H. Cortez}
 \affil[1]{Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah}
 \affil[2]{Department of Wildlife Ecology and Conservation, University of Florida, Gainesville, Florida}
-\affil[3]{Department of Biological Science, Florida State University, Tallahasee, Florida}
+\affil[3]{Informatics Institute, University of Florida, Gainesville, Florida}
+\affil[4]{Biodiversity Institute, University of Florida, Gainesville, Florida}
+\affil[5]{Department of Biological Science, Florida State University, Tallahasee, Florida}
 
@@
 We made forecasts based on two phenomenological models, each representing processes operating at different time scales.
-The first phenomenological model represents the space-for-time substiution approach, which we call the ``spatial approach'' for brevity. 
+One model represents the time-series or ``temporal approach.'' We correlated interannual variation in an ecological response with interannual variation in the environment at just one site.
+The other model represents the space-for-time substitution approach, which we call the ``spatial approach'' for brevity.
 We correlated the mean environment with the mean of an ecological state or rate across many sites.
-The second model represents the time-series or ``temporal approach.'' We correlated interannual variation in an ecological response with interannual variation in the environment at just one site.
-We compared forecasts from both models to the simulated dynamics to determine how well the two approaches perform at different forecast horizons.
-We also assessed the potential for combining the information available in both spatial and temporal patterns by using a weighted average of the forecasts from the spatial and temporal approaches optimized to best match the (simulated) observations.
+We compared forecasts from both models to the simulated dynamics to determine how well the two approaches performed at different forecast horizons.
+We also assessed the potential for combining the information available in temporal and spatial patterns by using a weighted average of the forecasts from the temporal and spatial approaches optimized to best match the (simulated) observations.
 We then studied how the optimal model weights changed over time.
-These spatial, temporal and combined statistical
-models are described in Appendix \ref{models}.
 We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, and that a combination of the two approaches would provide the best forecasts at transitional, intermediate time scales.
+The spatial, temporal and combined statistical models are described in Supporint Information (Appendix \ref{models}) and all code for both case studies is available at Github (https://github.com/pbadler/space-time-forecast).
 
@@
 But at what forecast horizon do altered species interactions become impossible to ignore?
-We explored this question using a metacommunity model developed by \cite{alexander_lags_2018}.
-The model describes Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
+We explored this question using a metacommunity model developed by \cite{alexander_lags_2018} to study how community responses to increasing temperature depend on the interplay between within-site demography and competitive interactions and the movement of species across sites.
+The model features Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
 Composition varies along the gradient because of a trade-off between growth rate and cold tolerance: cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by  fast-growing species that are cold intolerant.
@@
 Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
-We provide a more detailed description of the model in Appendix \ref{metacomm}.
+We provide a more detailed description of the model in SI Appendix \ref{metacomm}.
 
@@
 In contrast to the strong spatial patterns, population and community responses to interannual variation in temperature within sites were weak.
-At our focal site in the center of the gradient, the biomass of the focal species was quite insensitive to interannnual variation in temperature (Fig.
+At our focal site in the center of the gradient, the biomass of the focal species was quite insensitive to interannnual variation in temperature, but showed strong temporal autocorrelation (Fig.
 \ref{fig:species-patterns-models}B).
-Our ``temporal model" estimates this weak, linear temperature effect, along with strong lag effects of biomass in the previous year.
+Our ``temporal model" estimates this weak, linear temperature effect, along with the strong lag effect of biomass in the previous year.
 
@@
 
-We forecast the effect of a temperature increase (Fig. \ref{fig:community-forecast-species}A) on the focal species' biomass at one location in the center of the temperature gradient 
-using both the spatial model and the temporal model. 
+We fit both a temporal and a spatial statistical model to forecast the effect of a temperature increase (Fig.
+\ref{fig:community-forecast-species}A) on the focal species' biomass at one location in the center of the temperature gradient.
 The predictions from the spatial and temporal models contrasted markedly, with the temporal 
@@
 
-To combine the spatial and temporal model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, the combined population forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see Appendix \ref{models} for full description of the approach).
+To combine the temporal and spatial model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
+At any time point, $t$, the combined population forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
 This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}B), but 
@@
 The compositional turnover affecting our focal species also influences total biomass, linking community and ecosystem dynamics.
-We repeated our focal species analysis for total community biomass, and the results were similar: the temporal model initially made the best forecasts immediately following the onset of the temperature increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (Figs.
+We repeated our focal species analysis for total community biomass, and the results were similar: the temporal model initially made the best forecasts immediately following the onset of the temperature increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (SI Figs.
 \ref{fig:community-models-total} and \ref{fig:community-forecasts-total}).
@@
 
-Evolutionary adaptation is a key uncertainty in predicting how environmental change will impact a focal population at a given location \citep{Hoffmann2011}. Like the shifts
-in species composition illustrated in the previous example, shifts in genotype frequencies can also influence dynamics and forecasts at different time scales.
-We demonstrate this process with a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
+Evolutionary adaptation is a key uncertainty in predicting how environmental change will impact a focal population at a given location \citep{Hoffmann2011}.
+Like the shifts in species composition illustrated in the previous example, shifts in genotype frequencies can also influence dynamics and forecasts at different time scales.
+Although shifts in genotype frequencies at the population level are analogous to changes in species composition at the community level, the mechanisms are distinct: heterozygosity and genetic recombination have no analogue at the community level.
+We demonstrate how these processes influence short and long-term forecasts with a standard eco-evolutionary model for a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
 \ref{fig:evo_evo_spatial_temporal_models}A).
@@
 \ref{fig:evo_evo_spatial_temporal_models}B).
-The full model description is provided in Appendix \ref{eco-evo}.
+The full model description is provided in SI Appendix \ref{eco-evo}.
 
@@
 
-The change in model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
+We designed our simulation studies to illustrate how the change in model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
 Local time-series data capture demographic processes, lagged effects, and responses of current assemblages to small changes in environmental conditions.
@@
 The hypothesis that time-series models will be most effective for near-term forecasts is supported by recent empirical analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
-In contrast, datasets spanning spatial gradients capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
-Models based on data from spatial gradients may be a better choice for predicting long-term, steady-state outcomes of ecological change.
+In contrast, datasets spanning spatial gradients should capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
+Models based on data from spatial gradients may be a better choice for predicting long-term, steady-state outcomes of ecological change \citep{Blois2013}.
 
@@
 Even if the general form of the models was known, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well studied systems.
+% mention difficulty of dealing with error propogation?
 As a result, models used for ecological forecasting will include at least some phenomenological components.
@@
 While the recent emphasis on near-term iterative forecasting \citep{dietze_iterative_2018} is the logical and tractable starting point, we also need to build understanding and capacity for forecasting ecological dynamics across all temporal scales of interest.
+
+\subsubsection*{Acknowledgements}
+
+We thank Heather Lynch and Juan Manuel Morales for comments that improved early drafts of the paper.
+PBA was supported by the Utah Agriculture Experiment Station.
+EPW was supported by the Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative through Grant GBMF4563.
 
@@
 We assume that the temperature, $K(x,t)$, at each location can vary in time and space.
+To forecast the dynamics generated by these simulations models, we fit a series of statistical models.
 
@@
 where $g_{i}(K(t))$ is the probability of germination, $K(t)$ is the temperature at time $t$, $s_i$ is the seed survival probability for species $i$, and $\lambda_i$ is the seed production rate per plant.
-Below we refer to the $\alpha_{ij}$ as intra- and inter-specific competition coefficients.
+Below we refer to the $\alpha_{ij}$ as intra- and inter-genotype competition coefficients.
 
@@
 Including competition changes the way in which we compute $T$, $\phi_A$, and $\phi_a$.
-Specifically, because the total number of seeds produced per year by each genotypes is reduced based on intra- and inter-specific competition, the total number of gamete-pairs becomes
+Specifically, because the total number of seeds produced per year by each genotypes is reduced based on intra- and inter-genotype competition, the total number of gamete-pairs becomes
 \begin{align}\begin{split}

### 2019-09-17 genetic-diversity/genetic_diversity_sim_spatial.R
Minutia

@@ -34,4 +34,5 @@ for(iT in 1:length(Tmean)){
 simN_mix[is.na(simN_mix)] <- 0 # replace missing values with zero
 
 # fit a "spatial" model describing mean abundance as a function of mean temperature
-spatial_model <- lm(N_Pop~Tmean + I(Tmean^2),data=simN_mix)
+Tmean2 <- simN_mix$Tmean^2
+spatial_model <- lm(N_Pop~Tmean + Tmean2,data=simN_mix)

### 2019-09-18 doc/manuscript.tex
Uncertainty bits added

@@
 
+In the main text, we show the point forecasts but not the uncertainty around the forecasts.
+After exploring that uncertainty, we decided that presenting it would be misleading.
+For the spatial and, especially, the temporal statistical models, the uncertainty is unrealistically low, because the models are estimated with very large samples sizes from the simulations.
+Furthermore, the simulations do not include noise; the only reason there is any uncertainty is because the statistical models are slightly mis-specified with respect to the process models.
+Showing uncertainty for the combined model would be even less meaningful, because it is not a true, out-of-sample forecast (parameters are fit directly to the observations for which we make predictions).
+The R code to compute uncertainties for the spatial and temporal forecasts is available on our Github repository (https://github.com/pbadler/space-time-forecast), but is commented out.
+
 \section{Description of the meta-community model}\label{metacomm}

### 2019-09-18 community-turnover/comm_turn_forecast.R

@@ -23,6 +23,9 @@ forecast_Tmean <- Tmean[site] + c(seq(deltaT/warming_yrs,deltaT,deltaT/warming_y
 
 # make predictions from spatial model
 spatial_forecast <- coef(spatial_model)[1] + coef(spatial_model)[2]*forecast_Tmean 
+# make predictions from spatial model with uncertainty
+# spatial_forecast_spp <- predict(spatial_model_spp,newdata=data.frame(x_spatial=forecast_Tmean),
+#                                 interval="predict")
 
 # make predictions from temporal model
 temporal_forecast <- rep(NA,length(forecast_Tmean)+1)
@@ -33,6 +36,18 @@ for(iTime in 1:length(forecast_Tmean)){
 }
 temporal_forecast <- temporal_forecast[2:length(temporal_forecast)]  # drop first ("observed") year
 
+# make predictions from temporal model using a Monte Carlo approach for uncertainty
+# ndraws <- 100
+# pars <- rmvnorm(ndraws,coef(temporal_model_spp),vcov(temporal_model_spp))
+# temporal_forecast_spp <- matrix(NA,length(forecast_Tmean)+1,ndraws)
+# temporal_forecast_spp[1,] <- y_temporal_spp[length(y_temporal)] # initialize at biomass from last baseline year
+# for(iPar in 1:ndraws){
+#   for(iTime in 1:length(forecast_Tmean)){
+#     temporal_forecast_spp[iTime+1,iPar] <- pars[iPar,1] +pars[iPar,2]*temporal_forecast_spp[iTime,iPar] + 
+#       pars[iPar,3]*forecast_Tmean[iTime]
+#   }
+# }
+# temporal_forecast_spp <- temporal_forecast_spp[2:dim(temporal_forecast_spp)[1],]  # drop first ("observed") year
 
 # fit combined model
 y <- rowSums(spxp[(burnin_yrs+baseline_yrs+1):sim_yrs,site,])

### 2019-09-18 genetic-diversity/genetic_diversity_forecast.R

@@ -7,6 +7,9 @@ forecast_Tmean <-  Tmean[(tmp+1):length(Tmean)]
 
 # make predictions from spatial model
 spatial_forecast <- coef(spatial_model)[1] + coef(spatial_model)[2]*forecast_Tmean + coef(spatial_model)[3]*forecast_Tmean^2
+# next line to make predictions with uncertainty
+# spatial_forecast <- predict(spatial_model,newdata=data.frame(Tmean =forecast_Tmean,Tmean2 = forecast_Tmean^2),
+#                             interval="prediction")
 
 # make predictions from temporal model
 temporal_forecast <- rep(NA,length(forecast_Tmean)+1)
@@ -17,6 +20,21 @@ for(iTime in 1:length(forecast_Tmean)){
   temporal_forecast[iTime+1] <- exp(r_iTime)*temporal_forecast[iTime]
 }
 temporal_forecast <- temporal_forecast[2:length(temporal_forecast)]  # drop first ("observed") year
+
+# make predictions from temporal model with uncertainty (Monte Carlo approach)
+# ndraws <- 100
+# pars <- rmvnorm(ndraws,coef(temporal_model),vcov(temporal_model))
+# temporal_forecast <- matrix(NA,length(forecast_Tmean)+1,ndraws)
+# temporal_forecast[1,] <- N[baseline_yrs] # initialize at N from last baseline year
+# for(iPar in 1:ndraws){
+#   for(iTime in 1:length(forecast_Tmean)){
+#     r_iTime <- pars[iPar,1] +pars[iPar,2]*log(temporal_forecast[iTime,iPar]) + 
+#       pars[iPar,3]*forecast_Tmean[iTime] + pars[iPar,4]*forecast_Tmean[iTime]^2
+#     temporal_forecast[iTime+1,iPar] <- exp(r_iTime)*temporal_forecast[iTime,iPar]
+#   }
+# }
+# temporal_forecast <- temporal_forecast[2:dim(temporal_forecast)[1],]  # drop first ("observed") year
+
 
 # fit combined model
 y <- N[(baseline_yrs+1):sim_yrs]

### 2019-10-03 doc/manuscript.tex
Revised to address Drake suggestions

@@
 The influence of these slow processes is seen in very long time series, or in spatial gradients.
-Spatiotemporal datasets with long enough records to capture changes in spatial patterns are rare, but would be extremely valuable for understanding dynamics at intermediate time scales.
-The time scales shown here were chosen with vascular plants in mind; the same concepts would apply for much shorter-lived organisms but at shorter time scales.
+\new{Understanding dynamics at intermediate time scales requires integrating information from spatial and temporal data sources.
+We propose a model weighting approach; mechanistic spatiotemporal modeling is an alternative}.
+The time scales shown here were chosen with vascular plants in mind, but the same concepts would apply for much shorter-lived organisms but at shorter time scales.
 }
@@
 depend on the time horizon of the forecast, and 2) explore how time-series and space-for-time approaches might be 
-combined to make better forecasts at intermediate time scales. We conduct two simulation case studies, one 
+combined via weighted averaging to make better forecasts at intermediate time scales. We conduct two simulation case studies, one 
 focused on how interspecific interactions affect the population dynamics of a focal species, and the second focused on
@@
 We also assessed the potential for combining the information available in temporal and spatial patterns by using a weighted average of the forecasts from the temporal and spatial approaches optimized to best match the (simulated) observations.
-We then studied how the optimal model weights changed over time.
-We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, and that a combination of the two approaches would provide the best forecasts at transitional, intermediate time scales.
-The spatial, temporal and combined statistical models are described in Supporint Information (Appendix \ref{models}) and all code for both case studies is available at Github (https://github.com/pbadler/space-time-forecast).
+\new{We call this spatial-temporal-weighting approach the ``STeW'' model.} We then studied how the optimal model weights changed over time.
+We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, \new{while the STeW model wuld show potential} to provide the best forecasts at transitional, intermediate time scales.
+The spatial, temporal and STeW models are described in Supporting Information (Appendix \ref{models}) and all code for both case studies is available at Github (https://github.com/pbadler/space-time-forecast).
 
@@
 
-To combine the temporal and spatial model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
+To combine the temporal and spatial model into a single spatial-temporal-weighted (STeW) forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
 At any time point, $t$, the combined population forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
-This combined model accurately predicts the simulated dynamics across the full forecast horizon (Fig. \ref{fig:community-forecast-species}B), but 
-the reason it works so well is that the weights were determined by fitting directly to the data. Unlike our spatial and temporal model forecasts, 
-we did not generate out-of-sample predictions from the combined model; it merely provides 
-a convenient way to quantify how rapidly dynamics shift from being dominated by interannual 
-variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
+The STeW model accurately predicts the simulated dynamics across the full forecast horizon (Fig.
+\ref{fig:community-forecast-species}B), but the reason it works so well is that the weights were determined by fitting directly to the data.
+Unlike our spatial and temporal model forecasts, we did not generate out-of-sample predictions from the combined model; it merely provides a convenient way to quantify how rapidly dynamics shift from being dominated by interannual variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
 \ref{fig:community-forecast-species}B) to being dominated by the steady-state equilibrium captured by the spatial model (time $t \geq 2500$).
-A true, combined forecast would require a method to determine the model weights \emph{a priori}.
-The most rapid shifts in the model weights occurred during the period when warm-adapted, faster growing species were increasing most rapidly
-in abundance (Fig. \ref{fig:community-forecast-species}C).
+A true STeW forecast would require a method to determine the model weights \emph{a priori}.The most rapid shifts in the model weights occurred during the period when warm-adapted, faster growing species were increasing most rapidly in abundance (Fig.
+\ref{fig:community-forecast-species}C).
 
@@
 \caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated focal species biomass and forecasts from the spatial, temporal and combined models at the focal site in the metacommunity model.
+(B) Simulated focal species biomass and forecasts from the spatial, temporal and STeW models at the focal site in the metacommunity model.
 (C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal model for focal species biomass (blue).
@@
 
-As in the community turnover example, we also fit a combined model as a weighted average of the spatial and temporal model, with the weights changing over time.
-The combined model initially reflected the temporal model (decrease from $t=500$ to $t=600$), but then rapidly transitioned to reflect the spatial model ($t \geq 700$).
+As in the community turnover example, we also fit a weighted average of the spatial and temporal model, with the weights changing over time.
+This STeW model initially reflected the temporal model (decrease from $t=500$ to $t=600$), but then rapidly transitioned to reflect the spatial model ($t \geq 700$).
 The rapid transition in the weighting term, $\omega$, occurred during the period of most rapid change in genotype frequencies (Fig.
 \ref{fig:forecast_supp}).
-The combined model's predictions look impressively accurate, but, as in the community turnover example, that is because we used the full, simulated time series to fit the weighting term.
+The STeW model's predictions look impressively accurate, but, as in the community turnover example, that is because we used the full, simulated time series to fit the weighting term.
 To generate a true forecast from the combined model, we would need an independent method to predict how the model weights shift over time.
@@
 \caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
-(Bottom) Simulated population size and forecasts from the spatial, temporal and combined models.
+(Bottom) Simulated population size and forecasts from the spatial, temporal and STeW models.
 }
@@
 In our simulations, time-series approaches performed best for short-term forecasts, whereas models based on spatial data made more accurate long-term forecasts.
+In addition, our simulations demonstrate extended transitional periods during which neither the time-series or the spatial approach is effective on its own.
+The challenge is determining what is ``near-term'' and what is ``long-term,'' and how to handle the many forecasts we need in ecology which fall in between.
+We have proposed that a weighted combination of the time-series and space-for-time approaches may produce better forecasts at these intermediate forecast horizons.
 
 We designed our simulation studies to illustrate how the change in model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
-Local time-series data capture demographic processes, lagged effects, and responses of current assemblages to small changes in environmental conditions.
+\new{How could these hypotheses be tested with empirical data?
+The hypothesis that time-series models will be most effective for near-term forecasts already has empirical support, in the form of recent analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
+The result should not be surprising, since local time-series data capture demographic processes, lagged effects, and responses of current assemblages to small changes in environmental conditions.
 In addition, the state of the system in the near future depends heavily on the current state.
-Since short-term forecasts do not typically require extrapolating into novel conditions, a model based on the historical range of variation which incorporates lags and accurate initial conditions is likely to be the most generally useful approach for forecasting.
-The hypothesis that time-series models will be most effective for near-term forecasts is supported by recent empirical analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
-In contrast, datasets spanning spatial gradients should capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
-Models based on data from spatial gradients may be a better choice for predicting long-term, steady-state outcomes of ecological change \citep{Blois2013}.
+Since short-term forecasts do not typically require extrapolating into novel conditions, a model based on the historical range of variation which incorporates lags and accurate initial conditions is likely to be successful.
 
-The challenge is determining what is ``near-term'' and what is ``long-term,'' and how to handle the many forecasts we need in ecology which fall in between.
-Our simulations demonstrate extended transitional periods during which neither the time-series or the spatial approach is effective on its own, but a weighted combination of the two can produce highly accurate predictions.
-Unfortunately, we currently have no way to determine how to weight these two approaches for different time horizons.
-Our combined models used weights fitted directly to the simulated data, which is impossible to do for actual forecasting when the future is unknown.
+Space-for-time modeling approaches for predicting long-term, steady-state outcomes of ecological change have also been tested empirically, primarily via hind-casting.
+Overall, the results are mixed: some studies show fairly successful prediction of changes in community composition \citep{Blois2013} or species distributions (REFS), supporting the hypthesis that datasets spanning spatial gradients capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
+Other attempts to validate predictions from space-for-time models have been discouraging (REFS), suggesting violations of model assumptions or effects of transient dynamics.
+However, predictions from the space-for-time approaches are rarely compared directly to predictions from time-series models (Renwick).
+We need more such comparisons to identify the appropriate modeling approach for different forecast horizons.
+
+\new{The greatest empirical challenge will be testing our hypothesis that a weighted average of spatial and temporal models will make the best forecasts at intermediate time scales.
+There are two problems: finding appropriate data and determining the model weights \emph{a priori}.
+Many data sets with both a longitudinal and spatial dimension exist, but could not think of one which also featured a clear ecological response to significant environmental change.
+Surely such datasets exist, and we hope researchers who work with them will test our proposed STeW model.
+Determining model weights may be more difficult.
+In our simulations, we fitted the weights directly to the simulated data, which is impossible to do for actual forecasting when the future is unknown.
 We need new theory or empirical case studies in order to assign these weights \emph{a priori}.
+}
 
@@
 
-Empirical research could address this challenge by accumulating enough case studies to infer patterns in the weighting functions and guide applications in new systems.
+Empirical research could inform model weights by accumulating enough case studies to infer patterns in the weighting functions and guide applications in new systems.
 Developing  rules of thumb would require testing many forecasts from both time-series and spatial models across a range of time-horizons.
@@
 
-Another potential approach for intermediate scale forecast horizons is to use models that incorporate both spatial and temporal information simultaneously.
-Spatiotemporal modeling approaches are being developed to study patterns and processes of interest to ecological forecasters, such the spread of an invasive species or population status of a threatened species  \citep{Wikle2003,Williams2017,Schliep2018}.
+\new{While fully process-based models may not be practical, there are more mechanistic alternatives to our phenomenological STeW model for integrating spatial and temporal information}.
+Spatiotemporal statistical modeling approaches are being developed to study patterns and processes of interest to ecological forecasters, such the spread of an invasive species or population status of a threatened species  \citep{Wikle2003,Williams2017,Schliep2018}.
 Because these models include both fast processes, such as births and deaths, and slower processes, such as colonization and extinction dynamics, they have the potential to make better predictions at intermediate forecast horizons than purely spatial or temporal models.
@@
 
-Our results have important implications for the emerging field of ecological forecasting. First, they suggest that evaluating both near-term and long-term
-forecasts will be essential as research on forecasting methods accelerates.
+Our results have important implications for the emerging field of ecological forecasting.
+First, they suggest that evaluating both near-term and long-term forecasts will be essential as research on forecasting methods accelerates.
 Second, while single approaches may perform reasonably well at either short or long forecast horizons, skillful predictions at intermediate time horizons may require a combination of information from spatial and temporal patterns.
@@
 
-\section{Spatial, temporal and combined models}\label{models}
+\section{Spatial, temporal and spatial-temporal-weighted models}\label{models}
 
@@
 
-The combined model is a weighted average of predictions from the spatial and temporal models, with the weights changing as a function of time, here expressed as the forecast horizon.
+The spatial-temporal-weighted (STeW) model is a weighted average of predictions from the spatial and temporal models, with the weights changing as a function of time, here expressed as the forecast horizon.
 The weights change as a function of the square root of the forecast horizon, to allow rapid shifts in the model weights.
@@
 Furthermore, the simulations do not include noise; the only reason there is any uncertainty is because the statistical models are slightly mis-specified with respect to the process models.
-Showing uncertainty for the combined model would be even less meaningful, because it is not a true, out-of-sample forecast (parameters are fit directly to the observations for which we make predictions).
+Showing uncertainty for the STeW model would be even less meaningful, because it is not a true, out-of-sample forecast (parameters are fit directly to the observations for which we make predictions).
 The R code to compute uncertainties for the spatial and temporal forecasts is available on our Github repository (https://github.com/pbadler/space-time-forecast), but is commented out.

### 2019-10-03 doc/manuscript.tex
Intermediate commit, editing ms

@@
 model predictions \citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
-The combined benefits of informing policy and management and advancing basic research 
+The dual benefits of informing management and advancing basic knowledge 
 makes forecasting an important priority for ecological research.
@@
 time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of 
-environmental conditions represented in observed data.
+environmental conditions represented in observed data (Fig. \ref{fig:fast-slow-concept}).
 At short forecast horizons (days to years), dynamics will reflect the physiological and demographic responses of 
@@
 	\centering
-	\includegraphics[width=0.5 \textwidth] {fast-slow-figure.png}
+	\includegraphics[width=0.7 \textwidth] {fast-slow-figure.png}
 	\caption{Fast and slow processes operate at different time scales, and are reflected in different kinds of datasets.
@@
 To combine the temporal and spatial model into a single spatial-temporal-weighted (STeW) forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, the combined population forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
+At any time point, $t$, the STeW forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
 The STeW model accurately predicts the simulated dynamics across the full forecast horizon (Fig.
-\ref{fig:community-forecast-species}B), but the reason it works so well is that the weights were determined by fitting directly to the data.
-Unlike our spatial and temporal model forecasts, we did not generate out-of-sample predictions from the combined model; it merely provides a convenient way to quantify how rapidly dynamics shift from being dominated by interannual variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
+\ref{fig:community-forecast-species}B).
+It also shows that the most rapid shifts in the model weights occurred during the period when warm-adapted, faster growing species were increasing most rapidly in abundance (Fig.
+\ref{fig:community-forecast-species}C).
+However, the reason the STeW models works so well is that the weights were determined by fitting directly to the data.
+Unlike our spatial and temporal model forecasts, we did not generate out-of-sample predictions from the STeW model; it merely provides a convenient way to quantify how rapidly dynamics shift from being dominated by interannual variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
 \ref{fig:community-forecast-species}B) to being dominated by the steady-state equilibrium captured by the spatial model (time $t \geq 2500$).
-A true STeW forecast would require a method to determine the model weights \emph{a priori}.The most rapid shifts in the model weights occurred during the period when warm-adapted, faster growing species were increasing most rapidly in abundance (Fig.
-\ref{fig:community-forecast-species}C).
+A true STeW forecast would require a method to determine the model weights \emph{a priori}.
 
@@
 \caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated focal species biomass and forecasts from the spatial, temporal and STeW models at the focal site in the metacommunity model.
+(B) Simulated focal species biomass and forecasts from the spatial, temporal and combined (STeW) models at the focal site in the metacommunity model.
 (C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal model for focal species biomass (blue).
@@
 The STeW model's predictions look impressively accurate, but, as in the community turnover example, that is because we used the full, simulated time series to fit the weighting term.
-To generate a true forecast from the combined model, we would need an independent method to predict how the model weights shift over time.
+A true forecast would require an independent method to predict how the model weights shift over time.
 
@@
 \caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
-(Bottom) Simulated population size and forecasts from the spatial, temporal and STeW models.
+(Bottom) Simulated population size and forecasts from the spatial, temporal and combined (STeW) models.
 }
@@
 We designed our simulation studies to illustrate how the change in model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
-\new{How could these hypotheses be tested with empirical data?
-The hypothesis that time-series models will be most effective for near-term forecasts already has empirical support, in the form of recent analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
+\new{How could these hypotheses be tested with empirical data?} The hypothesis that time-series models will be most effective for near-term forecasts already has empirical support, in the form of recent analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
 The result should not be surprising, since local time-series data capture demographic processes, lagged effects, and responses of current assemblages to small changes in environmental conditions.
@@
 \end{equation}
-For the community turnover example, the predicted biomass from the combined model is:
+For the community turnover example, the predicted biomass from the STeW model is:
 \begin{equation}
@@
 \end{equation}
-Again, we suppress the spatial subscript ($x$) here because we are focused on densities at just one location.
-For the eco-evolutionary example, the predicted population size from the combined model is:
+Again, we suppress the spatial subscript ($x$) here because we are focused on densities at just one location. For the eco-evolutionary example, the predicted population size from the STeW model is:
 \begin{equation}
@@
 (A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated total biomass and forecasts from the spatial, temporal and combined models.
+(B) Simulated total biomass and forecasts from the spatial, temporal and combined (STeW) models.
 (C) Simulated changes in biomass of all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).

### 2019-10-04 doc/manuscript.tex
Finished ms and bib edits

@@
 \new{Understanding dynamics at intermediate time scales requires integrating information from spatial and temporal data sources.
-We propose a model weighting approach; mechanistic spatiotemporal modeling is an alternative}.
+We propose a model weighting approach; mechanistic spatiotemporal modeling is another alternative}.
 The time scales shown here were chosen with vascular plants in mind, but the same concepts would apply for much shorter-lived organisms but at shorter time scales.
@@
 We also assessed the potential for combining the information available in temporal and spatial patterns by using a weighted average of the forecasts from the temporal and spatial approaches optimized to best match the (simulated) observations.
-\new{We call this spatial-temporal-weighting approach the ``STeW'' model.} We then studied how the optimal model weights changed over time.
+\new{We call this Spatial and Temporal Weighting approach the ``STeW'' model.} We then studied how the optimal model weights changed over time.
 We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, \new{while the STeW model wuld show potential} to provide the best forecasts at transitional, intermediate time scales.
-The spatial, temporal and STeW models are described in Supporting Information (Appendix \ref{models}) and all code for both case studies is available at Github (https://github.com/pbadler/space-time-forecast).
+The three statistical models are described in Supporting Information (Appendix \ref{models}) and all code for both case studies is available at Github (https://github.com/pbadler/space-time-forecast).
 
@@
 
-To combine the temporal and spatial model into a single spatial-temporal-weighted (STeW) forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, the STeW forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
+To combine the temporal and spatial model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
+At any time point, $t$, this STeW forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
 The STeW model accurately predicts the simulated dynamics across the full forecast horizon (Fig.
@@
 In addition, our simulations demonstrate extended transitional periods during which neither the time-series or the spatial approach is effective on its own.
-The challenge is determining what is ``near-term'' and what is ``long-term,'' and how to handle the many forecasts we need in ecology which fall in between.
+The challenge is determining what is ``short-term,'' what is ``long-term,'' and how to handle the many forecasts we need in ecology which fall in between.
 We have proposed that a weighted combination of the time-series and space-for-time approaches may produce better forecasts at these intermediate forecast horizons.
@@
 
-Space-for-time modeling approaches for predicting long-term, steady-state outcomes of ecological change have also been tested empirically, primarily via hind-casting.
-Overall, the results are mixed: some studies show fairly successful prediction of changes in community composition \citep{Blois2013} or species distributions (REFS), supporting the hypthesis that datasets spanning spatial gradients capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
-Other attempts to validate predictions from space-for-time models have been discouraging (REFS), suggesting violations of model assumptions or effects of transient dynamics.
-However, predictions from the space-for-time approaches are rarely compared directly to predictions from time-series models (Renwick).
+\new{Space-for-time modeling approaches for predicting long-term, steady-state outcomes of ecological change have also been tested empirically, primarily via hind-casting.
+Overall, the results are mixed: some tests show skillful prediction of changes in community composition \citep{Blois2013,Illan2014} or species distributions \citep{Norberg2019}, supporting the hypothesis that datasets spanning spatial gradients capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
+Other attempts to validate predictions from space-for-time models have been discouraging \citep{Worth2014,Illan2014,Davis2014,Brun2016,Veloz2012}, indicating violations of model assumptions or effects of transient dynamics.
+However, predictions from the space-for-time approaches are rarely compared directly to predictions from time-series models (e.g., \citealt{Renwick2018}).
 We need more such comparisons to identify the appropriate modeling approach for different forecast horizons.
+}
 
@@
 There are two problems: finding appropriate data and determining the model weights \emph{a priori}.
-Many data sets with both a longitudinal and spatial dimension exist, but could not think of one which also featured a clear ecological response to significant environmental change.
+Many data sets have both a longitudinal and spatial dimension, but we could not think of one which also featured a clear ecological response to significant environmental change.
 Surely such datasets exist, and we hope researchers who work with them will test our proposed STeW model.
 Determining model weights may be more difficult.
-In our simulations, we fitted the weights directly to the simulated data, which is impossible to do for actual forecasting when the future is unknown.
+In our simulations, we fit the weights directly to the simulated data, which is impossible to do for actual forecasting when the future is unknown.
 We need new theory or empirical case studies in order to assign these weights \emph{a priori}.
@@
 Other parameters have less intuitive effects: we expected that increasing the temperature tolerance of genotypes in the evo-evolutionary example would accelerate the shift in model weights by maintaining higher genetic diversity.
-Our simulations showed the opposite effect, with wider tolerances slowing the shift in model weights, presumably by decreasing the strength of selection.
+Our simulations showed the opposite effect, with wider tolerances slowing the shift in model weights (Fig.
+\ref{fig:dispersal_niche_width}B), presumably by decreasing the strength of selection.
 Additional factors to consider include organism lifespans and the magnitude of directional environmental change relative to historical interannual variation.
@@
 
-\new{While fully process-based models may not be practical, there are more mechanistic alternatives to our phenomenological STeW model for integrating spatial and temporal information}.
+\new{While fully process-based models may not be practical, there are more mechanistic alternatives to our extremely phenomenological STeW model for integrating spatial and temporal information}.
 Spatiotemporal statistical modeling approaches are being developed to study patterns and processes of interest to ecological forecasters, such the spread of an invasive species or population status of a threatened species  \citep{Wikle2003,Williams2017,Schliep2018}.

### 2019-10-15 doc/manuscript.tex
more ms edits

@@
 The influence of these slow processes is seen in very long time series, or in spatial gradients.
-\new{Understanding dynamics at intermediate time scales requires integrating information from spatial and temporal data sources.
-We propose a model weighting approach; mechanistic spatiotemporal modeling is another alternative}.
+Understanding dynamics at intermediate time scales requires integrating information from spatial and temporal data sources.
+We propose a model weighting approach; mechanistic spatiotemporal modeling is another alternative.
 The time scales shown here were chosen with vascular plants in mind, but the same concepts would apply for much shorter-lived organisms but at shorter time scales.
@@
 We also assessed the potential for combining the information available in temporal and spatial patterns by using a weighted average of the forecasts from the temporal and spatial approaches optimized to best match the (simulated) observations.
-\new{We call this Spatial and Temporal Weighting approach the ``STeW'' model.} We then studied how the optimal model weights changed over time.
-We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, \new{while the STeW model wuld show potential} to provide the best forecasts at transitional, intermediate time scales.
+We then studied how the optimal model weights changed over time.
+We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, while the weighted model would show potential to provide the best forecasts at transitional, intermediate time scales.
 The three statistical models are described in Supporting Information (Appendix \ref{models}) and all code for both case studies is available at Github (https://github.com/pbadler/space-time-forecast).
@@
 To combine the temporal and spatial model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, this STeW forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
-The STeW model accurately predicts the simulated dynamics across the full forecast horizon (Fig.
+At any time point, $t$, this weighted forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
+The weighted model accurately predicts the simulated dynamics across the full forecast horizon (Fig.
 \ref{fig:community-forecast-species}B).
@@
 \ref{fig:community-forecast-species}C).
-However, the reason the STeW models works so well is that the weights were determined by fitting directly to the data.
-Unlike our spatial and temporal model forecasts, we did not generate out-of-sample predictions from the STeW model; it merely provides a convenient way to quantify how rapidly dynamics shift from being dominated by interannual variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
+However, the reason the weighted models works so well is that the weights were determined by fitting directly to the data.
+Unlike our spatial and temporal model forecasts, we did not generate out-of-sample predictions from the weighted model; it merely provides a convenient way to quantify how rapidly dynamics shift from being dominated by interannual variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
 \ref{fig:community-forecast-species}B) to being dominated by the steady-state equilibrium captured by the spatial model (time $t \geq 2500$).
-A true STeW forecast would require a method to determine the model weights \emph{a priori}.
+A true forecast from the weighted model would require a method to determine the model weights \emph{a priori}.
 
@@
 \caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated focal species biomass and forecasts from the spatial, temporal and combined (STeW) models at the focal site in the metacommunity model.
+(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted models at the focal site in the metacommunity model.
 (C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal model for focal species biomass (blue).
@@
 As in the community turnover example, we also fit a weighted average of the spatial and temporal model, with the weights changing over time.
-This STeW model initially reflected the temporal model (decrease from $t=500$ to $t=600$), but then rapidly transitioned to reflect the spatial model ($t \geq 700$).
+This weighted model initially reflected the temporal model (decrease from $t=500$ to $t=600$), but then rapidly transitioned to reflect the spatial model ($t \geq 700$).
 The rapid transition in the weighting term, $\omega$, occurred during the period of most rapid change in genotype frequencies (Fig.
 \ref{fig:forecast_supp}).
-The STeW model's predictions look impressively accurate, but, as in the community turnover example, that is because we used the full, simulated time series to fit the weighting term.
+The weighted model's predictions look impressively accurate, but, as in the community turnover example, that is because we used the full, simulated time series to fit the weighting term.
 A true forecast would require an independent method to predict how the model weights shift over time.
@@
 \caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
-(Bottom) Simulated population size and forecasts from the spatial, temporal and combined (STeW) models.
+(Bottom) Simulated population size and forecasts from the spatial, temporal and weighted models.
 }
@@
 We designed our simulation studies to illustrate how the change in model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
-\new{How could these hypotheses be tested with empirical data?} The hypothesis that time-series models will be most effective for near-term forecasts already has empirical support, in the form of recent analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
+How could these hypotheses be tested with empirical data?
+The hypothesis that time-series models will be most effective for near-term forecasts already has empirical support, in the form of recent analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
 The result should not be surprising, since local time-series data capture demographic processes, lagged effects, and responses of current assemblages to small changes in environmental conditions.
@@
 
-\new{Space-for-time modeling approaches for predicting long-term, steady-state outcomes of ecological change have also been tested empirically, primarily via hind-casting.
-Overall, the results are mixed: some tests show skillful prediction of changes in community composition \citep{Blois2013,Illan2014} or species distributions \citep{Norberg2019}, supporting the hypothesis that datasets spanning spatial gradients capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
+Space-for-time modeling approaches for predicting long-term, steady-state outcomes of ecological change have also been tested empirically, primarily via hind-casting.
+Overall, the results are mixed: some tests show reasonable prediction of changes in community composition \citep{Blois2013,Illan2014} or species distributions \citep{Norberg2019}, supporting the hypothesis that datasets spanning spatial gradients capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
 Other attempts to validate predictions from space-for-time models have been discouraging \citep{Worth2014,Illan2014,Davis2014,Brun2016,Veloz2012}, indicating violations of model assumptions or effects of transient dynamics.
-However, predictions from the space-for-time approaches are rarely compared directly to predictions from time-series models (e.g., \citealt{Renwick2018}).
+However, predictions from the space-for-time approaches are rarely compared directly to predictions from time-series models (\citealt{harris_forecasting_2018} but see \citealt{Renwick2018}).
 We need more such comparisons to identify the appropriate modeling approach for different forecast horizons.
-}
 
-\new{The greatest empirical challenge will be testing our hypothesis that a weighted average of spatial and temporal models will make the best forecasts at intermediate time scales.
+The greatest empirical challenge will be testing our hypothesis that a weighted average of spatial and temporal models will make the best forecasts at intermediate time scales.
 There are two problems: finding appropriate data and determining the model weights \emph{a priori}.
 Many data sets have both a longitudinal and spatial dimension, but we could not think of one which also featured a clear ecological response to significant environmental change.
-Surely such datasets exist, and we hope researchers who work with them will test our proposed STeW model.
+Surely such datasets exist, and we hope researchers who work with them will test our proposed weighted model.
 Determining model weights may be more difficult.
@@
 We need new theory or empirical case studies in order to assign these weights \emph{a priori}.
-}
 
@@
 
-\new{While fully process-based models may not be practical, there are more mechanistic alternatives to our extremely phenomenological STeW model for integrating spatial and temporal information}.
+While fully process-based models may not be practical, there are more mechanistic alternatives to our phenomenological, weighted model for integrating spatial and temporal information.
 Spatiotemporal statistical modeling approaches are being developed to study patterns and processes of interest to ecological forecasters, such the spread of an invasive species or population status of a threatened species  \citep{Wikle2003,Williams2017,Schliep2018}.
@@
 
-The spatial-temporal-weighted (STeW) model is a weighted average of predictions from the spatial and temporal models, with the weights changing as a function of time, here expressed as the forecast horizon.
+The weighted model is a weighted average of predictions from the spatial and temporal models, with the weights changing as a function of time, here expressed as the forecast horizon.
 The weights change as a function of the square root of the forecast horizon, to allow rapid shifts in the model weights.
@@
 \end{equation}
-For the community turnover example, the predicted biomass from the STeW model is:
+For the community turnover example, the predicted biomass from the weighted model is:
 \begin{equation}
@@
 \end{equation}
-Again, we suppress the spatial subscript ($x$) here because we are focused on densities at just one location. For the eco-evolutionary example, the predicted population size from the STeW model is:
+Again, we suppress the spatial subscript ($x$) here because we are focused on densities at just one location.
+For the eco-evolutionary example, the predicted population size from the weighted model is:
 \begin{equation}
@@
 Furthermore, the simulations do not include noise; the only reason there is any uncertainty is because the statistical models are slightly mis-specified with respect to the process models.
-Showing uncertainty for the STeW model would be even less meaningful, because it is not a true, out-of-sample forecast (parameters are fit directly to the observations for which we make predictions).
+Showing uncertainty for the weighted model would be even less meaningful, because it is not a true, out-of-sample forecast (parameters are fit directly to the observations for which we make predictions).
 The R code to compute uncertainties for the spatial and temporal forecasts is available on our Github repository (https://github.com/pbadler/space-time-forecast), but is commented out.
@@
 (A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated total biomass and forecasts from the spatial, temporal and combined (STeW) models.
+(B) Simulated total biomass and forecasts from the spatial, temporal and weighted models.
 (C) Simulated changes in biomass of all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).

### 2019-10-15 doc/manuscript.tex
small figure edits

@@
 (C) The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
-Colors show population size (the blue to red gradient depicting high to low population density), which influences the population growth rate through density dependence.
+Colors show population size (the green to brown gradient depicting low to high population density), which influences the population growth rate through density dependence.
 }

### 2019-10-15 genetic-diversity/genetic_diversity_figures.R

@@ -31,7 +31,7 @@ png("figures/spatial&temporal_model.png",height=10,width=4,res=400,units="in")
   # first make colors based on N0 density
   rr <- range(N_t0_baseline)
   svals <- (N_t0_baseline-rr[1])/diff(rr)
-  f <- colorRamp(c("red", "blue"), 0.4)
+  f <- colorRamp(c("green4", "tan2"), 1)
   density_colors <- rgb(f(svals)/255, alpha=0.6)
   # now plot
   plot(temperature_t1_baseline,r_baseline,xlab="Annual temperature",ylab="log Population growth rate",

### 2019-10-15 doc/manuscript.tex
final edits before submission

@@
 At intermediate time-scales, a weighted average of the two approaches shows promise.
-However, making this combined approach operational will require new research to predict the rate at which slow processes begin to influence dynamics.
+However, making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
 
@@
 
-Most models used for ecological forecasting have traditionally been based on either time-series approaches or space-for-time substitutions.
+Models used for ecological forecasting typically rely on either time-series approaches or space-for-time substitutions.
 The time-series approach involves fitting models to long-term datasets to describe the temporal dynamics of a system.
 We then use those dynamic models to make predictions about what will happen in the future.
-This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, or primary production as a function of precipitation \citep{lauenroth_long-term_1992}.
-Whether process-based or data-driven (e.g., \citealt{ward_complexity_2014}), time-series models typically capture ``fast processes'' operating on interannual time-scales, such as birth, death, individual growth, small-scale dispersal events, and short-term responses to environmental conditions (Fig.
+This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, or primary production as a function of annual precipitation \citep{lauenroth_long-term_1992}.
+Whether process-based or data-driven (e.g., \citealt{ward_complexity_2014}), time-series models capture ``fast processes'' operating on interannual time-scales, such as birth, death, individual growth, small-scale dispersal events, and short-term responses to environmental conditions (Fig.
 \ref{fig:fast-slow-concept}).
-However, models built using this approach typically cover a limited spatial extent (but see \citealt{Hefley2017,kleinhesselink_response_2018}), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could influence dynamics at longer time scales \citep{clark_ecological_2001}.
+However, models built using this approach normally cover a limited spatial extent (but see \citealt{Hefley2017,kleinhesselink_response_2018}), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could influence dynamics at longer time scales \citep{clark_ecological_2001}.
 
@@
 To make a forecast, we first predict the future environmental conditions and then determine the associated ecological response, based on the observed spatial relationship.
-This is the approach commonly used to predict population distribution or abundance as a function of climate \citep{elith_species_2009} or primary production as a function of mean precipitation \citep{Sala1988}.
+This is the approach commonly used to predict population distribution or abundance as a function of climate \citep{elith_species_2009} or mean primary production as a function of mean precipitation \citep{Sala1988}.
 Space-for-time models capture the outcome of interactions between fast processes and slower processes operating over long time periods, such as immigration, extinction, and responses to large or prolonged environmental changes (Fig.
@@
 In fact, transient dynamics could prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
-Although both time-series and space-for-time approaches are widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics (\citealt{harris_forecasting_2018}, but see \citealt{Renwick2018}).
+Although both time-series and space-for-time approaches are widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics \citep{harris_forecasting_2018,Renwick2018}.
 
@@
 \begin{equation}
-logit(\omega_t)=\beta^C_0 + \beta^C_1 \sqrt{t}
+logit(\omega_t)=\beta^W_0 + \beta^W_1 \sqrt{t}
 \label{eqn:weights}
@@
 
-We used the \texttt{optim} function to estimate the $\beta^C$s that minimize the sum of squared errors, $(\hat{N}(t+1) - N(t+1))^2$.
+We used the \texttt{optim} function to estimate the $\beta^W$s that minimize the sum of squared errors, $(\hat{N}(t+1) - N(t+1))^2$.
 

### 2019-10-15 community-turnover/comm_turn_figures.R

@@ -90,7 +90,7 @@ png("figures/community_forecast_species.png",height=7.5,width=4.5,res=400,units=
   lines(1:length(my_biomass_spp),combined_forecast_spp,col="purple",lwd=2,lty=2)
   mtext(side=3," B",line=-1.1,adj=0,cex=0.8)
   
-  legend("right",c("Observed","Spatial forecast","Temporal forecast","Combined forecast"),
+  legend("right",c("Observed","Spatial forecast","Temporal forecast","Weighted forecast"),
          col=c("darkgrey","red","blue","purple"),lwd=c(1,2,2,2),lty=c(1,1,1,2),bty="n",cex=0.8)
   
   # bottom panel: weights and species
@@ -131,7 +131,7 @@ png("figures/community_forecast_total.png",height=7.5,width=4.5,res=400,units="i
   lines(1:length(my_biomass),temporal_forecast,col="blue",lwd=2)
   lines(1:length(my_biomass),combined_forecast,col="purple",lwd=2,lty=2)
   mtext(side=3," B",line=-1.1,adj=0,cex=0.8)
-  legend("right",c("Observed","Spatial forecast","Temporal forecast","Combined forecast"),
+  legend("right",c("Observed","Spatial forecast","Temporal forecast","Weighted forecast"),
          col=c("darkgrey","red","blue","purple"),lwd=c(1,2,2,2),lty=c(1,1,1,2),bty="n",cex=0.8)
   
   # weights

### 2019-10-15 genetic-diversity/genetic_diversity_figures.R

@@ -57,7 +57,7 @@ png("figures/forecast.png",height=5,width=5,res=400,units="in")
   lines(1:length(N),spatial_forecast,col="red",lwd=2)
   lines(1:length(N),temporal_forecast,col="blue",lwd=2)
   lines(1:length(N),combined_forecast,col="purple",lwd=2,lty=2)
-  legend("bottomleft",c("Observed","Spatial forecast","Temporal forecast","Combined forecast"),
+  legend("bottomleft",c("Observed","Spatial forecast","Temporal forecast","Weighted forecast"),
          col=c("darkgrey","red","blue","purple"),lwd=c(1,2,2,2),lty=c(1,1,1,2),bty="n")
 
 dev.off()

### 2019-10-16 doc/manuscript.tex
Format changes for EcoLetts

@@
 
-\author[1]{Peter B. Adler}  %\thanks{Corresponding author. Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah Email: peter.adler@usu.edu}}
+\author[1]{Peter B. Adler}  
 \author[2,3,4]{Ethan P. White}
@@
 
+% rest of title page
+\noindent \emph{Statement of authorship:} PBA and EPW designed the study, PBA and MHC built the models, and PBA analyzed the models.
+PBA wrote the first draft of the manuscript, and all authors contributed substantially to revisions.
+
+\noindent \emph{Data accessibility statement:} The manuscript contains no original data.
+All computer code is available at https://github.com/pbadler/space-time-forecast.
+Upon acceptance of the manuscript, code will be archived at Zenodo.
+
+\noindent \emph{Running title:} Forecast horizon and ecological process
+
+\noindent \emph{Keywords:} dispersal, ecological forecasting, eco-evolutionary dynamics, global change, selection
+
+\noindent \emph{Word counts:} 174 words in abstract; 4044 words in main text.
+
+\noindent \emph{References}: 35
+
+\noindent \emph{References}: 6 figures, no tables or text boxes
+
+\noindent \emph{Corresponding author:} Peter Adler, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah, 84322, USA; email: peter.adler@usu.edu; tel: 435-797-1021
+
+\newpage
+
 \section*{Abstract}
@@
 However, making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
-
-\vspace{0.2in}
-
-\textbf{\large{Keywords:}} dispersal, ecological forecasting, eco-evolutionary dynamics, global change, selection
 
@@
 for near-term forecasts time-series approaches may capture the key dynamics and provide accurate predictions.
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.7 \textwidth] {fast-slow-figure.png}
-	\caption{Fast and slow processes operate at different time scales, and are reflected in different kinds of datasets.
-Fast processes, such as births, deaths, and individual growth, operate at all time scales, but are the exclusive drivers of the short-term dynamics captured in most time series datasets.
-Slower processes, such as evolutionary selection on genotype frequencies, ecological selection on species abundances, and colonization and extinction, interact with fast processes to drive dynamics over the long-term.
-The influence of these slow processes is seen in very long time series, or in spatial gradients.
-Understanding dynamics at intermediate time scales requires integrating information from spatial and temporal data sources.
-We propose a model weighting approach; mechanistic spatiotemporal modeling is another alternative.
-The time scales shown here were chosen with vascular plants in mind, but the same concepts would apply for much shorter-lived organisms but at shorter time scales.
-}
-	\label{fig:fast-slow-concept}
-\end{figure}
 
@@
 
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=1 \textwidth] {species_patterns_models.png}
-\caption{(A) Mean biomass by species (colors) across the temperature gradient during the baseline period.
-The focal species, dominant at the site in the center of the gradient (vertical gray line), is shown in dark blue.
-The dashed blue line shows predictions from the spatial model.
-(B) Annual biomass of the focal species at the central site during the baseline period.
-The dashed line shows predictions from the temporal model.
-}
-\label{fig:species-patterns-models}
-\end{figure}
-
 We fit both a temporal and a spatial statistical model to forecast the effect of a temperature increase (Fig.
@@
 
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_forecast_species.png}
-\caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted models at the focal site in the metacommunity model.
-(C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal model for focal species biomass (blue).
-Year 1000 in each panel corresponds to the start of the temperature increase.
-}
-\label{fig:community-forecast-species}
-\end{figure}
-
 The compositional turnover affecting our focal species also influences total biomass, linking community and ecosystem dynamics.
@@
 The full model description is provided in SI Appendix \ref{eco-evo}.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.4 \textwidth]{spatial&temporal_model.png}
-\caption{(A) Reaction norms of the three genotypes.
-(B) The spatial pattern of individual genotypes (colors) and total population abundance (black) at sites arrayed across a gradient of mean annual temperature.
-The dashed line shows predictions from an empirical ``spatial model," a linear regression that describes mean population size as a function of mean temperature.
-(C) The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
-Colors show population size (the green to brown gradient depicting low to high population density), which influences the population growth rate through density dependence.
-}
-\label{fig:evo_evo_spatial_temporal_models}
-\end{figure}
 
@@
 A true forecast would require an independent method to predict how the model weights shift over time.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {forecast.png}
-\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
-(Bottom) Simulated population size and forecasts from the spatial, temporal and weighted models.
-}
-\label{fig:forecast}
-\end{figure}
 
@@
 
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.7 \textwidth] {dispersal_niche_width.png}
-	\caption{The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
-Year 0 in these figures corresponds to the start of the temperature increase.
-}
-	\label{fig:dispersal_niche_width}
-\end{figure}
-
 An alternative to a weighted combination of predictions from time-series and space-for-time models is to rely on fully process-based models.
@@
 \bibliography{references}
+
+\newpage
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.7 \textwidth] {fast-slow-figure.png}
+	\caption{Fast and slow processes operate at different time scales, and are reflected in different kinds of datasets.
+Fast processes, such as births, deaths, and individual growth, operate at all time scales, but are the exclusive drivers of the short-term dynamics captured in most time series datasets.
+Slower processes, such as evolutionary selection on genotype frequencies, ecological selection on species abundances, and colonization and extinction, interact with fast processes to drive dynamics over the long-term.
+The influence of these slow processes is seen in very long time series, or in spatial gradients.
+Understanding dynamics at intermediate time scales requires integrating information from spatial and temporal data sources.
+We propose a model weighting approach; mechanistic spatiotemporal modeling is another alternative.
+The time scales shown here were chosen with vascular plants in mind, but the same concepts would apply for much shorter-lived organisms but at shorter time scales.
+}
+	\label{fig:fast-slow-concept}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=1 \textwidth] {species_patterns_models.png}
+	\caption{(A) Mean biomass by species (colors) across the temperature gradient during the baseline period.
+The focal species, dominant at the site in the center of the gradient (vertical gray line), is shown in dark blue.
+The dashed blue line shows predictions from the spatial model.
+(B) Annual biomass of the focal species at the central site during the baseline period.
+The dashed line shows predictions from the temporal model.
+}
+	\label{fig:species-patterns-models}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.7 \textwidth] {community_forecast_species.png}
+	\caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
+(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted models at the focal site in the metacommunity model.
+(C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal model for focal species biomass (blue).
+Year 1000 in each panel corresponds to the start of the temperature increase.
+}
+	\label{fig:community-forecast-species}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.4 \textwidth]{spatial&temporal_model.png}
+	\caption{(A) Reaction norms of the three genotypes.
+(B) The spatial pattern of individual genotypes (colors) and total population abundance (black) at sites arrayed across a gradient of mean annual temperature.
+The dashed line shows predictions from an empirical ``spatial model," a linear regression that describes mean population size as a function of mean temperature.
+(C) The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
+Colors show population size (the green to brown gradient depicting low to high population density), which influences the population growth rate through density dependence.
+}
+	\label{fig:evo_evo_spatial_temporal_models}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.7 \textwidth] {forecast.png}
+	\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
+(Bottom) Simulated population size and forecasts from the spatial, temporal and weighted models.
+}
+	\label{fig:forecast}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.7 \textwidth] {dispersal_niche_width.png}
+	\caption{The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
+Year 0 in these figures corresponds to the start of the temperature increase.
+}
+	\label{fig:dispersal_niche_width}
+\end{figure}
 

### 2019-12-09 doc/manuscript.tex
started new Intro paragr

@@
 
-\title{\textbf{ Matching the forecast horizon with the relevant ecological processes }}
+\title{\textbf{ Matching the forecast horizon with the relevant spatial and temporal processes and data sources }}
 
@@
 
-\noindent \emph{Data accessibility statement:} The manuscript contains no original data.
-All computer code is available at https://github.com/pbadler/space-time-forecast.
-Upon acceptance of the manuscript, code will be archived at Zenodo.
+\noindent \emph{Acknowledgements:} We thank Heather Lynch and Juan Manuel Morales for comments that improved early drafts of the paper.
+PBA was supported by the National Science Foundation (DEB-1927282) and the Utah Agriculture Experiment Station (grant 1322).
+EPW was supported by the Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative (GBMF4563).
 
@@
 
-\noindent \emph{Word counts:} 174 words in abstract; 4044 words in main text.
-
-\noindent \emph{References}: 35
-
-\noindent \emph{References}: 6 figures, no tables or text boxes
-
 \noindent \emph{Corresponding author:} Peter Adler, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah, 84322, USA; email: peter.adler@usu.edu; tel: 435-797-1021
-
-\newpage
 
@@
 
-Most models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
-Here we consider how the forecast horizon determines whether the most accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
-We use two simulation case studies to show that forecasts for short and long-time scales need to focus on different ecological processes, which are reflected in different kinds of data.
-In the short-term, dynamics reflect initial conditions and fast processes such as birth and death, and the phenomenological time-series approach makes the best predictions.
+Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
+Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
+We use two simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
+In the short-term, dynamics reflect initial conditions and fast processes such as birth and death, and the time-series approach makes the best predictions.
 In the long-term, dynamics reflect the additional influence of slower processes such as evolutionary and ecological selection, colonization and extinction, which the space-for-time approach can effectively capture.
@@
 However, making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
+
+\newpage
 
@@
 
-Models used for ecological forecasting typically rely on either time-series approaches or space-for-time substitutions.
+Statistical models used for ecological forecasting generally rely on either time-series approaches or space-for-time substitutions.
 The time-series approach involves fitting models to long-term datasets to describe the temporal dynamics of a system.
@@
 This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, or primary production as a function of annual precipitation \citep{lauenroth_long-term_1992}.
-Whether process-based or data-driven (e.g., \citealt{ward_complexity_2014}), time-series models capture ``fast processes'' operating on interannual time-scales, such as birth, death, individual growth, small-scale dispersal events, and short-term responses to environmental conditions (Fig.
+When time-series models are fit with typically short ecological data sets, they capture ``fast processes'' operating on interannual time-scales, such as birth, death, individual growth, small-scale dispersal events, and short-term responses to environmental conditions (Fig.
 \ref{fig:fast-slow-concept}).
-However, models built using this approach normally cover a limited spatial extent (but see \citealt{Hefley2017,kleinhesselink_response_2018}), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could influence dynamics at longer time scales \citep{clark_ecological_2001}.
+Statistical models built using this approach normally cover a limited spatial extent (but see \citealt{Hefley2017,kleinhesselink_response_2018} ADD Chevalier), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could influence dynamics at longer time scales \citep{clark_ecological_2001}.
 
@@
 environmental conditions represented in observed data (Fig. \ref{fig:fast-slow-concept}).
-At short forecast horizons (days to years), dynamics will reflect the physiological and demographic responses of 
-the organisms present at a site more than turnover of genotypes or species, 
+At short forecast horizons (days to years), dynamics will reflect physiological and demographic responses and interactions
+among the organisms present at a site more than temporal turnover of genotypes or species, 
 environmental conditions are likely to stay within the range of historical variation, 
@@
 historically observed are likely to not only occur but to persist long enough to drive significant turnover of genotypes and species 
-along with changes in the flux of energy and nutrients.  At these long scales, the current state of the system may be 
+through colonization and extinction as well as changes in the flux of energy and nutrients.  
+At these long forecast horizons, the state of the system at the time the forecast is issued may be 
 little help in predicting the future state. For the century-scale forecasts often featured in biodiversity and 
@@
 
-Here we use simulation models to 1) demonstrate that the best model-building approaches for ecological forecasting 
+Why not simply use process-based models to avoid the difficulties posed by these phenomenological models?
+If we could accurately characterize all of the processes governing a system, then a model based on that understanding should make accurate predictions at all forecast horizons.
+Process-based models should also be more robust for making predictions outside of historically observed conditions and even beyond the conditions observed across spatial gradients, which will be especially important for making predictions in a future with increasingly novel combinations of environment and species interactions \citep{Williams2007}.
+Unfortunately, in most cases this approach is not currently feasible because we lack a detailed knowledge of all the complex and interacting processes influencing the dynamics of real ecological systems.
+Even if the general form of the models was known, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well studied systems.
+Furthermore, the high complexity and corresponding parameter uncertainty of such models can increase predictive errors; simpler  time-series models may actually perform better \citep{ward_complexity_2014a}, though spatial replication can reduce the cost of complexity (ADD Chevalier).
+-very difficult to parameterize 
+-complexity is costly
+-spatiotemporal stats models more realistic
+-but seldom used for forecasting
+-for now, models that rely in some way on phenomenological, statistical relationships will play an important role
+
+Here we use two simulated case studies to 1) demonstrate that the best model-building approaches for ecological forecasting 
 depend on the time horizon of the forecast, and 2) explore how time-series and space-for-time approaches might be 
-combined via weighted averaging to make better forecasts at intermediate time scales. We conduct two simulation case studies, one 
-focused on how interspecific interactions affect the population dynamics of a focal species, and the second focused on
+combined via weighted averaging to make better forecasts at intermediate time scales. The first case study
+focuses on how interspecific interactions affect the population dynamics of a focal species, and the second focuses on
 an eco-evolutionary scenario. Our analyses show that: 
@@
 
-In two case studies, we simulated the effects of an increase in temperature on simple systems with known dynamics.
-The ``truth'' is represented by a model that is mechanistic for at least one important process, but we treat the model as unknown when
-analyzing the data and we assumed that perfectly recovering this model would not be possible in practice. We began each simulation
-under stationary temperature, allowing the system to equilibrate; we call this the baseline phase. 
+In each case study, we simulated the effects of an increase in temperature on simple systems with known dynamics.
+The truth is represented by a simulation model that is mechanistic for at least one important process, 
+but we treat the data-generating model as unknown when analyzing the data and we assumed that perfectly recovering the mechanisms
+it contains would not be possible in practice. We began each simulation
+under a stationary distribution of annual temperatures, allowing the system to equilibrate; we call this the baseline phase. 
 We then increased temperature progressively over a period of time, followed by a second period of stationary, now elevated, temperature. 
-The objective was to forecast the response of the system to the temperature increase based on data gathered during the baseline period.
+The objective was to forecast the response of the system to the temperature increase based on spatial and/or temporal data 
+``sampled" from the simulation during the baseline period.
 
-We made forecasts based on two phenomenological models, each representing processes operating at different time scales.
-One model represents the time-series or ``temporal approach.'' We correlated interannual variation in an ecological response with interannual variation in the environment at just one site.
-The other model represents the space-for-time substitution approach, which we call the ``spatial approach'' for brevity.
-We correlated the mean environment with the mean of an ecological state or rate across many sites.
-We compared forecasts from both models to the simulated dynamics to determine how well the two approaches performed at different forecast horizons.
+We made forecasts based on two phenomenological statistical models, each representing processes operating at different time scales.
+One statistical model represents the time-series or ``temporal approach.'' We correlated interannual variation in an ecological response with interannual variation in temperature at just one site.
+The other statistical model relies on a space-for-time substitution, which we call the ``spatial approach'' for brevity.
+We correlated the mean temperature with the mean of an ecological state or rate across many sites.
+We compared forecasts from both statistical models to the simulated dynamics to determine how well the two approaches performed at different forecast horizons.
 We also assessed the potential for combining the information available in temporal and spatial patterns by using a weighted average of the forecasts from the temporal and spatial approaches optimized to best match the (simulated) observations.
 We then studied how the optimal model weights changed over time.
-We expected the temporal model to best predict short-term dynamics, the spatial model to best predict long-term dynamics, while the weighted model would show potential to provide the best forecasts at transitional, intermediate time scales.
-The three statistical models are described in Supporting Information (Appendix \ref{models}) and all code for both case studies is available at Github (https://github.com/pbadler/space-time-forecast).
+We expected the temporal approach to best predict short-term dynamics, the spatial approach to best predict long-term dynamics, while the weighted model would show potential to provide the best forecasts at transitional, intermediate time scales.
+The three statistical models are described in Supporting Information (Appendix \ref{models}) and all code for both case studies is available for reviewers as Supporting Information (and will be archived at Zenodo upon acceptance).
 
@@
 But at what forecast horizon do altered species interactions become impossible to ignore?
-We explored this question using a metacommunity model developed by \cite{alexander_lags_2018} to study how community responses to increasing temperature depend on the interplay between within-site demography and competitive interactions and the movement of species across sites.
+We explored this question using a metacommunity simulation model developed by \cite{alexander_lags_2018} to study how community responses to increasing temperature depend on the interplay between within-site demography and competitive interactions and the movement of species across sites.
 The model features Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
@@
 Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
-We provide a more detailed description of the model in SI Appendix \ref{metacomm}.
+We provide a more detailed description of the simulation model in SI Appendix \ref{metacomm}.
 
@@
 Individual species, including our focal species, showed classic, unimodal ``Whittaker" patterns of abundances across the gradient (Fig. \ref{fig:species-patterns-models}A).
-These spatial patterns are the basis for our ``spatial model" of the temperature-biomass relationship for our focal species (Fig. \ref{fig:species-patterns-models}A). 
+These spatial patterns are the basis for our spatial statistical model of the temperature-biomass relationship for our focal species (Fig. \ref{fig:species-patterns-models}A). 
 In contrast to the strong spatial patterns, population and community responses to interannual variation in temperature within sites were weak.
@@
 \ref{fig:species-patterns-models}B).
-Our ``temporal model" estimates this weak, linear temperature effect, along with the strong lag effect of biomass in the previous year.
+Our temporal statistical model estimates this weak, linear temperature effect, along with the strong lag effect of biomass in the previous year.
 
-We fit both a temporal and a spatial statistical model to forecast the effect of a temperature increase (Fig.
+We used both the temporal and spatial statistical models to forecast the effect of a temperature increase (Fig.
 \ref{fig:community-forecast-species}A) on the focal species' biomass at one location in the center of the temperature gradient.
-The predictions from the spatial and temporal models contrasted markedly, with the temporal 
-model predicting a large increase in biomass and the spatial model predicting a decrease. Initially, the simulated abundances followed the increase predicted by the 
+The predictions from these two models contrasted markedly, with the temporal 
+statistical model predicting a large increase in biomass and the spatial statistical model predicting a decrease. Initially, the simulated abundances followed the increase predicted by the 
 temporal model, but as faster-growing species colonized and increased in abundance at the focal site, the biomass of the focal species decreased, eventually falling below its baseline level (Fig.
@@
 
-To combine the temporal and spatial model into a single forecast, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, this weighted forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
+To combine information from the temporal and spatial statistical models into a single prediction, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
+At any time point, $t$, this weighted forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal statistical model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial statistical model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
 The weighted model accurately predicts the simulated dynamics across the full forecast horizon (Fig.
@@
 However, the reason the weighted models works so well is that the weights were determined by fitting directly to the data.
-Unlike our spatial and temporal model forecasts, we did not generate out-of-sample predictions from the weighted model; it merely provides a convenient way to quantify how rapidly dynamics shift from being dominated by interannual variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
+Unlike the forecasts from the spatial and temporal statistical models, we did not generate out-of-sample predictions from the weighted model; it merely provides a convenient way to quantify how rapidly dynamics shift from being dominated by interannual variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
 \ref{fig:community-forecast-species}B) to being dominated by the steady-state equilibrium captured by the spatial model (time $t \geq 2500$).
@@
 The compositional turnover affecting our focal species also influences total biomass, linking community and ecosystem dynamics.
-We repeated our focal species analysis for total community biomass, and the results were similar: the temporal model initially made the best forecasts immediately following the onset of the temperature increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial model (SI Figs.
+We repeated our focal species analysis for total community biomass, and the results were similar: the temporal statistical model initially made the best forecasts immediately following the onset of the temperature increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial statistical model (SI Figs.
 \ref{fig:community-models-total} and \ref{fig:community-forecasts-total}).
@@
 Although shifts in genotype frequencies at the population level are analogous to changes in species composition at the community level, the mechanisms are distinct: heterozygosity and genetic recombination have no analogue at the community level.
-We demonstrate how these processes influence short and long-term forecasts with a standard eco-evolutionary model for a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
+We demonstrate how these processes influence short and long-term forecasts with a standard eco-evolutionary simulation model for a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
 \ref{fig:evo_evo_spatial_temporal_models}A).
 Our model describes how the local density of each genotype changes between years, which depends on temperature and genotype densities in the previous year.
-Transient temporal dynamics are computed directly from the model; these dynamics define our temporal forecast.
-To create ``spatial data'', we simulated the equilibrium density of each genotype under different mean temperatures.
-The pattern of equilibrium densities across a gradient in mean annual temperature defines our spatial forecast: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
+Transient temporal dynamics are computed directly from the model; these dynamics are the basis for the temporal statistical model.
+To create a spatial gradient, we simulated the equilibrium density of each genotype in a series of local populations experiencing different mean temperatures.
+The pattern of equilibrium densities across the mean annual temperature gradient is the basis for our spatial statistical model: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
 \ref{fig:evo_evo_spatial_temporal_models}B).
-The full model description is provided in SI Appendix \ref{eco-evo}.
+The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}.
 
@@
 During the baseline period, the population is dominated by the cold-adapted genotype.
-We used the simulated data from this baseline period to fit an empirical model that assumes no knowledge of the underlying eco-evolutionary process.
-This empirical temporal model (Appendix \ref{models}) predicts population growth rate as a function of annual temperature and population size (Fig.
-\ref{fig:evo_evo_spatial_temporal_models}C).
+We used the simulated data from this baseline period to fit a temporal statistical model (Appendix \ref{models}) that predicts population growth rate as a function of annual temperature and population size (Fig.
+\ref{fig:evo_evo_spatial_temporal_models}C), assuming no knowledge of the underlying eco-evolutionary dynamics.
 We then imposed a period of warming, followed by a final period of higher stationary temperature (Fig.
@@
 \ref{fig:forecast} bottom).
-The temporal model (solid blue line in Fig.
-\ref{fig:forecast}) accurately predicted the impact of the initial warming trend, but eventually became too pessimistic, while the spatial model (solid red line in Fig.
+The temporal statistical model (solid blue line in Fig.
+\ref{fig:forecast}) accurately predicted the impact of the initial warming trend, but eventually became too pessimistic, while the spatial statistical model (solid red line in Fig.
 \ref{fig:forecast}) did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
@@
 
-As in the community turnover example, we also fit a weighted average of the spatial and temporal model, with the weights changing over time.
+As in the community turnover example, we also fit a weighted average of predictions from the spatial and temporal statistical models, with the weights changing over time.
 This weighted model initially reflected the temporal model (decrease from $t=500$ to $t=600$), but then rapidly transitioned to reflect the spatial model ($t \geq 700$).
@@
 Which approach provides the most accurate forecasts depends on the forecast-horizon.
-In our simulations, time-series approaches performed best for short-term forecasts, whereas models based on spatial data made more accurate long-term forecasts.
-In addition, our simulations demonstrate extended transitional periods during which neither the time-series or the spatial approach is effective on its own.
+In our simulations, time-series approaches performed best for short forecast horizons, whereas models based on spatial data made more accurate forecasts at long horizons.
+In addition, our simulations demonstrate extended transitional periods during which neither the time-series or the spatial approach was effective on its own.
 The challenge is determining what is ``short-term,'' what is ``long-term,'' and how to handle the many forecasts we need in ecology which fall in between.
@@
 
-We designed our simulation studies to illustrate how the change in model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
+We designed our simulation studies to illustrate how the change in statistical model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
 How could these hypotheses be tested with empirical data?
@@
 
-The greatest empirical challenge will be testing our hypothesis that a weighted average of spatial and temporal models will make the best forecasts at intermediate time scales.
+The greatest empirical challenge will be testing our hypothesis that a weighted average of spatial and temporal statistical models will make the best forecasts at intermediate time scales.
 There are two problems: finding appropriate data and determining the model weights \emph{a priori}.
@@
 An alternative to a weighted combination of predictions from time-series and space-for-time models is to rely on fully process-based models. 
-If we could accurately characterize all of the processes governing a system, then a model based on that understanding should make accurate predictions at all time-horizons.
-For example, rather than fitting phenomenological models to our simulated time-series, we could have fit the exact process-based models that we used to generate those time-series.
-Assuming reasonable estimates of the parameter, those models would have accurately predicted the dynamics at all forecast horizons.
-Process-based models should also be more robust for making predictions outside of historically observed conditions and even beyond the conditions observed across spatial gradients, which will be especially important for making predictions in a future with increasingly novel combinations of environment and species interactions \citep{Williams2007}.
-Unfortunately, in most cases this approach is not currently feasible because we lack a detailed knowledge of all the complex and interacting processes influencing the dynamics of real ecological systems.
-Even if the general form of the models was known, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well studied systems.
 % mention difficulty of dealing with error propogation?
@@
 Our results have important implications for the emerging field of ecological forecasting.
-First, they suggest that evaluating both near-term and long-term forecasts will be essential as research on forecasting methods accelerates.
-Second, while single approaches may perform reasonably well at either short or long forecast horizons, skillful predictions at intermediate time horizons may require a combination of information from spatial and temporal patterns.
+First, they suggest that evaluating model performance at both short and long forecsat horizons will be essential as research on forecasting methods accelerates.
+Second, while single approaches may perform reasonably well for either short or long horizons, skillful predictions at intermediate forecast horizons may require a combination of information from spatial and temporal statistical models.
 Intermediate time horizons pose challenges in other forecasting contexts as well.
@@
 Forecasting these intermediate scales has been challenging in meteorology and will likely be challenging in ecology as well.
-While the recent emphasis on near-term iterative forecasting \citep{dietze_iterative_2018} is the logical and tractable starting point, we also need to build understanding and capacity for forecasting ecological dynamics across all temporal scales of interest.
+While the recent emphasis on near-term iterative forecasting \citep{dietze_iterative_2018} is the logical and tractable starting point, we also need to build understanding and capacity for forecasting ecological dynamics across all forecast horizons of interest.
 
-\subsubsection*{Acknowledgements}
-
-We thank Heather Lynch and Juan Manuel Morales for comments that improved early drafts of the paper.
-PBA was supported by the Utah Agriculture Experiment Station.
-EPW was supported by the Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative through Grant GBMF4563.
+\subsubsection*{Data accessibility statement:} The manuscript contains no original data.
+All computer code is uploaded for review in a zip file.
+Upon acceptance of the manuscript, code will be archived at Zenodo.
 
@@
 The focal species, dominant at the site in the center of the gradient (vertical gray line), is shown in dark blue.
-The dashed blue line shows predictions from the spatial model.
+The dashed blue line shows predictions from the spatial statistical model.
 (B) Annual biomass of the focal species at the central site during the baseline period.
-The dashed line shows predictions from the temporal model.
+The dashed line shows predictions from the temporal statistical model.
 }
@@
 	\caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted models at the focal site in the metacommunity model.
-(C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal model for focal species biomass (blue).
+(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
+(C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).
 Year 1000 in each panel corresponds to the start of the temperature increase.
@@
 (B) The spatial pattern of individual genotypes (colors) and total population abundance (black) at sites arrayed across a gradient of mean annual temperature.
-The dashed line shows predictions from an empirical ``spatial model," a linear regression that describes mean population size as a function of mean temperature.
+The dashed black line (almost entirely hidden by the slid black line) shows predictions from an empirical, spatial statistical model, a linear regression that describes mean population size as a function of mean temperature.
 (C) The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
@@
 	\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
-(Bottom) Simulated population size and forecasts from the spatial, temporal and weighted models.
+(Bottom) Simulated population size and forecasts from the spatial, temporal and weighted statistical models.
 }

### 2019-12-10 doc/manuscript.tex
Moved paragr to Intro

@@
 
-Why not simply use process-based models to avoid the difficulties posed by these phenomenological models?
+Why not simply use process-based models to avoid the difficulties posed by phenomenological time-series and space-for-time modeling approaches?
 If we could accurately characterize all of the processes governing a system, then a model based on that understanding should make accurate predictions at all forecast horizons.
-Process-based models should also be more robust for making predictions outside of historically observed conditions and even beyond the conditions observed across spatial gradients, which will be especially important for making predictions in a future with increasingly novel combinations of environment and species interactions \citep{Williams2007}.
+Process-based models should also be more robust for making predictions outside of historically observed conditions and even beyond the conditions observed across spatial gradients, which will be especially important in a future with increasingly novel combinations of environment and species interactions \citep{Williams2007}.
 Unfortunately, in most cases this approach is not currently feasible because we lack a detailed knowledge of all the complex and interacting processes influencing the dynamics of real ecological systems.
-Even if the general form of the models was known, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well studied systems.
+Even if the general form of the models were known, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well-studied systems.
 Furthermore, the high complexity and corresponding parameter uncertainty of such models can increase predictive errors; simpler  time-series models may actually perform better \citep{ward_complexity_2014a}, though spatial replication can reduce the cost of complexity (ADD Chevalier).
--very difficult to parameterize 
--complexity is costly
--spatiotemporal stats models more realistic
--but seldom used for forecasting
--for now, models that rely in some way on phenomenological, statistical relationships will play an important role
+As a result, models used for ecological forecasting will include at least some phenomenological components.
+But that does not mean that phenomenological forecast models cannot benefit from process-based understanding.
+Even if process-level understanding does not enable a fully mechanistic model, it can improve the specification of phenomenological models.
+Our hypothesis is that different processes may be relevant for different forecast horizons, and that we can act on this knowledge by fitting models to different kinds of datasets.
 
@@
 
-An alternative to a weighted combination of predictions from time-series and space-for-time models is to rely on fully process-based models. 
-% mention difficulty of dealing with error propogation?
-As a result, models used for ecological forecasting will include at least some phenomenological components.
-But that does not mean that phenomenological forecast models do not benefit from process-based understanding.
-The message from our simulations is that different processes should be considered for different forecast time-scales, and this can be done by fitting models to different kinds of datasets.
-Even when process-level understanding does not enable a fully mechanistic model, it can improve the specification of phenomenological models.
-
-While fully process-based models may not be practical, there are more mechanistic alternatives to our phenomenological, weighted model for integrating spatial and temporal information.
+Given the challenges of determining model weights \emph{a priori}, we should also pursue alternatives for intermediate forecast horizons.
+In the Introduction, we argued that fully process-based models are not feasible.
+However, a new class of statistical models offers a compromise between mechanistic detail and phenomenological feasibility.
 Spatiotemporal statistical modeling approaches are being developed to study patterns and processes of interest to ecological forecasters, such the spread of an invasive species or population status of a threatened species  \citep{Wikle2003,Williams2017,Schliep2018}.
@@
 In addition, new methods for spatiotemporal forecasting are being developed that address existing computational challenges \citep{McDermott2017}, and access to high performance computing resources is increasingly common.
-Given these developments, future ecological forecasting efforts should explore spatiotemporal approaches and assess whether they improve predictions at intermediate time scales relative to traditional spatial or temporal models.
+Given these developments, future ecological forecasting efforts should explore spatiotemporal approaches and assess whether they improve predictions at intermediate time scales relative to traditional time-series or space-for-time approaches.
 

### 2019-12-10 doc/manuscript.tex
Formatting

@@
 
-\title{\textbf{ Matching the forecast horizon with the relevant spatial and temporal processes and data sources }}
+\title{\textbf{Matching the forecast horizon with the relevant spatial and temporal processes and data sources}}
 
@@
 
-% \date{Last compile: \today}
+\date{}
 
@@
 % rest of title page
+
+\vspace{-1cm}
+
 \noindent \emph{Statement of authorship:} PBA and EPW designed the study, PBA and MHC built the models, and PBA analyzed the models.
@@
 
-\noindent \emph{Running title:} Forecast horizon and ecological process
+%\noindent \emph{Running title:} Forecast horizon and ecological process
+
+\vspace{0.25cm}
+
+\noindent \textbf{Abstract:}   Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
+Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
+We use two simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
+In the short-term, dynamics reflect initial conditions and fast processes such as birth and death, and the time-series approach makes the best predictions.
+In the long-term, dynamics reflect the additional influence of slower processes such as evolutionary and ecological selection, colonization and extinction, which the space-for-time approach can effectively capture.
+At intermediate time-scales, a weighted average of the two approaches shows promise.
+However, making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
+
+\vspace{0.25cm}
 
@@
 
-\noindent \emph{Corresponding author:} Peter Adler, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah, 84322, USA; email: peter.adler@usu.edu; tel: 435-797-1021
+\newpage
 
 \section*{Abstract} 
-
 Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
@@
 
-\newpage
+\vspace{0.25cm}
+
+\noindent \emph{Keywords:} dispersal, ecological forecasting, eco-evolutionary dynamics, global change, selection
 
@@
 Even if the general form of the models were known, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well-studied systems.
-Furthermore, the high complexity and corresponding parameter uncertainty of such models can increase predictive errors; simpler  time-series models may actually perform better \citep{ward_complexity_2014a}, though spatial replication can reduce the cost of complexity (ADD Chevalier).
+Furthermore, the high complexity and corresponding parameter uncertainty of such models can increase predictive errors; simpler  time-series models may actually perform better \citep{ward_complexity_2014}, though spatial replication can reduce the cost of complexity \citep{Chevalier}.
 As a result, models used for ecological forecasting will include at least some phenomenological components.

### 2019-12-11 doc/manuscript.tex
linked to code on google drive

@@
 We expected the temporal approach to best predict short-term dynamics, the spatial approach to best predict long-term dynamics, while the weighted model would show potential to provide the best forecasts at transitional, intermediate time scales.
-The three statistical models are described in Supporting Information (Appendix \ref{models}) and all code for both case studies is available for reviewers as Supporting Information (and will be archived at Zenodo upon acceptance).
+The three statistical models are described in Supporting Information (Appendix \ref{models}).
+Computer code for both case studies is available for reviewers online (https://drive.google.com/open?id=1aju04qtQvJmZG1mcpN-6hYBZILBhDfl2) and will be archived at Zenodo upon acceptance.
 
@@
 \subsubsection*{Data accessibility statement:} The manuscript contains no original data.
-All computer code is uploaded for review in a zip file.
+All computer code is available for reviewers in a zip archive that can be downloaded from Google Drive: https://drive.google.com/open?id=1aju04qtQvJmZG1mcpN-6hYBZILBhDfl2.
 Upon acceptance of the manuscript, code will be archived at Zenodo.

### 2019-12-13 doc/manuscript.tex
final edits

@@
 At short forecast horizons (days to years), dynamics will reflect physiological and demographic responses and interactions
-among the organisms present at a site more than temporal turnover of genotypes or species, 
-environmental conditions are likely to stay within the range of historical variation, 
+among the organisms present at a site more than temporal turnover of genotypes or species; 
+environmental conditions are likely to stay within the range of historical variation; 
 and the current state of the system is likely to capture the influence of unmeasured processes. As a result, 
@@
 In each case study, we simulated the effects of an increase in temperature on simple systems with known dynamics.
-The truth is represented by a simulation model that is mechanistic for at least one important process, 
-but we treat the data-generating model as unknown when analyzing the data and we assumed that perfectly recovering the mechanisms
+The truth was represented by a simulation model that was mechanistic for at least one important process, 
+but we treated the data-generating model as unknown when analyzing the data and we assumed that perfectly recovering the mechanisms
 it contains would not be possible in practice. We began each simulation
@@
 \ref{fig:evo_evo_spatial_temporal_models}A).
+
 Our model describes how the local density of each genotype changes between years, which depends on temperature and genotype densities in the previous year.

### 2019-12-13 doc/manuscript.tex
reformatted references

@@
 \ref{fig:fast-slow-concept}).
-Statistical models built using this approach normally cover a limited spatial extent (but see \citealt{Hefley2017,kleinhesselink_response_2018} ADD Chevalier), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could influence dynamics at longer time scales \citep{clark_ecological_2001}.
+Statistical models built using this approach normally cover a limited spatial extent (but see \citealt{Hefley2017,kleinhesselink_response_2018,Chevalier}), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could influence dynamics at longer time scales \citep{clark_ecological_2001}.
 
@@
 \renewcommand{\refname}{Literature cited}
-\bibliographystyle{Ecology}
+\bibliographystyle{ecography}
 \bibliography{references}

### 2020-03-10 doc/manuscript.tex
Ecography version
finished rewriting eco evo model descript.

@@
 \noindent \textbf{Abstract:}   Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
-Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
+\new{However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change.} Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
 We use two simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
@@
 Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
-Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
+\new{However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change.} Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
 We use two simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
@@
 In fact, transient dynamics could prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
+
 Although both time-series and space-for-time approaches are widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics \citep{harris_forecasting_2018,Renwick2018}.
+\new{Such a discussion is overdue given that these two approaches are likely to make very different predictions about ecological responses to future environmental change.
+For example, average forage production in the shortgrass steppe increases rapidly with increasing mean annual precipitation, but production is much less sensitive to interannual variation in precipitation at any one location \citep{lauenroth_long-term_1992}.
+The space-for-time approach would predict large changes in production in response to a future change in precipitation, while the time-series approach would predict smaller changes.
+Bird species' abundances in the United Kindgom also show different responses to spatial vs.
+temporal variation in weather covariates.
+Not only are abundances typically more sensitive to spatial variation in the covariates, but in many cases the temporal response takes the opposite sign \citep{Oedekoven2017}.
+Whenever the sign or magnitude of relationships based on time-series and space-for-time approaches differ, so will the resulting forecasts.}
 
-Whether historical dynamics, contemporary spatial patterns, or some combination of the two will serve as the best source of 
+Whether time-series models, space-for-time approaches, or some combination of the two will serve as the best source of 
 information for forecasting may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. 
@@
 
-\subsection*{Community turnover example}
+\section*{Community turnover example}
 
@@
 We provide a more detailed description of the simulation model in SI Appendix \ref{metacomm}.
+\new{Parameter values for the simulations described here in the main text are shown in Table \ref{tab:comm-turn-pars}.}
 
@@
 \ref{fig:community-forecast-species}B).
+\new{The qualitative results were the same when we repeated the simulation with a continuous, nonstationary temperature increase (SI Fig.
+\ref{fig:community-forecast-species-nonstationary}).}
 
@@
 The compositional turnover affecting our focal species also influences total biomass, linking community and ecosystem dynamics.
-We repeated our focal species analysis for total community biomass, and the results were similar: the temporal statistical model initially made the best forecasts immediately following the onset of the temperature increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial statistical model (SI Figs.
+We repeated our focal species analysis for total community biomass, and the results were similar: the temporal statistical model initially made the best forecasts immediately following the onset of the temperature increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial statistical model (Figs.
 \ref{fig:community-models-total} and \ref{fig:community-forecasts-total}).
@@
 
-\subsection*{Discussion}
+\section*{Discussion}
 
@@
 \cite{alexander_lags_2018} developed a meta-community model to represent dynamics of local communities arrayed along a one-dimensional elevation gradient, as influenced by three main processes: temperature-dependent growth, competition, and dispersal.
-Here we adapt their notation to be consistent our own.
+Here we adapt their notation to be consistent with our own.
 
@@
 
+\new{To assign species-specific parameter values, the number of species in the metacommunity is specified.
+Next, each species is assigned an optimal temperature within a specified temperature range by drawing from a uniform distribution.
+Sensitivity to interspecific competition is then determined as a decreasing function of optimal temperature.
+Calculations are performed in the script \texttt{SpeciesPoolGen.R}.
+}
+
 \section{Description of the eco-evolutionary annual plant model}\label{eco-evo}
@@
 \begin{align}\begin{split}
-N_{1,t+1} &= s_1 [1-g_1(K(t))]N_{1,t}+\frac{\lambda_1g_1(K(t))N_{1,t}}{1+ \alpha_{11}g_1(K(t))N_{1,t} + \alpha_{12}g_2(K(t))N_{2,t}}\\
-N_{2,t+1} &= s_2 [1-g_2(K(t))]N_{2,t}+\frac{\lambda_2g_2(K(t))N_{2,t}}{1+ \alpha_{21}g_1(K(t))N_{1,t} + \alpha_{22}g_2(K(t))N_{2,t}}
+N_{1,t+1} &= s_1 [1-g_1]N_{1,t}+\frac{\lambda_1(K(t))g_1N_{1,t}}{1+ \alpha_{11}g_1N_{1,t} + \alpha_{12}g_2N_{2,t}}\\
+N_{2,t+1} &= s_2 [1-g_2]N_{2,t}+\frac{\lambda_2(K(t))g_2N_{2,t}}{1+ \alpha_{21}g_1N_{1,t} + \alpha_{22}g_2N_{2,t}}
 \end{split}\end{align}
-where $g_{i}(K(t))$ is the probability of germination, $K(t)$ is the temperature at time $t$, $s_i$ is the seed survival probability for species $i$, and $\lambda_i$ is the seed production rate per plant.
+where $g_{i}$ is the probability of germination for species $i$, $s_i$ is the seed survival probability, $\lambda_{i}(K(t))$ is the seed production rate per plant, and $K(t)$ is the temperature at time $t$.
 Below we refer to the $\alpha_{ij}$ as intra- and inter-genotype competition coefficients.
@@
 The genotypes are denoted by $AA$, $Aa$, and $aa$.
-The number of each genotypes at time $t$ is $N_{AA}(t)$, $N_{Aa}(t)$, and $N_{aa}(t)$.
-The germination rates for each genotype are $g_{AA}(K(t))$, $g_{Aa}(K(t))$, and $g_{aa}(K(t))$.
-The seed survival probability and seed production rate for genotype $AA$ are $s_{AA}$ and $\lambda_{AA}$, respectively.
+The number of each genotype at time $t$ is $N_{AA}(t)$, $N_{Aa}(t)$, and $N_{aa}(t)$.
+The germination rates for each genotype are $g_{AA}$, $g_{Aa}$, and $g_{aa}$.
+The seed survival probability and seed production rate for genotype $AA$ are $s_{AA}$ and $\lambda_{AA}(K(t))$, respectively.
 The analogous parameters for the other genotypes are similarly denoted.
@@
 \begin{equation}
-T = \lambda_{AA}N_{AA}(t)g_{AA}(K(t))+ \lambda_{Aa}N_{Aa}(t)g_{Aa}(K(t))+\lambda_{aa}N_{aa}(t)g_{aa}(K(t)).
+T = \lambda_{AA}(K(t))N_{AA}(t)g_{AA} + \lambda_{Aa}(K(t))N_{Aa}(t)g_{Aa} +\lambda_{aa}(K(t))N_{aa}(t)g_{aa}.
 \end{equation}
@@
 \begin{align}\begin{split}
-\phi_{A} &= \frac{\lambda_{AA}N_{AA}(t)g_{AA}(K(t))+ \frac{1}{2}\lambda_{Aa}N_{Aa}(t)g_{Aa}(K(t))}{T} \hspace{10pt} \text{and} \hspace{10pt} \phi_a = 1-\phi_{A}.
+\phi_{A} &= \frac{\lambda_{AA}(K(t))N_{AA}(t)g_{AA}+ \frac{1}{2}\lambda_{Aa}(K(t))N_{Aa}(t)g_{Aa}}{T} \hspace{10pt} \text{and} \hspace{10pt} \phi_a = 1-\phi_{A}.
 \end{split}\end{align}
@@
 \begin{align}\begin{split}
-N_{AA}(t+1) &= s_{AA}[1-g_{AA}(K(t))]N_{AA}(t) + \phi_A^2T\\
-N_{Aa}(t+1) &= s_{Aa}[1-g_{Aa}(K(t))]N_{Aa}(t) + \phi_A\phi_aT\\
-N_{aa}(t+1) &= s_{aa}[1-g_{aa}(K(t))]N_{aa}(t) + \phi_a^2T
+N_{AA}(t+1) &= s_{AA}[1-g_{AA}]N_{AA}(t) + \phi_A^2T\\
+N_{Aa}(t+1) &= s_{Aa}[1-g_{Aa}]N_{Aa}(t) + \phi_A\phi_aT\\
+N_{aa}(t+1) &= s_{aa}[1-g_{aa}]N_{aa}(t) + \phi_a^2T
 \end{split}\end{align}
@@
 \begin{align}\begin{split}
-T &=  \frac{\lambda_{AA}N_{AA}(t)g_{AA}(K(t))}{1+ \alpha_{AA,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{AA,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{AA,aa}g_{aa}(K(t))N_{aa}(t)} \\ 
-&+ \frac{\lambda_{Aa}N_{Aa}(t)g_{Aa}(K(t))}{1+ \alpha_{Aa,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{Aa,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{Aa,aa}g_{aa}(K(t))N_{aa}(t)}\\
-&+\frac{\lambda_{aa}N_{aa}(t)g_{aa}(K(t))}{1+ \alpha_{aa,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{aa,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{aa,aa}g_{aa}(K(t))N_{aa}(t)}.
+T &=  \frac{\lambda_{AA}(K(t))N_{AA}(t)g_{AA}}{1+ \alpha_{AA,AA}g_{AA}N_{AA}(t) + \alpha_{AA,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{AA,aa}g_{aa}N_{aa}(t)} \\ 
+&+ \frac{\lambda_{Aa}(K(t))N_{Aa}(t)g_{Aa}}{1+ \alpha_{Aa,AA}g_{AA}N_{AA}(t) + \alpha_{Aa,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{Aa,aa}g_{aa}N_{aa}(t)}\\
+&+\frac{\lambda_{aa}(K(t))N_{aa}(t)g_{aa}}{1+ \alpha_{aa,AA}g_{AA}N_{AA}(t) + \alpha_{aa,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{aa,aa}g_{aa}N_{aa}(t)}.
 \label{eqn:defineT}
@@
 \begin{align}\begin{split}
-\phi_A &= \frac{1}{T}\frac{\lambda_{AA}N_{AA}(t)g_{AA}(K(t))}{1+ \alpha_{AA,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{AA,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{AA,aa}g_{aa}(K(t))N_{aa}(t)} \\
-&+ \frac{1}{2T}\frac{\lambda_{Aa}N_{Aa}(t)g_{Aa}(K(t))}{1+ \alpha_{Aa,AA}g_{AA}(K(t))N_{AA}(t) + \alpha_{Aa,Aa}g_{Aa}(K(t))N_{Aa}(t)+ \alpha_{Aa,aa}g_{aa}(K(t))N_{aa}(t)}\\
+\phi_A &= \frac{1}{T}\frac{\lambda_{AA}(K(t))N_{AA}(t)g_{AA}}{1+ \alpha_{AA,AA}g_{AA}N_{AA}(t) + \alpha_{AA,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{AA,aa}g_{aa}N_{aa}(t)} \\
+&+ \frac{1}{2T}\frac{\lambda_{Aa}(K(t))N_{Aa}(t)g_{Aa}}{1+ \alpha_{Aa,AA}g_{AA}N_{AA}(t) + \alpha_{Aa,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{Aa,aa}g_{aa}N_{aa}(t)}\\
 \phi_a &= 1- \phi_A
@@
 \begin{align}\begin{split}
-N_{AA}(t+1) &= s_{AA}[1-g_{AA}(K(t))]N_{AA}(t) + \phi_A^2T\\
-N_{Aa}(t+1) &= s_{Aa}[1-g_{Aa}(K(t))]N_{Aa}(t) + 2 \phi_A\phi_aT\\
-N_{aa}(t+1) &= s_{aa}[1-g_{aa}(K(t))]N_{aa}(t) + \phi_a^2T,
+N_{AA}(t+1) &= s_{AA}[1-g_{AA}]N_{AA}(t) + \phi_A^2T\\
+N_{Aa}(t+1) &= s_{Aa}[1-g_{Aa}]N_{Aa}(t) + 2 \phi_A\phi_aT\\
+N_{aa}(t+1) &= s_{aa}[1-g_{aa}]N_{aa}(t) + \phi_a^2T,
 \end{split}\end{align}
@@
 
-\section{Supplementary Figures}
+\section*{Supplementary Tables}
+
+\renewcommand{\thetable}{S-\arabic{table}}
+
+\begin{table}[htp]
+	\caption{\label{tab:comm-turn-pars}Parameters and parameter values for the community turnover case study. Values are assigned at
+		the start of \texttt{comm\_turn\_master.R}. ``Name" refers to the variable declared in the computer code. These names do not exactly 
+		match the symbols shown in the equations in Appendix \ref{metacomm}; rather, the species-specific values of those parameters are calculated in the computer code based on the values in this table.  }
+	\centering
+	\smallskip 
+	\begin{tabular}{llrl}
+		\hline
+		Name & Value & Definition \\
+		\hline
+		L\_land & 20 & Length of landscape \\
+		Tmin & 0 &  Minimum of spatial gradient in baseline temperature \\ 
+		Tmax & 15 &  Maximum of spatial gradient in baseline temperature \\
+		Tstdev & 2 &  Standard deviation of temperature (interannual variation) \\
+		deltaT & 4  & Magnitude of directional change in temperature \\
+		burnin\_yrs & 2000 & Number of years to initialize simulation \\
+		baseline\_yrs & 1000  & Number of years at baseline temperature used to fit statistical models \\
+		warming\_yrs  & 200 & Number of years over which temperature increases \\
+		final\_yrs & 2000  & Number of yrs at steady-state, elevated temperature \\
+		\hline
+		N & 40 & Number of species \\
+		Gmax & 0.5 & Maximum population growth rate \\
+		Gmin & 0.2 & Minimum population growth rate \\
+		Lmax & 1.5 & Maximum sensitivity to competition \\
+		Lmin & 0.7 & Minimum sensitivity to competition  \\
+		Cmax & 0.2 & Maximum additional sensitivity to conspecific competition \\
+		Cmin & 0.2 & Minimum additional sensitivity to conspecific competition \\
+		d & 0.01 &  Fraction of offspring dispersing from home site \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[htp]
+	\caption{\label{tab:eco-evo-pars}Parameters and parameter values for the eco-evolutionary case study. Values are assigned at
+		at the start of \texttt{genetic\_diversity\_master.R}. ``Name" refers to the variable declared in the computer code.  }
+	\centering
+	\smallskip 
+	\begin{tabular}{llrl}
+		\hline
+		Name & Values & Definition \\
+		\hline
+
+		\hline
+		fec\_Tmu & -1,0,1  & Optimal temperature for genotypes AA, Aa, and aa \\
+		fec\_Tsigma & 8  & Standard deviation in temperature niche for all genotypes \\
+		fec\_max & 100 &  Maximum fecundity for all genotypes \\
+		G & 1 & Germination rate for all genotypes \\
+		seedSurv &  0.5  & Survival rate for ungerminated seeds of all genotypes \\
+		alpha & 1  &  All competition coefficients (intra- and inter-) for all genotypes \\
+		\hline
+
+	\end{tabular}
+\end{table}
+
+\clearpage
+
+\section*{Supplementary Figures}
 
 \renewcommand{\thefigure}{S-\arabic{figure}}
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.7 \textwidth] {community_forecast_species_nonstationary.png}
+	\caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
+\new{In contrast to Fig.
+\ref{fig:community-forecast-species}, which shows results for a period of warming followed by stationary temperatures, for this simulation we spread the same temperature increase out over a longer period with no return to stationarity.} (B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
+(C) Simulated biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).
+Year 1000 in each panel corresponds to the start of the temperature increase.
+}
+	\label{fig:community-forecast-species-nonstationary}
+\end{figure}
 

### 2020-03-10 doc/manuscript.tex
finished param tables

@@
 \ref{fig:evo_evo_spatial_temporal_models}B).
-The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}.
+The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}, and \new{parameter values for simulations described here are shown in Table \ref{tab:eco-evo-pars}.}
 
@@
 	\smallskip 
-	\begin{tabular}{llrl}
+	\begin{tabular}{lrl}
 		\hline
@@
 		warming\_yrs  & 200 & Number of years over which temperature increases \\
-		final\_yrs & 2000  & Number of yrs at steady-state, elevated temperature \\
+		final\_yrs & 2000  & Number of years at steady-state, elevated temperature \\
 		\hline
@@
 	\caption{\label{tab:eco-evo-pars}Parameters and parameter values for the eco-evolutionary case study. Values are assigned at
-		at the start of \texttt{genetic\_diversity\_master.R}. ``Name" refers to the variable declared in the computer code.  }
+		at the start of \texttt{genetic\_diversity\_master.R}.
+``Name" refers to the variable declared in the computer code.
+Where appropriate, the corresponding symbols from equations in Appendix \ref{eco-evo} are shown in parentheses.
+}
 	\centering
 	\smallskip 
-	\begin{tabular}{llrl}
+	\begin{tabular}{lrl}
 		\hline
@@
 		\hline
-
+		Tstdev & 1 & Standard deviation of temperature (interannual variation) \\
+		baseT & -1 & Baseline temperature \\
+		deltaT & 5 & Total change in temperature \\
+		baseline\_yrs & 500 & Number of years at baseline temperature used to fit statistical models \\
+		warming\_yrs & 100  & Number of years over which temperature increases \\
+		final\_yrs & 300  & Number of years at steady-state, elevated temperature \\
 		\hline
-		fec\_Tmu & -1,0,1  & Optimal temperature for genotypes AA, Aa, and aa \\
-		fec\_Tsigma & 8  & Standard deviation in temperature niche for all genotypes \\
+		fec\_Tmu & -1,0,1  & Optimal temperature for genotypes AA, Aa, and aa, respectively \\
+		fec\_Tsigma & 8  & Standard deviation in fecundity for all genotypes \\
 		fec\_max & 100 &  Maximum fecundity for all genotypes \\
-		G & 1 & Germination rate for all genotypes \\
-		seedSurv &  0.5  & Survival rate for ungerminated seeds of all genotypes \\
-		alpha & 1  &  All competition coefficients (intra- and inter-) for all genotypes \\
+		G ($g$) & 1 & Germination rate for all genotypes \\
+		seedSurv ($s$)  &  0.5  & Survival rate for ungerminated seeds of all genotypes \\
+		alpha ($\alpha$) & 1  &  All competition coefficients for all genotypes \\
 		\hline
@@
 \renewcommand{\thefigure}{S-\arabic{figure}}
+
+\clearpage
 

### 2020-03-11 doc/manuscript.tex
trivia

@@
 		\hline
-		fec\_Tmu & -1,0,1  & Optimal temperature for genotypes AA, Aa, and aa, respectively \\
+		fec\_Tmu & -1,0,1  & Optimal fecundity temperature for genotypes AA, Aa, and aa \\
 		fec\_Tsigma & 8  & Standard deviation in fecundity for all genotypes \\

### 2020-03-11 doc/manuscript.tex
minor ms edits

@@
 Ecological forecasts are typically made using either a space-for-time substitution approach based on models fit to spatial data or using dynamic models fit to time-series data.
-Our results demonstrate that these two approaches can make very different predictions about the future state of ecological systems.
-Which approach provides the most accurate forecasts depends on the forecast-horizon.
+\new{Empirical studies show that the environment-response relationships detected by these approaches frequently differ in magnitude and even sign \citep{lauenroth_long-term_1992,Oedekoven2017,Amburgey2018,kleinhesselink_response_2018}.
+Our simulations illustrate how such differences may arise and then lead to} very different predictions about the future state of ecological systems.
+Which approach provides the most accurate forecasts \new{likely} depends on the forecast-horizon.
 In our simulations, time-series approaches performed best for short forecast horizons, whereas models based on spatial data made more accurate forecasts at long horizons.

### 2020-03-11 doc/manuscript.tex
minor ms edits again

@@
 There are two problems: finding appropriate data and determining the model weights \emph{a priori}.
-Many data sets have both a longitudinal and spatial dimension, but we could not think of one which also featured a clear ecological response to significant environmental change.
+Many data sets have both a longitudinal and spatial dimension, but we could not think of one which also featured a clear ecological response to directional environmental change.
 Surely such datasets exist, and we hope researchers who work with them will test our proposed weighted model.
@@
 
+\new{The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to phenomenological models.
+The same concept could work for any class of models that differ in predictive skill at different forecast horizons.
+Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.}
+
 Given the challenges of determining model weights \emph{a priori}, we should also pursue alternatives for intermediate forecast horizons.
@@
 
-Our results have important implications for the emerging field of ecological forecasting.
+Our \new{simulation studies} have important implications for the emerging field of ecological forecasting.
 First, they suggest that evaluating model performance at both short and long forecsat horizons will be essential as research on forecasting methods accelerates.

### 2020-03-12 doc/manuscript.tex
minor ms edits

@@
 The time-series approach involves fitting models to long-term datasets to describe the temporal dynamics of a system.
-We then use those dynamic models to make predictions about what will happen in the future.
+We then project those dynamic models to make predictions about what will happen in the future.
 This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, or primary production as a function of annual precipitation \citep{lauenroth_long-term_1992}.
@@
 
-Here we use two simulated case studies to 1) demonstrate that the best model-building approaches for ecological forecasting 
-depend on the time horizon of the forecast, and 2) explore how time-series and space-for-time approaches might be 
+Here we use two simulated case studies to \new{1) demonstrate why time-series and space-for-time approaches can make different predictions, 2) propose} that the best model-building approaches for ecological forecasting
+may depend on the time horizon of the forecast, and 3) explore how time-series and space-for-time approaches might be 
 combined via weighted averaging to make better forecasts at intermediate time scales. The first case study

### 2020-04-01 doc/manuscript.tex
revising code 4 nonstationarity

@@
 Our \new{simulation studies} have important implications for the emerging field of ecological forecasting.
-First, they suggest that evaluating model performance at both short and long forecsat horizons will be essential as research on forecasting methods accelerates.
+First, they suggest that evaluating model performance at both short and long forecast horizons will be essential as research on forecasting methods accelerates.
 Second, while single approaches may perform reasonably well for either short or long horizons, skillful predictions at intermediate forecast horizons may require a combination of information from spatial and temporal statistical models.

### 2020-04-01 community-turnover/comm_turn_figures.R

@@ -64,8 +64,11 @@ png("figures/community_models_total.png",height=3.5,width=4,res=400,units="in")
 dev.off()
 
 # plot forecast for focal species biomass and then total biomass
-my_expected_temp <- Tmean[site] + c(rep(0,baseline_yrs),seq(deltaT/warming_yrs,deltaT,deltaT/warming_yrs),
-                          rep(deltaT,final_yrs))
+my_expected_temp <- generateTemps(meanT=Tmean[site],changeT=deltaT, stationary=stationary_periods)
+
+# chop off burn in years
+my_expected_temp <- my_expected_temp[(burnin_yrs+1):sim_yrs]
+
 my_obs_temp <- temperature[(burnin_yrs+1):sim_yrs,site]
 my_biomass <- rowSums(spxp[(burnin_yrs+1):sim_yrs,site,])
 my_biomass_spp <- spxp[(burnin_yrs+1):sim_yrs,site,my_species]
@@ -83,7 +86,9 @@ png("figures/community_forecast_species.png",height=7.5,width=4.5,res=400,units=
   mtext(side=3," A",line=-1.1,adj=0,cex=0.8)
   
   # middle panel: focal species and forecasts
-  plot(my_biomass_spp,xlab="",ylab="Focal species biomass",type="l",col="darkgrey")
+  my_ylims <- c(0.9*min(c(my_biomass_spp,spatial_forecast_spp,temporal_forecast_spp),na.rm=T),
+                  1.05*max(c(my_biomass_spp,spatial_forecast_spp,temporal_forecast_spp),na.rm = T))
+  plot(my_biomass_spp,xlab="",ylab="Focal species biomass",type="l",col="darkgrey",ylim=my_ylims)
   abline(v=baseline_yrs,col="black",lty=3)
   lines(1:length(my_biomass_spp),spatial_forecast_spp,col="red",lwd=2)
   lines(1:length(my_biomass_spp),temporal_forecast_spp,col="blue",lwd=2)
@@ -125,13 +130,15 @@ png("figures/community_forecast_total.png",height=7.5,width=4.5,res=400,units="i
   mtext(side=3," A",line=-1.1,adj=0,cex=0.8)
   
   # total biomass
-  plot(my_biomass,xlab="",ylab="Total biomass",type="l",col="darkgrey")
+  my_ylims <- c(0.9*min(c(my_biomass,spatial_forecast,temporal_forecast),na.rm=T),
[…]

### 2020-04-01 community-turnover/comm_turn_forecast.R

@@ -18,11 +18,12 @@ y_lag <- rowSums(spxp[(burnin_yrs-1):(burnin_yrs+baseline_yrs-1),site,])
 temporal_model <- lm(y_temporal ~ y_lag + x_temporal)
 
 # get mean temperatures for forecast period (after baseline_yrs)
-forecast_Tmean <- Tmean[site] + c(seq(deltaT/warming_yrs,deltaT,deltaT/warming_yrs),
-                          rep(deltaT,final_yrs))
+forecast_Tmean <- generateTemps(meanT=Tmean[site],changeT=deltaT,stationary=stationary_periods)
+forecast_Tmean <- forecast_Tmean[(burnin_yrs+baseline_yrs+1):sim_yrs] 
 
 # make predictions from spatial model
 spatial_forecast <- coef(spatial_model)[1] + coef(spatial_model)[2]*forecast_Tmean 
+
 # make predictions from spatial model with uncertainty
 # spatial_forecast_spp <- predict(spatial_model_spp,newdata=data.frame(x_spatial=forecast_Tmean),
 #                                 interval="predict")

### 2020-04-01 community-turnover/comm_turn_master.R

@@ -24,6 +24,7 @@ Tmin <- 0; Tmax <- 15  # baseline temperature range
 Tmean <-seq(Tmin,Tmax,length=L_land)   # baseline mean temperatures across the landscape
 Tstdev <- 2   # st dev of temperature (interannual variation, stationary)
 deltaT <- 4  # total change in mean temperature
+stationary_periods <- F # F gives continuous temperature increase; T gives stationary baseline and final periods
 
 # length of simulation
 burnin_yrs <- 2000

### 2020-04-01 community-turnover/comm_turn_sim.R

@@ -3,8 +3,9 @@
 # generate one set of temperature time series to use in all simulations
 temperature<-matrix(NA,sim_yrs,length(Tmean))
 for(iT in 1:length(Tmean)){
-  tmp_mean <- Tmean[iT] + c(rep(0,burnin_yrs+baseline_yrs),seq(deltaT/warming_yrs,deltaT,deltaT/warming_yrs),
-                            rep(deltaT,final_yrs))
+  
+  tmp_mean <- generateTemps(meanT=Tmean[iT],changeT=deltaT,stationary=stationary_periods)
+  
   temperature[,iT] = rnorm(sim_yrs,tmp_mean,Tstdev)  # generate temperature time series
 }
 
@@ -14,7 +15,7 @@ landscape <- cbind(1:L_land , rep(1 ,L_land),Tmean)
 # 1st step: initialisation of the metacommunity
 spxp <- array(NA, dim=c(sim_yrs,L_land,N ))
 spxp[1,,] <- 2/N
-dt <- 0.025   # ????
+dt <- 0.025   
 d <- d/dt
 
 # Generation of the species pool
@@ -28,7 +29,7 @@ diag(seed_rain) <- 1-d*dt
 seed_rain[1,1] = 1-dt*d/2  
 seed_rain[L_land,L_land] = 1-dt*d/2
 
-# Run of the initialisation of community
+# Loop through time
 for(t in 2:sim_yrs){
   yr_temp <- temperature[t,]
   spxp[t,,] <- CommunityTempDis(spxp[t-1,,], tr, seed_rain, yr_temp, dt)

### 2020-04-01 community-turnover/lib/CommunityTempDis.R

@@ -1,7 +1,7 @@
 #
 #
 #	Help function to run community dynamics
-#	Main script author: Loïc Chalmandrier - lchalman@uwyo.edu, https://github.com/LoicChr
+#	Main script author: Loic Chalmandrier - lchalman@uwyo.edu, https://github.com/LoicChr
 #	Co-authors (and project leaders): J. Alexander - jake.alexander@unil.ch & L. Pellissier - loic.pellissier@usys.ethz.ch
 
 # Translated from Matlab to R by Peter Adler - peter.adler@usu.edu
@@ -26,3 +26,18 @@ CommunityTempDis <- function(spxp_ini, tr, seed_rain,temp,dt){
   
 }
 
+generateTemps <- function(meanT,changeT,stationary=T){
+  
+  if(stationary==T){
+    
+    out <- meanT + c(rep(0,burnin_yrs+baseline_yrs),seq(changeT/warming_yrs,changeT,changeT/warming_yrs),
+                        rep(changeT,final_yrs))
+  }else{
+    
+    out <- meanT + seq(changeT/sim_yrs,changeT,changeT/sim_yrs)
+    
+  }
+  
+  return(out)
+  
+}

### 2020-04-01 community-turnover/comm_turn_master.R
finished code revisions

@@ -24,7 +24,7 @@ Tmin <- 0; Tmax <- 15  # baseline temperature range
 Tmean <-seq(Tmin,Tmax,length=L_land)   # baseline mean temperatures across the landscape
 Tstdev <- 2   # st dev of temperature (interannual variation, stationary)
 deltaT <- 4  # total change in mean temperature
-stationary_periods <- F # F gives continuous temperature increase; T gives stationary baseline and final periods
+stationary_periods <- T # F gives continuous temperature increase; T gives stationary baseline and final periods
 
 # length of simulation
 burnin_yrs <- 2000

### 2020-04-01 doc/response_2_reviews_Ecography1.txt
first draft of response letter

Recommendation by the Subject Editor (anonymous):

Both reviewers consider the work interesting and potentially highly valuable. I fully agree with them. The reason why your manuscript is rejected (with possibility to resubmit) was the the reviewers considered that the present version does not include sufficient information to truly evaluate the scientific contents (the analyses are basically not reproducible due to lacking information). Thus I urge you to pay specific attention to describe the methods in sufficient detail so that the results can be critically evaluated by the reviewers. Naturally, please address also all other comments provided by the two reviewers.

Reviewer(s)' Comments to Author:
Reviewer: 1

Comments to the Author
The manuscript compares time series and a "space-for-time" approach for predicting species abundance and distribution in the future under two theoretical scenarios: meta-community dynamics and eco-evolutionary dynamics. The paper tackles an interesting and relevant question related to species distribution modeling and population assessment and provides some useful insight to it. The paper is well and clearly written.

My main concern with the paper is that it seems to exaggerate the importance and generality of its results. The abstract and introduction give impression that ecological forecast is tackled in very general manner. However, at the end authors study only two very simplistic (even though relavant) descriptions of ecological processes. Moreover, the set of statistical models compared is very narrow. Hence, the generality of their results is limited as well. This is not a weakness as such since with this simple example authors are able to show interesting results which are rather easily understandable. However, in my opinion the tone of the paper should be much more modest and authors should give relevant information about their simulation set up already in the abstract.

RESPONSE: Ethan and Michael, I am not sure what to say here...the approaches we study are very common and, in a sense, very general. And I can't believe that the details of our statistical models matter, especially since we are analyzing simulated data with a very high signal to noise ratio. Go back to the point that the key difference is in the shape/sign of environment-response relationships in space vs time? Also, I re-read with an eye on tone, and we have caveats throughout. Not sure what I can do.

I don't like the term "space-for-time". I have not encountered such term before and for me it is not informative. Why not use species distribution model or something similar that has already an established meaning.

RESPONSE: In our experience, "space-for-time" is a common approach. A quick search on Google Scholar shows many papers with "space-for-time" in the title or abstract, including papers in high impact journals. "Species distribution model" is a narrower term, which does not suit our purpose as well.

In lines 150 and 152 authors use term "correlate". I know that statistical models are often called correlative models in ecology but it is typically wrong and at least impresice. Hence, I would describe in more detail how the model inference was done - maximum likelihood, Bayesian inference or least squares or what.

RESPONSE: We replaced "correlated" with regressed. Details on the regression models are provided in Appendix A.

The description of the simulation study should include the specific values used for the simulator. Authors should also tell more clearly are their results from single simulation or from several simulations. Moreover, it would be good to know how sensitive the results are to the chosen simulator parameters. That is, if the simulated dynamics was made faster/slower and the interspecific interactions and dispersal made stronger/weaker how would the results change. This kind of "sensitivity" test would provide information about the generality of the results and insight on for which kind of real systems they might apply.

RESPONSE:
Both reviewers raised concerns about the lack of sensitivity analyses and the robustness of our results. We take these concerns seriously and have responded in a number of ways. 	1) We added new material and references to the Introduction to emphasize that our simulations are motivated by empirical examples in which spatial and temporal models detect different relationships between an environmental driver and an ecological response. Any difference in magnitude or sign in these relationships will lead to contrasting forecasts. This situation appears quite frequently in the empirical literature; it is not a theoretical curiosity. The purpose of our simulations is not to convince readers that this situation might possibly occur (we already know it does), but rather to demonstrate how this situation may arise and the challenges and opportunities it presents for ecological forecasting. Given the generality of the motivating empirical pattern, and the fact that we designed our case study to reproduce that pattern, and that the signal to noise ratio in our simulation is high, we do not believe that a detailed sensitivity analysis is necessary. However, we do agree with the reviewer that understanding how different parameters influence the rate of the shift in model weights is an important goal for future research. Fig. 6 seeks to make that point, but a more rigorous analysis is beyond the scope of this study.
	2) We have added supplementary tables showing parameter values for both case studies. 
	3) We now clarify that the results are based on one simulation run, but qualitative results were consistent.
	4) We addressed Reviewer 2's question about nonstationarity (see below).

Reviewer: 2

Comments to the Author
The manuscript asks an interesting question - is it better to use (short) time series or space-for-time substitutions. The results make intuitive sense, but they are difficult to judge because the manuscript does not give enough details about the work. There are no details about the simulations - how many years were simulated? What spatial extent? What parameter values were used? How was the (simulated) data collected? How many replicate simulations were run (if one, why were you unable to run more than one? How are we meant to know if the results replicate?)? etc. etc. Basically, it is impossible to re-create these analyses from the description. I would strongly suggest having separate Methods and Results sections, as this makes it easier to follow the description of the work (and also helps you focus on describing the methods, and the results).

RESPONSE: Our response to Reviewer 1's similar concerns addresses many of these points. We originally considered separate Methods and Results sections, but found that structure awkward given our two case studies. We could add Methods and Results subheaders within each of the case studies if the editor and reviewers think that would be helpful.

Beyond the writing, I think one important issue is the assumption of stationarity. My guess (and I have to guess, because we're not given this information) the spatial data is taken after the simulation has run to equilibrium. But what happens if the spatial data is taken when the climate is changing? I think that would be the important question for researchers using space-for-time with contemporary data.

RESPONSE: Interesting question! We modified our code for the community turnover example to study the situation in which temperature increases through the entire simulation, with no stationary period for model fitting. The key result is shown in SI Fig. 1. Qualitatively, the result is consistent: at short forecast horizons the temporal statistical model makes a better forecast while at longer forecast horizons the spatial model does better. However, as we explain in the main text (LINES), the forecast from the temporal model is not as accurate, even at short forecast horizons, as in the original simulation with a stationary period for model fitting. We believe this reflects that fact that the temporal statistical model does not account for the long-term trend. Doing so might recover the same kind of information about slow processes contained in the spatial statistical model.

I also had a quick look at the code. Unfortunately, it's written so as not to be usable by anyone other then the authors, or someone working on a Windows computer who wants to set up the same directory structure. Please, make the code portable. I would suggest making an R package, as this will give you a consistent structure, means you don't need to use setwd() and rm(list=ls()) (both abominations in portable code), and you can write vignettes which are human readable to demonstrate how the code was used to run the simulations.

RESPONSE: Ethan?

### 2020-04-01 doc/manuscript.tex
first draft of response letter

@@
 We made forecasts based on two phenomenological statistical models, each representing processes operating at different time scales.
-One statistical model represents the time-series or ``temporal approach.'' We correlated interannual variation in an ecological response with interannual variation in temperature at just one site.
+One statistical model represents the time-series or ``temporal approach.'' We regressed interannual variation in an ecological response with interannual variation in temperature at just one site.
 The other statistical model relies on a space-for-time substitution, which we call the ``spatial approach'' for brevity.
-We correlated the mean temperature with the mean of an ecological state or rate across many sites.
+We regressed the mean temperature on the mean of an ecological state or rate across many sites.
 We compared forecasts from both statistical models to the simulated dynamics to determine how well the two approaches performed at different forecast horizons.
@@
 We focused on the biomass dynamics of one focal species that dominated the central site during the baseline period.
+\new{We report results from one simulation run; results were qualitatively consistent for replicate runs.}
 
@@
 \ref{fig:community-forecast-species}B).
-\new{The qualitative results were the same when we repeated the simulation with a continuous, nonstationary temperature increase (SI Fig.
-\ref{fig:community-forecast-species-nonstationary}).}
 
@@
 A true forecast from the weighted model would require a method to determine the model weights \emph{a priori}.
+
+\new{When we repeated the simulation with a continuous, nonstationary temperature increase, we see a qualitatively similar shift in weights with increasing forecast horizon from the temporal to spatial statistical model (SI Fig.
+\ref{fig:community-forecast-species-nonstationary}).
+In this case, the forecast from the temporal statistical model is not as skillful in the near-term forecast horizon, because the model does not account for the temperature trend during the model fitting period.
+Separating the effect of annual temperature deviations from the temperature trend would distinguish between short and long-term patterns, much as our temporal and spatial statistical models do in the simulation with stationary temperature periods.
+}
 
@@
 \ref{fig:evo_evo_spatial_temporal_models}B).
-The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}, and \new{parameter values for simulations described here are shown in Table \ref{tab:eco-evo-pars}.}
+The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}, and \new{parameter values for simulations described here are shown in Table \ref{tab:eco-evo-pars}.
+As in the first case study, we report results from just one simulation run, but results were qualitatively consistent for replicate runs.}
 
@@
 \new{In contrast to Fig.
-\ref{fig:community-forecast-species}, which shows results for a period of warming followed by stationary temperatures, for this simulation we spread the same temperature increase out over a longer period with no return to stationarity.} (B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
+\ref{fig:community-forecast-species}, which shows results for a period of warming followed by stationary temperatures, for this simulation we spread the same temperature increase out over the entire simulation with no stationary periods.} (B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
 (C) Simulated biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).

### 2020-04-01 doc/response_2_reviews_Ecography1.txt
trivia

@@
 	3) We now clarify that the results are based on one simulation run, but qualitative results were consistent.
-	4) We addressed Reviewer 2's question about nonstationarity (see below).
+	4) We have added more detail to the meta-community model description in Appendix B.
+We believe the description of the evo-evo model is quite complete (Appendix C), but would be happy to add more detail in response to specific questions.
+	5) We addressed Reviewer 2's question about nonstationarity (see below).
 
@@
 We originally considered separate Methods and Results sections, but found that structure awkward given our two case studies.
-We could add Methods and Results subheaders within each of the case studies if the editor and reviewers think that would be helpful.
+We could add Methods and Results subheaders within each of the case study sections if the editor and reviewers think that would be helpful.
 
@@
 
-RESPONSE: Ethan?
+RESPONSE: Ethan? Would it be rude to point out that the code uses relative paths, so you don't need to "create" a directory structure? You just download, read the readme, and set one path?

### 2020-04-28 doc/response_2_reviews_Ecography1.txt
made code portable

@@
 
-RESPONSE: Ethan? Would it be rude to point out that the code uses relative paths, so you don't need to "create" a directory structure? You just download, read the readme, and set one path?
+RESPONSE: We made the following changes to the code to make it more portable: 1) We removed rm(list=ls()).
+2) We now use the "here" package to set the working directory.
+3) We now use the "pacman" package to install and load required R packages.
+4) Where we refer readers to the url for downloading the code, we have added a note directing readers to the README file in the main directory, which contains very simple instructions for running the code.
+The original version of our code used relative paths, so users should not need to "set up the same directory structure." As long as the user does not change the directory structure of the zip archive we distribute, the code should run (we have asked our students to test it and it has worked on a variety of platforms).

### 2020-04-28 community-turnover/comm_turn_master.R

@@ -1,16 +1,14 @@
 
-# clean up
-rm(list=ls())
+# install R packages (if necessary) and load them
+if (!require("pacman")) install.packages("pacman")
+pacman::p_load(boot, scales, here)
 
 # set working directory
-setwd("C:/repos/space-time-forecast/community-turnover")
+here()
 
-# load packages and functions
+# load custom functions
 source("lib/CommunityTempDis.R")
 source("lib/SpeciesPoolGen.R")
-library("boot")
-library("scales")
-#library("latticeExtra")
 
 ###
 ### 1. Simulate a metacommunity: first stationary conditions, then temperature increase, then stationary again

### 2020-04-28 genetic-diversity/genetic_diversity_master.R

@@ -1,14 +1,13 @@
 
-# clean up
-rm(list=ls())
+# install R packages (if necessary) and load them
+if (!require("pacman")) install.packages("pacman")
+pacman::p_load(boot, latticeExtra, here)
 
 # set working directory
-setwd("C:/Repos/space-time-forecast/genetic-diversity")
+here()
 
-# load packages and functions
+# load custom functions
 source("genetic_diversity_functions.R")
-library("boot")
-library("latticeExtra")
 
 ###
 ### 1. Simulate and describe spatial pattern in abundances at sites across a range of mean temperatures
@@ -33,8 +32,6 @@ seedSurv = 0.5  # survival of ungerminated seeds (same for both genotypes)
 alpha = matrix(1,3,3)      # all competition coefficients (intra- and inter-) are equal for both genotypes
 
 source("genetic_diversity_sim_spatial.R")
-
-#CHECK: why zeros for least fit genotype's mean abundances at lowest temps but not highest?
 
 ###
 ### 2. Simulate and describe temporal pattern in abundances at one site experiencing warming

### 2020-04-29 genetic-diversity/genetic_diversity_functions.R
removed seedbank from code

@@ -24,9 +24,9 @@ newSeeds = function(N,lambda,alpha,G){
   	# G = vector of length 3 of each genotype's germination rate
   	# output = vector (3) giving # seeds produced by each genotype at time t+1
     
-    AA = lambda[1]*N[1]*G/(1+alpha[1,1]*N[1]*G+alpha[1,2]*N[2]*G+alpha[1,3]*N[3]*G)
-  	Aa = lambda[2]*N[2]*G/(1+alpha[2,1]*N[1]*G+alpha[2,2]*N[2]*G+alpha[2,3]*N[3]*G)
-    aa = lambda[3]*N[3]*G/(1+alpha[3,1]*N[1]*G+alpha[3,2]*N[2]*G+alpha[3,3]*N[3]*G)
+    AA = lambda[1]*N[1]/(1+alpha[1,1]*N[1]+alpha[1,2]*N[2]+alpha[1,3]*N[3])
+  	Aa = lambda[2]*N[2]/(1+alpha[2,1]*N[1]+alpha[2,2]*N[2]+alpha[2,3]*N[3])
+    aa = lambda[3]*N[3]/(1+alpha[3,1]*N[1]+alpha[3,2]*N[2]+alpha[3,3]*N[3])
   	return(c(AA,Aa,aa))
 
 }
@@ -40,7 +40,7 @@ getPhi<-function(seeds){
   return(phi)
 }
 
-diploid = function(N,seedSurv,G,seeds,phi){
+diploid = function(N,seeds,phi){
     # This is an annual plant model, so it tracks
     # populations as numbers of seeds
 
@@ -51,12 +51,12 @@ diploid = function(N,seedSurv,G,seeds,phi){
     # phi = vector (2) giving frequency of homozygotes
   	# output = vector (3) giving density of each genotype at time t+1
     
-    Plants = G*N
+    Plants = N
     bigT = sum(seeds)  # total seed production
     Fec=c(phi[1]^2*bigT, 2*phi[1]*phi[2]*bigT, phi[2]^2*bigT)
-  	AA = seedSurv*(1-G)*N[1]+Fec[1]
-  	Aa = seedSurv*(1-G)*N[2]+Fec[2]
-    aa = seedSurv*(1-G)*N[3]+Fec[3]
+  	AA = Fec[1]
+  	Aa = Fec[2]
+    aa = Fec[3]
   	SB = c(AA,Aa,aa)
   	return(list(SB=SB,Plants=Plants,SP=Fec))
 }

### 2020-04-29 genetic-diversity/genetic_diversity_master.R

@@ -28,7 +28,7 @@ fec_Tmu = c(-1,0,1)  # optimal temperature for genotypes AA, Aa, and aa
 fec_Tsigma = rep(8,3)     # standard deviation in temperature response for genotypes AA, Aa, and aa
 fec_max = c(100,100,100)  # fecundity for genotypes AA, Aa, and aa
 G <- 1
-seedSurv = 0.5  # survival of ungerminated seeds (same for both genotypes)
+seedSurv = 0  # survival of ungerminated seeds (same for both genotypes)
 alpha = matrix(1,3,3)      # all competition coefficients (intra- and inter-) are equal for both genotypes
 
 source("genetic_diversity_sim_spatial.R")
@@ -55,7 +55,6 @@ temperature = rnorm(sim_yrs,Tmean,Tstdev)  # generate temperature time series
 # use same parameters from diploid model above 
 
 source("genetic_diversity_sim_temporal.R")
-
 
 ###
 ### 3. Generate forecasts for the temporal simulation (part 2)

### 2020-04-29 genetic-diversity/genetic_diversity_sim_spatial.R

@@ -15,9 +15,9 @@ for(iT in 1:length(Tmean)){
   lambdas = getLambdas(temperature[,iT],fec_Tmu,fec_Tsigma,fec_max)
   
   for(iYr in 2:(sim_yrs)){
-    seeds<-newSeeds(N[iYr-1,],lambdas[iYr,],alpha,G)
+    seeds<-newSeeds(N[iYr-1,],lambdas[iYr,],alpha)
     phi <- getPhi(seeds)
-    out <- diploid(N[iYr-1,],seedSurv,G,seeds,phi)
+    out <- diploid(N[iYr-1,],seeds,phi)
     N[iYr,] <- out$SB
     Plants[iYr,] <- out$Plants
     Fec[iYr,] <- out$SP

### 2020-04-29 genetic-diversity/genetic_diversity_sim_temporal.R

@@ -11,9 +11,9 @@ N[1,] = fec_max[1]/3
 lambdas = getLambdas(temperature,fec_Tmu,fec_Tsigma,fec_max)
 
 for(iYr in 2:(sim_yrs)){
-  seeds<-newSeeds(N[iYr-1,],lambdas[iYr,],alpha,G)
+  seeds<-newSeeds(N[iYr-1,],lambdas[iYr,],alpha)
   phi <- getPhi(seeds)
-  out <- diploid(N[iYr-1,],seedSurv,G,seeds,phi)
+  out <- diploid(N[iYr-1,],seeds,phi)
   N[iYr,] <- out$SB  # total seedbank
   Plants[iYr,] <- out$Plants  # germinated plants
   Fec[iYr,] <- out$SP # seed production

### 2020-04-29 doc/manuscript.tex
made code portable
removed s and g from manuscript

@@
 Computer code for both case studies is available for reviewers online (https://drive.google.com/open?id=1aju04qtQvJmZG1mcpN-6hYBZILBhDfl2) and will be archived at Zenodo upon acceptance.
+Instructions for running the code are in the README file in the main directory of the zip archive.
 
@@
 \ref{fig:evo_evo_spatial_temporal_models}A).
+All seeds germinate every year, preventing a seedbank from developing.
 
@@
 
-\noindent \textbf{Haploid Model:} Begin with a haploid model that describes the number of seeds present in a seed bank.  $N_{i,t}$ is the number of seeds of species $i$ at time $t$. The model is
+\noindent \textbf{Haploid Model:} Begin with a haploid model that describes the number of seeds present in a population.
+We model a scenario in which all seeds germinate, so we can ignore seedbank dynamics.
+$N_{i,t}$ is the number of seeds of species $i$ at time $t$.
+The model is
 \begin{align}\begin{split}
-N_{1,t+1} &= s_1 [1-g_1]N_{1,t}+\frac{\lambda_1(K(t))g_1N_{1,t}}{1+ \alpha_{11}g_1N_{1,t} + \alpha_{12}g_2N_{2,t}}\\
-N_{2,t+1} &= s_2 [1-g_2]N_{2,t}+\frac{\lambda_2(K(t))g_2N_{2,t}}{1+ \alpha_{21}g_1N_{1,t} + \alpha_{22}g_2N_{2,t}}
+N_{1,t+1} &= \frac{\lambda_1(K(t))N_{1,t}}{1+ \alpha_{11}N_{1,t} + \alpha_{12}N_{2,t}}\\
+N_{2,t+1} &= \frac{\lambda_2(K(t))N_{2,t}}{1+ \alpha_{21}N_{1,t} + \alpha_{22}N_{2,t}}
 \end{split}\end{align}
-where $g_{i}$ is the probability of germination for species $i$, $s_i$ is the seed survival probability, $\lambda_{i}(K(t))$ is the seed production rate per plant, and $K(t)$ is the temperature at time $t$.
-Below we refer to the $\alpha_{ij}$ as intra- and inter-genotype competition coefficients.
+where $\lambda_{i}(K(t))$ is the seed production rate per plant, and $K(t)$ is the temperature at time $t$.  Below we refer to the $\alpha_{ij}$ as intra- and inter-genotype competition coefficients.
 
-\noindent \textbf{Diploid Model:} Consider a one-species diploid model.
+%\noindent \textbf{Diploid Model:} Consider a one-species diploid model.
 The genotypes are denoted by $AA$, $Aa$, and $aa$.
@@
 
+\noindent \textbf{Diploid Model:} Consider a one-species diploid model.
+The genotypes are denoted by $AA$, $Aa$, and $aa$.
+The number of each genotype at time $t$ is $N_{AA}(t)$, $N_{Aa}(t)$, and $N_{aa}(t)$.
+%The germination rates for each genotype are $g_{AA}$, $g_{Aa}$, and $g_{aa}$.  
+%The seed survival probability and %
+The seed production rate for genotype $AA$ is $\lambda_{AA}(K(t))$, and the analogous parameters for the other genotypes are similarly denoted.
+The competition coefficients are denoted by $\alpha_{i,j}$, e.g., $\alpha_{AA,AA}$ or $\alpha_{AA,Aa}$.
+Throughout we assume that gametes mix randomly in the population.
+
 First consider the case where the competition coefficients are zero ($\alpha_{i,j}=0$).  Let $T$ denote the total number of gamete-pairs produced in a given year,
 \begin{equation}
-T = \lambda_{AA}(K(t))N_{AA}(t)g_{AA} + \lambda_{Aa}(K(t))N_{Aa}(t)g_{Aa} +\lambda_{aa}(K(t))N_{aa}(t)g_{aa}.
+T = \lambda_{AA}(K(t))N_{AA}(t) + \lambda_{Aa}(K(t))N_{Aa}(t) +\lambda_{aa}(K(t))N_{aa}(t).
 \end{equation}
@@
 \begin{align}\begin{split}
-\phi_{A} &= \frac{\lambda_{AA}(K(t))N_{AA}(t)g_{AA}+ \frac{1}{2}\lambda_{Aa}(K(t))N_{Aa}(t)g_{Aa}}{T} \hspace{10pt} \text{and} \hspace{10pt} \phi_a = 1-\phi_{A}.
+\phi_{A} &= \frac{2\lambda_{AA}(K(t))N_{AA}(t)+ \lambda_{Aa}(K(t))N_{Aa}(t)}{2T} \hspace{10pt} \text{and} \hspace{10pt} \phi_a = 1-\phi_{A}.
 \end{split}\end{align}
@@
 \begin{align}\begin{split}
-N_{AA}(t+1) &= s_{AA}[1-g_{AA}]N_{AA}(t) + \phi_A^2T\\
-N_{Aa}(t+1) &= s_{Aa}[1-g_{Aa}]N_{Aa}(t) + \phi_A\phi_aT\\
-N_{aa}(t+1) &= s_{aa}[1-g_{aa}]N_{aa}(t) + \phi_a^2T
+N_{AA}(t+1) &= \phi_A^2T\\
+N_{Aa}(t+1) &= \phi_A\phi_aT\\
+N_{aa}(t+1) &= \phi_a^2T
 \end{split}\end{align}
@@
 \begin{align}\begin{split}
-T &=  \frac{\lambda_{AA}(K(t))N_{AA}(t)g_{AA}}{1+ \alpha_{AA,AA}g_{AA}N_{AA}(t) + \alpha_{AA,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{AA,aa}g_{aa}N_{aa}(t)} \\ 
-&+ \frac{\lambda_{Aa}(K(t))N_{Aa}(t)g_{Aa}}{1+ \alpha_{Aa,AA}g_{AA}N_{AA}(t) + \alpha_{Aa,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{Aa,aa}g_{aa}N_{aa}(t)}\\
-&+\frac{\lambda_{aa}(K(t))N_{aa}(t)g_{aa}}{1+ \alpha_{aa,AA}g_{AA}N_{AA}(t) + \alpha_{aa,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{aa,aa}g_{aa}N_{aa}(t)}.
+T &=  \frac{\lambda_{AA}(K(t))N_{AA}(t)}{1+ \alpha_{AA,AA}N_{AA}(t) + \alpha_{AA,Aa}N_{Aa}(t)+ \alpha_{AA,aa}N_{aa}(t)} \\ 
+&+ \frac{\lambda_{Aa}(K(t))N_{Aa}(t)}{1+ \alpha_{Aa,AA}N_{AA}(t) + \alpha_{Aa,Aa}N_{Aa}(t)+ \alpha_{Aa,aa}N_{aa}(t)}\\
+&+\frac{\lambda_{aa}(K(t))N_{aa}(t)}{1+ \alpha_{aa,AA}N_{AA}(t) + \alpha_{aa,Aa}N_{Aa}(t)+ \alpha_{aa,aa}N_{aa}(t)}.
 \label{eqn:defineT}
@@
 \begin{align}\begin{split}
-\phi_A &= \frac{1}{T}\frac{\lambda_{AA}(K(t))N_{AA}(t)g_{AA}}{1+ \alpha_{AA,AA}g_{AA}N_{AA}(t) + \alpha_{AA,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{AA,aa}g_{aa}N_{aa}(t)} \\
-&+ \frac{1}{2T}\frac{\lambda_{Aa}(K(t))N_{Aa}(t)g_{Aa}}{1+ \alpha_{Aa,AA}g_{AA}N_{AA}(t) + \alpha_{Aa,Aa}g_{Aa}N_{Aa}(t)+ \alpha_{Aa,aa}g_{aa}N_{aa}(t)}\\
+\phi_A &= \frac{2}{2T}\frac{\lambda_{AA}(K(t))N_{AA}(t)}{1+ \alpha_{AA,AA}N_{AA}(t) + \alpha_{AA,Aa}N_{Aa}(t)+ \alpha_{AA,aa}N_{aa}(t)} \\
+&+ \frac{1}{2T}\frac{\lambda_{Aa}(K(t))N_{Aa}(t)}{1+ \alpha_{Aa,AA}N_{AA}(t) + \alpha_{Aa,Aa}N_{Aa}(t)+ \alpha_{Aa,aa}N_{aa}(t)}\\
 \phi_a &= 1- \phi_A
@@
 \begin{align}\begin{split}
-N_{AA}(t+1) &= s_{AA}[1-g_{AA}]N_{AA}(t) + \phi_A^2T\\
-N_{Aa}(t+1) &= s_{Aa}[1-g_{Aa}]N_{Aa}(t) + 2 \phi_A\phi_aT\\
-N_{aa}(t+1) &= s_{aa}[1-g_{aa}]N_{aa}(t) + \phi_a^2T,
+N_{AA}(t+1) &= \phi_A^2T\\
+N_{Aa}(t+1) &= 2 \phi_A\phi_aT\\
+N_{aa}(t+1) &= \phi_a^2T,
 \end{split}\end{align}
@@
 		fec\_max & 100 &  Maximum fecundity for all genotypes \\
-		G ($g$) & 1 & Germination rate for all genotypes \\
-		seedSurv ($s$)  &  0.5  & Survival rate for ungerminated seeds of all genotypes \\
 		alpha ($\alpha$) & 1  &  All competition coefficients for all genotypes \\

### 2020-04-29 doc/response_2_reviews_Ecography1.txt
finishing revisions

@@
 Naturally, please address also all other comments provided by the two reviewers.
+
+RESPONSE: Thank you for the opportunity to resubmit.  We have carefully addressed all of the reviewer's comments, making sure to add sufficient detail so that our results can be reproduced.
 
@@
 
-RESPONSE: Ethan and Michael, I am not sure what to say here...the approaches we study are very common and, in a sense, very general.
-And I can't believe that the details of our statistical models matter, especially since we are analyzing simulated data with a very high signal to noise ratio.
-Go back to the point that the key difference is in the shape/sign of environment-response relationships in space vs time?
-Also, I re-read with an eye on tone, and we have caveats throughout.
-Not sure what I can do.
+RESPONSE: We have rewritten the second half of the abstract to make it clear that all results come from two simple simulation studies.
+We also looked closely at the text in the Discussion to make sure we are not overselling our results.
+We think the revised version makes the limited scope of inference clear.
+As for the limited set of simple statistical models, that choice was intentional, as the reviewer recognizes.
+Because the statistical models are simple, they represent a wide variety of analogous models currently used in empirical research.
 
@@
 
-RESPONSE: In our experience, "space-for-time" is a common approach.
-A quick search on Google Scholar shows many papers with "space-for-time" in the title or abstract, including papers in high impact journals.
-"Species distribution model" is a narrower term, which does not suit our purpose as well.
+RESPONSE: In our experience, "space-for-time" is a common term.
+A quick search on Google Scholar shows many papers with "space-for-time" in the title or abstract, including papers cited hundreds of times.
+Stuart Pickett's 1989 book chapter in the Long-term Studies in Ecology, titled, "Space-for-Time Substitution as an Alternative to Long-Term Studies," has been cited almost 1300 times (Google Scholar).
+"Species distribution model" is a narrower term, which does encompass other response variable, such as primary production.
+We wanted it to be clear that space-for-time approaches are not limited to species distribution models.
 
@@
 The purpose of our simulations is not to convince readers that this situation might possibly occur (we already know it does), but rather to demonstrate how this situation may arise and the challenges and opportunities it presents for ecological forecasting.
-Given the generality of the motivating empirical pattern, and the fact that we designed our case study to reproduce that pattern, and that the signal to noise ratio in our simulation is high, we do not believe that a detailed sensitivity analysis is necessary.
+Given the generality of the motivating empirical pattern, the fact that we designed our case study to reproduce that pattern, and the fact that the signal to noise ratio in our deterministic simulations is high (only temperature is stochastic), we do not believe that a detailed sensitivity analysis is necessary.
 However, we do agree with the reviewer that understanding how different parameters influence the rate of the shift in model weights is an important goal for future research.
@@
 	2) We have added supplementary tables showing parameter values for both case studies. 
-	3) We now clarify that the results are based on one simulation run, but qualitative results were consistent.
+	3) We now clarify that the main text presents the results based on a single simulation run, but our results are qualitatively consistent for other runs.
 	4) We have added more detail to the meta-community model description in Appendix B.

### 2020-04-29 doc/manuscript.tex
finishing revisions

@@
 Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
-\new{However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change.} Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
-We use two simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
-In the short-term, dynamics reflect initial conditions and fast processes such as birth and death, and the time-series approach makes the best predictions.
-In the long-term, dynamics reflect the additional influence of slower processes such as evolutionary and ecological selection, colonization and extinction, which the space-for-time approach can effectively capture.
-At intermediate time-scales, a weighted average of the two approaches shows promise.
-However, making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
+However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change.
+Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
+We use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
+First, we simulated population or community dynamics under stationary temperature using two simple, mechanistic models.
+Second, we fit statistical models to the simulated data using a time-series approach, a space-for-time approach, or a weighted average.
+We then forecast the response to a temperature increase using the statistical models, and compared these forecasts to temperature effects simulated by the mechanistic models.
+We found that the time-series approach made accurate short-term predictions because it captured initial conditions and effects of fast processes such as birth and death.
+The space-for-time approach made more accurate long-term predictions because it better captured the influence of slower processes such as evolutionary and ecological selection.
+The weighted average made accurate predictions at all time scales, including intermediate time-scales where the other two approaches performed poorly.
+A weighted average of time-series and space-for-time approaches shows promise, but making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
 
@@
 We explored this question using a metacommunity simulation model developed by \cite{alexander_lags_2018} to study how community responses to increasing temperature depend on the interplay between within-site demography and competitive interactions and the movement of species across sites.
+
+\subsection*{Methods}
+
 The model features Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
@@
 \new{We report results from one simulation run; results were qualitatively consistent for replicate runs.}
+
+\subsection*{Results}
 
@@
 Although shifts in genotype frequencies at the population level are analogous to changes in species composition at the community level, the mechanisms are distinct: heterozygosity and genetic recombination have no analogue at the community level.
-We demonstrate how these processes influence short and long-term forecasts with a standard eco-evolutionary simulation model for a hypothetical annual plant population in which fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
+We demonstrate how these processes influence short and long-term forecasts with a standard eco-evolutionary simulation model for a hypothetical annual plant population.
+
+\subsection*{Methods}
+
+Our model assumes that fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
 \ref{fig:evo_evo_spatial_temporal_models}A).
 All seeds germinate every year, preventing a seedbank from developing.
-
-Our model describes how the local density of each genotype changes between years, which depends on temperature and genotype densities in the previous year.
+The model describes how the local density of each genotype changes between years, which depends on temperature and genotype densities in the previous year.
 Transient temporal dynamics are computed directly from the model; these dynamics are the basis for the temporal statistical model.
@@
 However, if warmer than normal conditions persist for many years, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
+
+\subsection*{Results}
 

### 2020-04-29 community-turnover/comm_turn_figures.R

@@ -49,16 +49,16 @@ png("figures/community_models_total.png",height=3.5,width=4,res=400,units="in")
   
   # temporal
   xx <- seq(min(x_temporal),max(x_temporal),0.1)
-  points(x_temporal,y_temporal,pch=1,col=myCols[2])
+  points(x_temporal,y_temporal,pch=1,col=myCols[1])
   y_hat <- coef(temporal_model)[1] + coef(temporal_model)[2]*mean(y_temporal) + coef(temporal_model)[3]*xx
-  lines(xx,y_hat, col=myCols[2], lwd=2, lty="solid" ) # plot marginal effect of temperature for temporal model
+  lines(xx,y_hat, col=myCols[1], lwd=2, lty="solid" ) # plot marginal effect of temperature for temporal model
   
   # total biomass spatial
-  points(x_spatial,y_spatial,pch=16,col=myCols[1])
-  abline(spatial_model,col=myCols[1],lwd=1)
+  points(x_spatial,y_spatial,pch=16,col=myCols[2])
+  abline(spatial_model,col=myCols[2],lwd=1)
   
-  legend("topleft",c("Spatial","Temporal"),
-         lty=c("solid","solid"),pch=c(16,1),lwd=c(1,1),
+  legend("topleft",c("Temporal","Spatial"),
+         lty=c("solid","solid"),pch=c(1,16),lwd=c(1,1),
          col=myCols,bty="n",cex=0.8)
 
 dev.off()

### 2020-04-29 community-turnover/comm_turn_master.R

@@ -4,7 +4,8 @@ if (!require("pacman")) install.packages("pacman")
 pacman::p_load(boot, scales, here)
 
 # set working directory
-here()
+setwd(here()) # set root directory
+setwd("./community-turnover")  # move to subdirectory
 
 # load custom functions
 source("lib/CommunityTempDis.R")

### 2020-04-29 community-turnover/comm_turn_sim.R

@@ -38,7 +38,7 @@ for(t in 2:sim_yrs){
 spxp_mean_baseline <- apply(spxp[burnin_yrs:(burnin_yrs+baseline_yrs),,],MARGIN = c(2,3),FUN=mean)
 
 # save the initial metacommunity
-dir.create("simulations/run_1")
+if(file.exists("simulations/run_1")==F) dir.create("simulations/run_1")
 write.csv(spxp_mean_baseline,"simulations/run_1/spxp.csv",row.names = F)
 write.csv(landscape,"simulations/run_1/landscape.csv",row.names = F)
 write.csv(tr,"simulations/run_1/tr.csv",row.names = F)

### 2020-04-29 genetic-diversity/genetic_diversity_figures.R

@@ -11,7 +11,7 @@ png("figures/spatial&temporal_model.png",height=10,width=4,res=400,units="in")
   # plot fecundity as function of temperature for all genotypes
   temp=seq(-7,7,0.01)
   react_norms=getLambdas(temp,fec_Tmu,fec_Tsigma,fec_max)
-  matplot(temp,react_norms,type="l",xlab="Temperature",ylab="Germination rate",col=myCols,lty="solid")
+  matplot(temp,react_norms,type="l",xlab="Temperature",ylab="Fecundity",col=myCols,lty="solid")
   text(fec_Tmu[1]-0.4,0.95*fec_max[1],"AA",col=myCols[1])
   text(fec_Tmu[2],0.95*fec_max[2],"Aa",col=myCols[2])
   text(fec_Tmu[3]+0.4,0.95*fec_max[3],"aa",col=myCols[3])

### 2020-04-29 genetic-diversity/genetic_diversity_master.R

@@ -4,7 +4,8 @@ if (!require("pacman")) install.packages("pacman")
 pacman::p_load(boot, latticeExtra, here)
 
 # set working directory
-here()
+setwd(here()) # set root directory
+setwd("./genetic-diversity")  # move to subdirectory
 
 # load custom functions
 source("genetic_diversity_functions.R")

### 2020-04-29 doc/manuscript.tex
removed red ink

@@
 \noindent \textbf{Abstract:}   Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
-\new{However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change.} Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
+However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change.
+Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
 We use two simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
@@
 Although both time-series and space-for-time approaches are widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics \citep{harris_forecasting_2018,Renwick2018}.
-\new{Such a discussion is overdue given that these two approaches are likely to make very different predictions about ecological responses to future environmental change.
-For example, average forage production in the shortgrass steppe increases rapidly with increasing mean annual precipitation, but production is much less sensitive to interannual variation in precipitation at any one location \citep{lauenroth_long-term_1992}.
+Such a discussion is overdue given that these two approaches are likely to make very different predictions about ecological responses to future environmental change.
+For example, average forage production in the U.S.
+shortgrass steppe increases rapidly with increasing mean annual precipitation, but production is much less sensitive to interannual variation in precipitation at any one location \citep{lauenroth_long-term_1992}.
 The space-for-time approach would predict large changes in production in response to a future change in precipitation, while the time-series approach would predict smaller changes.
@@
 Not only are abundances typically more sensitive to spatial variation in the covariates, but in many cases the temporal response takes the opposite sign \citep{Oedekoven2017}.
-Whenever the sign or magnitude of relationships based on time-series and space-for-time approaches differ, so will the resulting forecasts.}
+Whenever the sign or magnitude of relationships based on time-series and space-for-time approaches differ, so will the resulting forecasts.
 
@@
 
-Here we use two simulated case studies to \new{1) demonstrate why time-series and space-for-time approaches can make different predictions, 2) propose} that the best model-building approaches for ecological forecasting
+Here we use two simulated case studies to 1) demonstrate why time-series and space-for-time approaches can make different predictions, 2) propose that the best model-building approaches for ecological forecasting
 may depend on the time horizon of the forecast, and 3) explore how time-series and space-for-time approaches might be 
@@
 We provide a more detailed description of the simulation model in SI Appendix \ref{metacomm}.
-\new{Parameter values for the simulations described here in the main text are shown in Table \ref{tab:comm-turn-pars}.}
+Parameter values for the simulations described here in the main text are shown in Table \ref{tab:comm-turn-pars}.
 
@@
 We focused on the biomass dynamics of one focal species that dominated the central site during the baseline period.
-\new{We report results from one simulation run; results were qualitatively consistent for replicate runs.}
+We report results from one simulation run; results were qualitatively consistent for replicate runs.
 
@@
 
-\new{When we repeated the simulation with a continuous, nonstationary temperature increase, we see a qualitatively similar shift in weights with increasing forecast horizon from the temporal to spatial statistical model (SI Fig.
+When we repeated the simulation with a continuous, nonstationary temperature increase, we see a qualitatively similar shift in weights with increasing forecast horizon from the temporal to spatial statistical model (SI Fig.
 \ref{fig:community-forecast-species-nonstationary}).
@@
 Separating the effect of annual temperature deviations from the temperature trend would distinguish between short and long-term patterns, much as our temporal and spatial statistical models do in the simulation with stationary temperature periods.
-}
 
@@
 \ref{fig:evo_evo_spatial_temporal_models}B).
-The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}, and \new{parameter values for simulations described here are shown in Table \ref{tab:eco-evo-pars}.
-As in the first case study, we report results from just one simulation run, but results were qualitatively consistent for replicate runs.}
+The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}, and parameter values for simulations described here are shown in Table \ref{tab:eco-evo-pars}.
+As in the first case study, we report results from just one simulation run, but results were qualitatively consistent for replicate runs.
 
@@
 Ecological forecasts are typically made using either a space-for-time substitution approach based on models fit to spatial data or using dynamic models fit to time-series data.
-\new{Empirical studies show that the environment-response relationships detected by these approaches frequently differ in magnitude and even sign \citep{lauenroth_long-term_1992,Oedekoven2017,Amburgey2018,kleinhesselink_response_2018}.
-Our simulations illustrate how such differences may arise and then lead to} very different predictions about the future state of ecological systems.
-Which approach provides the most accurate forecasts \new{likely} depends on the forecast-horizon.
+Empirical studies show that the environment-response relationships detected by these approaches frequently differ in magnitude and even sign \citep{lauenroth_long-term_1992,Oedekoven2017,Amburgey2018,kleinhesselink_response_2018}.
+Our simulations illustrate how such differences may arise and then lead to very different predictions about the future state of ecological systems.
+Which approach provides the most accurate forecasts likely depends on the forecast-horizon.
 In our simulations, time-series approaches performed best for short forecast horizons, whereas models based on spatial data made more accurate forecasts at long horizons.
@@
 
-\new{The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to phenomenological models.
+The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to phenomenological models.
 The same concept could work for any class of models that differ in predictive skill at different forecast horizons.
-Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.}
+Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.
 
@@
 
-Our \new{simulation studies} have important implications for the emerging field of ecological forecasting.
+Our simulation studies have important implications for the emerging field of ecological forecasting.
 First, they suggest that evaluating model performance at both short and long forecast horizons will be essential as research on forecasting methods accelerates.
@@
 
-\new{To assign species-specific parameter values, the number of species in the metacommunity is specified.
+To assign species-specific parameter values, the number of species in the metacommunity is specified.
 Next, each species is assigned an optimal temperature within a specified temperature range by drawing from a uniform distribution.
@@
 Calculations are performed in the script \texttt{SpeciesPoolGen.R}.
-}
 
@@
 	\caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-\new{In contrast to Fig.
-\ref{fig:community-forecast-species}, which shows results for a period of warming followed by stationary temperatures, for this simulation we spread the same temperature increase out over the entire simulation with no stationary periods.} (B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
+In contrast to Fig.
+\ref{fig:community-forecast-species}, which shows results for a period of warming followed by stationary temperatures, for this simulation we spread the same temperature increase out over the entire simulation with no stationary periods.
+(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
 (C) Simulated biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).

### 2020-04-29 doc/response_2_reviews_Ecography1.txt
revising response letter

@@
 RESPONSE: Our response to Reviewer 1's similar concerns addresses many of these points.
-We originally considered separate Methods and Results sections, but found that structure awkward given our two case studies.
-We could add Methods and Results subheaders within each of the case study sections if the editor and reviewers think that would be helpful.
+We implemented the suggestion for separate Methods and Results sections by creating separate Methods and Results subsections within each of the two case studies.
 

### 2020-04-30 doc/response_2_reviews_Ecography1.txt
final small revisions

@@
+We have completed our revision, and we thank the editor and reviewers for suggestions which (we think) have strengthened our paper.
+Our point-by-point responses are below; all caps "RESPONSE" indicates the start of each.
+
 Recommendation by the Subject Editor (anonymous):
@@
 
-RESPONSE: Thank you for the opportunity to resubmit.  We have carefully addressed all of the reviewer's comments, making sure to add sufficient detail so that our results can be reproduced.
+RESPONSE: 
+Thank you for the opportunity to resubmit.  We have carefully addressed all of the reviewer's comments, making sure to add sufficient detail so that our results can be reproduced.
 
@@
 
-RESPONSE: We have rewritten the second half of the abstract to make it clear that all results come from two simple simulation studies.
+RESPONSE: 
+We have rewritten the second half of the abstract to make it clear that all results come from two simple simulation studies.
 We also looked closely at the text in the Discussion to make sure we are not overselling our results.
@@
 
-RESPONSE: In our experience, "space-for-time" is a common term.
+RESPONSE: 
+In our experience, "space-for-time" is a common term.
 A quick search on Google Scholar shows many papers with "space-for-time" in the title or abstract, including papers cited hundreds of times.
 Stuart Pickett's 1989 book chapter in the Long-term Studies in Ecology, titled, "Space-for-Time Substitution as an Alternative to Long-Term Studies," has been cited almost 1300 times (Google Scholar).
-"Species distribution model" is a narrower term, which does encompass other response variable, such as primary production.
+"Species distribution model" is a narrower term, which does not encompass other response variable, such as primary production.
 We wanted it to be clear that space-for-time approaches are not limited to species distribution models.
@@
 
-RESPONSE: We replaced "correlated" with regressed. Details on the regression models are provided in Appendix A.
+RESPONSE: 
+We replaced "correlated" with "regressed." Details on the regression models are provided in Appendix A.
 
@@
 
-RESPONSE: Our response to Reviewer 1's similar concerns addresses many of these points.
+RESPONSE: 
+Our response to Reviewer 1's similar concerns addresses many of these points.
 We implemented the suggestion for separate Methods and Results sections by creating separate Methods and Results subsections within each of the two case studies.
@@
 
-RESPONSE: Interesting question!
+RESPONSE: 
+Interesting question!
 We modified our code for the community turnover example to study the situation in which temperature increases through the entire simulation, with no stationary period for model fitting.
@@
 Qualitatively, the result is consistent: at short forecast horizons the temporal statistical model makes a better forecast while at longer forecast horizons the spatial model does better.
-However, as we explain in the main text (LINES), the forecast from the temporal model is not as accurate, even at short forecast horizons, as in the original simulation with a stationary period for model fitting.
-We believe this reflects that fact that the temporal statistical model does not account for the long-term trend.
+However, as we explain in the main text (lines 260-264), the forecast from the temporal model is not as accurate, even at short forecast horizons, as in the original simulation with a stationary period for model fitting.
+We believe this reflects that fact that the temporal statistical model does not account for a long-term trend.
 Doing so might recover the same kind of information about slow processes contained in the spatial statistical model.
@@
 
-RESPONSE: We made the following changes to the code to make it more portable: 1) We removed rm(list=ls()).
+RESPONSE: 
+We made the following changes to the code to make it more portable: 1) We removed rm(list=ls()).
 2) We now use the "here" package to set the working directory.

### 2020-04-30 doc/manuscript.tex
final small revisions

@@
 Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
-We use two simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
-In the short-term, dynamics reflect initial conditions and fast processes such as birth and death, and the time-series approach makes the best predictions.
-In the long-term, dynamics reflect the additional influence of slower processes such as evolutionary and ecological selection, colonization and extinction, which the space-for-time approach can effectively capture.
-At intermediate time-scales, a weighted average of the two approaches shows promise.
-However, making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
+We use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
+First, we simulated population or community dynamics under stationary temperature using two simple, mechanistic models.
+Second, we fit statistical models to the simulated data using a time-series approach, a space-for-time approach, or a weighted average.
+We then forecast the response to a temperature increase using the statistical models, and compared these forecasts to temperature effects simulated by the mechanistic models.
+We found that the time-series approach made accurate short-term predictions because it captured initial conditions and effects of fast processes such as birth and death.
+The space-for-time approach made more accurate long-term predictions because it better captured the influence of slower processes such as evolutionary and ecological selection.
+The weighted average made accurate predictions at all time scales, including intermediate time-scales where the other two approaches performed poorly.
+A weighted average of time-series and space-for-time approaches shows promise, but making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
 
@@
 We made forecasts based on two phenomenological statistical models, each representing processes operating at different time scales.
-One statistical model represents the time-series or ``temporal approach.'' We regressed interannual variation in an ecological response with interannual variation in temperature at just one site.
+One statistical model represents the time-series or ``temporal approach.'' We regressed interannual variation in an ecological response on interannual variation in temperature at just one site.
 The other statistical model relies on a space-for-time substitution, which we call the ``spatial approach'' for brevity.
@@
 We provide a more detailed description of the simulation model in SI Appendix \ref{metacomm}.
-Parameter values for the simulations described here in the main text are shown in Table \ref{tab:comm-turn-pars}.
 
@@
 We focused on the biomass dynamics of one focal species that dominated the central site during the baseline period.
+Parameter values for the simulations described in the main text are shown in Table \ref{tab:comm-turn-pars}.
 We report results from one simulation run; results were qualitatively consistent for replicate runs.

### 2020-06-05 community-turnover/comm_turn_expts.R
working on replicate runs

@@ -1,9 +1,11 @@
 # Call from comm_turn_master.R. 
 # Assumes many objects already created.
+
 # Completes a couple more runs at higher dispersal rates
 # and compares rate of change in model weights for the focal species.
-
 # Uses existing species pool
+
+# Also runs one set of parameters ten times to check run-to-run variation
 
 ###
 ### 1. Increase dispersal rates
@@ -40,9 +42,53 @@ for(iRun in 2:length(d_vals)){
   
 }
 
+# write results to file
+colnames(saved_weights_d) <- d_vals
+write.csv(saved_weights_d,"simulations/saved_weights.csv",row.names=F)
+
 ###
-### 2. Increase strength of selection by increasing sensitivity to interspp competition
+### 2. Do ten replicate runs to check replicability
 ###
+
+d_baseline <- 0.01   # this should match value specified on comm_turn_master.R
+d_vals <- rep(d_baseline,10) # specify additional values of d to try
+saved_weights_d <- matrix(NA, nrow=length(weight_spp),ncol=length(d_vals)) # matrix to store fitted model weights
+
+for(iRun in 1:length(d_vals)){
+  
+  d <- d_vals[iRun]
+  
+  # dispersal matrix (seeds disperse only to nearest neighbor locations)
+  d <- d/dt
+  seed_rain <- as.matrix(dist(landscape[,1:2],diag=T,upper=T))
+  seed_rain[seed_rain > 1] <- 0
+  seed_rain <- seed_rain*(d*dt*0.5)
+  diag(seed_rain) <- 1-d*dt
+  seed_rain[1,1] = 1-dt*d/2  
+  seed_rain[L_land,L_land] = 1-dt*d/2
+  
+  # Initialize community
+  
+  # Simulate community
+  for(t in 2:sim_yrs){
+    yr_temp <- temperature[t,]
+    spxp[t,,] <- CommunityTempDis(spxp[t-1,,], tr, seed_rain, yr_temp, dt)
+  } 
+  spxp_mean_baseline <- apply(spxp[burnin_yrs:(burnin_yrs+baseline_yrs),,],MARGIN = c(2,3),FUN=mean)
+  
+  source("comm_turn_forecast.R")
+  
+  saved_weights_d[,iRun] <- weight_spp
+  
+}
+
+# write results to file
+colnames(saved_weights_d) <- d_vals
+write.csv(saved_weights_d,"simulations/saved_weights_reps.csv",row.names=F)
+
+###
[…]

### 2020-06-05 genetic-diversity/final_figure.R

@@ -2,7 +2,7 @@
 # call this script from genetic_diversity_master.R
 
 
-# read in weights from commuynity turnover example
+# read in weights from community turnover example
 saved_weights_comm <- read.csv("./../community-turnover/simulations/saved_weights.csv",header=T)
 d_vals <- colnames(saved_weights_comm)
 d_vals <- gsub("X","",d_vals)

### 2020-06-05 community-turnover/comm_turn_expts.R
finished rep runs fig

@@ -58,8 +58,16 @@ for(iRun in 1:length(d_vals)){
   
   d <- d_vals[iRun]
   
-  # dispersal matrix (seeds disperse only to nearest neighbor locations)
+  # 1st step: initialisation of the metacommunity
+  spxp <- array(NA, dim=c(sim_yrs,L_land,N ))
+  spxp[1,,] <- 2/N
+  dt <- 0.025   
   d <- d/dt
+  
+  # Generation of the species pool
+  tr <- SpeciesPoolGen(N, Tmin, Tmax+deltaT, Gmin, Gmax, Lmin, Lmax, Cmin, Cmax)
+  
+  # dispersal matrix (seeds disperse only to nearest neighbor locations)
   seed_rain <- as.matrix(dist(landscape[,1:2],diag=T,upper=T))
   seed_rain[seed_rain > 1] <- 0
   seed_rain <- seed_rain*(d*dt*0.5)
@@ -67,17 +75,18 @@ for(iRun in 1:length(d_vals)){
   seed_rain[1,1] = 1-dt*d/2  
   seed_rain[L_land,L_land] = 1-dt*d/2
   
-  # Initialize community
-  
-  # Simulate community
+  # Loop through time
   for(t in 2:sim_yrs){
     yr_temp <- temperature[t,]
     spxp[t,,] <- CommunityTempDis(spxp[t-1,,], tr, seed_rain, yr_temp, dt)
   } 
+  
   spxp_mean_baseline <- apply(spxp[burnin_yrs:(burnin_yrs+baseline_yrs),,],MARGIN = c(2,3),FUN=mean)
   
+  # Generate forecasts
   source("comm_turn_forecast.R")
   
+  # record weights
   saved_weights_d[,iRun] <- weight_spp
   
 }
@@ -90,36 +99,36 @@ write.csv(saved_weights_d,"simulations/saved_weights_reps.csv",row.names=F)
 ### 3. Increase strength of selection by increasing sensitivity to interspp competition
 ### (not reported in manuscript)
 
-# reset dispersal to baseline value
-d <- d_baseline  
-d <- d/dt
-seed_rain <- as.matrix(dist(landscape[,1:2],diag=T,upper=T))
-seed_rain[seed_rain > 1] <- 0
-seed_rain <- seed_rain*(d*dt*0.5)
-diag(seed_rain) <- 1-d*dt
-seed_rain[1,1] = 1-dt*d/2  
[…]

### 2020-06-05 genetic-diversity/final_figure.R

@@ -1,6 +1,9 @@
 
 # call this script from genetic_diversity_master.R
 
+###
+### 1. Fig. 6 
+###
 
 # read in weights from community turnover example
 saved_weights_comm <- read.csv("./../community-turnover/simulations/saved_weights.csv",header=T)
@@ -31,3 +34,25 @@ png("figures/dispersal_niche_width.png",height=3.5,width=8.5,res=400,units="in")
 dev.off()
 
 
+###
+### 2. Supp. Fig on variation among replicate runs
+###
+
+# read in weights from community turnover example
+saved_weights_comm <- read.csv("./../community-turnover/simulations/saved_weights_reps.csv",header=T)
+
+png("figures/replicates_weights.png",height=3.5,width=8.5,res=400,units="in")
+
+par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,2,1))
+
+#community example
+matplot(saved_weights_comm,type="l",lwd=2,lty=1,col="black",
+        xlab="Time step",ylab="Weight")
+mtext("(A)",side=3,line=0.5,adj=0)
+
+#eco-evo example
+matplot(saved_weights_reps,type="l",lwd=2,lty=1,col="black",
+        xlab="Time step",ylab="Weight")
+mtext("(B)",side=3,line=0.5,adj=0)
+
+dev.off()

### 2020-06-05 genetic-diversity/genetic_diversity_expt.R

@@ -1,8 +1,10 @@
 
 # Call from genetic_diversity_master.R, with all objects in memory
 
-# See how rate of shift in models weights responds to
-# an increased temperature niche of each genotype.
+
+###
+### 1. See how rate of shift in models weights responds to
+###  an increased temperature niche of each genotype.
 
 fec_Tsigma_baseline <- fec_Tsigma # baseline temperature niche widths
 delta_fec_sigma <- c(0.5,1,1.5)  
@@ -39,4 +41,43 @@ for(iRun in 1:length(delta_fec_sigma)){
   saved_weights[,iRun] <- weight
   
 }
+
+###
+### 2. Do replicates of the run featured in main text
+###
+reps <- 10
+saved_weights_reps <- matrix(NA,nrow=length(weight),ncol=reps)
+fec_Tsigma <- fec_Tsigma_baseline
+
+for(iRun in 1:reps){
+  
+  # simulate baseline populations across temperature gradient
+  sim_yrs<-2000
+  Tmean = seq(-5,5,0.25)   # set of mean temperatures to explore
+  Tstdev <- 1   # st dev of temperature
+  temperature<-matrix(NA,sim_yrs,length(Tmean))
+  for(iT in 1:length(Tmean)){
+    temperature[,iT] = rnorm(sim_yrs,Tmean[iT],Tstdev)  # generate temperature time series
+  } 
+  source("genetic_diversity_sim_spatial.R")
+  
+  # simulate long-term run for focal population
+  # baseline_yrs<-500  # number of yrs at baseline temperature
+  # warming_yrs <- 100  # number of yrs with warming occurring
+  # final_yrs <- 300  # number of yrs at steadty-state, warmed temperature
+  sim_yrs <- baseline_yrs + warming_yrs + final_yrs # total number of years
+  Tmean <- baseT + c(rep(0,baseline_yrs),seq(deltaT/warming_yrs,deltaT,deltaT/warming_yrs),
+                     rep(deltaT,final_yrs))
+  temperature = rnorm(sim_yrs,Tmean,Tstdev)  # generate temperature time series
+  source("genetic_diversity_sim_temporal.R")
+  
+  # generate forecasts
+  source("genetic_diversity_forecast.R")
+  
+  saved_weights_reps[,iRun] <- weight
+  
+}
+
+colnames(saved_weights_reps)<-paste0("rep",1:reps)
+write.csv(saved_weights_reps,"saved_weights_reps.csv",row.names = F)
 

### 2020-06-08 doc/response_2_reviews_Ecography2.txt
revisions round 2

Recommendation by the Subject Editor (anonymous):

Both reviewers found the manuscript to be much improved from the previous version, I fully agree with them. While reviewer 1 has only minor comments, reviewer 2 still has some more major concerns that need to be addressed. Thus I invite you to prepare a revision which implements those suggestions.

Reviewer(s)' Comments to Author:
Reviewer: 1

Comments to the Author
The manuscript has improved since its previous round. The work is interesting and carefully executed even though authors should be a bit more critical on the generality of their results. After all this is a very theoretical study where the simulation model is rather simplistic description of the reality.

ADD CAVEATS?

Few specific points:
- space-for-time model. I don't understand the motivation behind this term. Moreover, I don't think this is a commonly used term in ecology. I would encourage revising it.

RESPONSE: We responded to this comment in our first revision, pointing out that "a quick search on Google Scholar shows many papers with "space-for-time" in the title or abstract, including papers cited hundreds of times. Stuart Pickett's 1989 book chapter in the Long-term Studies in Ecology, titled, "Space-for-Time Substitution as an Alternative to Long-Term Studies," has been cited almost 1300 times (Google Scholar

- On line 254, figures and elsewhere. Please provide units (year?) for time

RESPONSE: We added text to this line, and to most figure legends, to clarify that time is measured in years.

Reviewer: 2

Comments to the Author
I reviewed this previously, and asked that more details be included about the methods. Whilst there is some improvement, it is still not up to the standard I would hope for. For example, the text still does not say which years are used in the estimation. It is now clear that only a single simulation is used, which is not really acceptable. We need to be shown (not just told) that the results replicate. being able to replicate the simulations and analysis would also make it easier to run the models under different conditions. This would enable the exploration of a wider range of conditions, which should help to explain the main results.

I would recommend finding a single statistic that can be used to summarise the predictive success, e.g. the probability of the data given the predictions. This can be calculated across some or all of the sites (rather than having to chose a single site to look at). So, for a single simulation, data can be taken at time t, with spatial data across the whole region, and the same amount of temporal data (from either one or several sites). The predictions can be compared to the "real" data at (say) 10 or 1000 years into the future. This basic set-up can be replicated repeatedly under different scenarios, e.g. no climate change, climate change before the sampling, climate change after the sampling, etc.

RESPONSE: We agree, "showing" is better than "telling." We chose not to take the reviewer's suggestion of finding a single measure of model accuracy, because we want readers to focus on the most important general result, which is the shift in model weight from time-series to space-for-time with increasing forecast time scale. We have added a new supplementary figure (SI Fig. 1) showing the temporal trend in the model weight term for 10 independent realizations of the community tunover example and the eco-evolutionary example. We believe this figure shows how consistent the pattern is across simulation runs.

One issue that I think needs to be addressed is the problem of extrapolation outside of the environmental conditions when the sampling is done. In practice, this is a major concern. If, say, methods work well until they extrapolate, then that would be an important message for studies of real organisms.

RESPONSE: We added text to the discussion to flag the important issue of extrapolation. Third to last discussion paragraph now reads as follow, where \comments{} show the new text. 
"The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to \comment{extremely simple statistical models like the ones we used in this study}. The same concept could work for any class of models that differ in predictive skill at different forecast horizons. \comment{One possibility might be more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation. A set of process-based models focusing on processes operating at different time scales would be another.}  Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales."
We note that the extrapolation issue is also addressed in the Introduction (lines __).

### 2020-06-08 doc/manuscript.tex
revisions round 2

@@
 The truth was represented by a simulation model that was mechanistic for at least one important process, 
-but we treated the data-generating model as unknown when analyzing the data and we assumed that perfectly recovering the mechanisms
+but we treated this data-generating model as unknown when analyzing the data and we assumed that perfectly recovering the mechanisms
 it contains would not be possible in practice. We began each simulation
@@
 Parameter values for the simulations described in the main text are shown in Table \ref{tab:comm-turn-pars}.
-We report results from one simulation run; results were qualitatively consistent for replicate runs.
+We report results from one simulation run; results were qualitatively consistent for replicate runs (\comment{Fig.
+\ref{fig:replicates}A}).
 
@@
 To combine information from the temporal and spatial statistical models into a single prediction, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, $t$, this weighted forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal statistical model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial statistical model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
+At any time point, \comment{or year}, $t$, this weighted forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal statistical model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial statistical model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
 The weighted model accurately predicts the simulated dynamics across the full forecast horizon (Fig.
@@
 The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}, and parameter values for simulations described here are shown in Table \ref{tab:eco-evo-pars}.
-As in the first case study, we report results from just one simulation run, but results were qualitatively consistent for replicate runs.
+As in the first case study, we report results from just one simulation run, but results were qualitatively consistent for replicate runs (\comment{Fig.
+\ref{fig:replicates}B}).
 
@@
 \ref{fig:evo_evo_spatial_temporal_models}A), even though the heat-adapted homozygote might perform optimally at that temperature.
-However, if warmer than normal conditions persist for many years, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
+However, if warmer than normal conditions persist for many \comment{years}, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
 
@@
 
-The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to phenomenological models.
+The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to \comment{extremely simple statistical models like the ones we used in this study}.
 The same concept could work for any class of models that differ in predictive skill at different forecast horizons.
-Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.
+\comment{One possibility might be more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation.
+A set of process-based models focusing on processes operating at different time scales would be another.}  Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.
 
@@
 (C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).
-Year 1000 in each panel corresponds to the start of the temperature increase.
+Time 1000 (years) in each panel corresponds to the start of the temperature increase.
 }
@@
 	\caption{The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
-Year 0 in these figures corresponds to the start of the temperature increase.
+Time 0 (years) in these figures corresponds to the start of the temperature increase.
 }
@@
 
-\renewcommand{\thetable}{S-\arabic{table}}
+\renewcommand{\thetable}{SM-\arabic{table}}
 
@@
 
-\renewcommand{\thefigure}{S-\arabic{figure}}
+\renewcommand{\thefigure}{SM-\arabic{figure}}
 
-\clearpage
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1 \textwidth] {replicates_weights.png}
+	\caption{(A) Temporal shifts in the model weighting term for 10 independent simulations of (A) the community turnover model, and (B) the eco-evolutionary model.
+For the community turnover model, each simulation began with initialization of a new regional species pool.
+For the eco-evolutionary model, genotype parameters were fixed, and only the sequence of annual temperatures varied between runs.
+In all cases, the combined forecast is heavily weighted towards the time-series model at short forecast time scales, and towards the space-for-time model at long forecast time scales.
+}
+	\label{fig:replicates}
+\end{figure}
 
@@
 (C) Simulated biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).
-Year 1000 in each panel corresponds to the start of the temperature increase.
+Time 1000 (years) in each panel corresponds to the start of the temperature increase.
 }
@@
 \includegraphics[width=0.7 \textwidth] {community_models_total.png}
-\caption{(Results for total biomass from the community turnover model.
+\caption{Results for total biomass from the community turnover model.
 Blue points show mean total biomass during the baseline period at locations across the temperature gradient, and the blue line shows predictions from the spatial model.
@@
 (C) Simulated changes in biomass of all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).
-Year 1000 in this figure corresponds to the start of the temperature increase.
+Time 1000 (years) in this figure corresponds to the start of the temperature increase.
 }
@@
 \caption{Simulated shifts in genotype abundances, and the model weighting term, $\omega$, during the warming phase and the following stationary temperature phase.
-Year 0 in this figure corresponds to the start of the temperature increase.}
+Time 0 (years) in this figure corresponds to the start of the temperature increase.}
 \label{fig:forecast_supp}

### 2020-06-08 genetic-diversity/final_figure.R

@@ -19,14 +19,14 @@ png("figures/dispersal_niche_width.png",height=3.5,width=8.5,res=400,units="in")
   
   #community example
   matplot(saved_weights_comm,type="l",lwd=2,lty=1,col=myCols,
-          xlab="Time step",ylab="Weight")
+          xlab="Time",ylab="Weight")
   legend("topright",legend=d_vals,
        title="dispersal",lwd=2,lty=1,col=myCols,bty="n")
   mtext("(A)",side=3,line=0.5,adj=0)
   
   #eco-evo example
   matplot(saved_weights,type="l",lwd=2,lty=1,col=myCols,
-          xlab="Time step",ylab="Weight")
+          xlab="Time",ylab="Weight")
   legend("topright",legend=fec_Tsigma_baseline[1]*delta_fec_sigma,
          title=expression(sigma[T]),lwd=2,lty=1,col=myCols,bty="n")
   mtext("(B)",side=3,line=0.5,adj=0)
@@ -47,12 +47,12 @@ par(mfrow=c(1,2),tcl=-0.2,mgp=c(2,0.5,0),mar=c(3,4,2,1))
 
 #community example
 matplot(saved_weights_comm,type="l",lwd=2,lty=1,col="black",
-        xlab="Time step",ylab="Weight")
+        xlab="Time",ylab="Weight")
 mtext("(A)",side=3,line=0.5,adj=0)
 
 #eco-evo example
 matplot(saved_weights_reps,type="l",lwd=2,lty=1,col="black",
-        xlab="Time step",ylab="Weight")
+        xlab="Time",ylab="Weight")
 mtext("(B)",side=3,line=0.5,adj=0)
 
 dev.off()

### 2020-06-08 doc/response_2_reviews_Ecography2.txt
finished response

@@
 Thus I invite you to prepare a revision which implements those suggestions.
+
+RESPONSE: We thank the Editor for the opportunity to address the reviewers' concerns with another revision/
 
@@
 After all this is a very theoretical study where the simulation model is rather simplistic description of the reality.
+
+RESPONSE: We re-read the Discussion to make sure we are not over-selling our results.
+We believe it is clear that our results are based only on simulations of a couple theoretical models, and that empirical tests are still required.
+Furthermore, the text is littered with qualifying words such as "may", "likely", and "suggest."
 
@@
 
-RESPONSE: We agree, "showing" is better than "telling." We chose not to take the reviewer's suggestion of finding a single measure of model accuracy, because we want readers to focus on the most important general result, which is the shift in model weight from time-series to space-for-time with increasing forecast time scale.
+RESPONSE: We agree that "showing" is better than "telling." We chose not to take the reviewer's suggestion of finding a single measure of model accuracy, because we want readers to focus on the most important general result, which is the shift in model weight from time-series to space-for-time with increasing forecast time scale.
 We have added a new supplementary figure (SI Fig.
@@
 
-RESPONSE: We added text to the discussion to flag the important issue of extrapolation. Third to last discussion paragraph now reads as follow, where \comments{} show the new text. 
+RESPONSE: We added text to the discussion to flag the important issue of extrapolation. The third to last discussion paragraph now reads as follow, where \comments{} show the new text. 
 "The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to \comment{extremely simple statistical models like the ones we used in this study}.
-The same concept could work for any class of models that differ in predictive skill at different forecast horizons.
-\comment{One possibility might be more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation.
-A set of process-based models focusing on processes operating at different time scales would be another.}  Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales."
-We note that the extrapolation issue is also addressed in the Introduction (lines __).
+The same concept could work for any class of models that differ in predictive skill at different forecast horizons, \comment{such as more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation, or a set of process-based models focusing on mechanisms operating at different time scales.}  Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales."
+We note that the extrapolation issue is also addressed in the Introduction (lines 132-135).

### 2020-06-08 doc/manuscript.tex
finished response

@@
 
-Space-for-time modeling approaches for predicting long-term, steady-state outcomes of ecological change have also been tested empirically, primarily via hind-casting.
+Space-for-time modeling approaches for predicting long-term, steady-state outcomes of ecological change have also been tested empirically, primarily via hindcasting.
 Overall, the results are mixed: some tests show reasonable prediction of changes in community composition \citep{Blois2013,Illan2014} or species distributions \citep{Norberg2019}, supporting the hypothesis that datasets spanning spatial gradients capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
 Other attempts to validate predictions from space-for-time models have been discouraging \citep{Worth2014,Illan2014,Davis2014,Brun2016,Veloz2012}, indicating violations of model assumptions or effects of transient dynamics.
-However, predictions from the space-for-time approaches are rarely compared directly to predictions from time-series models (\citealt{harris_forecasting_2018} but see \citealt{Renwick2018}).
+However, predictions from the space-for-time approaches are rarely compared directly to predictions from time-series models (\citealt{harris_forecasting_2018}, but see \citealt{Renwick2018}).
 We need more such comparisons to identify the appropriate modeling approach for different forecast horizons.
@@
 The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to \comment{extremely simple statistical models like the ones we used in this study}.
-The same concept could work for any class of models that differ in predictive skill at different forecast horizons.
-\comment{One possibility might be more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation.
-A set of process-based models focusing on processes operating at different time scales would be another.}  Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.
+The same concept could work for any class of models that differ in predictive skill at different forecast horizons, \comment{such as more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation, or a set of process-based models focusing on mechanisms operating at different time scales.}  Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.
 

### 2020-06-09 doc/response_2_reviews_Ecography2.txt
tiny changes

@@
 
-ADD CAVEATS?
-
 Few specific points:
@@
 RESPONSE: We responded to this comment in our first revision, pointing out that "a quick search on Google Scholar shows many papers with "space-for-time" in the title or abstract, including papers cited hundreds of times.
-Stuart Pickett's 1989 book chapter in the Long-term Studies in Ecology, titled, "Space-for-Time Substitution as an Alternative to Long-Term Studies," has been cited almost 1300 times (Google Scholar
+Stuart Pickett's 1989 book chapter in the Long-term Studies in Ecology, titled, "Space-for-Time Substitution as an Alternative to Long-Term Studies," has been cited almost 1300 times (Google Scholar)."
 
@@
 
-RESPONSE: We added text to this line, and to most figure legends, to clarify that time is measured in years.
+RESPONSE: We added revised this line, and most figure legends, to clarify that time is measured in years.
 

### 2020-06-09 doc/manuscript.tex
tiny changes

@@
 
-When we repeated the simulation with a continuous, nonstationary temperature increase, we see a qualitatively similar shift in weights with increasing forecast horizon from the temporal to spatial statistical model (SI Fig.
+When we repeated the simulation with a continuous, nonstationary temperature increase, we see a qualitatively similar shift in weights with increasing forecast horizon from the temporal to spatial statistical model (Fig.
 \ref{fig:community-forecast-species-nonstationary}).
@@
 \ref{fig:forecast}) accurately predicted the impact of the initial warming trend, but eventually became too pessimistic, while the spatial statistical model (solid red line in Fig.
-\ref{fig:forecast}) did not handle the initial trend but accurately predicted the eventual, new steady state (Fig.
-\ref{fig:forecast} bottom).
+\ref{fig:forecast}) did not handle the initial trend but accurately predicted the eventual, new steady state.
 
-As in the community turnover example, we also fit a weighted average of predictions from the spatial and temporal statistical models, with the weights changing over time.
+As in the community turnover example, we also fit a weighted average of predictions from the spatial and temporal statistical models (purple line in Fig.
+\ref{fig:forecast}), with the weights changing over time.
 This weighted model initially reflected the temporal model (decrease from $t=500$ to $t=600$), but then rapidly transitioned to reflect the spatial model ($t \geq 700$).

### 2020-07-03 doc/manuscript.tex
Round 3 revisions

@@
 focuses on how interspecific interactions affect the population dynamics of a focal species, and the second focuses on
-an eco-evolutionary scenario. Our analyses show that: 
+an eco-evolutionary scenario. Our \comment{simulations illustrate} that: 
 \begin{enumerate}
-	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require accounting for the influence of slow processes such as evolutionary and ecological selection as well as dispersal.
+	\item For short-term forecasts, phenomenological time-series approaches \comment{may be} hard to beat, whereas longer-term forecasts \comment{may} require accounting for the influence of slow processes such as evolutionary and ecological selection as well as dispersal.
 	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and fast responses of current assemblages to interannual environmental variation, while data spanning spatial gradients capture the long-term outcome of interactions between fast and slow processes.
-Whether predictive models should be trained using longitudinal or spatial data sets, or both, depends on the time-scale of the desired forecast.
+Whether predictive models should be trained using longitudinal or spatial data sets, or both, \comment{likely} depends on the time-scale of the desired forecast.
 	\item A key challenge for future research is determining the rate at which slow processes begin to influence dynamics.
@@
 Parameter values for the simulations described in the main text are shown in Table \ref{tab:comm-turn-pars}.
-We report results from one simulation run; results were qualitatively consistent for replicate runs (\comment{Fig.
-\ref{fig:replicates}A}).
+We report results from one simulation run; results were qualitatively consistent for replicate runs (Fig.
+\ref{fig:replicates}A).
 
@@
 To combine information from the temporal and spatial statistical models into a single prediction, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, \comment{or year}, $t$, this weighted forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal statistical model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial statistical model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
+At any time point, or year, $t$, this weighted forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal statistical model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial statistical model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
 The weighted model accurately predicts the simulated dynamics across the full forecast horizon (Fig.
@@
 The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}, and parameter values for simulations described here are shown in Table \ref{tab:eco-evo-pars}.
-As in the first case study, we report results from just one simulation run, but results were qualitatively consistent for replicate runs (\comment{Fig.
-\ref{fig:replicates}B}).
+As in the first case study, we report results from just one simulation run, but results were qualitatively consistent for replicate runs (Fig.
+\ref{fig:replicates}B).
 
@@
 \ref{fig:evo_evo_spatial_temporal_models}A), even though the heat-adapted homozygote might perform optimally at that temperature.
-However, if warmer than normal conditions persist for many \comment{years}, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
+However, if warmer than normal conditions persist for many years, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
 
@@
 
-The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to \comment{extremely simple statistical models like the ones we used in this study}.
-The same concept could work for any class of models that differ in predictive skill at different forecast horizons, \comment{such as more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation, or a set of process-based models focusing on mechanisms operating at different time scales.}  Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.
+The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to extremely simple statistical models like the ones we used in this study.
+The same concept could work for any class of models that differ in predictive skill at different forecast horizons, such as more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation, or a set of process-based models focusing on mechanisms operating at different time scales.
+Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.
+
+\comment{On the other hand, there is no guarantee that our proposed model weighting scheme will work when applied in real ecosystems.
+The most obvious potential problem is that space-for-time approaches may fail to predict long-term dynamics if model assumptions are violated, transient dynamics are strong, or future environmental conditions have no current analog \citep{Worth2014,Veloz2012}.
+The notion that model weighting can improve forecasts at intermediate and long forecast horizons must be a viewed as a hypothesis to be tested with empirical data.}
 
@@
 \end{equation}
-In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression is 
+In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression, \comment{which includes a quadratic effect of temperature,} is 
 \begin{equation}

### 2020-07-03 community-turnover/comm_turn_forecast.R

@@ -2,6 +2,63 @@
 # call this script from comm_turn_master.R after running the previous scripts
 
 ###
+###  Model biomass of the dominant species 
+###
+
+site <- L_land/2   # focal site
+
+my_species <- which(spxp_mean_baseline[site,]==max(spxp_mean_baseline[site,]))
+
+# spatial model
+x_spatial <- Tmean
+y_spatial_spp <- spxp_mean_baseline[,my_species]
+spatial_model_spp <- lm(y_spatial_spp~x_spatial+I(x_spatial^2))
+
+# temporal model
+x_temporal <- temperature[burnin_yrs:(burnin_yrs+baseline_yrs),site]
+y_temporal_spp <- spxp[burnin_yrs:(burnin_yrs+baseline_yrs),site,my_species]
+y_lag_spp <- spxp[(burnin_yrs-1):(burnin_yrs+baseline_yrs-1),site,my_species]
+temporal_model_spp <- lm(y_temporal_spp ~ y_lag_spp + x_temporal)
+
+# get mean temperatures for forecast period (after baseline_yrs)
+forecast_Tmean <- Tmean[site] + c(seq(deltaT/warming_yrs,deltaT,deltaT/warming_yrs),
+                                  rep(deltaT,final_yrs))
+
+# make predictions from spatial model
+spatial_forecast_spp <- coef(spatial_model_spp)[1] + coef(spatial_model_spp)[2]*forecast_Tmean +
+                            coef(spatial_model_spp)[3]*forecast_Tmean^2
+
+# make predictions from temporal model
+temporal_forecast_spp <- rep(NA,length(forecast_Tmean)+1)
+temporal_forecast_spp[1] <- y_temporal_spp[length(y_temporal_spp)] # initialize at biomass from last baseline year
+for(iTime in 1:length(forecast_Tmean)){
+  temporal_forecast_spp[iTime+1] <- coef(temporal_model_spp)[1] +coef(temporal_model_spp)[2]*temporal_forecast_spp[iTime] + 
+    coef(temporal_model_spp)[3]*forecast_Tmean[iTime]
+}
+temporal_forecast_spp <- temporal_forecast_spp[2:length(temporal_forecast_spp)]  # drop first ("observed") year
+
+# fit combined model
+y <- spxp[(burnin_yrs+baseline_yrs+1):sim_yrs,site,my_species]
[…]

### 2020-07-07 doc/response_2_reviews_Ecography3.txt
final accepted version

03-Jul-2020

Dear Dr. Adler:

Manuscript ID ECOG-05271.R1 entitled "Matching the forecast horizon with the relevant spatial and temporal processes and data sources" which you submitted to Ecography, has been reviewed.  The comments of the reviewer(s) and recommendation by the Subject Editor are included at the bottom of this letter.

The reviewer(s) and the Subject Editor have recommended some minor revisions to your manuscript.  Therefore, I invite you to respond to the reviewer(s)' comments and revise your manuscript.

Recommendation by the Subject Editor (anonymous):
Your manuscript was evaluated by the reviewer who was the more critical reviewer of the original submission (the more positive reviewer was satisfied already with the revisions made earlier).

This was not an easy decision to make: reviewer 2 recommends rejection, because the same fundamental issues remain that were present basically in the original submission. Based on my own reading, I agree with the reviewer that your study has substantial limitations. At the same time, based on my own reading, I consider that the manuscript makes a very important point, and that it is generally very nicely written. Thus, in my view the models presented are simply "proofs of concepts" that help to discuss the (quite obvious but yet maybe not sufficiently acknowledged) fact that time-series models should do better in short-term predictions and space-for-time models should not better in long-term predictions. Given that making the examples more realistic and general would lead to a completely new manuscript, my options were basically to reject the manuscript, or accept it with these limitations. Given the above, I ended up to a decision that was favorable for your side. Thus, I invite you to prepare a minor revision, where you make one more effort to double-check that you don't overstate the generality of the findings, but rather present your models as proofs of concepts that help to illustrate your verbal argument. Related to this, I agree with the reviewer that you may have overstated the value of the simplistic linear weighting approach (even if you discuss its limitations nicely in the main text), as based on your text some readers might think that "just by figuring out the weighting functions one can make accurate predictions for any times-scale". This is of course not expected to be the case for any real case study, even if the optimal weighting would be known. So I would propose that you discuss even more critically the prospects of the weighted-model approach in the revised version.

RESPONSE: We thank the editor for their favorable decision, and for the opportunity to revise. We have addressed the reasonable concern about the potential value of the model weighting approach by adding a new, short paragraph to the Discussion: "On the other hand, there is no guarantee that our proposed model weighting scheme will work when applied in real ecosystems. The most obvious potential problem is that space-for-time approaches may fail to predict long-term dynamics if model assumptions are violated, transient dynamics are strong, or future environmental conditions have no current analog \citep{Worth2014,Veloz2012}. The notion that model weighting can improve forecasts at intermediate and long forecast horizons must be a viewed as a hypothesis to be tested with empirical data."

RESPONSE CONT'D: We also added caveats to the section at the end of the Introduction where we briefly describe our findings (changes tracked in red font).

Reviewer(s)' Comments to Author:
Reviewer: 2

Comments to the Author
This is a resubmission. It still suffers from the same problem as the previous version - the authors are not willing to explore the methods fully. The authors state they find the weighting interesting, but even this is only explored under one set of conditions. How do we know it will act the same if (for example) there is no change in climate?

RESPONSE: If there is no change in climate, there will be no need for model weighting. As our results show, the time series approach will work fine on its own.  Pointing this out seems unnecessary because the premise of the whole exercise was to predict the impact of environmental change.

I don't agree that the shift in weighting is so important, and I can't see that it is general. Only one specific scenario is used, so what will happen under different scenarios? And what if different models were used for the projections, e.g. adding a quadratic term?

RESPONSE: We are sorry the reviewer sees so little value in the analysis. However, it is not strictly accurate to say that we only explored one scenario. We studied two different model systems, and, at the reviewers' request, we also explored a non-stationary scenario. Perhaps the reviewer believes we are proposing a solution to a problem. Our real objective is to highlight a previously under-appreciated problem and propose one potential, and novel, solution. We believe we have added enough caveats and qualifiers to make this clear.

RESPONSE: As for the comment about adding a quadratic term, the spatial models for both the metacommunity example and the eco-evo example do include a quadratic effect of temperature (Eq. 1 in Appendix A, also shown in Figs. 2A and 4B). While the temporal model for the metacommunity example does not contain a quadratic effect of temperature (Appendix A Eq. 2, and Fig. 2B), the temporal model for the eco-evo example does (Appendix A Eq. 3). We added text in Appendix A to clarify this distinction. The reason the temporal model for the metacommunity example does not include a quadratic term is that the data from the baseline period show a linear relationship between temperature and biomass (Fig. 2B). Even if we fit a quadratic term, the coefficient would be close to zero and would not improve long-term forecasts. This illustrates a key point of the paper: different kinds of data represent different processes operating at different time scales.

A minor point: check that the descriptions of the models match the code. I was wondering why there isn't  a quadratic term on K(t) in the temporal model for biomass, but according to the code there isn't one on the spatial model either.

RESPONSE: We address this comment in the paragraph above. We double-checked our computer code and it is correct. The reviewer may have been looking at the spatial model for total biomass in the community turnover example, rather than the biomass of the focal species. That model does not include a quadratic term because the relationship is monotonic increasing. To minimize the chance of this confusion, we moved this block of code to the end of the script to follow the order in which results are presented.

### 2020-07-07 doc/manuscript_acceptedJuly2020.tex
final accepted version

--- 2020-07-03 doc/manuscript.tex
@@
 
-\noindent \emph{Statement of authorship:} PBA and EPW designed the study, PBA and MHC built the models, and PBA analyzed the models.
-PBA wrote the first draft of the manuscript, and all authors contributed substantially to revisions.
-
-\noindent \emph{Acknowledgements:} We thank Heather Lynch and Juan Manuel Morales for comments that improved early drafts of the paper.
-PBA was supported by the National Science Foundation (DEB-1927282) and the Utah Agriculture Experiment Station (grant 1322).
-EPW was supported by the Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative (GBMF4563).
-
 %\noindent \emph{Running title:} Forecast horizon and ecological process
@@
 Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
-We use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
-First, we simulated population or community dynamics under stationary temperature using two simple, mechanistic models.
-Second, we fit statistical models to the simulated data using a time-series approach, a space-for-time approach, or a weighted average.
-We then forecast the response to a temperature increase using the statistical models, and compared these forecasts to temperature effects simulated by the mechanistic models.
-We found that the time-series approach made accurate short-term predictions because it captured initial conditions and effects of fast processes such as birth and death.
-The space-for-time approach made more accurate long-term predictions because it better captured the influence of slower processes such as evolutionary and ecological selection.
-The weighted average made accurate predictions at all time scales, including intermediate time-scales where the other two approaches performed poorly.
-A weighted average of time-series and space-for-time approaches shows promise, but making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
-
-\vspace{0.25cm}
-
-\noindent \emph{Keywords:} dispersal, ecological forecasting, eco-evolutionary dynamics, global change, selection
-
-\newpage
-
-\section*{Abstract} 
-Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
-However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change.
-Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
-We use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
+As proof of concept, we use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
 First, we simulated population or community dynamics under stationary temperature using two simple, mechanistic models.
@@
 focuses on how interspecific interactions affect the population dynamics of a focal species, and the second focuses on
-an eco-evolutionary scenario. Our \comment{simulations illustrate} that: 
+an eco-evolutionary scenario. Our simulations illustrate that: 
 \begin{enumerate}
-	\item For short-term forecasts, phenomenological time-series approaches \comment{may be} hard to beat, whereas longer-term forecasts \comment{may} require accounting for the influence of slow processes such as evolutionary and ecological selection as well as dispersal.
+	\item For short-term forecasts, phenomenological time-series approaches may be hard to beat, whereas longer-term forecasts may require accounting for the influence of slow processes such as evolutionary and ecological selection as well as dispersal.
 	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and fast responses of current assemblages to interannual environmental variation, while data spanning spatial gradients capture the long-term outcome of interactions between fast and slow processes.
-Whether predictive models should be trained using longitudinal or spatial data sets, or both, \comment{likely} depends on the time-scale of the desired forecast.
+Whether predictive models should be trained using longitudinal or spatial data sets, or both, likely depends on the time-scale of the desired forecast.
 	\item A key challenge for future research is determining the rate at which slow processes begin to influence dynamics.
@@
 
-\comment{On the other hand, there is no guarantee that our proposed model weighting scheme will work when applied in real ecosystems.
+On the other hand, there is no guarantee that our proposed model weighting scheme will work when applied in real ecosystems.
 The most obvious potential problem is that space-for-time approaches may fail to predict long-term dynamics if model assumptions are violated, transient dynamics are strong, or future environmental conditions have no current analog \citep{Worth2014,Veloz2012}.
-The notion that model weighting can improve forecasts at intermediate and long forecast horizons must be a viewed as a hypothesis to be tested with empirical data.}
+The notion that model weighting can improve forecasts at intermediate and long forecast horizons must be a viewed as a hypothesis to be tested with empirical data.
 
@@
 
+\vspace{0.25cm}
+
 \subsubsection*{Data accessibility statement:} The manuscript contains no original data.
@@
 Upon acceptance of the manuscript, code will be archived at Zenodo.
+
+\subsubsection*{Statement of authorship:} PBA and EPW designed the study, PBA and MHC built the models, and PBA analyzed the models.
+PBA wrote the first draft of the manuscript, and all authors contributed substantially to revisions.
+
+\subsubsection*{Acknowledgements:} We thank Heather Lynch and Juan Manuel Morales for comments that improved early drafts of the paper.
+PBA was supported by the National Science Foundation (DEB-1927282) and the Utah Agriculture Experiment Station (grant 1322).
+EPW was supported by the Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative (GBMF4563).
 
@@
 \end{equation}
-In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression, \comment{which includes a quadratic effect of temperature,} is 
+In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression, which includes a quadratic effect of temperature, is 
 \begin{equation}

### 2020-07-07 doc/final_files/manuscript.tex
final files setup
final files finished

--- 2020-07-07 doc/manuscript_acceptedJuly2020.tex
@@
 
+% refer to labels in supplement
+\usepackage{xr}
+\externaldocument{supplement}
+
 \usepackage[authoryear]{natbib}
 
-\graphicspath{ {./../genetic-diversity/figures/} {./../community-turnover/figures/} }
+\graphicspath{ {./} }
+\usepackage{caption}
+\captionsetup[figure]{labelsep=space}
 
@@
 Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
-We provide a more detailed description of the simulation model in SI Appendix \ref{metacomm}.
+We provide a more detailed description of the simulation model in Supplementary Material (Appendix \ref{metacomm}).
 
@@
 
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.7 \textwidth] {fast-slow-figure.png}
-	\caption{Fast and slow processes operate at different time scales, and are reflected in different kinds of datasets.
+\subsection*{Figure legends}
+
+\paragraph{Fig.
+1.} Fast and slow processes operate at different time scales, and are reflected in different kinds of datasets.
 Fast processes, such as births, deaths, and individual growth, operate at all time scales, but are the exclusive drivers of the short-term dynamics captured in most time series datasets.
@@
 The time scales shown here were chosen with vascular plants in mind, but the same concepts would apply for much shorter-lived organisms but at shorter time scales.
-}
-	\label{fig:fast-slow-concept}
+
+\paragraph{Fig.
+2.}  Mean biomass by species (colors) across the temperature gradient during the baseline period.
+The focal species, dominant at the site in the center of the gradient (vertical gray line), is shown in dark blue.
+The dashed blue line shows predictions from the spatial statistical model.
+(B) Annual biomass of the focal species at the central site during the baseline period.
+The dashed line shows predictions from the temporal statistical model.
+
+\paragraph{Fig.
+3.}  (A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
+(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
+(C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).
+Time 1000 (years) in each panel corresponds to the start of the temperature increase.
+
+\paragraph{Fig.
+4.}  (A) Reaction norms of the three genotypes.
+(B) The spatial pattern of individual genotypes (colors) and total population abundance (black) at sites arrayed across a gradient of mean annual temperature.
+The dashed black line (almost entirely hidden by the slid black line) shows predictions from an empirical, spatial statistical model, a linear regression that describes mean population size as a function of mean temperature.
+(C) The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
+Colors show population size (the green to brown gradient depicting low to high population density), which influences the population growth rate through density dependence.
+
+\paragraph{Fig.
+5.}  (Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
+(Bottom) Simulated population size and forecasts from the spatial, temporal and weighted statistical models.
+
+\paragraph{Fig.
+6.}  The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
+Time 0 (years) in these figures corresponds to the start of the temperature increase.
+
+\newpage
+
+\subsection*{Figures}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7 \textwidth] {Fig1.png}
+	\caption{}
+	\label{fig:Fig1}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=1 \textwidth] {species_patterns_models.png}
-	\caption{(A) Mean biomass by species (colors) across the temperature gradient during the baseline period.
-The focal species, dominant at the site in the center of the gradient (vertical gray line), is shown in dark blue.
-The dashed blue line shows predictions from the spatial statistical model.
-(B) Annual biomass of the focal species at the central site during the baseline period.
-The dashed line shows predictions from the temporal statistical model.
-}
+	\includegraphics[width=1 \textwidth] {Fig2.png}
+	\caption{}
 	\label{fig:species-patterns-models}
@@
 	\centering
-	\includegraphics[width=0.7 \textwidth] {community_forecast_species.png}
-	\caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
-(C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).
-Time 1000 (years) in each panel corresponds to the start of the temperature increase.
-}
+	\includegraphics[width=0.7 \textwidth] {Fig3.png}
+	\caption{}
 	\label{fig:community-forecast-species}
@@
 	\centering
-	\includegraphics[width=0.4 \textwidth]{spatial&temporal_model.png}
-	\caption{(A) Reaction norms of the three genotypes.
-(B) The spatial pattern of individual genotypes (colors) and total population abundance (black) at sites arrayed across a gradient of mean annual temperature.
-The dashed black line (almost entirely hidden by the slid black line) shows predictions from an empirical, spatial statistical model, a linear regression that describes mean population size as a function of mean temperature.
-(C) The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
-Colors show population size (the green to brown gradient depicting low to high population density), which influences the population growth rate through density dependence.
-}
+	\includegraphics[width=0.4 \textwidth]{Fig4.png}
+	\caption{}
 	\label{fig:evo_evo_spatial_temporal_models}
@@
 	\centering
-	\includegraphics[width=0.7 \textwidth] {forecast.png}
-	\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
-(Bottom) Simulated population size and forecasts from the spatial, temporal and weighted statistical models.
-}
+	\includegraphics[width=0.7 \textwidth] {Fig5.png}
+	\caption{}
 	\label{fig:forecast}
@@
 	\centering
-	\includegraphics[width=0.7 \textwidth] {dispersal_niche_width.png}
-	\caption{The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
-Time 0 (years) in these figures corresponds to the start of the temperature increase.
-}
+	\includegraphics[width=0.7 \textwidth] {Fig6.png}
+	\caption{}
 	\label{fig:dispersal_niche_width}
@@
 
-%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-% APPENDICES !
-%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-\clearpage 
-\newpage
-
-\setcounter{page}{1}
-\setcounter{equation}{0}
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\setcounter{table}{0}
-
-\centerline{\Large \textbf{Appendices}}
-
-\renewcommand{\thesection}{\Alph{section}}
-
-\section{Spatial, temporal and spatial-temporal-weighted models}\label{models}
-
-The two simulation models in the main text describe how population size, $N(x,t)$, at location $x$ changes over time ($t$).
-We assume that the temperature, $K(x,t)$, at each location can vary in time and space.
-To forecast the dynamics generated by these simulations models, we fit a series of statistical models.
-
-The spatial model, which we refer to as $S$, is a quadratic regression of the mean long-term population density at a location ($\bar{N}(x)$) against the mean temperature at that location ($\bar{K}(x)$).
-The quadratic term describes the unimodal relationship between $\bar{N}$ and $\bar{K}$.
-The spatial statistical model is
-\begin{equation}
-\bar{N}(x) = S(\bar{K}(x)) = \beta^S_0 +  \beta^S_1 \bar{K}(x) +\beta^S_2 {\bar{K}(x)}^2 + \varepsilon
-\label{eqn:spatial_regression}
-\end{equation}
-
-The temporal model, which we call $T$, starts with a time-series of ``observed" population sizes, or total biomasses, at one location, $N(t)$, for $t=1...n$ (the spatial index is suppressed because we only focus on one location at a time).
-In the community turnover example, we fit the following regression, which predicts biomass at time $t+1$ as a function of biomass ($N(t)$) and annual temperature ($K(t)$) at time $t$,
-\begin{equation}
-\ln(N(t+1)) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 \ln(N(t)) +\beta^T_2 K(t)  +  \varepsilon
-\label{eqn:temporal_regression_community1}
-\end{equation}
-In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression, which includes a quadratic effect of temperature, is 
-\begin{equation}
-\ln\left(\frac{N(t+1)}{N(t)}\right) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 \ln(N(t)) +\beta^T_2 K(t)  +\beta^T_3 K(t)^2 +  \varepsilon
-\label{eqn:temporal_regression_ecoevo2}
-\end{equation}
-This version of the temporal model returns a per capita growth rate on the log scale.
-To predict population size at the next time step, we exponentiate the growth rate and multiply it by the current population size: $\exp(T(N(t),K(t))) N(t)$.
-
-The weighted model is a weighted average of predictions from the spatial and temporal models, with the weights changing as a function of time, here expressed as the forecast horizon.
-The weights change as a function of the square root of the forecast horizon, to allow rapid shifts in the model weights.
-\begin{equation}
-logit(\omega_t)=\beta^W_0 + \beta^W_1 \sqrt{t}
-\label{eqn:weights}
-\end{equation}
-For the community turnover example, the predicted biomass from the weighted model is:
-\begin{equation}
-\hat{N}(t+1)= \omega \cdot T(N(t),K(t)) + (1-\omega) \cdot S(K(t)) 
-\label{eqn:combined_model1}
-\end{equation}
-Again, we suppress the spatial subscript ($x$) here because we are focused on densities at just one location.
-For the eco-evolutionary example, the predicted population size from the weighted model is:
-\begin{equation}
-\hat{N}(t+1) = \omega \cdot \exp(T(N(t),K(t))) N(t) + (1-\omega) \cdot S(K(t)) 
-\label{eqn:combined_model2}
-\end{equation}
-
-We used the \texttt{optim} function to estimate the $\beta^W$s that minimize the sum of squared errors, $(\hat{N}(t+1) - N(t+1))^2$.
-
-In the main text, we show the point forecasts but not the uncertainty around the forecasts.
-After exploring that uncertainty, we decided that presenting it would be misleading.
-For the spatial and, especially, the temporal statistical models, the uncertainty is unrealistically low, because the models are estimated with very large samples sizes from the simulations.
-Furthermore, the simulations do not include noise; the only reason there is any uncertainty is because the statistical models are slightly mis-specified with respect to the process models.
-Showing uncertainty for the weighted model would be even less meaningful, because it is not a true, out-of-sample forecast (parameters are fit directly to the observations for which we make predictions).
-The R code to compute uncertainties for the spatial and temporal forecasts is available on our Github repository (https://github.com/pbadler/space-time-forecast), but is commented out.
-
-\section{Description of the meta-community model}\label{metacomm}
-
-\cite{alexander_lags_2018} developed a meta-community model to represent dynamics of local communities arrayed along a one-dimensional elevation gradient, as influenced by three main processes: temperature-dependent growth, competition, and dispersal.
-Here we adapt their notation to be consistent with our own.
-
-The population size of species $i$ in cell $x$ at time $t + 1$, $N_{i}(x,t+1)$, is computed in two steps.
-The first step accounts for changes in local population sizes due to dispersal.
-In each local community, all species export a fraction ($d$) of their local population to the two adjacent communities in the 1-dimensional landscape:
-\begin{equation}
-N'_{i}(x,t) = (1-d) \cdot N_{i}(x,t) + \frac{d}{2} \cdot (N_{i}(x+1,t) + N_{i}(x-1,t))
-\end{equation}
-Here $N'$ distinguishes the post-dispersal population size from the pre-dispersal population size.
-
-The second step computes population growth, taking into account competition:
-\begin{equation}
-N_{i}(x,t+1) = N'_{i}(x,t) + N'_{i}(x,t)[g_i(K(x) - Kmin_i) - c_i N'_{i}(x,t) - l_i \sum_{k} N'_{k}(x,t)]
-\end{equation}
-In the absence of competition, the growth rate ($g_i$) is determined by the difference between the temperature at site $x$ ($K(x)$) and the focal species' minimum temperature tolerance, $Kmin_i$, the lowest temperature at which a species can maintain a positive growth rate.
-Growth is further reduced by intraspecific and interspecific competition, parameterized by $c_i$ and $l_i$.
-All species are assigned the same value of $c_i$, which represents an additional effect of intraspecific competition on top of interspecific competition.
-This stabilizes coexistence, since every species will exert stronger intra- than interspecific competition.
-However, values of $l$ vary among species to create a trade-off between growth rates and competitive ability versus low temperature tolerance: fast-growing species (high $g_i$) are more tolerant of interspecific competition (low $l_i$) but are more limited by temperature (high $Kmin_i$).
-
-To assign species-specific parameter values, the number of species in the metacommunity is specified.
-Next, each species is assigned an optimal temperature within a specified temperature range by drawing from a uniform distribution.
-Sensitivity to interspecific competition is then determined as a decreasing function of optimal temperature.
-Calculations are performed in the script \texttt{SpeciesPoolGen.R}.
-
-\section{Description of the eco-evolutionary annual plant model}\label{eco-evo}
-
-%\renewcommand{\theequation}{B-\arabic{equation}}
-%\renewcommand{\thetable}{B-\arabic{table}}
-%\renewcommand{\thefigure}{B-\arabic{figure}}
-
-\noindent \textbf{Haploid Model:} Begin with a haploid model that describes the number of seeds present in a population.
-We model a scenario in which all seeds germinate, so we can ignore seedbank dynamics.
-$N_{i,t}$ is the number of seeds of species $i$ at time $t$.
-The model is
-\begin{align}\begin{split}
-N_{1,t+1} &= \frac{\lambda_1(K(t))N_{1,t}}{1+ \alpha_{11}N_{1,t} + \alpha_{12}N_{2,t}}\\
-N_{2,t+1} &= \frac{\lambda_2(K(t))N_{2,t}}{1+ \alpha_{21}N_{1,t} + \alpha_{22}N_{2,t}}
-\end{split}\end{align}
-where $\lambda_{i}(K(t))$ is the seed production rate per plant, and $K(t)$ is the temperature at time $t$.  Below we refer to the $\alpha_{ij}$ as intra- and inter-genotype competition coefficients.
-
-%\noindent \textbf{Diploid Model:} Consider a one-species diploid model.
-The genotypes are denoted by $AA$, $Aa$, and $aa$.
-The number of each genotype at time $t$ is $N_{AA}(t)$, $N_{Aa}(t)$, and $N_{aa}(t)$.
-The germination rates for each genotype are $g_{AA}$, $g_{Aa}$, and $g_{aa}$.
-The seed survival probability and seed production rate for genotype $AA$ are $s_{AA}$ and $\lambda_{AA}(K(t))$, respectively.
-The analogous parameters for the other genotypes are similarly denoted.
-The competition coefficients are denoted by $\alpha_{i,j}$, e.g., $\alpha_{AA,AA}$ or $\alpha_{AA,Aa}$.
-Throughout we assume that gametes mix randomly in the population.
-
-\noindent \textbf{Diploid Model:} Consider a one-species diploid model.
-The genotypes are denoted by $AA$, $Aa$, and $aa$.
-The number of each genotype at time $t$ is $N_{AA}(t)$, $N_{Aa}(t)$, and $N_{aa}(t)$.
-%The germination rates for each genotype are $g_{AA}$, $g_{Aa}$, and $g_{aa}$.  
-%The seed survival probability and %
-The seed production rate for genotype $AA$ is $\lambda_{AA}(K(t))$, and the analogous parameters for the other genotypes are similarly denoted.
-The competition coefficients are denoted by $\alpha_{i,j}$, e.g., $\alpha_{AA,AA}$ or $\alpha_{AA,Aa}$.
-Throughout we assume that gametes mix randomly in the population.
-
-First consider the case where the competition coefficients are zero ($\alpha_{i,j}=0$).  Let $T$ denote the total number of gamete-pairs produced in a given year,
-\begin{equation}
-T = \lambda_{AA}(K(t))N_{AA}(t) + \lambda_{Aa}(K(t))N_{Aa}(t) +\lambda_{aa}(K(t))N_{aa}(t).
-\end{equation}
-The first term is the number of gamete-pairs produced by $AA$ individuals.
-The second and third terms are the numbers of gamete-pairs produced by $Aa$ and $aa$ individuals, respectively.
-The proportion of $A$ gametes ($\phi_A$) and the proportion of $a$ gametes ($\phi_a$) are given by
-\begin{align}\begin{split}
-\phi_{A} &= \frac{2\lambda_{AA}(K(t))N_{AA}(t)+ \lambda_{Aa}(K(t))N_{Aa}(t)}{2T} \hspace{10pt} \text{and} \hspace{10pt} \phi_a = 1-\phi_{A}.
-\end{split}\end{align}
-Note that the $T$ in the denominator of $\phi_A$ shows up because we are computing proportions.  Combining all of these we get the dynamics for each genotype,
-\begin{align}\begin{split}
-N_{AA}(t+1) &= \phi_A^2T\\
-N_{Aa}(t+1) &= \phi_A\phi_aT\\
-N_{aa}(t+1) &= \phi_a^2T
-\end{split}\end{align}
-
-Now consider the case where the competition coefficients are non-zero ($\alpha_{i,j}\neq0$).
-Including competition changes the way in which we compute $T$, $\phi_A$, and $\phi_a$.
-Specifically, because the total number of seeds produced per year by each genotypes is reduced based on intra- and inter-genotype competition, the total number of gamete-pairs becomes
-\begin{align}\begin{split}
-T &=  \frac{\lambda_{AA}(K(t))N_{AA}(t)}{1+ \alpha_{AA,AA}N_{AA}(t) + \alpha_{AA,Aa}N_{Aa}(t)+ \alpha_{AA,aa}N_{aa}(t)} \\ 
-&+ \frac{\lambda_{Aa}(K(t))N_{Aa}(t)}{1+ \alpha_{Aa,AA}N_{AA}(t) + \alpha_{Aa,Aa}N_{Aa}(t)+ \alpha_{Aa,aa}N_{aa}(t)}\\
-&+\frac{\lambda_{aa}(K(t))N_{aa}(t)}{1+ \alpha_{aa,AA}N_{AA}(t) + \alpha_{aa,Aa}N_{Aa}(t)+ \alpha_{aa,aa}N_{aa}(t)}.
-\label{eqn:defineT}
-\end{split}\end{align}
-The first line is the number of gamete-pairs produced by $AA$ individuals after accounting for the effects of competition.
-The second and third lines are the numbers of gamete-pairs produced by $Aa$ and $aa$ individuals, respectively.
-The proportions of $A$ gametes and $a$ gametes are
-\begin{align}\begin{split}
-\phi_A &= \frac{2}{2T}\frac{\lambda_{AA}(K(t))N_{AA}(t)}{1+ \alpha_{AA,AA}N_{AA}(t) + \alpha_{AA,Aa}N_{Aa}(t)+ \alpha_{AA,aa}N_{aa}(t)} \\
-&+ \frac{1}{2T}\frac{\lambda_{Aa}(K(t))N_{Aa}(t)}{1+ \alpha_{Aa,AA}N_{AA}(t) + \alpha_{Aa,Aa}N_{Aa}(t)+ \alpha_{Aa,aa}N_{aa}(t)}\\
-\phi_a &= 1- \phi_A
-\label{eqn:definePhi}
-\end{split}\end{align}
-Combining all of this results in the same model as above,
-\begin{align}\begin{split}
-N_{AA}(t+1) &= \phi_A^2T\\
-N_{Aa}(t+1) &= 2 \phi_A\phi_aT\\
-N_{aa}(t+1) &= \phi_a^2T,
-\end{split}\end{align}
-but the definitions of $T$, $\phi_A$, and $\phi_a$ are given by equations (\ref{eqn:defineT}) and (\ref{eqn:definePhi}) .
-
-\newpage
-
-\section*{Supplementary Tables}
-
-\renewcommand{\thetable}{SM-\arabic{table}}
-
-\begin{table}[htp]
-	\caption{\label{tab:comm-turn-pars}Parameters and parameter values for the community turnover case study. Values are assigned at
-		the start of \texttt{comm\_turn\_master.R}. ``Name" refers to the variable declared in the computer code. These names do not exactly 
-		match the symbols shown in the equations in Appendix \ref{metacomm}; rather, the species-specific values of those parameters are calculated in the computer code based on the values in this table.  }
-	\centering
-	\smallskip 
-	\begin{tabular}{lrl}
-		\hline
-		Name & Value & Definition \\
-		\hline
-		L\_land & 20 & Length of landscape \\
-		Tmin & 0 &  Minimum of spatial gradient in baseline temperature \\ 
-		Tmax & 15 &  Maximum of spatial gradient in baseline temperature \\
-		Tstdev & 2 &  Standard deviation of temperature (interannual variation) \\
-		deltaT & 4  & Magnitude of directional change in temperature \\
-		burnin\_yrs & 2000 & Number of years to initialize simulation \\
-		baseline\_yrs & 1000  & Number of years at baseline temperature used to fit statistical models \\
-		warming\_yrs  & 200 & Number of years over which temperature increases \\
-		final\_yrs & 2000  & Number of years at steady-state, elevated temperature \\
-		\hline
-		N & 40 & Number of species \\
-		Gmax & 0.5 & Maximum population growth rate \\
-		Gmin & 0.2 & Minimum population growth rate \\
-		Lmax & 1.5 & Maximum sensitivity to competition \\
-		Lmin & 0.7 & Minimum sensitivity to competition  \\
-		Cmax & 0.2 & Maximum additional sensitivity to conspecific competition \\
-		Cmin & 0.2 & Minimum additional sensitivity to conspecific competition \\
-		d & 0.01 &  Fraction of offspring dispersing from home site \\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[htp]
-	\caption{\label{tab:eco-evo-pars}Parameters and parameter values for the eco-evolutionary case study. Values are assigned at
-		at the start of \texttt{genetic\_diversity\_master.R}.
-``Name" refers to the variable declared in the computer code.
-Where appropriate, the corresponding symbols from equations in Appendix \ref{eco-evo} are shown in parentheses.
-}
-	\centering
-	\smallskip 
-	\begin{tabular}{lrl}
-		\hline
-		Name & Values & Definition \\
-		\hline
-		Tstdev & 1 & Standard deviation of temperature (interannual variation) \\
-		baseT & -1 & Baseline temperature \\
-		deltaT & 5 & Total change in temperature \\
-		baseline\_yrs & 500 & Number of years at baseline temperature used to fit statistical models \\
-		warming\_yrs & 100  & Number of years over which temperature increases \\
-		final\_yrs & 300  & Number of years at steady-state, elevated temperature \\
-		\hline
-		fec\_Tmu & -1,0,1  & Optimal fecundity temperature for genotypes AA, Aa, and aa \\
-		fec\_Tsigma & 8  & Standard deviation in fecundity for all genotypes \\
-		fec\_max & 100 &  Maximum fecundity for all genotypes \\
-		alpha ($\alpha$) & 1  &  All competition coefficients for all genotypes \\
-		\hline
-
-	\end{tabular}
-\end{table}
-
-\clearpage
-
-\section*{Supplementary Figures}
-
-\renewcommand{\thefigure}{SM-\arabic{figure}}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=1 \textwidth] {replicates_weights.png}
-	\caption{(A) Temporal shifts in the model weighting term for 10 independent simulations of (A) the community turnover model, and (B) the eco-evolutionary model.
-For the community turnover model, each simulation began with initialization of a new regional species pool.
-For the eco-evolutionary model, genotype parameters were fixed, and only the sequence of annual temperatures varied between runs.
-In all cases, the combined forecast is heavily weighted towards the time-series model at short forecast time scales, and towards the space-for-time model at long forecast time scales.
-}
-	\label{fig:replicates}
-\end{figure}
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.7 \textwidth] {community_forecast_species_nonstationary.png}
-	\caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-In contrast to Fig.
-\ref{fig:community-forecast-species}, which shows results for a period of warming followed by stationary temperatures, for this simulation we spread the same temperature increase out over the entire simulation with no stationary periods.
-(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
-(C) Simulated biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).
-Time 1000 (years) in each panel corresponds to the start of the temperature increase.
-}
-	\label{fig:community-forecast-species-nonstationary}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_models_total.png}
-\caption{Results for total biomass from the community turnover model.
-Blue points show mean total biomass during the baseline period at locations across the temperature gradient, and the blue line shows predictions from the spatial model.
-Red points show annual total biomass during the baseline period as a function of annual temperature at the central site on the gradient.
-The red line shows predictions from the temporal model.
-}
-\label{fig:community-models-total}
-\end{figure}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {community_forecast_total.png}
-\caption{Results for total biomass from the community turnover model.
-(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated total biomass and forecasts from the spatial, temporal and weighted models.
-(C) Simulated changes in biomass of all species (grey) at the focal site in the metacommunity model, and the weight given to the temporal model for total biomass (blue).
-Time 1000 (years) in this figure corresponds to the start of the temperature increase.
-}
-\label{fig:community-forecasts-total}
-\end{figure}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=0.7 \textwidth] {forecast_supplement.png}
-\caption{Simulated shifts in genotype abundances, and the model weighting term, $\omega$, during the warming phase and the following stationary temperature phase.
-Time 0 (years) in this figure corresponds to the start of the temperature increase.}
-\label{fig:forecast_supp}
-\end{figure}
-
 \end{document}

### 2020-07-07 doc/final_files/supplement.tex
final files finished

--- 2020-06-09 doc/manuscript.tex
@@
 
-\graphicspath{ {./../genetic-diversity/figures/} {./../community-turnover/figures/} }
+\graphicspath{ {./} }
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-%%% Just for commenting
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[usenames]{color}
-\newcommand{\new}{\textcolor{red}}
-\newcommand{\spe}{\textcolor{blue}}
-\newcommand{\comment}{\textcolor{red}}
-
 \newcommand{\be}{\begin{equation}}
@@
 
-\title{\textbf{Matching the forecast horizon with the relevant spatial and temporal processes and data sources}}
+\title{\textbf{Supplementary Materials for: Matching the forecast horizon with the relevant spatial and temporal processes and data sources}}
 
@@
 
-\linenumbers
-
-% rest of title page
-
-\vspace{-1cm}
-
-\noindent \emph{Statement of authorship:} PBA and EPW designed the study, PBA and MHC built the models, and PBA analyzed the models.
-PBA wrote the first draft of the manuscript, and all authors contributed substantially to revisions.
-
-\noindent \emph{Acknowledgements:} We thank Heather Lynch and Juan Manuel Morales for comments that improved early drafts of the paper.
-PBA was supported by the National Science Foundation (DEB-1927282) and the Utah Agriculture Experiment Station (grant 1322).
-EPW was supported by the Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative (GBMF4563).
-
-%\noindent \emph{Running title:} Forecast horizon and ecological process
-
-\vspace{0.25cm}
-
-\noindent \textbf{Abstract:}   Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
-However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change.
-Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
-We use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
-First, we simulated population or community dynamics under stationary temperature using two simple, mechanistic models.
-Second, we fit statistical models to the simulated data using a time-series approach, a space-for-time approach, or a weighted average.
-We then forecast the response to a temperature increase using the statistical models, and compared these forecasts to temperature effects simulated by the mechanistic models.
-We found that the time-series approach made accurate short-term predictions because it captured initial conditions and effects of fast processes such as birth and death.
-The space-for-time approach made more accurate long-term predictions because it better captured the influence of slower processes such as evolutionary and ecological selection.
-The weighted average made accurate predictions at all time scales, including intermediate time-scales where the other two approaches performed poorly.
-A weighted average of time-series and space-for-time approaches shows promise, but making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
-
-\vspace{0.25cm}
-
-\noindent \emph{Keywords:} dispersal, ecological forecasting, eco-evolutionary dynamics, global change, selection
-
-\newpage
-
-\section*{Abstract} 
-Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients.
-However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change.
-Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two.
-We use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data.
-First, we simulated population or community dynamics under stationary temperature using two simple, mechanistic models.
-Second, we fit statistical models to the simulated data using a time-series approach, a space-for-time approach, or a weighted average.
-We then forecast the response to a temperature increase using the statistical models, and compared these forecasts to temperature effects simulated by the mechanistic models.
-We found that the time-series approach made accurate short-term predictions because it captured initial conditions and effects of fast processes such as birth and death.
-The space-for-time approach made more accurate long-term predictions because it better captured the influence of slower processes such as evolutionary and ecological selection.
-The weighted average made accurate predictions at all time scales, including intermediate time-scales where the other two approaches performed poorly.
-A weighted average of time-series and space-for-time approaches shows promise, but making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.
-
-\vspace{0.25cm}
-
-\noindent \emph{Keywords:} dispersal, ecological forecasting, eco-evolutionary dynamics, global change, selection
-
-\section*{Introduction}
-
-Forecasting is increasingly recognized as important to the application and advancement of ecological research.
-Forecasts are necessary to guide environmental policy and management
-decisions about mitigation and adaption to global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
-But forecasts can also advance understanding of the processes governing ecological systems by providing rigorous tests of 
-model predictions \citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
-The dual benefits of informing management and advancing basic knowledge 
-makes forecasting an important priority for ecological research.
-
-Statistical models used for ecological forecasting generally rely on either time-series approaches or space-for-time substitutions.
-The time-series approach involves fitting models to long-term datasets to describe the temporal dynamics of a system.
-We then project those dynamic models to make predictions about what will happen in the future.
-This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, or primary production as a function of annual precipitation \citep{lauenroth_long-term_1992}.
-When time-series models are fit with typically short ecological data sets, they capture ``fast processes'' operating on interannual time-scales, such as birth, death, individual growth, small-scale dispersal events, and short-term responses to environmental conditions (Fig.
-\ref{fig:fast-slow-concept}).
-Statistical models built using this approach normally cover a limited spatial extent (but see \citealt{Hefley2017,kleinhesselink_response_2018,Chevalier}), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, that could influence dynamics at longer time scales \citep{clark_ecological_2001}.
-
-Space-for-time substitution approaches begin by describing how an ecological variable of interest, such as occupancy or productivity,  varies across sites experiencing different environmental conditions.
-These spatial relationships between environment and ecological response are assumed to also hold for changes at a site through time.
-To make a forecast, we first predict the future environmental conditions and then determine the associated ecological response, based on the observed spatial relationship.
-This is the approach commonly used to predict population distribution or abundance as a function of climate \citep{elith_species_2009} or mean primary production as a function of mean precipitation \citep{Sala1988}.
-Space-for-time models capture the outcome of interactions between fast processes and slower processes operating over long time periods, such as immigration, extinction, and responses to large or prolonged environmental changes (Fig.
-\ref{fig:fast-slow-concept}).
-However, space-for-time models provide no information about how quickly the system will move from the current state to the predicted, future state.
-In fact, transient dynamics could prevent the system from ever reaching the predicted steady state \citep{Urban2012}.
-
-Although both time-series and space-for-time approaches are widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or advancing our understanding of ecological dynamics \citep{harris_forecasting_2018,Renwick2018}.
-Such a discussion is overdue given that these two approaches are likely to make very different predictions about ecological responses to future environmental change.
-For example, average forage production in the U.S.
-shortgrass steppe increases rapidly with increasing mean annual precipitation, but production is much less sensitive to interannual variation in precipitation at any one location \citep{lauenroth_long-term_1992}.
-The space-for-time approach would predict large changes in production in response to a future change in precipitation, while the time-series approach would predict smaller changes.
-Bird species' abundances in the United Kindgom also show different responses to spatial vs.
-temporal variation in weather covariates.
-Not only are abundances typically more sensitive to spatial variation in the covariates, but in many cases the temporal response takes the opposite sign \citep{Oedekoven2017}.
-Whenever the sign or magnitude of relationships based on time-series and space-for-time approaches differ, so will the resulting forecasts.
-
-Whether time-series models, space-for-time approaches, or some combination of the two will serve as the best source of 
-information for forecasting may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. 
-This potential dependency on the ``forecast horizon'' (\textit{sensu} \citealt{Hyndman2018})
-reflects lags in the response of ecological conditions to environmental change, shifts in the importance of ecological processes with
-time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of 
-environmental conditions represented in observed data (Fig. \ref{fig:fast-slow-concept}).
-At short forecast horizons (days to years), dynamics will reflect physiological and demographic responses and interactions
-among the organisms present at a site more than temporal turnover of genotypes or species; 
-environmental conditions are likely to stay within the range of historical variation; 
-and the current state of the system is likely to capture the influence of unmeasured processes. As a result, 
-for near-term forecasts time-series approaches may capture the key dynamics and provide accurate predictions.
-
-In contrast, at long forecast horizons (decades to centuries),  environmental conditions that have not been 
-historically observed are likely to not only occur but to persist long enough to drive significant turnover of genotypes and species 
-through colonization and extinction as well as changes in the flux of energy and nutrients.  
-At these long forecast horizons, the state of the system at the time the forecast is issued may be 
-little help in predicting the future state. For the century-scale forecasts often featured in biodiversity and 
-species-distribution modeling, space-for-time approaches may effectively capture the response of ecosystems to major shifts 
-in climate over long periods, producing better long-term forecasts than time-series approaches. 
-Using different modeling approaches for different forecast horizons is common in other disciplines.
-For example, meteorological models for short-term weather forecasts differ substantially in spatial
-and temporal resolution and extent from the global circulation models used to predict long-term changes
-in climate.
-
-Why not simply use process-based models to avoid the difficulties posed by phenomenological time-series and space-for-time modeling approaches?
-If we could accurately characterize all of the processes governing a system, then a model based on that understanding should make accurate predictions at all forecast horizons.
-Process-based models should also be more robust for making predictions outside of historically observed conditions and even beyond the conditions observed across spatial gradients, which will be especially important in a future with increasingly novel combinations of environment and species interactions \citep{Williams2007}.
-Unfortunately, in most cases this approach is not currently feasible because we lack a detailed knowledge of all the complex and interacting processes influencing the dynamics of real ecological systems.
-Even if the general form of the models were known, estimating the high number of parameters and quantifying how they vary across ecosystems typically requires more data than is currently available even for well-studied systems.
-Furthermore, the high complexity and corresponding parameter uncertainty of such models can increase predictive errors; simpler  time-series models may actually perform better \citep{ward_complexity_2014}, though spatial replication can reduce the cost of complexity \citep{Chevalier}.
-As a result, models used for ecological forecasting will include at least some phenomenological components.
-But that does not mean that phenomenological forecast models cannot benefit from process-based understanding.
-Even if process-level understanding does not enable a fully mechanistic model, it can improve the specification of phenomenological models.
-Our hypothesis is that different processes may be relevant for different forecast horizons, and that we can act on this knowledge by fitting models to different kinds of datasets.
-
-Here we use two simulated case studies to 1) demonstrate why time-series and space-for-time approaches can make different predictions, 2) propose that the best model-building approaches for ecological forecasting
-may depend on the time horizon of the forecast, and 3) explore how time-series and space-for-time approaches might be 
-combined via weighted averaging to make better forecasts at intermediate time scales. The first case study
-focuses on how interspecific interactions affect the population dynamics of a focal species, and the second focuses on
-an eco-evolutionary scenario. Our analyses show that: 
-\begin{enumerate}
-	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require accounting for the influence of slow processes such as evolutionary and ecological selection as well as dispersal.
-	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and fast responses of current assemblages to interannual environmental variation, while data spanning spatial gradients capture the long-term outcome of interactions between fast and slow processes.
-Whether predictive models should be trained using longitudinal or spatial data sets, or both, depends on the time-scale of the desired forecast.
-	\item A key challenge for future research is determining the rate at which slow processes begin to influence dynamics.
-\end{enumerate}
-
-\section*{Modeling approach}
-
-In each case study, we simulated the effects of an increase in temperature on simple systems with known dynamics.
-The truth was represented by a simulation model that was mechanistic for at least one important process, 
-but we treated this data-generating model as unknown when analyzing the data and we assumed that perfectly recovering the mechanisms
-it contains would not be possible in practice. We began each simulation
-under a stationary distribution of annual temperatures, allowing the system to equilibrate; we call this the baseline phase. 
-We then increased temperature progressively over a period of time, followed by a second period of stationary, now elevated, temperature. 
-The objective was to forecast the response of the system to the temperature increase based on spatial and/or temporal data 
-``sampled" from the simulation during the baseline period.
-
-We made forecasts based on two phenomenological statistical models, each representing processes operating at different time scales.
-One statistical model represents the time-series or ``temporal approach.'' We regressed interannual variation in an ecological response on interannual variation in temperature at just one site.
-The other statistical model relies on a space-for-time substitution, which we call the ``spatial approach'' for brevity.
-We regressed the mean temperature on the mean of an ecological state or rate across many sites.
-We compared forecasts from both statistical models to the simulated dynamics to determine how well the two approaches performed at different forecast horizons.
-We also assessed the potential for combining the information available in temporal and spatial patterns by using a weighted average of the forecasts from the temporal and spatial approaches optimized to best match the (simulated) observations.
-We then studied how the optimal model weights changed over time.
-We expected the temporal approach to best predict short-term dynamics, the spatial approach to best predict long-term dynamics, while the weighted model would show potential to provide the best forecasts at transitional, intermediate time scales.
-The three statistical models are described in Supporting Information (Appendix \ref{models}).
-Computer code for both case studies is available for reviewers online (https://drive.google.com/open?id=1aju04qtQvJmZG1mcpN-6hYBZILBhDfl2) and will be archived at Zenodo upon acceptance.
-Instructions for running the code are in the README file in the main directory of the zip archive.
-
-\section*{Community turnover example}
-
-Conservation biologists and natural resource managers often need to anticipate the impact of environmental change on the abundance of 
-endangered species, biological invaders, and harvested species.
-Although the managers may be primarily interested in just one focal species, skillful prediction might require considering interactions with many other species, greatly complicating the problem.
-But at what forecast horizon do altered species interactions become impossible to ignore?
-We explored this question using a metacommunity simulation model developed by \cite{alexander_lags_2018} to study how community responses to increasing temperature depend on the interplay between within-site demography and competitive interactions and the movement of species across sites.
-
-\subsection*{Methods}
-
-The model features Lotka-Volterra competitive interactions among plants within sites that are arrayed along an elevation and temperature gradient.
-Composition varies along the gradient because of a trade-off between growth rate and cold tolerance: cold sites are dominated by slow-growing species that can tolerate low temperatures, while warm sites are dominated by  fast-growing species that are cold intolerant.
-Multiple species can coexist within sites because all species experience stronger competition from conspecifics than from heterospecifics.
-Sites are linked by dispersal: a specified fraction of each species' offspring leaves the site where they were produced and reaches all other sites with equal probability.
-We provide a more detailed description of the simulation model in SI Appendix \ref{metacomm}.
-
-We first simulated a baseline period with variable but stationary temperature, followed by a period of rapid temperature increase, and then a final period of stationary temperature.
-Interannual variation in temperature is the same at all sites, but mean temperature varies among sites.
-All sites experienced the same absolute increase in mean temperature.
-We focused on the biomass dynamics of one focal species that dominated the central site during the baseline period.
-Parameter values for the simulations described in the main text are shown in Table \ref{tab:comm-turn-pars}.
-We report results from one simulation run; results were qualitatively consistent for replicate runs (\comment{Fig.
-\ref{fig:replicates}A}).
-
-\subsection*{Results}
-
-During the baseline period there were strong spatial patterns across the mean temperature gradient. 
-Individual species, including our focal species, showed classic, unimodal ``Whittaker" patterns of abundances across the gradient (Fig. \ref{fig:species-patterns-models}A).
-These spatial patterns are the basis for our spatial statistical model of the temperature-biomass relationship for our focal species (Fig. \ref{fig:species-patterns-models}A). 
-In contrast to the strong spatial patterns, population and community responses to interannual variation in temperature within sites were weak.
-At our focal site in the center of the gradient, the biomass of the focal species was quite insensitive to interannnual variation in temperature, but showed strong temporal autocorrelation (Fig.
-\ref{fig:species-patterns-models}B).
-Our temporal statistical model estimates this weak, linear temperature effect, along with the strong lag effect of biomass in the previous year.
-
-We used both the temporal and spatial statistical models to forecast the effect of a temperature increase (Fig.
-\ref{fig:community-forecast-species}A) on the focal species' biomass at one location in the center of the temperature gradient.
-The predictions from these two models contrasted markedly, with the temporal 
-statistical model predicting a large increase in biomass and the spatial statistical model predicting a decrease. Initially, the simulated abundances followed the increase predicted by the 
-temporal model, but as faster-growing species colonized and increased in abundance at the focal site, the biomass of the focal species decreased, eventually falling below its baseline level (Fig.
-\ref{fig:community-forecast-species}B).
-
-To combine information from the temporal and spatial statistical models into a single prediction, we fit a weighting parameter, $\omega$, which varies over time and is bounded between 0 and 1.
-At any time point, \comment{or year}, $t$, this weighted forecast is $\omega \cdot T(N_{t-1},K_t) + (1-\omega) \cdot S(K_t) $ where $T$ is the temporal statistical model, which depends on population size, $N$, and expected temperature, $K$, and $S$ is the spatial statistical model, which depends only on $K$ (see SI Appendix \ref{models} for a full description of the approach).
-The weighted model accurately predicts the simulated dynamics across the full forecast horizon (Fig.
-\ref{fig:community-forecast-species}B).
-It also shows that the most rapid shifts in the model weights occurred during the period when warm-adapted, faster growing species were increasing most rapidly in abundance (Fig.
-\ref{fig:community-forecast-species}C).
-However, the reason the weighted models works so well is that the weights were determined by fitting directly to the data.
-Unlike the forecasts from the spatial and temporal statistical models, we did not generate out-of-sample predictions from the weighted model; it merely provides a convenient way to quantify how rapidly dynamics shift from being dominated by interannual variation captured in the temporal model (time $t=0$ to $t \approx 1250$ in Fig.
-\ref{fig:community-forecast-species}B) to being dominated by the steady-state equilibrium captured by the spatial model (time $t \geq 2500$).
-A true forecast from the weighted model would require a method to determine the model weights \emph{a priori}.
-
-When we repeated the simulation with a continuous, nonstationary temperature increase, we see a qualitatively similar shift in weights with increasing forecast horizon from the temporal to spatial statistical model (Fig.
-\ref{fig:community-forecast-species-nonstationary}).
-In this case, the forecast from the temporal statistical model is not as skillful in the near-term forecast horizon, because the model does not account for the temperature trend during the model fitting period.
-Separating the effect of annual temperature deviations from the temperature trend would distinguish between short and long-term patterns, much as our temporal and spatial statistical models do in the simulation with stationary temperature periods.
-
-The compositional turnover affecting our focal species also influences total biomass, linking community and ecosystem dynamics.
-We repeated our focal species analysis for total community biomass, and the results were similar: the temporal statistical model initially made the best forecasts immediately following the onset of the temperature increase, but as the identity and abundances of species at the study site changed, the model weights rapidly shifted to the spatial statistical model (Figs.
-\ref{fig:community-models-total} and \ref{fig:community-forecasts-total}).
-
-\section*{Eco-evolutionary example}
-
-Evolutionary adaptation is a key uncertainty in predicting how environmental change will impact a focal population at a given location \citep{Hoffmann2011}.
-Like the shifts in species composition illustrated in the previous example, shifts in genotype frequencies can also influence dynamics and forecasts at different time scales.
-Although shifts in genotype frequencies at the population level are analogous to changes in species composition at the community level, the mechanisms are distinct: heterozygosity and genetic recombination have no analogue at the community level.
-We demonstrate how these processes influence short and long-term forecasts with a standard eco-evolutionary simulation model for a hypothetical annual plant population.
-
-\subsection*{Methods}
-
-Our model assumes that fecundity is temperature dependent, and different genotypes have different temperature optima (Fig.
-\ref{fig:evo_evo_spatial_temporal_models}A).
-All seeds germinate every year, preventing a seedbank from developing.
-The model describes how the local density of each genotype changes between years, which depends on temperature and genotype densities in the previous year.
-Transient temporal dynamics are computed directly from the model; these dynamics are the basis for the temporal statistical model.
-To create a spatial gradient, we simulated the equilibrium density of each genotype in a series of local populations experiencing different mean temperatures.
-The pattern of equilibrium densities across the mean annual temperature gradient is the basis for our spatial statistical model: cold sites will be dominated by the cold-adapted homozygous genotype, warm sites will be dominated by the heat-adapted homozygous genotype, and intermediate sites will be dominated by the heterozygous genotype (Fig.
-\ref{fig:evo_evo_spatial_temporal_models}B).
-The full description of the eco-evolutionary simulation model is provided in SI Appendix \ref{eco-evo}, and parameter values for simulations described here are shown in Table \ref{tab:eco-evo-pars}.
-As in the first case study, we report results from just one simulation run, but results were qualitatively consistent for replicate runs (\comment{Fig.
-\ref{fig:replicates}B}).
-
-The spatial pattern shown in Fig.
-\ref{fig:evo_evo_spatial_temporal_models}B is the outcome of steady-state conditions.
-But at any one site, the population's short-term response to temperature will be determined by the dominant genotype's reaction norm (Fig.
-\ref{fig:evo_evo_spatial_temporal_models}A).
-For example, at a cold site dominated by the cold-adapted homozygous genotype, a warmer than average year would cause a decrease in population size due to decreases in fecundity (blue line in Fig.
-\ref{fig:evo_evo_spatial_temporal_models}A), even though the heat-adapted homozygote might perform optimally at that temperature.
-However, if warmer than normal conditions persist for many \comment{years}, then genotype frequencies should shift, and the heat-adapted homozygote will compensate for the decreases of the cold-adapted genotype.
-
-\subsection*{Results}
-
-To demonstrate these dynamics, we simulated a diploid annual plant population at a colder than average site.
-During the baseline period, the population is dominated by the cold-adapted genotype.
-We used the simulated data from this baseline period to fit a temporal statistical model (Appendix \ref{models}) that predicts population growth rate as a function of annual temperature and population size (Fig.
-\ref{fig:evo_evo_spatial_temporal_models}C), assuming no knowledge of the underlying eco-evolutionary dynamics.
-We then imposed a period of warming, followed by a final period of higher stationary temperature (Fig.
-\ref{fig:forecast} top).
-
-With the onset of warming, the population crashed as the cold-adapted genotype decreased in abundance.
-Eventually, frequencies of the heterozygous genotype and the warm-adapted homozygous genotype began to increase and the population recovered (Fig.
-\ref{fig:forecast} bottom).
-The temporal statistical model (solid blue line in Fig.
-\ref{fig:forecast}) accurately predicted the impact of the initial warming trend, but eventually became too pessimistic, while the spatial statistical model (solid red line in Fig.
-\ref{fig:forecast}) did not handle the initial trend but accurately predicted the eventual, new steady state.
-
-As in the community turnover example, we also fit a weighted average of predictions from the spatial and temporal statistical models (purple line in Fig.
-\ref{fig:forecast}), with the weights changing over time.
-This weighted model initially reflected the temporal model (decrease from $t=500$ to $t=600$), but then rapidly transitioned to reflect the spatial model ($t \geq 700$).
-The rapid transition in the weighting term, $\omega$, occurred during the period of most rapid change in genotype frequencies (Fig.
-\ref{fig:forecast_supp}).
-The weighted model's predictions look impressively accurate, but, as in the community turnover example, that is because we used the full, simulated time series to fit the weighting term.
-A true forecast would require an independent method to predict how the model weights shift over time.
-
-\section*{Discussion}
-
-Ecological forecasts are typically made using either a space-for-time substitution approach based on models fit to spatial data or using dynamic models fit to time-series data.
-Empirical studies show that the environment-response relationships detected by these approaches frequently differ in magnitude and even sign \citep{lauenroth_long-term_1992,Oedekoven2017,Amburgey2018,kleinhesselink_response_2018}.
-Our simulations illustrate how such differences may arise and then lead to very different predictions about the future state of ecological systems.
-Which approach provides the most accurate forecasts likely depends on the forecast-horizon.
-In our simulations, time-series approaches performed best for short forecast horizons, whereas models based on spatial data made more accurate forecasts at long horizons.
-In addition, our simulations demonstrate extended transitional periods during which neither the time-series or the spatial approach was effective on its own.
-The challenge is determining what is ``short-term,'' what is ``long-term,'' and how to handle the many forecasts we need in ecology which fall in between.
-We have proposed that a weighted combination of the time-series and space-for-time approaches may produce better forecasts at these intermediate forecast horizons.
-
-We designed our simulation studies to illustrate how the change in statistical model performance with increasing forecast horizon reflects differences in the types and scales of processes captured by spatial and temporal data sets.
-How could these hypotheses be tested with empirical data?
-The hypothesis that time-series models will be most effective for near-term forecasts already has empirical support, in the form of recent analyses of biodiversity forecasts at time scales from one to ten years \citep{harris_forecasting_2018}.
-The result should not be surprising, since local time-series data capture demographic processes, lagged effects, and responses of current assemblages to small changes in environmental conditions.
-In addition, the state of the system in the near future depends heavily on the current state.
-Since short-term forecasts do not typically require extrapolating into novel conditions, a model based on the historical range of variation which incorporates lags and accurate initial conditions is likely to be successful.
-
-Space-for-time modeling approaches for predicting long-term, steady-state outcomes of ecological change have also been tested empirically, primarily via hindcasting.
-Overall, the results are mixed: some tests show reasonable prediction of changes in community composition \citep{Blois2013,Illan2014} or species distributions \citep{Norberg2019}, supporting the hypothesis that datasets spanning spatial gradients capture the long-term outcome of interactions between fast processes and slower processes such as ecological and evolutionary selection, dispersal, and responses to large changes in the environment.
-Other attempts to validate predictions from space-for-time models have been discouraging \citep{Worth2014,Illan2014,Davis2014,Brun2016,Veloz2012}, indicating violations of model assumptions or effects of transient dynamics.
-However, predictions from the space-for-time approaches are rarely compared directly to predictions from time-series models (\citealt{harris_forecasting_2018}, but see \citealt{Renwick2018}).
-We need more such comparisons to identify the appropriate modeling approach for different forecast horizons.
-
-The greatest empirical challenge will be testing our hypothesis that a weighted average of spatial and temporal statistical models will make the best forecasts at intermediate time scales.
-There are two problems: finding appropriate data and determining the model weights \emph{a priori}.
-Many data sets have both a longitudinal and spatial dimension, but we could not think of one which also featured a clear ecological response to directional environmental change.
-Surely such datasets exist, and we hope researchers who work with them will test our proposed weighted model.
-Determining model weights may be more difficult.
-In our simulations, we fit the weights directly to the simulated data, which is impossible to do for actual forecasting when the future is unknown.
-We need new theory or empirical case studies in order to assign these weights \emph{a priori}.
-
-Theory could explore the influence of different parameters on the rate at which slow processes begin to influence dynamics.
-The effects of some parameters are intuitive: in the community turnover example, increasing the fraction of dispersing individuals caused a more rapid shift in species composition and in model weights (Fig.
-\ref{fig:dispersal_niche_width}A).
-Other parameters have less intuitive effects: we expected that increasing the temperature tolerance of genotypes in the evo-evolutionary example would accelerate the shift in model weights by maintaining higher genetic diversity.
-Our simulations showed the opposite effect, with wider tolerances slowing the shift in model weights (Fig.
-\ref{fig:dispersal_niche_width}B), presumably by decreasing the strength of selection.
-Additional factors to consider include organism lifespans and the magnitude of directional environmental change relative to historical interannual variation.
-
-Empirical research could inform model weights by accumulating enough case studies to infer patterns in the weighting functions and guide applications in new systems.
-Developing  rules of thumb would require testing many forecasts from both time-series and spatial models across a range of time-horizons.
-This effort may require a novel integration of typically disparate approaches, such as analyses of paleoecological data (e.g., \citealt{Worth2014}), long-term observational (e.g., \citealt{Nice2019}) or experimental data (e.g., \citealt{Silvertown2006}), and model systems with short-generation times (e.g., \citealt{Good2017}).
-
-The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to \comment{extremely simple statistical models like the ones we used in this study}.
-The same concept could work for any class of models that differ in predictive skill at different forecast horizons, \comment{such as more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation, or a set of process-based models focusing on mechanisms operating at different time scales.}  Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales.
-
-Given the challenges of determining model weights \emph{a priori}, we should also pursue alternatives for intermediate forecast horizons.
-In the Introduction, we argued that fully process-based models are not feasible.
-However, a new class of statistical models offers a compromise between mechanistic detail and phenomenological feasibility.
-Spatiotemporal statistical modeling approaches are being developed to study patterns and processes of interest to ecological forecasters, such the spread of an invasive species or population status of a threatened species  \citep{Wikle2003,Williams2017,Schliep2018}.
-Because these models include both fast processes, such as births and deaths, and slower processes, such as colonization and extinction dynamics, they have the potential to make better predictions at intermediate forecast horizons than purely spatial or temporal models.
-However, these spatiotemporal models have rarely been used in a forecasting context, due to a combination of data limitation and computational challenges.
-Many data sources contain either spatial or temporal variation, but not both, and when spatiotemporal datasets are available they often involve irregular sampling, creating challenges for modeling.
-Fitting and generating predictions from spatiotemporal models is also computationally intensive, especially with large datasets \citep{McDermott2017}.
-Fortunately, thanks to large-scale monitoring efforts from remote sensing platforms, the National Ecological Observatory Network (https://www.neonscience.org/), and community science projects (e.g., eBird), large scale spatiotemporal data is increasingly available.
-In addition, new methods for spatiotemporal forecasting are being developed that address existing computational challenges \citep{McDermott2017}, and access to high performance computing resources is increasingly common.
-Given these developments, future ecological forecasting efforts should explore spatiotemporal approaches and assess whether they improve predictions at intermediate time scales relative to traditional time-series or space-for-time approaches.
-
-Our simulation studies have important implications for the emerging field of ecological forecasting.
-First, they suggest that evaluating model performance at both short and long forecast horizons will be essential as research on forecasting methods accelerates.
-Second, while single approaches may perform reasonably well for either short or long horizons, skillful predictions at intermediate forecast horizons may require a combination of information from spatial and temporal statistical models.
-Intermediate time horizons pose challenges in other forecasting contexts as well.
-Weather forecasts based on regional-scale meteorological models are very effective for forecasting a week to ten days in advance, but then become largely uninformative.
-Forecasting these intermediate scales has been challenging in meteorology and will likely be challenging in ecology as well.
-While the recent emphasis on near-term iterative forecasting \citep{dietze_iterative_2018} is the logical and tractable starting point, we also need to build understanding and capacity for forecasting ecological dynamics across all forecast horizons of interest.
-
-\subsubsection*{Data accessibility statement:} The manuscript contains no original data.
-All computer code is available for reviewers in a zip archive that can be downloaded from Google Drive: https://drive.google.com/open?id=1aju04qtQvJmZG1mcpN-6hYBZILBhDfl2.
-Upon acceptance of the manuscript, code will be archived at Zenodo.
-
-\newpage
-\renewcommand{\refname}{Literature cited}
-\bibliographystyle{ecography}
-\bibliography{references}
-
-\newpage
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.7 \textwidth] {fast-slow-figure.png}
-	\caption{Fast and slow processes operate at different time scales, and are reflected in different kinds of datasets.
-Fast processes, such as births, deaths, and individual growth, operate at all time scales, but are the exclusive drivers of the short-term dynamics captured in most time series datasets.
-Slower processes, such as evolutionary selection on genotype frequencies, ecological selection on species abundances, and colonization and extinction, interact with fast processes to drive dynamics over the long-term.
-The influence of these slow processes is seen in very long time series, or in spatial gradients.
-Understanding dynamics at intermediate time scales requires integrating information from spatial and temporal data sources.
-We propose a model weighting approach; mechanistic spatiotemporal modeling is another alternative.
-The time scales shown here were chosen with vascular plants in mind, but the same concepts would apply for much shorter-lived organisms but at shorter time scales.
-}
-	\label{fig:fast-slow-concept}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=1 \textwidth] {species_patterns_models.png}
-	\caption{(A) Mean biomass by species (colors) across the temperature gradient during the baseline period.
-The focal species, dominant at the site in the center of the gradient (vertical gray line), is shown in dark blue.
-The dashed blue line shows predictions from the spatial statistical model.
-(B) Annual biomass of the focal species at the central site during the baseline period.
-The dashed line shows predictions from the temporal statistical model.
-}
-	\label{fig:species-patterns-models}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.7 \textwidth] {community_forecast_species.png}
-	\caption{(A) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts, at the focal site.
-(B) Simulated focal species biomass and forecasts from the spatial, temporal and weighted statistical models at the focal site in the metacommunity model.
-(C) Simulated changes in biomass of the focal species (black) and all other species (grey), and the weight given to the temporal statistical model for focal species biomass (blue).
-Time 1000 (years) in each panel corresponds to the start of the temperature increase.
-}
-	\label{fig:community-forecast-species}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.4 \textwidth]{spatial&temporal_model.png}
-	\caption{(A) Reaction norms of the three genotypes.
-(B) The spatial pattern of individual genotypes (colors) and total population abundance (black) at sites arrayed across a gradient of mean annual temperature.
-The dashed black line (almost entirely hidden by the slid black line) shows predictions from an empirical, spatial statistical model, a linear regression that describes mean population size as a function of mean temperature.
-(C) The relationship between annual temperature and per capita growth rate at a location with a mean temperature that favors the cold-adapted genotype.
-Colors show population size (the green to brown gradient depicting low to high population density), which influences the population growth rate through density dependence.
-}
-	\label{fig:evo_evo_spatial_temporal_models}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.7 \textwidth] {forecast.png}
-	\caption{(Top) Simulated annual temperatures (grey) and expected temperature (black), which was used to make forecasts.
-(Bottom) Simulated population size and forecasts from the spatial, temporal and weighted statistical models.
-}
-	\label{fig:forecast}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.7 \textwidth] {dispersal_niche_width.png}
-	\caption{The rate of change in the weight of the temporal forecast (y-axis) depends on (A) the fraction of propagules dispersing in the community turnover example and (B) on the temperature tolerance of genotypes, given by $\sigma_T$ (larger values indicate wider thermal niches) in the eco-evolutionary example.
-Time 0 (years) in these figures corresponds to the start of the temperature increase.
-}
-	\label{fig:dispersal_niche_width}
-\end{figure}
-
 %~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
@@
 %~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-\clearpage 
-\newpage
 
@@
 \end{equation}
-In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression is 
+In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression, which includes a quadratic effect of temperature, is 
 \begin{equation}
@@
 
+\renewcommand{\refname}{Literature cited}
+\bibliographystyle{ecography}
+\bibliography{references}
+
 \newpage

### 2020-07-08 doc/manuscript_acceptedJuly2020.tex
cleaning up

@@
 The three statistical models are described in Supporting Information (Appendix \ref{models}).
-Computer code for both case studies is available for reviewers online (https://drive.google.com/open?id=1aju04qtQvJmZG1mcpN-6hYBZILBhDfl2) and will be archived at Zenodo upon acceptance.
+Computer code for both case studies will be archived at Zenodo upon acceptance.
 Instructions for running the code are in the README file in the main directory of the zip archive.
@@
 
-\subsubsection*{Data accessibility statement:} The manuscript contains no original data.
-All computer code is available for reviewers in a zip archive that can be downloaded from Google Drive: https://drive.google.com/open?id=1aju04qtQvJmZG1mcpN-6hYBZILBhDfl2.
-Upon acceptance of the manuscript, code will be archived at Zenodo.
+\subsubsection*{Data accessibility statement:} The manuscript contains no original data. All computer code will be archived at Zenodo upon acceptance of the paper.
 

### 2020-07-08 community-turnover/example.m

@@ -1,68 +1,0 @@
-%
-%
-%	Example script to run the model of Alexander et al. 2018 'Lags in the response of mountain plant communities to climate change' - Global Change Biology
-%	Main script author: Loïc Chalmandrier - lchalman@uwyo.edu, https://github.com/LoicChr
-%
-
-
-% Sourcing the help functions
-addpath('./lib')
-
-%generation of the Landscape aka a table with the following characteristics: X coordinate, Y coordinate, temperature
-L_land = 500;
-landscape = [1:L_land ; ones(1 ,L_land); linspace(0,20, L_land)]';
-
-% Parameters to generate the species pool
-N = 100;
-Gmax = 0.5;
-Gmin = 0.2;
-Lmax = 1.5;
-Lmin = 0.7;
-Cmax = 0.2;
-Cmin = 0.2; 
-d = 10^-3;
- 
-% 1st step: initialisation of the metacommunity
-   Tmax = 40000; dt = 0.025;
-   spxp = ones(L_land, N).*(2/N);
-   d = d./dt;
-
- % Generation of the species pool
- tr = SpeciesPoolGen(N, 4, 16.5, Gmin, Gmax, Lmin, Lmax, Cmin, Cmax);
- temp = landscape(:,3);
- 
- % dispersal matrix
-   seed_rain =  squareform(pdist(landscape(:,1:2)) < 2).*(d*dt*0.5);
-   indx = logical( diag(ones(1,size(landscape,1)),0) );
-   seed_rain(indx) = 1-d*dt;
-   seed_rain(1,1) = 1-dt*d/2;
-   seed_rain(size(landscape,1),size(landscape,1)) = 1-dt*d/2;
-   
- %Run of the initialisation of community
- for t = 1:Tmax
-     spxp = CommunityTempDis(spxp, tr, seed_rain, temp, dt);
- end 
- spxp_ini = spxp;
-
- %save the initial metacommunity
- mkdir('simulations/run_1')
- dlmwrite(strcat('simulations/run_1/spxp_ini.txt'),spxp_ini)
-dlmwrite(strcat('simulations/run_1/landscape.txt'),landscape)
-dlmwrite(strcat('simulations/run_1/tr.txt'),tr)
-
-spxp_ini(spxp_ini < 10e-12) = 0;
-
-spxp = spxp_ini;
- %From the initial situation, re-run community dynamic while temperature is increasing.
- temp = landscape(:,3);
- time_to_max = 20000;
[…]
